### 2017-07-11 resex.tex
first start

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

%\usepackage[showframe,%Uncomment any one of the following lines to test 
%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
%%text={7in,10in},centering,
%%margin=1.5in,
%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
%%height=10in,a5paper,hmargin={3cm,0.8in},
%]{geometry}

\begin{document}

\title{Effects of pulsed hollow electron lens operation on the beam core in HL-LHC: \\First experimental studies and simulations.}% Force line breaks with \\
\thanks{Fermilab is operated by Fermi Research Alliance, LLC under
	Contract No.~DE-AC02-07CH11359 with the United States Department of
	Energy. This work was partially supported by the US DOE LHC
	Accelerator Research Program (LARP) and by the European FP7 HiLumi
	LHC Design Study, Grant Agreement 284404.}

\author{Miriam Fitterer}
 \email{mfittere@fnal.gov}
\author{Giulio Stancari}%
\author{Alexander Valishev}%
\affiliation{Fermi National Accelerator Laboratory, Batavia, Illinois, USA
% This line break forced with \textbackslash\textbackslash
}%

\author{Giulia Papotti}
\author{Stefano Redaelli}
\author{Daniel Valuch}
\affiliation{CERN, Geneva, Switzerland}%

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
In the HL-LHC a considerable amount of energy is stored in the beam tails due to the high beam intensity and an overpopulation of the tails compared to a Gaussian distribution. To control and clean the tail population, the installation of two hollow electron lenses, one per beam, is considered. Beside the DC operation, also a pulsed operation of the hollow electron lens is considered, which would considerably increase the diffusion speed by putting noise on the halo particles. In the ideal case, that is in case of no field at the beam core, only the halo particles are excited while leaving the core unperturbed. The picture though changes, if a residual field is present also at the location of the beam core putting noise also on the beam core. In this paper we present for estimates of the residual field at the beam core expected from the HL-LHC hollow electron lens and first experimental results of the effect of this excitation on the beam core together with the supporting simulations.
\end{abstract}

% PACS 2008:
% 29.20.db Storage rings and colliders

\pacs{29.20.D-}% PACS, the Physics and Astrtheonomy
                             % Classification Scheme.
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

%\tableofcontents

\section{\label{sec:intro}Introduction}%force line break with \protect\\
Looking back and forward at the last, current and future high energy and intensity accelerators, each new machine represents a considerable leap in stored beam energy with rising values for future accelerators and Colliders (see Table~\ref{tab:stored_energy})
\begin{table*}%The best place to locate the table environment is directly after its first reference in text
	\caption{\label{tab:stored_energy}%
		Stored beam energy for different past, present and future colliders. Each new machine represents a leap in stored beam energy.
	}
	\begin{ruledtabular}
		\begin{tabular}{lccccc}
			Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 & LHC nominal & HL-LHC & FCC\\
			\colrule
			Beam energy [GeV] & 980 & ? & ? & ? & \\
			Number of bunches & 36 & ? & ? & ? & \\
			Number of particles per bunch & $2.9\times 10^{11}$ & ? & ? & ? & \\
			Stored beam energy [MJ] & 1.6 & ? & ? & ? & \\
		\end{tabular}
	\end{ruledtabular}
\end{table*}

- why do we need active halo control -> refer to different methods -> say that e-lens is superior as shown in review
- layout
\begin{figure}[h]
	\includegraphics[width=1.0\linewidth]{hel_layout}% Here is how to import EPS art
	\caption{\label{fig:hel_layout} Layout of HEL \textbf{Ask Diego how to acknowledge correctly}.}
\end{figure}

\begin{acknowledgments}
We wish to acknowledge the support of the author community in using
REV\TeX{}, offering suggestions and encouragement, testing new versions,
\dots.
\end{acknowledgments}

\appendix

\section{Appendixes}

\bibliography{bibliography}% Produces the bibliography via BibTeX.

\end{document}
%
% ****** End of file apssamp.tex ******

### 2017-07-12 resex.tex
introduction

@@
 %rmp,
-%prstab,
+prstab,
 %prstper,
@@
 %\linenumbers\relax % Commence numbering lines
+\usepackage{color}
 
@@
 \section{\label{sec:intro}Introduction}%force line break with \protect\\
-Looking back and forward at the last, current and future high energy and intensity accelerators, each new machine represents a considerable leap in stored beam energy with rising values for future accelerators and Colliders (see Table~\ref{tab:stored_energy})
-\begin{table*}%The best place to locate the table environment is directly after its first reference in text
+Considering past, current and future high energy and intensity colliders, each new machine has represented a considerable leap in stored beam energy with rising values for future accelerators and colliders (see Table~\ref{tab:stored_energy}).
+\begin{table*}
 	\caption{\label{tab:stored_energy}%
@@
 		\begin{tabular}{lccccc}
-			Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 & LHC nominal & HL-LHC & FCC\\
+			Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{} \footnote{values from \url{https://lhc-commissioning.web.cern.ch/lhc-commissioning/performance/2016-performance.htm}}& LHC nominal \cite{lhc_design} & HL-LHC \cite{} \footnote{values from parameter and layout committee website \url{https://espace.cern.ch/HiLumi/PLC/default.aspx}}& FCC \cite{fcc_param_2017} \footnote{get right reference and values}\\
 			\colrule
-			Beam energy [GeV] & 980 & ? & ? & ? & \\
-			Number of bunches & 36 & ? & ? & ? & \\
-			Number of particles per bunch & $2.9\times 10^{11}$ & ? & ? & ? & \\
-			Stored beam energy [MJ] & 1.6 & ? & ? & ? & \\
+			Beam energy [TeV] & 0.98 & 6.5 & 7.0 & 7.0 & 50.0\\
+			Number of bunches & 36 & 2076 & 2808 & 2748 & ? \\
+			Number of particles per bunch & $2.90\times 10^{11}$ & $1.18\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
+			Stored beam energy [MJ] & 1.6 & 255.1 & 362.2 & 678.0 & 8400 \\
 		\end{tabular}
@@
 \end{table*}
+Recent measurements at the LHC furthermore show that the tails of the transverse beam distribution are overpopulated compared to a Gaussian distribution resulting in a considerable amount of energy being stored in the beam tails alone.
+In case of the LHC explicitly around 5\% of the beam population is stored in the tails above 3.5~$\sigma$ compared to 0.22\% in case of a Gaussian distribution leading to 19~MJ of stored energy in the tails in case of nominal LHC parameters and 34~MJ in case of HL-LHC~\cite{helreview_valentino}.
 
-- why do we need active halo control -> refer to different methods -> say that e-lens is superior as shown in review
-- layout
+All of the above reasons lead to the conclusion that a mechanism is needed to deplete the beam tails in a controlled manner (see for example \cite{helreview} in case of HL-LHC).
+The most obvious idea is to decrease the collimator gaps or scrape the tails with a collimator type device \textcolor{red}{Why is that not possible?}.
+This is however not possible due to impedance issues.
+Other mechanisms must therefore be found to actively deplete the tails.
+Most promising are methods, which considerably increase the diffusion speed in the region of the tail particles while leaving the beam core unperturbed.
+The diffusing tail particles are then intercepted by the collimation system and removed (see FIG.~\ref{fig:active_halo_control} for illustration).
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{passive_halo_control.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{active_halo_control.png}
+	\end{minipage}	
+	\caption{\label{fig:active_halo_control} Sketch of passive halo control as with the collimation system (top) and active halo control using in addition for example a hollow electron lens (HEL) to control the diffusion speed in the region tail region without affected the beam core (bottom).
+\textbf{Any acknowledgement needed for the plots?}.}
+\end{figure}
+In view of the need for LHC and HL-LHC in particular and in general for future high power accelerators, different methods have been studied in the last recent years \cite{helreview_bruce}, of which the HEL is considered the superior device at least in case of the HL-LHC \cite{helreview} and is also considered for other future high energy colliders like HE-LHC and FCC-hh \cite{,}.
+
+The concept of active halo control however breaks down if also the beam core is affected, which would ultimately lead to a degradation of the performance.
+In this paper, we will concentrate on this aspect for the HEL foreseen to be installed in the HL-LHC.
+We will summarize possible sources of perturbations of the beam core concentrating in particular on the case of pulsed operation and with the focus on the beam experiments at the LHC performed in the context of these studies.
+
+This paper is structured as follows: Section~\ref{sec:hel} gives an introduction to the concept of HELs and summarizes the design parameters of the HL-LHC HEL.
+Section~\ref{sec:core} is dedicated to describing the sources of a residual field from the HEL in the core region.
+Section~\ref{sec:exp} presents the results of the LHC experiment to study the effects on the beam core in case of a pulsed operation of the HEL, explicitly a resonant excitation.
+This includes the detailed analysis of the observed losses, emittance and transverse beam distribution changes.
+To the knowledge of the authors, the observed distribution changes presented in this paper have never been measured before in the context of a resonant excitation.
+A resonant excitation has been previously studied at the Tevatron in the context of the HEL experiments \cite{hel_tevatron_stancari} and the abort gap cleaning used in operation \cite{hel_tevatron_abortgap_zhang}.
+Both studies however only concentrated on the losses and emittance changes without measuring the detailed changes of the distribution.
+In addition the presented experiment also provides scaling of the losses and emittance growth with the excitation amplitude allowing for a comparison with simulations and ultimately an extrapolation to HL-LHC.
+
+\section{\label{sec:hel}Hollow electron lens for HL-LHC}
 \begin{figure}[h]
 	\includegraphics[width=1.0\linewidth]{hel_layout}% Here is how to import EPS art
-	\caption{\label{fig:hel_layout} Layout of HEL \textbf{Ask Diego how to acknowledge correctly}.}
+	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL \textbf{Ask Diego how to acknowledge correctly}.}
 \end{figure}
+
+\section{\label{sec:core}Sources of residual field in the proton beam core region}
+
+\section{\label{sec:exp}LHC experiment}
 

### 2017-07-14 resex.tex
completed introduction + section about estimate of kick at beam core

@@
 \usepackage{color}
+\usepackage{siunitx}
+\usepackage{url}
 
@@
 %]{geometry}
+
+\DeclareFontFamily{U}{wncy}{}
+\DeclareFontShape{U}{wncy}{m}{n}{<->wncyr10}{}
+\DeclareSymbolFont{mcy}{U}{wncy}{m}{n}
+\DeclareMathSymbol{\Sh}{\mathord}{mcy}{"58}
+
+\newcommand{\q}[2]{\ensuremath{#1\ \mathrm{#2}}} % quantity with units
 
@@
 Most promising are methods, which considerably increase the diffusion speed in the region of the tail particles while leaving the beam core unperturbed.
-The diffusing tail particles are then intercepted by the collimation system and removed (see FIG.~\ref{fig:active_halo_control} for illustration).
+The diffusing tail particles are then intercepted by the collimation system and removed (see Fig.~\ref{fig:active_halo_control} for illustration).
 \begin{figure}[h]
@@
 We will summarize possible sources of perturbations of the beam core concentrating in particular on the case of pulsed operation and with the focus on the beam experiments at the LHC performed in the context of these studies.
+As there is currently no hollow electron lens installed in the LHC, the kick on the beam core is emulated by a dipole kick applied with the transverse damper (see Section~\ref{sec:adt}).
+Explicitly, all orders higher than the first order are neglected.
 
@@
 \section{\label{sec:hel}Hollow electron lens for HL-LHC}
+Electron lenses in general consist of a DC or pulsed low energy electron beam, which is generated with an electron gun, then guided and confined by strong solenoids and finally collected with a collector.
+Exemplary for the conceptual design of all electron lenses, the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
 \begin{figure}[h]
@@
 \end{figure}
+The circulating beam, in case of the LHC the proton beam, is then affected by the electromagnetic field of the electron beam.
+For the application of active halo control, the electron beam needs to generate an electromagnetic field at the location of the halo particles while leaving the core untouched.
+This can be achieved by using a uniform hollow distribution in radius $r=\sqrt{x^2+y^2}$ with inner radius $R_1$ and outer radius $R_2$.
+In this case, the circulating proton beam experiences a radial kick $\theta(r)$
+\begin{equation}\label{eq:field_1}
+\theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max}
+\end{equation}
+where $f(r)$ is a shape function with
+\begin{equation}\label{eq:field_2}
+f(r) =
+\begin{cases} 0 &,\quad r< R_1\\
+\frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2\\
+1 &,\quad R_2 \leq r
+\end{cases}
+\end{equation}
+and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle with
+\begin{equation}
+\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{q}{p_0}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2}
+\end{equation}
+where $L$ is the length of the e-lens, $I_T$ the total electron beam current, $\beta_{e}$ and $\beta_{p}$ the relativistic $\beta$ of electron and proton beam, $\frac{q}{p_0}=\left(B\rho\right)_p$ the magnetic rigidity for the reference particle, $c$ the speed of light and $\epsilon_0$ the vacuum permittivity.
+\begin{figure}[b]
+	\includegraphics[width=0.8\linewidth]{hel_field}% Here is how to import EPS art
+	\caption{\label{fig:hel_field} Illustration of the hollow electron beam distribution (blue), the kick experienced by the proton beam (red) and the collimators (gray).}
+\end{figure}
+\begin{table}[t]
+	\caption{\label{tab:hel_param}%
+		HL-LHC hollow electron lens parameters as in \cite{hel_cdr}.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lcc}
+			Geometry & Value& Unit\\
+			\colrule
+		Length $L$    &  3 & m\\
+		Desired range of scraping positions\footnote{$\sigma$ denotes here the Beam sigma and not the collimation $\sigma$ which assumes a normalized emittance of 3.5~$\mu$m.} & 4-8 &$\sigma_p$\\
+		\colrule
+		Magnetic fields & & \\
+		\colrule
+		Gun solenoid, $B_g$ & 0.2-0.4 & T\\
+		Main solenoid, $B_m$ & 2-6 & T\\
+		Collector solenoid, $B_c$ & 0.2-0.4 & T\\
+		Compression factor ($k=\sqrt{B_m/B_g}$) & 2.2-5.5 & -\\
+		\colrule
+		Electron gun & & \\
+		\colrule
+		Peak yield $I_e$ at 10~keV & 5.0 & A\\
+		Gun perveance $P$ & 5 & $\mu$perv\\
+		Inner/outer cathode radius, $R_1/R_2$ & 6.75/12.7 & mm\\
+		\colrule
+		High voltage modulator & & \\
+		\colrule
+		Cathode-anode voltage & 10.0 & kV\\
+		Rise time (10\%-90\%) & 200 & ns \\
+		Repetition rate & 35 & kHz
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
 
-\section{\label{sec:core}Sources of residual field in the proton beam core region}
+The distribution together with the resulting kick are illustrated in Fig.~\ref{fig:hel_field}.
+The $\pm$-sign represents the two cases of the electron beam traveling in the direction of the proton beam ($v_e v_p>0$) leading to ``$-$" or in the opposite direction ($v_e v_p<0$) leading to ``$+$".
+For hollow electron beam collimation, electron and proton beam travel in opposite directions.
+Assuming HL-LHC and HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) the maximum kick of the HEL is:
+\begin{eqnarray}
+\theta_{\rm max,B1} = 392~\rm{nrad},\\
+\theta_{\rm max,B2} = 341~\rm{nrad}.
+\end{eqnarray}
+for an inner radius $R_1=4\sigma_p$, peak current $I_e=$\SI{5.0}{A}. Similar values are obtained for Beam~2.
+
+As depicted in Fig.~\ref{fig:hel_field} and in evidence from Eq.~\ref{eq:field_1}--\ref{eq:field_2}, the field at the beam core vanishes in the ideal case.
+Effects on the beam core thus only arise in case of imperfections, where possible sources are the bends of the HEL, as the electron beam crosses directly the proton beam, and distortions in the electron beam profile (see Section~\ref{sec:core}).
+Both sources result in non-linear kicks (see for example \cite{hel_bends_stancari,hel_model_polynomial_morozov} (\textcolor{red}{is there a better reference for the profile imperfections, updated note from Giulio???})).
+
+In DC operation of the HEL, the non-linear kicks from the HEL stay in the shadow of the non-linearities otherwise present in the machine, which are at injection energy mainly magnet errors and at collision energy the non-linear head-on and long-range beam-beam kicks.
+Tolerances on imperfections are therefore not particular stringent and not a concern.
+The picture however changes significantly in case of pulsed operation of the HEL, meaning that the electron gun voltage is modulated using a white or colored noise spectrum.
+If the electromagnetic field of the HEL does not vanish at the proton beam core, this noise is transferred not only to the halo particles, as intended, but also the beam core, with naturally much smaller amplitude.
+Tolerances in this case rapidly become much more stringent than in DC operation and studies of the effect of the HEL on the beam core are therefore focusing on this mode of operation.
+\begin{table}[t]
+	\caption{\label{tab:hllhc_param}%
+		HL-LHC design parameters at top energy \cite{} (\textcolor{red}{values from parameter and layout committee website - replace by reference} and parameters relevant in connection with the HEL.
+Optics parameters at HEL are based on a position of the HEL of $-40$~m for Beam~1 (B1) and $+40$~m for Beam~2 (B2) from IP4 using HL-LHC optics V1.3 \cite{hllhcv13} with $\beta^{*}=0.15$~m.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lccc}
+			Beam parameters & Value(B1) & Value(B2) & Unit\\
+			\colrule
+			Beam energy  $E_{p}$  &  \multicolumn{2}{c}{7} & TeV\\
+			Number of bunches $n_b$ & \multicolumn{2}{c}{2748} & - \\
+			Number of particles per bunch $N_b$ & \multicolumn{2}{c}{$2.2\times 10^{11}$} & -\\
+			Normalized emittance $\epsilon_{N,x/y}$ & \multicolumn{2}{c}{2.5} & $\mu$m\\
+			Bunch spacing & \multicolumn{2}{c}{25} & ns\\
+			\colrule
+			\multicolumn{4}{l}{Optics paramters at HEL (Beam~1) \footnote{As the twiss parameters at IP4 do not change during the entire squeeze, and IP4 and the HEL are only separated by a drift space, the twiss parameters stay constant also at the HEL during the entire squeeze.}} \\
+			\colrule
+			$\beta_{x}$ at HEL  & 197.5 & 280.6 & m\\
+			$\beta_{y}$ at HEL & 211.9 & 262.6 m\\
+			Dispersion $D_{x}$ at HEL & 0.0& 0.0 & m\\
+			Dispersion $D_{y}$ at HEL & 0.0& 0.0 & m\\
+			Proton beam size $\sigma_{p,x}$ at HEL & 0.26 & 0.31& mm \\			
+			Proton beam size $\sigma_{p,y}$ at HEL & 0.27 & 0.30 &mm \\
+			\multicolumn{4}{l}{scraping position}\\ \hspace{1cm}$\sigma_{p}=\max(\sigma_{p,x},\sigma_{p,y})$ & 0.27& 0.31 & mm\\
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
+
+Pulsing the HEL is currently considered as optional mode of operation to further increase the halo diffusion rates if needed.
+Pulsing increases the halo diffusion rate as in addition to the non-linear kick, noise is exerted on the halo particles of the proton beam.
+Two different pulsing patterns are being considered at the moment:
+\begin{itemize}
+	\item \textbf{random excitation:} The electron gun voltage is modulated between:
+	\begin{eqnarray}
+		U_{\mathrm{e-gun}}&=&a\cdot U_{\mathrm{max}}\\
+		& &+(1-a)\cdot \mathrm{ran}(0,1)\cdot U_{\mathrm{max}}
+	\end{eqnarray}
+	with $U_{\mathrm{max}}$ the equivalent voltage in DC operation, $a$ the modulation strength with $a\in[0,1]$, and $\mathrm{ran}(0,1)$ a uniformly distributed random number between~[0,1].
+	\item \textbf{resonant excitation:} The HEL is switched on only every $k^{\mathrm{th}}$ turn. The excitation can be represented by:
+	\begin{eqnarray}\label{intro:eqn:1}
+	f(t)&=&\sum_{p=-\infty}^{+\infty}\delta(t-n\cdot(kT)),
+	\end{eqnarray}
+	where $p$ is the turn number and $T$ the revolution time. The Fourier series is then given by:
+	\begin{eqnarray}\label{intro:eqn:2}
+	f(t)=\Sh_{kT}(t)
+	&=&\frac{1}{kT}\sum_{n=-\infty}^{+\infty}e^{2\pi i f_nt} \\
+	& & \quad \text{with} \ f_n=\frac{n}{k}f_{\rm rev}.
+	\end{eqnarray}
+	where $\Sh_{kT}$ is the Dirac comb.
+As can be seen from th Fourier spectrum, a $k^{\mathrm{th}}$~turn pulsing in general drives $k^{\mathrm{th}}$ order resonances (see \cite{md_sim_hel_res_ex_fitterer} for examples) and was used in regular operation for abort cleaning in the Tevatron \cite{hel_tevatron_abortgap_zhang}.
+\end{itemize}
+In the LHC experiment and the preparatory simulations presented in this paper, only the effect of a resonant excitation was studied as analytical as well as experimental results already exist for a random excitation equivalent to the case of exerting white noise on the beam \cite{noise_2007_ohmi,noise_alexahin_lhc,noise_lebedev_ssc,noise_2014_ohmi,md1433_noise_top_energy,md400_noise_injection}.
+Furthermore, the non-linear kick is approximated only to first order, i.e.
+by a dipole kick.
+The expected amplitude of the dipole kick is derived in Section~\ref{sec:core}.
+In future simulation studies, the dipole kick will have to be substituted with the expected full non-linear kick.
+For the HEL bends, a non-linear symplectic map has already been derived \cite{hel_bends_stancari}, however the kick is estimated to be small compared to the one expected from profile imperfections (see Section~\ref{fig:hel_field}).
+A model for the non-linear kick from profile imperfections is currently under study based on the approach taken in~\cite{hel_model_polynomial_morozov}.
+
+\section{\label{sec:core}Sources of residual field in the proton beam core region and first order estimates}
+With the current e-lens layout (see Fig.~\ref{fig:hel_layout}), parasitic kicks on the proton beam core can arise in the central region (main solenoid) due to profile imperfections in the electron beam and at the entrance and exit of the HEL, where electron and proton beam directly overlap.
+
+As there is currently no HEL installed in the LHC, the kick on the proton beam core in pulsed operation can experimentally only be approximated to first order by applying a dipole kick with the corresponding excitation pattern using the LHC transverse damper (see Section~\ref{sec:adt}).
+As derived in Sec.~\ref{core:sec:1}--\ref{core:sec:2} based on the measurements of the newest electron gun prototype \cite{} and in more detail in \cite{md_sim_hel_res_ex_fitterer} for the first electron gun prototype \cite{}, the expected kick to first order is for the HEL bends~(Eqn.~\ref{eqn:kick_bends}):
+\begin{equation}
+	\Delta x'_{\rm{bends}},\Delta y'_{\rm{bends}}\leq \SI{0.5}{nrad},
+\end{equation}
+and for the central region due to profile imperfections~(Eqn.~\ref{}):
+\begin{equation}
+	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \leq \SI{15.0}{nrad}
+\end{equation}
+where the current HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) were assumed yielding the maximum kick, explicitly $B_{m}=\SI{5}{T}$, $I_e=\q{5.0}{A}$, $E_{e} = \SI{10.0}{keV}$, $L=\SI{3}{m}$ and $E_{p} = \SI{7.0}{TeV}$.
+The contribution from the central region is clearly dominating.
+
+\subsection{Uncompensated kicks from HEL bends}
+\label{core:sec:1}
+The symplectic map for the HL-LHC HEL bends is derived in detail in \cite{hel_bends_stancari}.
+In this case the e-lens bends are modeled as a cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons bend in the horizontal plane.
+In case of a S-shaped HEL, the transverse dipole kicks at entrance and exit compensate each other to first order, in case of an U-shaped HEL, they would add up.
+Out of this reason a S-shape has been chosen for the HL-LHC HEL (see Fig.~\ref{fig:hel_layout}).
+Uncompensated kicks therefore only arise due to imperfections.
+As a first estimate, we assume 10\% fluctuations between the entrance and exit, which can originate from for example profile imperfections and current fluctuations.
+Furthermore, we assume, that the kicks from entrance and exit due to these imperfections add up.
+Using the electric field calculations in \cite{hel_bends_stancari}, the maximum values for the integrated electric field are around
+\begin{equation}
+\int_{z_1}^{z_2} E_{x,y} dz= 10 \ \mathrm{kV}
+\end{equation}
+assuming an \SI{1}{A}, \SI{5}{keV} electron beam. For \SI{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
+\begin{equation}
+\Delta x'=\Delta y'= \SI{1.4}{nrad},
+\end{equation}
+Scaling to the HEL design parameters of a \SI{5}{A} and \SI{10}{keV} electron beam (Table~\ref{tab:hel_param}), one obtains:
+\begin{equation}
+\int_{z_1}^{z_2} E_{x} dz= 36 \ \mathrm{kV} \Rightarrow \Delta x'= \q{5.1}{nrad}
+\end{equation}
+Assuming now 10\% fluctuation between entrance and exit kick, the expected kick from the HEL bends is
+\begin{equation}\label{eqn:kick_bends}
+\Delta x', \Delta y'= \q{0.5}{nrad}.
+\end{equation}
+
+\subsection{Kicks in the central region (main solenoid)}
+\label{core:sec:2}
+For a perfectly uniform, annular and radially symmetric profile, the field in the region of the proton beam core vanishes (see Eqn.~\ref{eq:field_2}).
+In case of electron beam profile imperfections the radial symmetry is broken, leading to a residual field at the beam core.
+Fig.~\ref{core:fig:1} shows an example of the electric field calculated from a measured asymmetric profile.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.0\linewidth]{e-beam_profile.png}
+	\caption{Calculated hollow electron beam field from measured profile of 9~kV, 2.49~A e-gun and different main solenoidal fields $Bm$.
+The field has been calculated using the code WARP \cite{warp}.
+\textcolor{red}{change $Bm$ to $B_m$ +substitute with new picture.}}
+	\label{core:fig:1}
+\end{figure}
+A first estimate of the residual kick for the HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) with a main solenoid field of $B_{m}=\SI{5}{T}$, $I_e=\SI{5.0}{A}$, $E_{e} = \SI{10.0}{keV}$, $L=\SI{3}{m}$ and $E_{p} = \SI{7.0}{TeV}$ can be obtained by scaling the electric field at the center as obtained in Fig.~\ref{core:fig:1} with the main solenoid field $B_{m}$ and the electron beam current $I_e$ and energy $E_{e}$, yielding:
+\begin{equation}
+\Delta x', \Delta y'=\q{15}{nrad}
+\end{equation}
+\textcolor{red}{New gun measurements, can we maybe have a WARP simulation for the new design parameters instead of scaling?}
+
+\section{\label{sec:adt}Excitation with transverse damper}
 
@@
 \begin{acknowledgments}
-We wish to acknowledge the support of the author community in using
-REV\TeX{}, offering suggestions and encouragement, testing new versions,
-\dots.
+We wish to acknowledge Roderik Bruce and Gianluca Valentino for there very helpful suggestions in preparing the experiment at the LHC and we would like to thank all participants of the experiment for helping acquire the data presented in this paper.
+We would like to acknowledge Dmitry Shatilov for his help with the Lifetrac code, and St\'{e}phane Fartoukh, Riccardo De Maria and Rogelio Tomás for their help with generating the appropiate optics model.
+We are grateful also for the support from the beam instrumentation team, in particular Enrico Bravin and Georges Trad, with the BSRT and are thankful for the collaboration with St\'{e}phania Papadopoulou and Fanouria Antonio for the analysis of the BSRT profiles.
 \end{acknowledgments}

### 2017-07-14 scripts/optics_kick_hel.py

@@ -1,0 +1,75 @@
+
+from pyoptics import *
+from matplotlib.pyplot import *
+from numpy import *
+from beams import *
+
+#simulation parameters
+EGeV=7000
+epsn0=2.5 # emittance in mum
+eps0=beamparam.emitrms(epsn0,EGeV) # geometric emittance in mum
+print eps0
+
+I=5.0 # A
+
+#calculate the energy spread
+beam='b2'
+fn = 'twiss_lhc%s_ref.tfs'%beam
+opt=optics.open(fn) # ref for reference file
+print 'optics file = %s'%(fn)
+print 'beta_x/y(IP5)=%2.4f/%2.4f m'%(opt.betx[opt//'IP5'],opt.bety[opt//'IP5'])
+print 'p_x/y(IP5)=%4.2f/%4.2f urad'%(opt.px[opt//'IP5']*1.e6,opt.py[opt//'IP5']*1.e6)
+
+# plot optics, HEL is marked with hel1
+opt4=opt.select('E.DS.L4.%s'%beam.upper(),'S.DS.R4.%s'%beam.upper(),shift=False)
+hel_name='HEBC'
+sip4=opt4.s[opt4//'IP4']
+
+shebip4=opt4.s[opt4//hel_name]-sip4
+close('all')
+opt4.plotbeta()
+plot(list((opt4.s[opt4//hel_name],)*2),[-30,30],'k--',label='HEL: %s m'%shebip4)
+plot(sip4,0,marker='o',color='orange',label='IP4')
+handles,labels=gca().get_legend_handles_labels()
+legend(handles[-2:],labels[-2:])
+savefig('opthelip4%s.png'%beam,bbox_inches='tight')
+
+# calculate e-lens parameters
+sigx=sqrt(eps0*1.e-6*opt4.betx[opt4//hel_name]) # m
+sigy=sqrt(eps0*1.e-6*opt4.bety[opt4//hel_name]) # m
+sig=max([sigx,sigy])
+r1=4*sig # m
+## parameters CERN gun
+r1gun=6.75e-3 # m
+r2gun=12.7e-3 # m
+r2ovr1 = r2gun/r1gun # r2/r1, r2 = outer radius, r1 = inner radius
+r2=r2ovr1*r1
+l=3 # m
+EkeV_e=10 # keV
[…]

### 2017-07-14 resex.tex
add structure for simulation and experimental results

@@
 \end{equation}
-and for the central region due to profile imperfections~(Eqn.~\ref{}):
+and for the central region due to profile imperfections~(Eqn.~\ref{eqn:kick_central}):
 \begin{equation}
@@
 A first estimate of the residual kick for the HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) with a main solenoid field of $B_{m}=\SI{5}{T}$, $I_e=\SI{5.0}{A}$, $E_{e} = \SI{10.0}{keV}$, $L=\SI{3}{m}$ and $E_{p} = \SI{7.0}{TeV}$ can be obtained by scaling the electric field at the center as obtained in Fig.~\ref{core:fig:1} with the main solenoid field $B_{m}$ and the electron beam current $I_e$ and energy $E_{e}$, yielding:
-\begin{equation}
+\begin{equation}\label{eqn:kick_central}
 \Delta x', \Delta y'=\q{15}{nrad}
@@
 
-\section{\label{sec:exp}LHC experiment}
+\section{\label{sec:exp_sum}Overview of LHC experiment}
+
+\section{\label{sec:sim}Summary of simulation results}
+
+\section{\label{sec:exp}Summary of experimental results}
+
+\section{\label{sec:exp_sim}Comparison of simulation and experimental results}
 

### 2017-12-19 resex.tex
update 18 Dec 2017

@@
 
-\section{\label{sec:intro}Introduction}%force line break with \protect\\
+\section{Introduction\label{sec:intro}}%force line break with \protect\\
 Considering past, current and future high energy and intensity colliders, each new machine has represented a considerable leap in stored beam energy with rising values for future accelerators and colliders (see Table~\ref{tab:stored_energy}).
@@
 		\begin{tabular}{lccccc}
-			Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{} \footnote{values from \url{https://lhc-commissioning.web.cern.ch/lhc-commissioning/performance/2016-performance.htm}}& LHC nominal \cite{lhc_design} & HL-LHC \cite{} \footnote{values from parameter and layout committee website \url{https://espace.cern.ch/HiLumi/PLC/default.aspx}}& FCC \cite{fcc_param_2017} \footnote{get right reference and values}\\
+			Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{chamonix2017param}
+			& LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{fcc_param_2017} \\
 			\colrule
 			Beam energy [TeV] & 0.98 & 6.5 & 7.0 & 7.0 & 50.0\\
-			Number of bunches & 36 & 2076 & 2808 & 2748 & ? \\
-			Number of particles per bunch & $2.90\times 10^{11}$ & $1.18\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
-			Stored beam energy [MJ] & 1.6 & 255.1 & 362.2 & 678.0 & 8400 \\
+			Number of bunches & 36 & 2220 & 2808 & 2748 & ? \\
+			Number of particles per bunch & $2.90\times 10^{11}$ & $1.15\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
+			Stored beam energy [MJ] & 1.6 & 265.9 & 362.2 & 678.0 & 8400 \\
 		\end{tabular}
@@
 All of the above reasons lead to the conclusion that a mechanism is needed to deplete the beam tails in a controlled manner (see for example \cite{helreview} in case of HL-LHC).
-The most obvious idea is to decrease the collimator gaps or scrape the tails with a collimator type device \textcolor{red}{Why is that not possible?}.
-This is however not possible due to impedance issues.
+The most obvious idea is to decrease the collimator gaps or scrape the tails with a collimator type device.
+The minimum gap size of the collimators is however limited by the impedance of the machine and small enough gap sizes can therefore not be reached without risking an instability of the beam.
 Other mechanisms must therefore be found to actively deplete the tails.
-Most promising are methods, which considerably increase the diffusion speed in the region of the tail particles while leaving the beam core unperturbed.
-The diffusing tail particles are then intercepted by the collimation system and removed (see Fig.~\ref{fig:active_halo_control} for illustration).
+
+Most promising are methods, which considerably increase the diffusion speed in the region of the halo particles while leaving the beam core unperturbed.
+The diffusing halo particles are then intercepted by the collimation system and removed (see Fig.~\ref{fig:active_halo_control} for illustration).
+This concept is also referred to as active halo control compared to a passive system, which only intercepts the halo particles without actively controlling the diffusion speed.
+Currently, the LHC collimation system is a passive halo control system.
 \begin{figure}[h]
@@
 	\caption{\label{fig:active_halo_control} Sketch of passive halo control as with the collimation system (top) and active halo control using in addition for example a hollow electron lens (HEL) to control the diffusion speed in the region tail region without affected the beam core (bottom).
-\textbf{Any acknowledgement needed for the plots?}.}
+\textcolor{red}{Any acknowledgement needed for the plots?}.}
 \end{figure}
-In view of the need for LHC and HL-LHC in particular and in general for future high power accelerators, different methods have been studied in the last recent years \cite{helreview_bruce}, of which the HEL is considered the superior device at least in case of the HL-LHC \cite{helreview} and is also considered for other future high energy colliders like HE-LHC and FCC-hh \cite{,}.
 
-The concept of active halo control however breaks down if also the beam core is affected, which would ultimately lead to a degradation of the performance.
-In this paper, we will concentrate on this aspect for the HEL foreseen to be installed in the HL-LHC.
-We will summarize possible sources of perturbations of the beam core concentrating in particular on the case of pulsed operation and with the focus on the beam experiments at the LHC performed in the context of these studies.
-As there is currently no hollow electron lens installed in the LHC, the kick on the beam core is emulated by a dipole kick applied with the transverse damper (see Section~\ref{sec:adt}).
-Explicitly, all orders higher than the first order are neglected.
+In view of the need of an active halo control system for LHC and HL-LHC in particular and in general for future high power accelerators, different methods have been studied in the last recent years \cite{helreview_bruce}, of which the HEL is considered the superior device at least in case of the HL-LHC \cite{helreview} and is also considered for other future high energy colliders like HE-LHC and FCC-hh \cite{helhcparam2011,fccparam2016}.
+
+The beneficial effect of an active halo control system is however abrogated if also the beam core is affected as this would ultimately lead to a degradation of the performance due to losses of particles in the core region and emittance growth.
+In this paper, we will concentrate on this aspect in view of future HEL to be installed in the HL-LHC.
+We will summarize possible sources of perturbations of the beam core concentrating in particular on the case of pulsed operation and with the focus on the beam experiments at the LHC performed in this context.
 
@@
 Section~\ref{sec:core} is dedicated to describing the sources of a residual field from the HEL in the core region.
-Section~\ref{sec:exp} presents the results of the LHC experiment to study the effects on the beam core in case of a pulsed operation of the HEL, explicitly a resonant excitation.
+Section~\ref{sec:exp_sum}--\ref{sec:exp_sim} describes then in detail the two beam experiments conducted at the LHC to study the effect on the beam core in case of a pulsed HEL operation.
 This includes the detailed analysis of the observed losses, emittance and transverse beam distribution changes.
@@
 A resonant excitation has been previously studied at the Tevatron in the context of the HEL experiments \cite{hel_tevatron_stancari} and the abort gap cleaning used in operation \cite{hel_tevatron_abortgap_zhang}.
-Both studies however only concentrated on the losses and emittance changes without measuring the detailed changes of the distribution.
-In addition the presented experiment also provides scaling of the losses and emittance growth with the excitation amplitude allowing for a comparison with simulations and ultimately an extrapolation to HL-LHC.
+Both studies however only concentrated on the losses and emittance changes without measuring the detailed changes of the distribution itself.
+In addition, the presented experiments also provide scalings of the losses and emittance growth with the excitation amplitude.
+This allows to provide first tolerances for the residual field of the HL-LHC HEL at the beam core and also to compare the obtained results with simulations and thus benchmark the same.
 
-\section{\label{sec:hel}Hollow electron lens for HL-LHC}
-Electron lenses in general consist of a DC or pulsed low energy electron beam, which is generated with an electron gun, then guided and confined by strong solenoids and finally collected with a collector.
+\section{Hollow electron lens for HL-LHC\label{sec:hel}}
+\subsection{General Overview\label{sec:hel:intro}}
+In general, electron lenses are DC or pulsed low energy electron beams.
+The electron beam is generated in an electron gun, then guided and confined by strong solenoids and finally dumped onto a collector.
 Exemplary for the conceptual design of all electron lenses, the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
 \begin{figure}[h]
-	\includegraphics[width=1.0\linewidth]{hel_layout}% Here is how to import EPS art
-	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL \textbf{Ask Diego how to acknowledge correctly}.}
+	\includegraphics[width=1.0\linewidth]{hel_layout_epbeam}% Here is how to import EPS art
+	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL.}
 \end{figure}
@@
 \begin{equation}\label{eq:field_1}
-\theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max}
+\theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max},
 \end{equation}
@@
 f(r) =
-\begin{cases} 0 &,\quad r< R_1\\
-\frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2\\
+\begin{cases} 0 &,\quad r< R_1,\\
+\frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2,\\
 1 &,\quad R_2 \leq r
@@
 \end{equation}
-and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle with
+and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle given by
 \begin{equation}
-\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{q}{p_0}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2}
+\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{q}{p_0}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2},
 \end{equation}
@@
 	\caption{\label{tab:hel_param}%
-		HL-LHC hollow electron lens parameters as in \cite{hel_cdr}.
+		HL-LHC hollow electron lens parameters as in~\cite{hel_cdr}.
 	}
@@
 Assuming HL-LHC and HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) the maximum kick of the HEL is:
-\begin{eqnarray}
-\theta_{\rm max,B1} = 392~\rm{nrad},\\
-\theta_{\rm max,B2} = 341~\rm{nrad}.
-\end{eqnarray}
-for an inner radius $R_1=4\sigma_p$, peak current $I_e=$\SI{5.0}{A}. Similar values are obtained for Beam~2.
+\begin{equation}
+\theta_{\rm max,B1} = 392~\rm{nrad}
+\end{equation}
+%\begin{eqnarray}
+%\theta_{\rm max,B1} = 392~\rm{nrad},\\
+%\theta_{\rm max,B2} = 341~\rm{nrad}.
+%\end{eqnarray}
+for an inner radius $R_1=4\sigma_p$, outer radius $R_2=7.5\sigma_p$ and a peak current of $I_e=$\SI{5.0}{A}. Similar values are obtained for Beam~2.
 
+\subsection{Operation modes and effects on the beam core\label{sec:hel:core}}
 As depicted in Fig.~\ref{fig:hel_field} and in evidence from Eq.~\ref{eq:field_1}--\ref{eq:field_2}, the field at the beam core vanishes in the ideal case.
@@
 
-In DC operation of the HEL, the non-linear kicks from the HEL stay in the shadow of the non-linearities otherwise present in the machine, which are at injection energy mainly magnet errors and at collision energy the non-linear head-on and long-range beam-beam kicks.
+In DC operation of the HEL, the non-linear kicks from the HEL stay in the shadow of the non-linearities otherwise present in the machine.
 Tolerances on imperfections are therefore not particular stringent and not a concern.
-The picture however changes significantly in case of pulsed operation of the HEL, meaning that the electron gun voltage is modulated using a white or colored noise spectrum.
-If the electromagnetic field of the HEL does not vanish at the proton beam core, this noise is transferred not only to the halo particles, as intended, but also the beam core, with naturally much smaller amplitude.
+The picture however changes significantly in case of pulsed operation of the HEL, in which case the electron gun voltage is modulated using a white or colored noise spectrum.
+If the electromagnetic field of the HEL does not vanish at the proton beam core, noise is transferred not only to the halo particles, as intended, but also to the beam core with naturally much smaller amplitude.
 Tolerances in this case rapidly become much more stringent than in DC operation and studies of the effect of the HEL on the beam core are therefore focusing on this mode of operation.
@@
 	\caption{\label{tab:hllhc_param}%
-		HL-LHC design parameters at top energy \cite{} (\textcolor{red}{values from parameter and layout committee website - replace by reference} and parameters relevant in connection with the HEL.
-Optics parameters at HEL are based on a position of the HEL of $-40$~m for Beam~1 (B1) and $+40$~m for Beam~2 (B2) from IP4 using HL-LHC optics V1.3 \cite{hllhcv13} with $\beta^{*}=0.15$~m.
+		HL-LHC design parameters at top energy \cite{hlcdr} and parameters relevant in connection with the HEL.
+Optics parameters at HEL are based on a position of the HEL of $-40$~m for Beam~1 (B1) and $+40$~m for Beam~2 (B2) from IP4 using HL-LHC optics V1.3 with $\beta^{*}=0.15$~m \cite{hlv13}.
 	}
@@
 
-Pulsing the HEL is currently considered as optional mode of operation to further increase the halo diffusion rates if needed.
-Pulsing increases the halo diffusion rate as in addition to the non-linear kick, noise is exerted on the halo particles of the proton beam.
-Two different pulsing patterns are being considered at the moment:
+The main benefit of pulsing the HEL is to increase the halo removal rate in case a fast depletion of the halo is desired.
+The halo diffusion rate is increased in this case as in addition to the non-linear kick, noise is exerted on the halo particles of the proton beam.
+For the HL-LHC two different pulsing patterns are currently being considered:
 \begin{itemize}
@@
 	with $U_{\mathrm{max}}$ the equivalent voltage in DC operation, $a$ the modulation strength with $a\in[0,1]$, and $\mathrm{ran}(0,1)$ a uniformly distributed random number between~[0,1].
-	\item \textbf{resonant excitation:} The HEL is switched on only every $k^{\mathrm{th}}$ turn. The excitation can be represented by:
+	\item \textbf{resonant excitation:} The HEL is switched on only every $k^{\mathrm{th}}$ turn. The excitation can then be represented by:
 	\begin{eqnarray}\label{intro:eqn:1}
@@
 	\end{eqnarray}
-	where $p$ is the turn number and $T$ the revolution time. The Fourier series is then given by:
+	where $p$ is the turn number and $T$ the revolution time, and its Fourier series by:
 	\begin{eqnarray}\label{intro:eqn:2}
@@
 	&=&\frac{1}{kT}\sum_{n=-\infty}^{+\infty}e^{2\pi i f_nt} \\
-	& & \quad \text{with} \ f_n=\frac{n}{k}f_{\rm rev}.
+	& & \quad \text{with} \ f_n=\frac{n}{k}f_{\rm rev},
 	\end{eqnarray}
 	where $\Sh_{kT}$ is the Dirac comb.
-As can be seen from th Fourier spectrum, a $k^{\mathrm{th}}$~turn pulsing in general drives $k^{\mathrm{th}}$ order resonances (see \cite{md_sim_hel_res_ex_fitterer} for examples) and was used in regular operation for abort cleaning in the Tevatron \cite{hel_tevatron_abortgap_zhang}.
+As can be seen from the Fourier series, $k^{\mathrm{th}}$~turn pulsing in general drives $k^{\mathrm{th}}$ order resonances \cite{md_sim_hel_res_ex_fitterer}.
+Historically, the $k^{\mathrm{th}}$~turn pulsing was used in regular operation in the Tevatron for abort cleaning \cite{hel_tevatron_abortgap_zhang}.
 \end{itemize}
-In the LHC experiment and the preparatory simulations presented in this paper, only the effect of a resonant excitation was studied as analytical as well as experimental results already exist for a random excitation equivalent to the case of exerting white noise on the beam \cite{noise_2007_ohmi,noise_alexahin_lhc,noise_lebedev_ssc,noise_2014_ohmi,md1433_noise_top_energy,md400_noise_injection}.
-Furthermore, the non-linear kick is approximated only to first order, i.e.
-by a dipole kick.
+In general, the random excitation mode is the most effective and operationally easiest pulsing mode, but also the most stringent in terms of tolerances on the residual field at the beam core as it induces white noise and thus excites all beam frequencies.
+As the resonant excitation acts on specific resonances, the pulsing pattern has to be chosen according to the machine and beam parameters.
+On the one hand, this makes the operation more difficult as one has to find the most efficient $k^{\mathrm{th}}$~turn pulsing for each operation mode, on the other hand, it also relaxes the tolerances on the residual field at the beam core as, depending on the machine and beam configuration, it might be possible to find a resonant excitation pattern, which mainly affects the halo and not the core.
+%In the LHC experiment and the preparatory simulations presented in this paper, only the effect of a resonant excitation was studied as analytical as well as experimental results already exist for a random excitation equivalent to the case of exerting white noise on the beam \cite{noise_2007_ohmi,noise_alexahin_lhc,noise_lebedev_ssc,noise_2014_ohmi,md1433_noise_top_energy,md400_noise_injection}.
+Furthermore, the
+%As no HEL is currently installed in the LHC, the non-linear kick of the HEL is approximated only to first order, i.e.
+by a dipole kick, during the experiments.
 The expected amplitude of the dipole kick is derived in Section~\ref{sec:core}.
@@
 
-\section{\label{sec:core}Sources of residual field in the proton beam core region and first order estimates}
-With the current e-lens layout (see Fig.~\ref{fig:hel_layout}), parasitic kicks on the proton beam core can arise in the central region (main solenoid) due to profile imperfections in the electron beam and at the entrance and exit of the HEL, where electron and proton beam directly overlap.
+\section{Sources of residual field in the proton beam core region and first order estimates\label{sec:core}}
+With the current HEL layout (see Fig.~\ref{fig:hel_layout}), parasitic kicks on the proton beam core can arise in the central region (main solenoid) due to profile imperfections in the electron beam and at the entrance and exit of the HEL, where electron and proton beam overlap.
+Currently, no HEL is installed in the LHC, and the kick on the proton beam core must therefore be approximated to first order by a dipole kick for the experiments at the LHC.
+Explicitly, the dipole kick with the corresponding excitation pattern is applied with the LHC transverse damper (see Section~\ref{sec:adt}).
 
-As there is currently no HEL installed in the LHC, the kick on the proton beam core in pulsed operation can experimentally only be approximated to first order by applying a dipole kick with the corresponding excitation pattern using the LHC transverse damper (see Section~\ref{sec:adt}).
-As derived in Sec.~\ref{core:sec:1}--\ref{core:sec:2} based on the measurements of the newest electron gun prototype \cite{} and in more detail in \cite{md_sim_hel_res_ex_fitterer} for the first electron gun prototype \cite{}, the expected kick to first order is for the HEL bends~(Eqn.~\ref{eqn:kick_bends}):
+The expected kick to first order from the HEL bends~(Eqn.~\ref{eqn:kick_bends}) is
 \begin{equation}
@@
 \end{equation}
-where the current HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) were assumed yielding the maximum kick, explicitly $B_{m}=\SI{5}{T}$, $I_e=\q{5.0}{A}$, $E_{e} = \SI{10.0}{keV}$, $L=\SI{3}{m}$ and $E_{p} = \SI{7.0}{TeV}$.
-The contribution from the central region is clearly dominating.
-
+where the current HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) are used.
+% yielding the maximum kick, explicitly $B_{m}=\SI{5}{T}$, $I_e=\q{5.0}{A}$, $E_{e} = \SI{10.0}{keV}$, $L=\SI{3}{m}$ and $E_{p} = \SI{7.0}{TeV}$.
+The contribution from the central region is clearly dominating. A more detailed derivation of both values is given in Sec.~\ref{core:sec:1}--\ref{core:sec:2}.
 \subsection{Uncompensated kicks from HEL bends}
 \label{core:sec:1}
-The symplectic map for the HL-LHC HEL bends is derived in detail in \cite{hel_bends_stancari}.
-In this case the e-lens bends are modeled as a cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons bend in the horizontal plane.
-In case of a S-shaped HEL, the transverse dipole kicks at entrance and exit compensate each other to first order, in case of an U-shaped HEL, they would add up.
-Out of this reason a S-shape has been chosen for the HL-LHC HEL (see Fig.~\ref{fig:hel_layout}).
-Uncompensated kicks therefore only arise due to imperfections.
-As a first estimate, we assume 10\% fluctuations between the entrance and exit, which can originate from for example profile imperfections and current fluctuations.
-Furthermore, we assume, that the kicks from entrance and exit due to these imperfections add up.
-Using the electric field calculations in \cite{hel_bends_stancari}, the maximum values for the integrated electric field are around
+As reference for the estimate of the dipole component originating from the HEL bends, the symplectic map derived in \cite{hel_bends_stancari} is used.
+In this case the HEL bends are modeled as a cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons bend in the horizontal plane.
+In case of an U-shaped HEL, the transverse dipole kicks at entrance and exit add up, while for a S-shaped HEL the kicks compensated each other.
+For this reason a S-shape has been chosen for the HL-LHC HEL (see Fig.~\ref{fig:hel_layout}).
+In this case, uncompensated kicks therefore only arise from imperfections, which can originate from profile imperfections and current fluctuations.
+As a first rough estimate, we assume in this paper 10\% fluctuations between the entrance and exit and that the kicks from entrance and exit due to these imperfections add up.
+Using the electric field calculations in \cite{hel_bends_stancari}, the maximum values for the integrated electric field are then around
 \begin{equation}
@@
 \end{equation}
-assuming an \SI{1}{A}, \SI{5}{keV} electron beam. For \SI{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
-\begin{equation}
-\Delta x'=\Delta y'= \SI{1.4}{nrad},
-\end{equation}
-Scaling to the HEL design parameters of a \SI{5}{A} and \SI{10}{keV} electron beam (Table~\ref{tab:hel_param}), one obtains:
+assuming an \SI{1}{A}, \SI{5}{keV} electron beam.
+%For \SI{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
+%\begin{equation}
+%\Delta x'=\Delta y'= \SI{1.4}{nrad},
+%\end{equation}
+%Scaling to the HEL design parameters of a \SI{5}{A} and \SI{10}{keV} electron beam (Table~\ref{tab:hel_param}), one obtains:
+%\begin{equation}
+%\int_{z_1}^{z_2} E_{x} dz= 36 \ \mathrm{kV} \Rightarrow \Delta x'= \q{5.1}{nrad}
+%\end{equation}
+Neglecting magnetic effects and scaling to HL-LHC and HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) this yields a kick of:
 \begin{equation}
@@
 For a perfectly uniform, annular and radially symmetric profile, the field in the region of the proton beam core vanishes (see Eqn.~\ref{eq:field_2}).
-In case of electron beam profile imperfections the radial symmetry is broken, leading to a residual field at the beam core.
-Fig.~\ref{core:fig:1} shows an example of the electric field calculated from a measured asymmetric profile.
+In case of electron beam profile imperfections in particular the radial symmetry is broken, leading to a residual field at the beam core (see Fig.~\ref{core:fig:1}).
 \begin{figure}[h]
@@
 \end{figure}
-A first estimate of the residual kick for the HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) with a main solenoid field of $B_{m}=\SI{5}{T}$, $I_e=\SI{5.0}{A}$, $E_{e} = \SI{10.0}{keV}$, $L=\SI{3}{m}$ and $E_{p} = \SI{7.0}{TeV}$ can be obtained by scaling the electric field at the center as obtained in Fig.~\ref{core:fig:1} with the main solenoid field $B_{m}$ and the electron beam current $I_e$ and energy $E_{e}$, yielding:
+A first estimate of the residual kick for the HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) with a main solenoid field of $B_{m}=\SI{5}{T}$ can be obtained by scaling the electric field at the center from the measurements and simulations shown in Fig.~\ref{core:fig:1}, yielding:
 \begin{equation}\label{eqn:kick_central}
@@
 
-\section{\label{sec:adt}Excitation with transverse damper}
+\section{LHC experiments\label{sec:exp}}
+\subsection{Overview of LHC experiments\label{sec:exp_sum}}
+The experiments at the LHC presented in this paper were aiming at deriving tolerances on the residual HEL field at the location of the beam core in case of pulsed operation.
+In total, two experiments were conducted, the first one in 2016 \cite{resexmd2016} and the second one in 2017 \cite{resexmd2017}.
+The beam and machine parameters for both experiments are summarized in Table~\ref{tab:md_param}.
+\begin{table*}
+	\caption{\label{tab:md_param}%
+		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and 2017 \cite{resexmd2016,resexmd2017}.
+The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V horizontal and vertical at the same time.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lcc}
+			Parameter & Experiment 2017 & Experiment 2016  \\
+			\colrule
+			beam &\multicolumn{2}{c}{Beam~1} \\
+			beam energy &\multicolumn{2}{c}{injection energy, 450 GeV} \\\hline
+			single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
+			normalized emittance &\multicolumn{2}{c}{$2.5-3.5~\mathrm{\mu m}$} \\
+			$4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
+			$1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
+			number of bunches & $3\times72=216$ bunches & $12\times4=48$ single bunches\\
+			& (+ 1 pilot + 12 nominal) & \\\hline
+			injection optics, $\beta^*=11$~m & standard optics 2017 & standard optics 2016\\
+			Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=\pm19.6~\mathrm{A}$, explicitly +19.6 A for MOF circuit and}\\
+			& \multicolumn{2}{c}{-19.6~A for MOD circuit (standard 2016 settings)}\\\hline
+			working point $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\
+			chromaticity $(Q'_x,Q'_y)$ & \multicolumn{2}{c}{(+15,+15)}\\\hline
+			pulsing patterns  & $7^{\mathrm{th}}$~turn H,V,H+V;& $7^{\mathrm{th}}$~turn H;\\
+			& $8^{\mathrm{th}}$~turn H,V,H+V;& $10^{\mathrm{th}}$~turn V;\\
+			& random  H,V,H+V;& \\
+		\end{tabular}
+	\end{ruledtabular}
+\end{table*}
+As there is currently no HEL installed in the LHC, the kick on the beam core is emulated by a dipole kick applied with the transverse damper of which more details are given in Section~\ref{sec:adt}.
+Explicitly, all orders higher than the first order are neglected.
+Preceding the beam experiments, simulations were performed to identify the most effective pulsing pattern for the $k^{\mathrm{th}}$~turn pulsing (see Sec.~\ref{sec:pattern}) and yield estimates for losses and emittance growth.
+The two $k^{\mathrm{th}}$~turn pulsing patterns with the largest effect, one $k^{\mathrm{th}}$~turn pulsing pattern with no effect and the random excitation could then be experimentally tested in the LHC limited obviously by the beam time available.
+To at least test the reproducibility of the results on the example of one pulsing pattern, the $7^{\mathrm{th}}$~turn turn pulsing was tested in the 2016 experiments and then repeated in the 2017.
+During the experiments the losses were measured with the fast beam current transformers (FBCT) and the emittance and transverse beam profiles with the beam synchrotron radiation telescope (BSRT).
+The analysis of the BSRT profiles is quite involved and we therefore concentrate in this paper only on the main results.
+A detailed description of the analysis can be found in \cite{bsrtprofinj} and for the individual experiments in \cite{resexmd2016,resexmd2017}.
 
-\section{\label{sec:exp_sum}Overview of LHC experiment}
+\subsection{Excitation with transverse damper and filling scheme\label{sec:adt}}
+\textcolor{red}{Daniel's part}
+\begin{itemize}
+	\item error on amplitude
+	\item synchronization of H and V kicker
+	\item random noise on amplitude - is there any?
+	\item bandwidth - why does the damper not affect the nth turn pulsing????
+	\item filling scheme
+\end{itemize}
 
-\section{\label{sec:sim}Summary of simulation results}
+\subsection{Overview of simulations\label{sec:sim}}
+For the preparation and interpretation of the experiments two different types of simulations were performed:
+\begin{enumerate}
+	\item tracking of a Gaussian distribution (Distribution Tracking) to obtain loss rates and emittance growth,
+	\item Frequency map analysis (FMA)~\cite{fmalaskar} for the visualization of resonances.
+\end{enumerate}
+For both simulation types, the tracking code Lifetrac \cite{lifetrac} was used and the simulation parameters are summarized in Table~\ref{tab:sim_param}.
+Further details on the simulations can be found in \cite{md_sim_hel_res_ex_fitterer,md_sim_hel_res_ex_fitterer_update}.
+\begin{table}[b]
+	\caption{\label{tab:sim_param}%
+		Summary of simulation parameters for the Distribution Tracking and FMA analysis. For details see \cite{md_sim_hel_res_ex_fitterer,md_sim_hel_res_ex_fitterer_update}.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lcc}
+			Parameter & Distribution Tracking & FMA \\
+			\colrule
+			beam &\multicolumn{2}{c}{Beam~1} \\
+			beam energy &\multicolumn{2}{c}{450 GeV} \\
+			emittance & $2.5~\mu$m or $3.5~\mu$m& $2.5~\mu$m\\
+			$4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
+			$1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
+			particle & 6D Gaussian & equally spaced grid\\
+			distribution &  distribution & in $x,y$ up to $10~\sigma$\\
+			&  ($10^4$ particles) & with $\frac{\Delta p}{p_0}=0$\\
+			turns tracked & $10^6$ turns & $10^4$ turns \\\hline
+			optics & \multicolumn{2}{c}{2016 or 2017 injection optics,}\\
+			& \multicolumn{2}{c}{$\beta^*=11$~m}\\
+			machine  & standard errors & no errors \\
+			imperfections &with $a_1=b_1=a_2=0$\footnote{Orbit errors are disabled due to different implementation of the $a_1,b_1,a_2$ errors in Lifetrac and MAD-X.
+$b_2$ errors are adjusted to yield an average (over 60 seeds) peak $\beta$-beat of 15\% as expected from optics measurements in the LHC.} &  \\
+			octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MOF}}=+19.6~\mathrm{A}$, $I_{\mathrm{MOD}}=-19.6~\mathrm{A}$}\\
+			tune $(Q_x,Q_y)$ & \multicolumn{2}{c}{(64.28,59.31) for 2016,}\\
+			& \multicolumn{2}{c}{(62.27,60.295) for 2017}\\
+			chromaticity & \multicolumn{2}{c}{(+15,+15)}\\
+			 $(Q'_x,Q'_y)$ & & \\\hline
+			 transv. aperture & $5.7~\sigma$ & - \\
+			 long. aperture & $10~\sigma$ & - \\
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
 
-\section{\label{sec:exp}Summary of experimental results}
+\subsection{Simulation and experimental results\label{sec:simex}}
+In the following we will summarize the simulation and experimental results. We will jump forth and back between both as the experimental results are best understood on the base of the simulations.
 
-\section{\label{sec:exp_sim}Comparison of simulation and experimental results}
+In summary, the most efficient pulsing patterns and the general tendency of the distribution changes could be successfully predicted in simulations (Sec.~\ref{sec:pattern}).
+Quantitative predictions of loss rates and emittance growth though failed as both observables are influenced also by the natural noise present in the LHC like ground motion and tune modulation.
+At injection this natural noise component is now known for the LHC.
+The observed effect is then the result of the interaction of the external excitation with the natural noise present.
+
+In case of a resonant excitation, the beam distribution adjusts to a new equilibrium distribution.
+In the presence of only a resonant excitation, this distribution is stable and stays unchanged.
+If the resonant excitation interacts with the natural noise sources, the beam distribution still changes to a new equilibrium distribution but this change is also accompanied by a resonant excitation amplitude dependent emittance growth.
+Both effects were experimentally observed in case of the $10^{\mathrm{th}}$~turn pulsing (Sec.~\ref{sec:simex10}).
+
+In case of an excitation with random uniform noise, the shape of the distribution does stay Gaussian.
+The interaction with the natural noise source then only manifests itself in an underestimation of loss rates and emittance growth (see Sec.~\ref{sec:simexran}).
+This effect was also studied in more detail in \cite{md1433_noise_top_energy,md_noise_bbLHC} for colliding beams at the LHC.
+
+\subsubsection{Predicting the most efficient pulsing pattern with simulations\label{sec:pattern}}
+In order to predict the most efficient pulsing pattern, the loss rate and emittance growth from the Distribution Tracking are compared for the different pulsing patterns.
+The results for the 2016 optics and tune are shown in Fig.~\ref{fig:patternemit}.
+
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_emit1.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_emit2.png}
+	\end{minipage}	
+	\caption{\label{fig:patternemit} Horizontal (left) and vertical (right) emittance obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with standard errors.
+The excitation was applied in both planes with an amplitude of 96~nrad.}
+\end{figure}
+Only for the $10^{\mathrm{th}}$~turn pulsing and random excitation a considerable emittance growth is visible.
+For the $10^{\mathrm{th}}$~turn pulsing the vertical emittance also appears to start at an increased initial value and then stays almost constant.
+This behavior is due to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution.
+Further details on this adjustment will be given in Sec.~\ref{sec:simex10}.
+As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself only in an increased initial emittance value in Fig.~\ref{fig:patternemit}.
+
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_intensity.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_sigm.png}
+	\end{minipage}	
+	\caption{\label{fig:patternlosssig} Relative intensity (left) and bunch length (right) obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with standard errors.
+The excitation was applied in both planes with an amplitude of 96~nrad.}
+\end{figure}
+The largest losses are observed for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and random excitation.
+As the bunch length decreases accordingly, the losses are mostly associated to off-momentum particles.
+
+The sensitivity to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are the strong sextupoles and octupoles \cite{md_sim_hel_res_ex_fitterer}.
+The FMA analysis shown in Fig.~\ref{fig:patternfma} reveals that
+the $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		no excitation
+		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$10^{\mathrm{th}}$ H+V
+		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$7^{\mathrm{th}}$ H
+		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut7skh_dp0_ord7.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$7^{\mathrm{th}}$ V
+		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut7skv_dp0_ord7.png}
+	\end{minipage}
+	\caption{\label{fig:patternfma} FMA analysis for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing based on the 2016 injection optics with no machine errors.
+The excitation is 120~nrad in the corresponding plane.
+The absence of a strong excitation of any resonance in case of $7^{\mathrm{th}}$~turn pulsing only in V and the strong excitation in case of pulsing only in H further confirms that indeed the $7Q_x$ resonance is driven.
+For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H and V.}
+\end{figure}
+
+\begin{itemize}
+	\item repeat simulations for each kth turn pattern with only b2 errors
+	\item show results when adding random noise -> amplitude dependent emittance growth + losses
+	\item FMA plot of 7th (2017) -> explain why in H and in V both losses, FMA plot of 10th (2016), FMA plot of 8th and random
+	\item beam distribution change for 10th
+\end{itemize}
+\begin{itemize}
+	\item 7th turn: small emittance but high losses
+	\item 8th turn: almost no effect
+	\item 10th turn: adjustment of distribution -> emittance growth + small losses
+	\item random: constant filamentation -> emittance growth + later strong losses when beam hits aperature
+
+\end{itemize}
+
+\subsubsection{7th turn pulsing\label{sec:simex7}}
+\subsubsection{10th turn pulsing\label{sec:simex10}}
+\subsubsection{8th turn pulsing\label{sec:simex8}}
+\subsubsection{random excitation\label{sec:simexran}}
+
+\section{Summary\label{sec:sum}}
 

### 2017-12-19 scripts/lifetrac_sim_results/plot_loss_emit_pulsing.py

@@ -1,0 +1,100 @@
+from LifeDesk import *
+
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+# plot settings
+mpl.rcParams['xtick.labelsize'] = 14
+mpl.rcParams['ytick.labelsize'] = 14
+mpl.rcParams['axes.labelsize']  = 14
+mpl.rcParams['legend.fontsize']  = 12
+mpl.rcParams['legend.fontsize']  = 12
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+# 2016 optics
+# no pulsing
+opt0 = "2016injnocolc15o+19_6errb2u/1/eps3.5/p1e4t1e4s100weight"
+# with pulsing
+optsk = "2016injnocolc15o+19_6errb2ut%ssk/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
+# random
+optran8 = "2016injnocolc15o+19_6errb2uran/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
+optran9 = "2016injnocolc15o+19_6errb2uran/1/eps3.5/hv9_6e-9/p1e4t1e4s100weight"
+
+# no excitation
+t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
+
+plt.close('all')
+# losses
+plt.figure("losses 2016, 3.5 um emittance")
+plt.plot(t0.data['time'],t0.data['intensity'],'k-')
+for sk,csk in zip(['2','3','4','6','8','9','10'],['b','r','g','pink','cyan','indigo','lime']):
+#for sk,csk in zip(range(2,11),['b','r','g','m','orange','pink','cyan','indigo','lime']):
+  t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%sk),plt_dir=plt_dir)
+  if sk == '2':
+    pltlbl = '2nd'
+  elif sk == '3':
+    pltlbl = '3rd'
+  else:
+    pltlbl = '%sth'%sk
+  plt.plot(t.data['time'],t.data['intensity'],linestyle='-',color=csk,label=pltlbl)
+plt.grid(b=True)
+plt.xlabel('time [s]')
+plt.ylabel('relative intensity')
[…]

### 2017-12-22 scripts/lifetrac_sim_results/plot_distribution_7th_10th.py
update scripts and pictures, add analysis for 10th turn pulsing 2016

@@ -1,0 +1,22 @@
+from LifeDesk import *
+import os
+from matplotlib.pyplot import *
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+close('all')
+for sk in ['7','10']:
+  opt = '2016injnocolc15o+19_6errb2ut%ssk/1/eps3.5/hv9_6e-8/p1e4t1e2s100weight'%sk
+  t=LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt),plt_dir=plt_dir)
+  t.get_hist()
+  t.plot_hist_xyz(title=sk)
+  for p in 'xyz':
+    figure('%s %s'%(sk,p))
+    savefig('%s/2016injerra2b2u_%sthhv_3_5um_hist_%s.png'%(plt_dir,sk,p),bbox_inches='tight')
+
+draw()
+show()
+

### 2017-12-22 scripts/lifetrac_sim_results/plot_loss_emit_10V_2016.py

@@ -1,0 +1,114 @@
+from LifeDesk import *
+
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+# plot settings
+mpl.rcParams['xtick.labelsize'] = 14
+mpl.rcParams['ytick.labelsize'] = 14
+mpl.rcParams['axes.labelsize']  = 14
+mpl.rcParams['legend.fontsize']  = 12
+mpl.rcParams['legend.fontsize']  = 12
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+# 2016 optics
+# no noise
+opt0 = "2016injnocolc15o+19_6errb2u/1/eps3.5/p1e4t1e4s100weight"
+# only with random noise
+opt0ran = "2016injnocolc15o+19_6errb2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+# random noise + 10th V pulsing
+optskran = "2016injnocolc15o+19_6errb2ut10skran/1/eps3.5/v%se-8ran1_2e-3/p1e4t1e4s100weight"
+# 10th V pulsing
+optsk    = "2016injnocolc15o+19_6errb2ut10sk/1/eps3.5/v%se-8/p1e4t1e4s100weight"
+
+# no 10th turn pulsing
+t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
+t0ran = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0ran),plt_dir=plt_dir)
+
+plt.close('all')
+# losses
+plt.figure("losses 2016, 3.5 um emittance")
+plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
+plt.plot(t0ran.data['time'],t0ran.data['intensity'],'k:',label='0 nrad')
+for amp,camp in zip(['4_8','9_6'],['b','r','g','m','orange','pink','cyan','indigo','lime']):
+  t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%amp),plt_dir=plt_dir)
+  tran = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optskran%amp),plt_dir=plt_dir)
+  pltlbl = amp.replace('_','')+' nrad'
[…]

### 2017-12-22 scripts/lifetrac_sim_results/plot_loss_emit_7H_2016.py

@@ -1,0 +1,89 @@
+from LifeDesk import *
+
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+# plot settings
+mpl.rcParams['xtick.labelsize'] = 14
+mpl.rcParams['ytick.labelsize'] = 14
+mpl.rcParams['axes.labelsize']  = 14
+mpl.rcParams['legend.fontsize']  = 12
+mpl.rcParams['legend.fontsize']  = 12
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+# 2016 optics
+# only with random noise
+opt0 = "2016injnocolc15o+19_6errb2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+# random noise + 7th pulsing
+optsk = "2016injnocolc15o+19_6errb2ut7skran/1/eps3.5/h%se-8ran1_2e-3/p1e4t1e4s100weight"
+
+# no excitation
+t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
+
+plt.close('all')
+# losses
+plt.figure("losses 2016, 3.5 um emittance")
+plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
+for amp,camp in zip(['1_2','2_4','4_8'],['b','r','g','m','orange','pink','cyan','indigo','lime']):
+  t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%amp),plt_dir=plt_dir)
+  pltlbl = amp.replace('_','')+' nrad'
+  plt.plot(t.data['time'],t.data['intensity'],linestyle='-',color=camp,label=pltlbl)
+plt.grid(b=True)
+plt.xlabel('time [s]')
+plt.ylabel('relative intensity')
+leg = plt.legend(title='6 nrad random + 7th turn',ncol=2,fontsize=12)
+plt.setp(leg.get_title(),fontsize=12)
+plt.savefig(os.path.join(plt_dir,'2016injerrb2uran1_2e-3_7th_3_5um_intensity.png'),bbox_inches='tight')
+
+# bunch length
+plt.figure("bunch length 2016, 3.5 um emittance")
[…]

### 2017-12-22 scripts/lifetrac_sim_results/plot_loss_emit_7H_V_H+V_2017.py

@@ -1,0 +1,94 @@
+from LifeDesk import *
+
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+# plot settings
+mpl.rcParams['xtick.labelsize'] = 14
+mpl.rcParams['ytick.labelsize'] = 14
+mpl.rcParams['axes.labelsize']  = 14
+mpl.rcParams['legend.fontsize']  = 12
+mpl.rcParams['legend.fontsize']  = 12
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+# 2016 optics
+# only with random noise
+opt0 = "2017injnocolc15o+19_6erra2b2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+# random noise + 7th pulsing
+optsk = "2017injnocolc15o+19_6erra2b2ut7skran/1/eps3.5/%s%se-8ran1_2e-3/p1e4t1e4s100weight"
+
+# no excitation
+t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
+
+plt.close('all')
+# losses
+plt.figure("losses 2017, 3.5 um emittance")
+plt.clf()
+plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
+for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
+  for amp,lamp in zip(['2_4','4_8'],['--',':']):
+    t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%(plane.replace('+',''),amp)),plt_dir=plt_dir)
+    pltlbl = plane.upper()+' '+amp.replace('_','')+' nrad'
+    plt.plot(t.data['time'],t.data['intensity'],linestyle=lamp,color=cplane,label=pltlbl)
+plt.grid(b=True)
+plt.xlabel('time [s]')
+plt.ylabel('relative intensity')
+leg = plt.legend(title='excitation\n7th turn',ncol=1,fontsize=12)
+plt.setp(leg.get_title(),fontsize=12)
+plt.savefig(os.path.join(plt_dir,'2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png'),bbox_inches='tight')
+
+# bunch length
[…]

### 2017-12-22 scripts/lifetrac_sim_results/plot_loss_emit_pulsing.py

@@ -32,8 +32,7 @@ plt.close('all')
 # losses
 plt.figure("losses 2016, 3.5 um emittance")
 plt.plot(t0.data['time'],t0.data['intensity'],'k-')
-for sk,csk in zip(['2','3','4','6','8','9','10'],['b','r','g','pink','cyan','indigo','lime']):
-#for sk,csk in zip(range(2,11),['b','r','g','m','orange','pink','cyan','indigo','lime']):
+for sk,csk in zip(['2','3','4','5','6','7','8','9','10'],['b','r','g','m','orange','pink','cyan','indigo','lime']):
   t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%sk),plt_dir=plt_dir)
   if sk == '2':
     pltlbl = '2nd'
@@ -45,15 +44,23 @@ for sk,csk in zip(['2','3','4','6','8','9','10'],['b','r','g','pink','cyan','ind
 plt.grid(b=True)
 plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
-leg = plt.legend(title='pulsing pattern',ncol=2,fontsize=12)
+ax1=plt.gca()
+# second axis for random
+t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optran8),plt_dir=plt_dir)
+ax2 = plt.gca().twinx()
+ax2.plot(t.data['time'],t.data['intensity'],linestyle='-',color='grey',label='random')
+ax2.set_ylabel('relative intensity random',color='grey')
+ax2.tick_params('y', colors='grey')
+leg = ax1.legend(title='pulsing pattern',ncol=2,fontsize=12,loc='lower left')
+plt.setp(leg.get_title(),fontsize=12)
+leg = ax2.legend(title='pulsing pattern',ncol=1,fontsize=12,loc='lower right')
 plt.setp(leg.get_title(),fontsize=12)
 plt.savefig(os.path.join(plt_dir,'2016injerrb2u_pattern_3_5um_intensity.png'),bbox_inches='tight')
 
 # bunch length
 plt.figure("bunch length 2016, 3.5 um emittance")
 plt.plot(t0.data['time'],t0.data['sigm'],'k-')
-for sk,csk in zip(['2','3','4','6','8','9','10'],['b','r','g','pink','cyan','indigo','lime']):
[…]

### 2017-12-22 resex.tex
update 22 Dec 2017

@@
 On the one hand, this makes the operation more difficult as one has to find the most efficient $k^{\mathrm{th}}$~turn pulsing for each operation mode, on the other hand, it also relaxes the tolerances on the residual field at the beam core as, depending on the machine and beam configuration, it might be possible to find a resonant excitation pattern, which mainly affects the halo and not the core.
-%In the LHC experiment and the preparatory simulations presented in this paper, only the effect of a resonant excitation was studied as analytical as well as experimental results already exist for a random excitation equivalent to the case of exerting white noise on the beam \cite{noise_2007_ohmi,noise_alexahin_lhc,noise_lebedev_ssc,noise_2014_ohmi,md1433_noise_top_energy,md400_noise_injection}.
-Furthermore, the
-%As no HEL is currently installed in the LHC, the non-linear kick of the HEL is approximated only to first order, i.e.
-by a dipole kick, during the experiments.
-The expected amplitude of the dipole kick is derived in Section~\ref{sec:core}.
-In future simulation studies, the dipole kick will have to be substituted with the expected full non-linear kick.
-For the HEL bends, a non-linear symplectic map has already been derived \cite{hel_bends_stancari}, however the kick is estimated to be small compared to the one expected from profile imperfections (see Section~\ref{fig:hel_field}).
-A model for the non-linear kick from profile imperfections is currently under study based on the approach taken in~\cite{hel_model_polynomial_morozov}.
 
@@
 			beam energy &\multicolumn{2}{c}{450 GeV} \\
-			emittance & $2.5~\mu$m or $3.5~\mu$m& $2.5~\mu$m\\
+			emittance & $3.5~\mu$m& $2.5~\mu$m\\
 			$4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
@@
 			machine  & standard errors & no errors \\
-			imperfections &with $a_1=b_1=a_2=0$\footnote{Orbit errors are disabled due to different implementation of the $a_1,b_1,a_2$ errors in Lifetrac and MAD-X.
+			imperfections &with $a_1=b_1=a_2=0$\footnote{Orbit errors are disabled due to different implementation of the $a_1,b_1$ errors in Lifetrac and MAD-X.
 $b_2$ errors are adjusted to yield an average (over 60 seeds) peak $\beta$-beat of 15\% as expected from optics measurements in the LHC.} &  \\
@@
 
-\subsubsection{Predicting the most efficient pulsing pattern with simulations\label{sec:pattern}}
-In order to predict the most efficient pulsing pattern, the loss rate and emittance growth from the Distribution Tracking are compared for the different pulsing patterns.
-The results for the 2016 optics and tune are shown in Fig.~\ref{fig:patternemit}.
-
-\begin{figure}[h]
+\subsubsection{Prediction of the most efficient pulsing pattern\label{sec:pattern}}
+\begin{figure}[b]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_intensity.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_sigm.png}
+	\end{minipage}	
 	\begin{minipage}[t]{0.49\linewidth}
@@
 	\end{minipage}	
-	\caption{\label{fig:patternemit} Horizontal (left) and vertical (right) emittance obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with standard errors.
+	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
 The excitation was applied in both planes with an amplitude of 96~nrad.}
 \end{figure}
-Only for the $10^{\mathrm{th}}$~turn pulsing and random excitation a considerable emittance growth is visible.
-For the $10^{\mathrm{th}}$~turn pulsing the vertical emittance also appears to start at an increased initial value and then stays almost constant.
-This behavior is due to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution.
-Further details on this adjustment will be given in Sec.~\ref{sec:simex10}.
-As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself only in an increased initial emittance value in Fig.~\ref{fig:patternemit}.
+In order to predict the most efficient pulsing pattern, the loss rate and emittance growth from the Distribution Tracking are compared for the different pulsing patterns.
+The results for the 2016 optics and tune are shown in Fig.~\ref{fig:patternsim}.
+The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and the random excitation.
+As the bunch length decreases accordingly, the losses are mostly associated to off-momentum particles.
+Only for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as random excitation a considerable emittance growth is visible, where the random excitation has by far the strongest effect compared to any $k^{\mathrm{th}}$~turn pulsing pattern.
 
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_intensity.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_sigm.png}
-	\end{minipage}	
-	\caption{\label{fig:patternlosssig} Relative intensity (left) and bunch length (right) obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with standard errors.
-The excitation was applied in both planes with an amplitude of 96~nrad.}
-\end{figure}
-The largest losses are observed for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and random excitation.
-As the bunch length decreases accordingly, the losses are mostly associated to off-momentum particles.
-
-The sensitivity to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are the strong sextupoles and octupoles \cite{md_sim_hel_res_ex_fitterer}.
-The FMA analysis shown in Fig.~\ref{fig:patternfma} reveals that
-the $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing.
 \begin{figure}[h]
@@
 	\end{minipage}
-	\caption{\label{fig:patternfma} FMA analysis for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing based on the 2016 injection optics with no machine errors.
+	\caption{\label{fig:patternfma} FMA analysis for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
 The excitation is 120~nrad in the corresponding plane.
 The absence of a strong excitation of any resonance in case of $7^{\mathrm{th}}$~turn pulsing only in V and the strong excitation in case of pulsing only in H further confirms that indeed the $7Q_x$ resonance is driven.
-For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H and V.}
+For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V.}
+\end{figure}
+The sensitivity to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are the strong sextupoles and octupoles \cite{md_sim_hel_res_ex_fitterer}.
+The FMA analysis shown in Fig.~\ref{fig:patternfma} reveals that
+the $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing.
+As octupoles can only drive even resonances, the source of the $7Q_x$ resonances are the sextupoles while the octupoles only generate the tune footprint.
+The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect thus is based on the magnetic field errors assigned, explicitly also the seed chosen for the simulations.
+
+\begin{figure}[t]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$7^{\mathrm{th}}$ H+V
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_7thhv_3_5um_hist_x.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$10^{\mathrm{th}}$ H+V
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_10thhv_3_5um_hist_y.png}
+	\end{minipage}	
+	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn (left) and vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (right) from simulation results (Distribution Tracking) based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
+The excitation was applied in both planes with an amplitude of 96~nrad.}
+\end{figure}
+For the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing the emittance also appears to start at an increased initial value and then stays almost constant.
+This behavior is due to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution, and as the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}.
+Besides an increased emittance, the new distribution also becomes non-Gaussian as can be seen from the histograms shown in Fig.~\ref{fig:patternhist} for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
+This initial fast change of the beam distribution was also to the knowledge of the authors for the first time experimentally observed (Sec.~\ref{sec:simex10} and Sec.~\ref{sec:simex7}), however on a time scale of minutes instead of seconds.
+
+\subsubsection{10th turn pulsing\label{sec:simex10}}
+The $10^{\mathrm{th}}$~turn pulsing was tested in 2016 with an excitation in the vertical plane only. Experimentally, the following observations were made (Fig.~\ref{fig:10thexp}):
+\begin{itemize}
+	\item increase of the loss rate with the excitation amplitude,
+	\item increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane.
+\end{itemize}
+\begin{figure}[t]
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.49\linewidth]{bunch_intensity_v10th_no_damper_avg_annotate.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{emith_avg_rel_v10th_no_damper_annotate.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{emitv_avg_rel_v10th_no_damper_annotate.png}
+	\end{minipage}	
+	\caption{\label{fig:10thexp} Relative intensity (top) and emittance (bottom) obtained during the 2016 experiments.
+For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Sec.~\ref{sec:adt}).
+The different colored solid lines labeled with $n\Delta A$ indicate the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation as their envelope.}
+\end{figure}
+Comparing these experimental results with the simulations shown in Fig.~\ref{fig:10thsim} with only $10^{\mathrm{th}}$~turn pulsing (solid line), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth seen experimentally.
+Instead, the simulations yield a constant emittance after an initial increase during the first $10^4$~turns or first second.
+So obviously there is an essential ingredient missing.
+As it turns out, the constant emittance can be changed to an excitation amplitude dependent emittance growth by adding a random dipole noise component with uniform distribution (dotted line in Fig.~\ref{fig:10thsim}).
+As a first estimate the value obtained in \cite{md1433_noise_top_energy,md_noise_bbLHC} for 6.5~TeV was used and scaled by the magnetic rigidity to the injection energy of 450~GeV, as no noise estimate is available at injection, yielding a maximum kick amplitude at the transverse damper of approximately
+\begin{equation}
+\theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = 6~\mathrm{nrad},
+\end{equation}
+Adding this random noise component, the simulations then yield correctly an excitation amplitude dependent emittance growth in the vertical plane only after a first fast adjustment of the beam distribution visible as an increased initial value.
+In addition, also the loss rates increase.
+As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_intensity.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit2_rel.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit1.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit2.png}
+	\end{minipage}	
+	\caption{\label{fig:10thsim} Relative intensity (top left) and vert.
+emittance (top right), and hor.
+(bottom left) and vert.
+(bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$.
+The solid line indicates an excitation with only $10^{\mathrm{th}}$~turn pulsing in V and the dotted line with $10^{\mathrm{th}}$~turn pulsing in V plus a random dipole noise component in H+V of 6~nrad.}
+\end{figure}
+
+\subsubsection{8th turn pulsing\label{sec:simex8}}
+\subsubsection{random excitation\label{sec:simexran}}
+
+\subsubsection{7th turn pulsing\label{sec:simex7}}
+The $7^{\mathrm{th}}$~turn pulsing pattern is the only pulsing pattern which has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the results.
+However, in 2017 the machine tune was changed from (64.28,59.31) to (62.27,60.295) in standard operation.
+The attempt to change back to the fractional tune of (.28,.31) as chosen in 2016 however lead to large losses during the experiments in 2017 and could thus not be changed.
+???This change in tune also entailed a change in the driving resonances (Fig.~\ref{fig:7th2017fma}).
+As will be detailed in this section, the qualitative evolution of losses and emittance growth are still similar despite the different tunes, but the quantitative results differ.???
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		no excitation
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$7^{\mathrm{th}}$ H+V
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png}
+	\end{minipage}	
+	\caption{\label{fig:7th2017fma} FMA analysis for $7^{\mathrm{th}}$~turn pulsing based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
+The excitation is 96~nrad in both planes.
+The $7Q_x$ and $7Q_y$ resonance are both excited and in addition the 14th order $7Q_x+7Q_y$.}
 \end{figure}
@@
 \begin{itemize}
-	\item repeat simulations for each kth turn pattern with only b2 errors
+	\item Don't understand why there is emittance growth in H for 2016. There is no emittance growth in the experiment.
+	\item repeat simulations for 2017 for all three planes and with random noise => check if there is only emittance growth in V.
+	\item experimentally: 2016 + 2017 losses in all cases, only emittance growth in V in 2017
+\end{itemize}
+\subsubsection{Effect of the transverse damper\label{sec:damp}}
+Compare random and nth turn -> no damping for nth, only for random.
+
+In general:
+\begin{itemize}
+	\item repeat simulations for each kth turn pattern with only b2 errors (\textcolor{green}{done})
 	\item show results when adding random noise -> amplitude dependent emittance growth + losses
@@
 \end{itemize}
-
-\subsubsection{7th turn pulsing\label{sec:simex7}}
-\subsubsection{10th turn pulsing\label{sec:simex10}}
-\subsubsection{8th turn pulsing\label{sec:simex8}}
-\subsubsection{random excitation\label{sec:simexran}}
 

### 2017-12-26 scripts/lifetrac_sim_results/plot_distribution_7th_10th.py
add plots for 10th and 7th turn pulsing

@@ -1,22 +1,0 @@
-from LifeDesk import *
-import os
-from matplotlib.pyplot import *
-
-# directory with simulation results
-basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
-# plot directory for paper
-plt_dir = "plots"
-
-close('all')
-for sk in ['7','10']:
-  opt = '2016injnocolc15o+19_6errb2ut%ssk/1/eps3.5/hv9_6e-8/p1e4t1e2s100weight'%sk
-  t=LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt),plt_dir=plt_dir)
-  t.get_hist()
-  t.plot_hist_xyz(title=sk)
-  for p in 'xyz':
-    figure('%s %s'%(sk,p))
-    savefig('%s/2016injerra2b2u_%sthhv_3_5um_hist_%s.png'%(plt_dir,sk,p),bbox_inches='tight')
-
-draw()
-show()
-

### 2017-12-26 scripts/lifetrac_sim_results/plot_hist_7th_10th.py

@@ -1,0 +1,58 @@
+from LifeDesk import *
+import os
+from matplotlib.pyplot import *
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+# use weighted Gaussian (p1e4t1e2s100weight) or not (p1e4t1e2s100)
+weighted = False
+# pulsing pattern H+V
+sk = 't7sk'
+ylimresdiff = [-7,7]
+sk='t10sk'
+ylimresdiff = [-3.1,3.1]
+sk='ran'
+ylimresdiff = [-3.1,3.1]
+
+#if weighted:
+## weighted Gaussian
+#  opt = '2016injnocolc15o+19_6errb2u%s/1/eps3.5/hv9_6e-8/p1e4t1e2s100weight'%sk
+#else:
+## normal Gaussian
+#  opt = '2016injnocolc15o+19_6errb2u%s/1/eps3.5/hv9_6e-8/p1e4t1e2s100'%sk
+#
+#close('all')
+#t=LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt),plt_dir=plt_dir)
+#t.get_hist()
+#t.plot_hist_xyz(title=sk,ylimhist=[1.e-4,1.1],ylimresdiff=ylimresdiff,ylimresrat=[0.1,10])
+#for p in 'xyz':
+#  figure('%s %s'%(sk,p))
+#  if weighted:
+#    savefig('%s/2016injerra2b2u_%shv_3_5um_hist_weighted_%s.png'%(plt_dir,sk,p),bbox_inches='tight')
+#  else:
+#    savefig('%s/2016injerra2b2u_%shv_3_5um_hist_%s.png'%(plt_dir,sk,p),bbox_inches='tight')
+
+# no excitation
+if weighted:
+# weighted Gaussian
+  opt = '2016injnocolc15o+19_6errb2u/1/eps3.5/p1e4t1e2s100weight'
+else:
+# normal Gaussian
+  opt = '2016injnocolc15o+19_6errb2u/1/eps3.5/p1e4t1e2s100'
+close('all')
+t=LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt),plt_dir=plt_dir)
+t.get_hist()
[…]

### 2017-12-26 scripts/lifetrac_sim_results/plot_loss_emit_10V_2016.py

@@ -32,83 +32,93 @@ t0ran = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0ran),plt_dir=plt_dir
 
 plt.close('all')
 # losses
+# open file to dump loss rates [%/h]
+f = open(os.path.join(basedir,'2016injnocolc15o+19_6errb2ut10skran/1/eps3.5/','losses_ran_10V'), 'w+')
+f.write("# amplitude [nrad] loss rate [%/h]\n")
 plt.figure("losses 2016, 3.5 um emittance")
 plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
 plt.plot(t0ran.data['time'],t0ran.data['intensity'],'k:',label='0 nrad')
+lossrate = -(t0ran.data['intensity'][-1]-t0ran.data['intensity'][0])/(t0ran.data['time'][-1]-t0ran.data['time'][0])*60*60*100
+f.write("0 %4.8e\n"%(lossrate))
 for amp,camp in zip(['4_8','9_6'],['b','r','g','m','orange','pink','cyan','indigo','lime']):
   t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%amp),plt_dir=plt_dir)
   tran = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optskran%amp),plt_dir=plt_dir)
   pltlbl = amp.replace('_','')+' nrad'
   plt.plot(t.data['time'],t.data['intensity'],linestyle='-',color=camp,label=pltlbl)
   plt.plot(tran.data['time'],tran.data['intensity'],linestyle=':',color=camp,label=pltlbl)
+  # dump the loss rates into the file 2016injnocolc15o+19_6errb2ut10skran/1/eps3.5/losses_10V
+  lossrate = -(t.data['intensity'][-1]-t.data['intensity'][0])/(t.data['time'][-1]-t.data['time'][0])*60*60*100
+  f.write("%s %4.8e\n"%(amp.replace('_',''),lossrate))
 plt.grid(b=True)
 plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
+f.close()
 
-# bunch length
-plt.figure("bunch length 2016, 3.5 um emittance")
-plt.plot(t0.data['time'],t0.data['sigm'],'k-',label='0 nrad')
-plt.plot(t0ran.data['time'],t0ran.data['sigm'],'k:',label='0 nrad')
[…]

### 2017-12-26 scripts/lifetrac_sim_results/plot_loss_emit_7H_V_H+V_2017.py

@@ -16,7 +16,7 @@ basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
 # plot directory for paper
 plt_dir = "plots"
 
-# 2016 optics
+# 2017 optics
 # only with random noise
 opt0 = "2017injnocolc15o+19_6erra2b2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
 # random noise + 7th pulsing

### 2017-12-26 scripts/lifetrac_sim_results/plot_loss_emit_7H_V_H+V_2017_7H_2016.py

@@ -1,0 +1,142 @@
+from LifeDesk import *
+
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+# plot settings
+mpl.rcParams['xtick.labelsize'] = 14
+mpl.rcParams['ytick.labelsize'] = 14
+mpl.rcParams['axes.labelsize']  = 14
+mpl.rcParams['legend.fontsize']  = 12
+mpl.rcParams['legend.fontsize']  = 12
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+# 2016 optics
+# only with random noise
+opt0_16 = "2016injnocolc15o+19_6erra2b2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+# random noise + 7th pulsing H
+optsk_16 = "2016injnocolc15o+19_6erra2b2ut7skran/1/eps3.5/h%se-8ran1_2e-3/p1e4t1e4s100weight"
+
+# 2017 optics
+# only with random noise
+opt0 = "2017injnocolc15o+19_6erra2b2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+# random noise + 7th pulsing
+optsk = "2017injnocolc15o+19_6erra2b2ut7skran/1/eps3.5/%s%se-8ran1_2e-3/p1e4t1e4s100weight"
+
+# no excitation
+t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
+t0_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0_16),plt_dir=plt_dir)
+
+plt.close('all')
+# losses
+plt.figure("losses, 3.5 um emittance")
+plt.clf()
+# 2017
+plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
+for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
+  for amp,lamp in zip(['2_4','4_8'],['--',':']):
+    t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%(plane.replace('+',''),amp)),plt_dir=plt_dir)
+    pltlbl = plane.upper()+' '+amp.replace('_','')+' nrad'
[…]

### 2017-12-26 resex.tex
update 26 Dec 2017

@@
 	\end{minipage}	
-	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn (left) and vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (right) from simulation results (Distribution Tracking) based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
-The excitation was applied in both planes with an amplitude of 96~nrad.}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		random H+V
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		no excitation
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_3_5um_hist_y.png}
+	\end{minipage}	
+	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from simulation results (Distribution Tracking) based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
+The excitation was applied in both planes with an amplitude of 96~nrad.
+The residual is defined as the final distribution (here $10^4$~turns) minus the initial distribution (here $10^2$~turns), and the ratio as the final distribution divided by the initial distribution.
+The red line in the residual shows the difference between the Gaussian fit of the distribution and the distribution itself.}
 \end{figure}
 For the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing the emittance also appears to start at an increased initial value and then stays almost constant.
-This behavior is due to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution, and as the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}.
-Besides an increased emittance, the new distribution also becomes non-Gaussian as can be seen from the histograms shown in Fig.~\ref{fig:patternhist} for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
-This initial fast change of the beam distribution was also to the knowledge of the authors for the first time experimentally observed (Sec.~\ref{sec:simex10} and Sec.~\ref{sec:simex7}), however on a time scale of minutes instead of seconds.
+This behavior is due to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution.
+As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}.
+Besides an increased emittance, the new distribution also becomes non-Gaussian as can be seen from the residual in respect to the Gaussian distribution in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:patternhist}).
+For comparison, also the distribution in both planes for a random excitation is shown This initial fast change of the beam distribution was also to the knowledge of the authors for the first time experimentally observed (Sec.~\ref{sec:simex10} and Sec.~\ref{sec:simex7}).
+The timescale of this fast adjustment is experimentally however much slower than predicted in simulations - on the timescale of minutes compared to seconds.
 
 \subsubsection{10th turn pulsing\label{sec:simex10}}
-The $10^{\mathrm{th}}$~turn pulsing was tested in 2016 with an excitation in the vertical plane only. Experimentally, the following observations were made (Fig.~\ref{fig:10thexp}):
-\begin{itemize}
-	\item increase of the loss rate with the excitation amplitude,
-	\item increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane.
-\end{itemize}
-\begin{figure}[t]
+The $10^{\mathrm{th}}$~turn pulsing was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
+\begin{enumerate}
+	\item increase of the loss rate with the excitation amplitude (Fig.~\ref{fig:10thexp}, \ref{fig:10thexpsim}),
+	\item increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane (Fig.~\ref{fig:10thexp}).
+	\item change of the beam distribution due to the resonant excitation (Fig.~\ref{fig:10thexp}, \ref{fig:10thexpprof})
+\end{enumerate}
+As loss rates, emittance growth and distribution stayed unchanged for the 4~reference bunches, the above observations can be truly associated with the effect of the resonant excitation.
+\begin{figure}[h]
 	\begin{minipage}[t]{1.0\linewidth}
@@
 \end{figure}
-Comparing these experimental results with the simulations shown in Fig.~\ref{fig:10thsim} with only $10^{\mathrm{th}}$~turn pulsing (solid line), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth seen experimentally.
-Instead, the simulations yield a constant emittance after an initial increase during the first $10^4$~turns or first second.
-So obviously there is an essential ingredient missing.
-As it turns out, the constant emittance can be changed to an excitation amplitude dependent emittance growth by adding a random dipole noise component with uniform distribution (dotted line in Fig.~\ref{fig:10thsim}).
-As a first estimate the value obtained in \cite{md1433_noise_top_energy,md_noise_bbLHC} for 6.5~TeV was used and scaled by the magnetic rigidity to the injection energy of 450~GeV, as no noise estimate is available at injection, yielding a maximum kick amplitude at the transverse damper of approximately
-\begin{equation}
-\theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = 6~\mathrm{nrad},
-\end{equation}
-Adding this random noise component, the simulations then yield correctly an excitation amplitude dependent emittance growth in the vertical plane only after a first fast adjustment of the beam distribution visible as an increased initial value.
-In addition, also the loss rates increase.
-As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced.
+
+Besides the determination of the scaling of the loss rate and emittance growth with the excitation amplitude as required for the specification of the hollow electron lens, the most novel observation is the direct measurement of the change of the beam distribution, here with the Beam Synchrotron Radiation Telescope (BSRT) \cite{bsrtprofinj}.
+As an example for the change in distribution, the vertical profile of a reference bunch and one bunch experiencing the maximum excitation are compared in Fig.~\ref{fig:10thexpprof}.
+The distribution of the excited bunch clearly changes and becomes non-Gaussian, while the reference bunch stays almost unchanged.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_50.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_1300.png}
+	\end{minipage}
+	\caption{\label{fig:10thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2016 experiments.
+The profiles are taken at the end of the $10^{\mathrm{th}}$~turn pulsing in V.
+The black line in the plot of the residual is the current profile minus the initial profile and is a measure for the overall change of the distribution.
+The red line is the current profile minus its Gaussian fit, which gives and indication for the deviation of the distribution from a Gaussian distribution.
+For details on the analysis see \cite{bsrtprofinj}.
+The distribution of the reference bunch (left), here slot 50, stays unchanged, while the bunch experiencing the maximum excitation (right), here slot 1300, clearly shows a change to a non-Gaussian distribution.}
+\end{figure}
+
+Comparing these experimental results with the simulations including only the $10^{\mathrm{th}}$~turn pulsing (solid line, Fig.~\ref{fig:10thsim}), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth seen experimentally.
 \begin{figure}[h]
@@
 	\end{minipage}	
-	\caption{\label{fig:10thsim} Relative intensity (top left) and vert.
+	\caption{\label{fig:10thsim} Relative intensity (top left) and relative vert.
 emittance (top right), and hor.
@@
 The solid line indicates an excitation with only $10^{\mathrm{th}}$~turn pulsing in V and the dotted line with $10^{\mathrm{th}}$~turn pulsing in V plus a random dipole noise component in H+V of 6~nrad.}
+\end{figure}
+Instead, the simulations yield a constant emittance after an initial increase during the first $10^4$~turns or first second.
+So obviously there is an essential ingredient missing.
+As it turns out, the constant emittance can be changed to an excitation amplitude dependent emittance growth by adding a random dipole noise component with uniform distribution (dotted line in Fig.~\ref{fig:10thsim}).
+As a first estimate the value obtained in \cite{md1433_noise_top_energy,md_noise_bbLHC} for 6.5~TeV was used and scaled by the magnetic rigidity to the injection energy of 450~GeV, as no noise estimate is available at injection.
+This yields a maximum kick amplitude at the transverse damper of approximately
+\begin{equation}
+\theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = 6~\mathrm{nrad},
+\end{equation}
+Adding this random noise component, the simulations then yield correctly an excitation amplitude dependent emittance growth in the vertical plane only after a first fast adjustment of the beam distribution visible as an increased initial value.
+
+Adding a random noise component has in addition the effect of increasing the loss rate.
+As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced, but a large difference still prevails (Fig.~\ref{fig:10thexpsim}).
+\begin{figure}[h]
+		\centering
+		\includegraphics[width=1.0\linewidth]{scale_amp_10v_ran_10V_sim.png}
+	\caption{\label{fig:10thexpsim} Comparison of scaling of loss rates with the $10^{\mathrm{th}}$~turn pulsing excitation amplitude.
+The experimental results are shown in black (maximum excitation with 48~nrad) and blue (maximum excitation with 96~nrad) and the simulation results (2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad) in red.}
 \end{figure}
@@
 The $7^{\mathrm{th}}$~turn pulsing pattern is the only pulsing pattern which has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the results.
-However, in 2017 the machine tune was changed from (64.28,59.31) to (62.27,60.295) in standard operation.
-The attempt to change back to the fractional tune of (.28,.31) as chosen in 2016 however lead to large losses during the experiments in 2017 and could thus not be changed.
-???This change in tune also entailed a change in the driving resonances (Fig.~\ref{fig:7th2017fma}).
-As will be detailed in this section, the qualitative evolution of losses and emittance growth are still similar despite the different tunes, but the quantitative results differ.???
+However, in 2017 the machine tune was changed from (64.28,59.31) to (62.27,60.295) in standard operation accompanied by a slight change in optics.
+The attempt to change back to the fractional tune of (.28,.31) as chosen in 2016 however lead to large losses during the experiments in 2017.
+This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016  (Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (Fig.~\ref{fig:7th2017fma}) and in addition also the effected amplitudes of the beam distribution.
 \begin{figure}[h]
@@
 \end{figure}
+Therefore the effects on the beam also differ for both experiments (see Table~\ref{tab:7thexp} for a general overview and Fig.~\ref{} and Appendix~\ref{} for more details).
+\begin{table}[h]
+	\caption{\label{tab:7thexp}%
+		Overview of the effect of $7^{\mathrm{th}}$~turn pulsing during the two resonant excitation experiments in 2016 and 2017~\cite{resexmd2016,resexmd2017}.
+The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V for horizontal and vertical at the same time.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lcc}
+			Parameter & Experiment 2017 & Experiment 2016  \\
+			\colrule
+			tune $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\\hline
+			excitation plane  & H,V and H+V&  H\\
+			losses & similar loss rates for  & losses for pulsing \\
+			& pulsing in H,V & in H \\
+			&  and H+V & \\
+			hor. emit. growth & no & yes, very small \\
+			vert. emit. growth & yes, for pulsing in V & no\\
+			&  and H+V & \\
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
+Simulations more or less fail completely (Fig.~\ref{fig:7thsim}). Higher amplitudes have been used for simulations to see noticable effect.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png}
+	\end{minipage}	
+	\caption{\label{fig:7thsim} Relative intensity (top left) and relative vert.
+emittance (top right), and hor.
+(bottom left) and vert.
+(bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$ and 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $7^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
+\end{figure}
 
@@
 	\item Don't understand why there is emittance growth in H for 2016. There is no emittance growth in the experiment.
-	\item repeat simulations for 2017 for all three planes and with random noise => check if there is only emittance growth in V.
+	\item repeat simulations for 2017 for all three planes and with random noise => check if there is only emittance growth in V \textcolor{red}{done}.
 	\item experimentally: 2016 + 2017 losses in all cases, only emittance growth in V in 2017
@@
 \section{Appendixes}
+\section{Results of $7^{\mathrm{th}}$~turn pulsing during 2016 and 2017 experiments}
 

### 2018-01-11 resex.tex
remove grid in plots

@@
 \begin{figure}[b]
-	\includegraphics[width=0.8\linewidth]{hel_field}% Here is how to import EPS art
+	\includegraphics[width=0.8\linewidth]{kick_hel_lhc_no_grid}% Here is how to import EPS art
 	\caption{\label{fig:hel_field} Illustration of the hollow electron beam distribution (blue), the kick experienced by the proton beam (red) and the collimators (gray).}
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_intensity.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_intensity.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_sigm.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_sigm.png}
 	\end{minipage}	
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_emit1.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit1.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_emit2.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit2.png}
 	\end{minipage}	
@@
 		$7^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_7thhv_3_5um_hist_x.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
 	\end{minipage}
@@
 		$10^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_10thhv_3_5um_hist_y.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png}
 	\end{minipage}	
@@
 		\centering
-		\includegraphics[width=0.49\linewidth]{bunch_intensity_v10th_no_damper_avg_annotate.png}
+		\includegraphics[width=0.49\linewidth]{bunch_intensity_v10th_no_damper_avg.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{emith_avg_rel_v10th_no_damper_annotate.png}
+		\includegraphics[width=1.0\linewidth]{emith_avg_rel_v10th_no_damper.png}
 	\end{minipage}	
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{emitv_avg_rel_v10th_no_damper_annotate.png}
+		\includegraphics[width=1.0\linewidth]{emitv_avg_rel_v10th_no_damper.png}
 	\end{minipage}	
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_intensity.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_intensity.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit2_rel.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2_rel.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit1.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit1.png}
 	\end{minipage}	
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit2.png}
+		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2.png}
 	\end{minipage}	
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{scale_amp_10v_ran_10V_sim.png}
+		\includegraphics[width=1.0\linewidth]{scale_amp_10v_ran_lbllong_sim.png}
 	\caption{\label{fig:10thexpsim} Comparison of scaling of loss rates with the $10^{\mathrm{th}}$~turn pulsing excitation amplitude.
@@
 However, in 2017 the machine tune was changed from (64.28,59.31) to (62.27,60.295) in standard operation accompanied by a slight change in optics.
-The attempt to change back to the fractional tune of (.28,.31) as chosen in 2016 however lead to large losses during the experiments in 2017.
-This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016  (Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (Fig.~\ref{fig:7th2017fma}) and in addition also the effected amplitudes of the beam distribution.
+The attempt to change back to the fractional tune of (.28,.31) as chosen in 2016 lead to large losses during the experiments in 2017 and therefore the 2017 working point of (62.27,60.295) was used during the experiment.
+This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016  (Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (Fig.~\ref{fig:7th2017fma}) and also a change in the affected amplitudes of the beam distribution.
 \begin{figure}[h]
@@
 \end{figure}
-Therefore the effects on the beam also differ for both experiments (see Table~\ref{tab:7thexp} for a general overview and Fig.~\ref{} and Appendix~\ref{} for more details).
+Consequently the effects on the beam also differ for both experiments (see Table~\ref{tab:7thexp} for a general overview and Fig.~\ref{} and Appendix~\ref{} for more details).
 \begin{table}[h]

### 2018-01-11 scripts/lifetrac_sim_results/plot_hist_7th_10th.py

@@ -12,46 +12,46 @@ weighted = False
 # pulsing pattern H+V
 sk = 't7sk'
 ylimresdiff = [-7,7]
-sk='t10sk'
-ylimresdiff = [-3.1,3.1]
-sk='ran'
-ylimresdiff = [-3.1,3.1]
+#sk='t10sk'
+#ylimresdiff = [-3.1,3.1]
+#sk='ran'
+#ylimresdiff = [-3.1,3.1]
 
-#if weighted:
-## weighted Gaussian
-#  opt = '2016injnocolc15o+19_6errb2u%s/1/eps3.5/hv9_6e-8/p1e4t1e2s100weight'%sk
-#else:
-## normal Gaussian
-#  opt = '2016injnocolc15o+19_6errb2u%s/1/eps3.5/hv9_6e-8/p1e4t1e2s100'%sk
-#
-#close('all')
-#t=LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt),plt_dir=plt_dir)
-#t.get_hist()
-#t.plot_hist_xyz(title=sk,ylimhist=[1.e-4,1.1],ylimresdiff=ylimresdiff,ylimresrat=[0.1,10])
-#for p in 'xyz':
-#  figure('%s %s'%(sk,p))
-#  if weighted:
-#    savefig('%s/2016injerra2b2u_%shv_3_5um_hist_weighted_%s.png'%(plt_dir,sk,p),bbox_inches='tight')
-#  else:
-#    savefig('%s/2016injerra2b2u_%shv_3_5um_hist_%s.png'%(plt_dir,sk,p),bbox_inches='tight')
-
-# no excitation
 if weighted:
 # weighted Gaussian
-  opt = '2016injnocolc15o+19_6errb2u/1/eps3.5/p1e4t1e2s100weight'
+  opt = '2016injnocolc15o+19_6erra2b2u%s/1/eps3.5/hv9_6e-8/p1e4t1e2s100weight'%sk
 else:
 # normal Gaussian
-  opt = '2016injnocolc15o+19_6errb2u/1/eps3.5/p1e4t1e2s100'
+  opt = '2016injnocolc15o+19_6erra2b2u%s/1/eps3.5/hv9_6e-8/p1e4t1e2s100'%sk
+
 close('all')
 t=LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt),plt_dir=plt_dir)
 t.get_hist()
[…]

### 2018-01-11 scripts/lifetrac_sim_results/plot_loss_emit_10V_2016.py

@@ -18,13 +18,13 @@ plt_dir = "plots"
 
 # 2016 optics
 # no noise
-opt0 = "2016injnocolc15o+19_6errb2u/1/eps3.5/p1e4t1e4s100weight"
+opt0 = "2016injnocolc15o+19_6erra2b2u/1/eps3.5/p1e4t1e4s100weight"
 # only with random noise
-opt0ran = "2016injnocolc15o+19_6errb2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+opt0ran = "2016injnocolc15o+19_6erra2b2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
 # random noise + 10th V pulsing
-optskran = "2016injnocolc15o+19_6errb2ut10skran/1/eps3.5/v%se-8ran1_2e-3/p1e4t1e4s100weight"
+optskran = "2016injnocolc15o+19_6erra2b2ut10skran/1/eps3.5/v%se-8ran1_2e-3/p1e4t1e4s100weight"
 # 10th V pulsing
-optsk    = "2016injnocolc15o+19_6errb2ut10sk/1/eps3.5/v%se-8/p1e4t1e4s100weight"
+optsk    = "2016injnocolc15o+19_6erra2b2ut10sk/1/eps3.5/v%se-8/p1e4t1e4s100weight"
 
 # no 10th turn pulsing
 t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
@@ -33,7 +33,7 @@ t0ran = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0ran),plt_dir=plt_dir
 plt.close('all')
 # losses
 # open file to dump loss rates [%/h]
-f = open(os.path.join(basedir,'2016injnocolc15o+19_6errb2ut10skran/1/eps3.5/','losses_ran_10V'), 'w+')
+f = open(os.path.join(basedir,'2016injnocolc15o+19_6erra2b2ut10skran/1/eps3.5/','losses_ran_10V'), 'w+')
 f.write("# amplitude [nrad] loss rate [%/h]\n")
 plt.figure("losses 2016, 3.5 um emittance")
 plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
@@ -49,75 +49,75 @@ for amp,camp in zip(['4_8','9_6'],['b','r','g','m','orange','pink','cyan','indig
[…]

### 2018-01-11 scripts/lifetrac_sim_results/plot_loss_emit_7H_V_H+V_2017_7H_2016.py

@@ -49,7 +49,7 @@ for amp,lamp in zip(['2_4','4_8'],['--',':']):
   t_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk_16%(amp)),plt_dir=plt_dir)
   pltlbl = 'H '+amp.replace('_','')+' nrad'
   plt.plot(t_16.data['time'],t_16.data['intensity'],linestyle=lamp,color='orange',label=pltlbl)
-plt.grid(b=True)
+#plt.grid(b=True)
 plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
 handles,labels = plt.gca().get_legend_handles_labels()
@@ -74,7 +74,7 @@ for amp,lamp in zip(['2_4','4_8'],['--',':']):
   t_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk_16%(amp)),plt_dir=plt_dir)
   pltlbl = 'H '+amp.replace('_','')+' nrad'
   plt.plot(t_16.data['time'],t_16.data['sigm'],linestyle=lamp,color='orange',label=pltlbl)
-plt.grid(b=True)
+#plt.grid(b=True)
 plt.xlabel('time [s]')
 plt.ylabel('bunch length [cm]')
 handles,labels = plt.gca().get_legend_handles_labels()
@@ -100,7 +100,7 @@ for phv,p12 in zip(['hor.','vert.'],'12'):
     t_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk_16%(amp)),plt_dir=plt_dir)
     pltlbl = 'H '+amp.replace('_','')+' nrad'
     plt.plot(t_16.data['time'],t_16.data['emit%s'%p12],linestyle=lamp,color='orange',label=pltlbl)
-  plt.grid(b=True)
+#  plt.grid(b=True)
   plt.xlabel('time [s]')
   plt.ylabel('%s normalized emittance [$\mu$m]'%phv)
   handles,labels = plt.gca().get_legend_handles_labels()
@@ -126,7 +126,7 @@ for phv,p12 in zip(['hor.','vert.'],'12'):
     t_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk_16%(amp)),plt_dir=plt_dir)
     pltlbl = 'H '+amp.replace('_','')+' nrad'
[…]

### 2018-01-11 scripts/lifetrac_sim_results/plot_loss_emit_pulsing.py

@@ -18,12 +18,12 @@ plt_dir = "plots"
 
 # 2016 optics
 # no pulsing
-opt0 = "2016injnocolc15o+19_6errb2u/1/eps3.5/p1e4t1e4s100weight"
+opt0 = "2016injnocolc15o+19_6erra2b2u/1/eps3.5/p1e4t1e4s100weight"
 # with pulsing
-optsk = "2016injnocolc15o+19_6errb2ut%ssk/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
+optsk = "2016injnocolc15o+19_6erra2b2ut%ssk/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
 # random
-optran8 = "2016injnocolc15o+19_6errb2uran/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
-optran9 = "2016injnocolc15o+19_6errb2uran/1/eps3.5/hv9_6e-9/p1e4t1e4s100weight"
+optran8 = "2016injnocolc15o+19_6erra2b2uran/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
+optran9 = "2016injnocolc15o+19_6erra2b2uran/1/eps3.5/hv9_6e-9/p1e4t1e4s100weight"
 
 # no excitation
 t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
@@ -41,7 +41,7 @@ for sk,csk in zip(['2','3','4','5','6','7','8','9','10'],['b','r','g','m','orang
   else:
     pltlbl = '%sth'%sk
   plt.plot(t.data['time'],t.data['intensity'],linestyle='-',color=csk,label=pltlbl)
-plt.grid(b=True)
+#plt.grid(b=True)
 plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
 ax1=plt.gca()
@@ -55,7 +55,7 @@ leg = ax1.legend(title='pulsing pattern',ncol=2,fontsize=12,loc='lower left')
 plt.setp(leg.get_title(),fontsize=12)
 leg = ax2.legend(title='pulsing pattern',ncol=1,fontsize=12,loc='lower right')
 plt.setp(leg.get_title(),fontsize=12)
-plt.savefig(os.path.join(plt_dir,'2016injerrb2u_pattern_3_5um_intensity.png'),bbox_inches='tight')
[…]

### 2018-01-12 resex.tex
version 2018-01-11

@@
 			machine  & standard errors & no errors \\
-			imperfections &with $a_1=b_1=a_2=0$\footnote{Orbit errors are disabled due to different implementation of the $a_1,b_1$ errors in Lifetrac and MAD-X.
+			imperfections &with $a_1=b_1=0$\footnote{Orbit errors are disabled due to different implementation of the $a_1,b_1$ errors in Lifetrac and MAD-X.
 $b_2$ errors are adjusted to yield an average (over 60 seeds) peak $\beta$-beat of 15\% as expected from optics measurements in the LHC.} &  \\

### 2018-01-17 resex.tex
add Daniels part

@@
 \subsection{Excitation with transverse damper and filling scheme\label{sec:adt}}
-\textcolor{red}{Daniel's part}
+The primary function of the transverse feedback system in LHC (also referenced to as ADT) is to provide injection oscillation damping and actively damp the coupled bunch instabilities driven by the machine impedance.
+The main building blocks of the ADT system are stipline pickups at the position Q7 and Q9 at Point 4 of the LHC, connected to the beam position measurement modules at the surface, the digital signal processing modules (mDSPU) and a set of tetrode power amplifiers feeding electrostatic kickers in the RF zone of LHC \cite{}.
+
+Though originally designed to damp the undesired transverse activities only, thanks to the state of the art digital signal processing hardware the damper is being routinely used for sophisticated beam excitations as well (e.g.
+Abort gap and injection gap cleaning, single bunch excitation for tune and linear coupling measurements, machine development etc.).
+
+The transverse feedback is active during the whole LHC cycle, therefore it is of very high importance, that the noise introduced by the system does not contribute to any emittance growth.
+The typical machine cycle requires a short damping time of ~10-20 turns (high feedback gain) for injection oscillation damping, which is increased during the ramp and stable beams to ~50-100 turns (lower feedback gain).
+It had been demonstrated that noise introduced by the transverse feedback system itself is sufficiently low to introduce any measurable emittance growth \cite{}.
+All excitation signals referenced to in this paper, are digitally synthesized in the damper's digital signal processing units and therefore can be considered as ``noise-free''.
+
+The resonant excitation experiment involved simultaneous measurements on three groups of 72 bunches with a dedicated excitation pattern and transverse feedback configuration on each sub-group of 6 bunches as shown in Fig.~\ref{fig:fill}.
+\begin{figure*}
+		\centering
+		\includegraphics[width=1.0\linewidth]{bunchfilling.png}	
+	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2017.}
+\end{figure*}
+
+The excitation in the horizontal and the vertical plane is generated by a different set of signal processing hardware, however during this experiment the signals were synchronized at a turn level.
+This means the bunches of the third group (of 72 bunches) experienced the kick in both horizontal and vertical plane during the same turn.
+
+Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor.
+The measured activity is well in line with the expected excitation pattern.
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{bunchfilling_measured.png}	
+	\caption{\label{fig:fill_meas} Excitation amplitude in horizontal and vertical measured during the 2017 experiment during the $8^{\mathrm{th}}$~turn pulsing.}
+\end{figure}
+
 \begin{itemize}
-	\item error on amplitude
-	\item synchronization of H and V kicker
-	\item random noise on amplitude - is there any?
-	\item bandwidth - why does the damper not affect the nth turn pulsing????
-	\item filling scheme
+	\item error on amplitude \textcolor{red}{missing}
+	\item synchronization of H and V kicker \textcolor{green}{done}
+	\item random noise on amplitude - is there any? \textcolor{green}{done}
+	\item bandwidth - why does the damper not affect the nth turn pulsing???? \textcolor{red}{missing}
+	\item filling scheme \textcolor{green}{done for 2017}, \textcolor{red}{missing for 2016}
 \end{itemize}

### 2018-01-24 resex.tex
add Daniels comments

@@
 The primary function of the transverse feedback system in LHC (also referenced to as ADT) is to provide injection oscillation damping and actively damp the coupled bunch instabilities driven by the machine impedance.
-The main building blocks of the ADT system are stipline pickups at the position Q7 and Q9 at Point 4 of the LHC, connected to the beam position measurement modules at the surface, the digital signal processing modules (mDSPU) and a set of tetrode power amplifiers feeding electrostatic kickers in the RF zone of LHC \cite{}.
+The main building blocks of the ADT system are stipline pickups at the position Q7 and Q9 at Point 4 of the LHC, connected to the beam position measurement modules at the surface, the digital signal processing modules (mDSPU) and a set of tetrode power amplifiers feeding electrostatic kickers in the RF zone of LHC \cite{adt_sum_2008,adt_sum_2011}.
 
@@
 The typical machine cycle requires a short damping time of ~10-20 turns (high feedback gain) for injection oscillation damping, which is increased during the ramp and stable beams to ~50-100 turns (lower feedback gain).
-It had been demonstrated that noise introduced by the transverse feedback system itself is sufficiently low to introduce any measurable emittance growth \cite{}.
+It had been demonstrated that noise introduced by the transverse feedback system itself is sufficiently low to introduce any measurable emittance growth \cite{adt_noise_emit_2017}.
 All excitation signals referenced to in this paper, are digitally synthesized in the damper's digital signal processing units and therefore can be considered as ``noise-free''.
@@
 \begin{figure*}
+	\begin{minipage}[t]{1.0\linewidth}
 		\centering
+		2016 experiment
 		\includegraphics[width=1.0\linewidth]{bunchfilling.png}	
-	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2017.}
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+	\centering
+	2017 experiment
+	\includegraphics[width=1.0\linewidth]{bunchfilling.png}	
+	\end{minipage}
+	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).}
 \end{figure*}
@@
 	\includegraphics[width=1.0\linewidth]{bunchfilling_measured.png}	
-	\caption{\label{fig:fill_meas} Excitation amplitude in horizontal and vertical measured during the 2017 experiment during the $8^{\mathrm{th}}$~turn pulsing.}
+	\caption{\label{fig:fill_meas} Bunch by bunch excitation detected by the real time transverse activity monitor during the 2017 experiment (8th turn pulsing}
 \end{figure}

### 2018-01-24 scripts/lifetrac_sim_results/plot_loss_emit_8H_V_H+V_2017.py
add plotting for random and 8th turn pulsing simulations

@@ -1,0 +1,95 @@
+from LifeDesk import *
+
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+# plot settings
+mpl.rcParams['xtick.labelsize'] = 14
+mpl.rcParams['ytick.labelsize'] = 14
+mpl.rcParams['axes.labelsize']  = 14
+mpl.rcParams['legend.fontsize']  = 12
+mpl.rcParams['legend.fontsize']  = 12
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+# 2017 optics
+# only with random noise
+opt0 = "2017injnocolc15o+19_6erra2b2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+# random noise + 7th pulsing
+optsk = "2017injnocolc15o+19_6erra2b2ut8skran/1/eps3.5/%s%se-8ran1_2e-3/p1e4t1e4s100weight"
+
+# no excitation
+t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
+
+plt.close('all')
+# losses
+plt.figure("losses 2017, 3.5 um emittance")
+plt.clf()
+plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
+for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
+  for amp,lamp in zip(['9_6','19_2'],['--',':']):
+    t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%(plane.replace('+',''),amp)),plt_dir=plt_dir)
+    print os.path.join(basedir,optsk%(plane.replace('+',''),amp))
+    pltlbl = plane.upper()+' '+amp.replace('_','')+' nrad'
+    plt.plot(t.data['time'],t.data['intensity'],linestyle=lamp,color=cplane,label=pltlbl)
+plt.grid(b=True)
+plt.xlabel('time [s]')
+plt.ylabel('relative intensity')
+leg = plt.legend(title='excitation\n8th turn',ncol=1,fontsize=12)
+plt.setp(leg.get_title(),fontsize=12)
[…]

### 2018-01-24 scripts/lifetrac_sim_results/plot_loss_emit_ranH_V_H+V_2017.py

@@ -1,0 +1,95 @@
+from LifeDesk import *
+
+import os
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+
+# plot settings
+mpl.rcParams['xtick.labelsize'] = 14
+mpl.rcParams['ytick.labelsize'] = 14
+mpl.rcParams['axes.labelsize']  = 14
+mpl.rcParams['legend.fontsize']  = 12
+mpl.rcParams['legend.fontsize']  = 12
+
+# directory with simulation results
+basedir = "/home/mfittere/work/MDs/mdelens/simulations/lifetrac"
+# plot directory for paper
+plt_dir = "plots"
+
+# 2017 optics
+# only with random noise
+#opt0 = "2017injnocolc15o+19_6erra2b2u/1/eps3.5/p1e4t1e4s100weight"
+opt0 = "2017injnocolc15o+19_6erra2b2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+# random noise + 7th pulsing
+optsk = "2017injnocolc15o+19_6erra2b2uranadt/1/eps3.5/%s%se-8ran1_2e-3/p1e4t1e4s100weight"
+
+# no excitation
+t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
+
+plt.close('all')
+# losses
+plt.figure("losses 2017, 3.5 um emittance")
+plt.clf()
+plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
+for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
+  for amp,lamp in zip(['1_2','2_4'],['--',':']):
+    t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%(plane.replace('+',''),amp)),plt_dir=plt_dir)
+    pltlbl = plane.upper()+' '+amp.replace('_','')+' nrad'
+    plt.plot(t.data['time'],t.data['intensity'],linestyle=lamp,color=cplane,label=pltlbl)
+plt.grid(b=True)
+plt.xlabel('time [s]')
+plt.ylabel('relative intensity')
+leg = plt.legend(title='random\nexcitation',ncol=1,fontsize=12)
+plt.setp(leg.get_title(),fontsize=12)
[…]

### 2018-01-25 resex.tex
revised until 10th turn pulsing

--- 2018-01-12 resex.tex
@@
 \subsection{Excitation with transverse damper and filling scheme\label{sec:adt}}
-\textcolor{red}{Daniel's part}
-\begin{itemize}
-	\item error on amplitude
-	\item synchronization of H and V kicker
-	\item random noise on amplitude - is there any?
-	\item bandwidth - why does the damper not affect the nth turn pulsing????
-	\item filling scheme
-\end{itemize}
+The primary function of the transverse feedback system in LHC, also referenced to as ADT, is to provide injection oscillation damping and actively damp the coupled bunch instabilities driven by the machine impedance.
+The main building blocks of the ADT system are stipline pickups at the position Q7 and Q9 at Point~4 of the LHC, connected to the beam position measurement modules at the surface, the digital signal processing modules (mDSPU) and a set of tetrode power amplifiers feeding electrostatic kickers in the RF zone of the LHC \cite{adt_sum_2008,adt_sum_2011}.
+
+Though originally designed to damp the undesired transverse activities only, the damper is being routinely used for sophisticated beam excitations as well (e.g.
+Abort gap and injection gap cleaning, single bunch excitation for tune and linear coupling measurements, machine development etc.) thanks to the state of the art digital signal processing hardware.
+
+The transverse feedback is in general active during the whole LHC cycle, therefore it is of very high importance, that the noise introduced by the system does not contribute to any emittance growth.
+The typical machine cycle requires a short damping time of ~10-20 turns (high feedback gain) for injection oscillation damping, which is increased during the ramp and stable beams to ~50-100 turns (lower feedback gain).
+It has been demonstrated that noise introduced by the transverse feedback system itself is sufficiently low to introduce any measurable emittance growth \cite{adt_noise_emit_2017}.
+All excitation signals referenced to in this paper, are digitally synthesized in the damper's digital signal processing units and therefore can be considered as ``noise-free''.
+
+The resonant excitation experiment in 2017 involved simultaneous measurements on three groups of 72 bunches with a dedicated excitation pattern and transverse feedback configuration on each sub-group of 6 bunches.
+In 2016 a similar configuration was used, but with only 48 bunches in total.
+Both schemes are illustrated in Fig.~\ref{fig:fill}.
+\begin{figure*}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		2016 experiment
+		\includegraphics[width=1.0\linewidth]{bunchfilling.png}	
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+	\centering
+	2017 experiment
+	\includegraphics[width=1.0\linewidth]{bunchfilling.png}	
+	\end{minipage}
+	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).}
+\end{figure*}
+
+The excitation in the horizontal and the vertical plane is generated by a different set of signal processing hardware, however during this experiment the signals were synchronized at a turn level.
+This means the bunches of the third group (of 72 bunches) experienced the kick in both horizontal and vertical plane during the same turn.
+
+The ADT kicker deflection angle is calculated from the kicker geometry and the excitation voltage on the deflection plates.
+The excitation voltage used in the experiments is a complicated waveform, generated by a very complex hardware (digital signal processor, chain of transmission lines, low and high power amplifiers).
+The peak excitation voltage (and subsequently the kick) seen by the beam can be only indirectly measured using the probes on each kicker.
+The measured values are therefore believed to be known with an error margin of 10-15\% for the experiments in 2017 and 50\% in 2016.
+The error for 2016 is much larger as no calibration of the ADT was performed beforehand, which was done in 2017.
+As an example Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor.
+The measured activity is well in line with the expected excitation pattern.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.0\linewidth]{bunchfilling_measured.png}	
+	\caption{\label{fig:fill_meas} Bunch by bunch excitation detected by the real time transverse activity monitor during the 2017 experiment (8th turn pulsing}
+\end{figure}
 
@@
 For the preparation and interpretation of the experiments two different types of simulations were performed:
-\begin{enumerate}
-	\item tracking of a Gaussian distribution (Distribution Tracking) to obtain loss rates and emittance growth,
-	\item Frequency map analysis (FMA)~\cite{fmalaskar} for the visualization of resonances.
-\end{enumerate}
-For both simulation types, the tracking code Lifetrac \cite{lifetrac} was used and the simulation parameters are summarized in Table~\ref{tab:sim_param}.
-Further details on the simulations can be found in \cite{md_sim_hel_res_ex_fitterer,md_sim_hel_res_ex_fitterer_update}.
 \begin{table}[b]
 	\caption{\label{tab:sim_param}%
-		Summary of simulation parameters for the Distribution Tracking and FMA analysis. For details see \cite{md_sim_hel_res_ex_fitterer,md_sim_hel_res_ex_fitterer_update}.
+		Summary of simulation parameters for the Distribution Tracking and FMA analysis. For details see \cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
 	}
@@
 \end{table}
+\begin{enumerate}
+	\item tracking of a Gaussian distribution (Distribution Tracking) to obtain loss rates and emittance growth,
+	\item Frequency map analysis (FMA)~\cite{fmalaskar} for the visualization of resonances.
+\end{enumerate}
+For both simulation types, the tracking code Lifetrac \cite{lifetrac} was used.
+The simulation parameters are summarized in Table~\ref{tab:sim_param}.
+Further details on the simulations can be found in \cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
 
 \subsection{Simulation and experimental results\label{sec:simex}}
-In the following we will summarize the simulation and experimental results. We will jump forth and back between both as the experimental results are best understood on the base of the simulations.
+In the following we will summarize the simulation and experimental results.
+We will jump forth and back between both as the experimental results are best understood on the base of the simulations.
+In summary, the most efficient pulsing patterns could be successfully predicted in simulations (Sec.~\ref{sec:pattern}).
+Quantitative predictions of loss rates and emittance growth proved however challenging as both observables are influenced in addition by the natural noise present in the LHC originating for example from ground motion and magnetic field ripple.
+The observed effect is then the result of the interaction of the external excitation with the natural noise.
+At injection, this natural noise component is not known for the LHC.
+Beside the natural noise, also collective effects like intra-beam scattering and electron cloud influence the time evolution of emittance and losses.
+Both are not taken into account in the simulations.
 
-In summary, the most efficient pulsing patterns and the general tendency of the distribution changes could be successfully predicted in simulations (Sec.~\ref{sec:pattern}).
-Quantitative predictions of loss rates and emittance growth though failed as both observables are influenced also by the natural noise present in the LHC like ground motion and tune modulation.
-At injection this natural noise component is now known for the LHC.
-The observed effect is then the result of the interaction of the external excitation with the natural noise present.
-
-In case of a resonant excitation, the beam distribution adjusts to a new equilibrium distribution.
-In the presence of only a resonant excitation, this distribution is stable and stays unchanged.
-If the resonant excitation interacts with the natural noise sources, the beam distribution still changes to a new equilibrium distribution but this change is also accompanied by a resonant excitation amplitude dependent emittance growth.
-Both effects were experimentally observed in case of the $10^{\mathrm{th}}$~turn pulsing (Sec.~\ref{sec:simex10}).
-
-In case of an excitation with random uniform noise, the shape of the distribution does stay Gaussian.
-The interaction with the natural noise source then only manifests itself in an underestimation of loss rates and emittance growth (see Sec.~\ref{sec:simexran}).
-This effect was also studied in more detail in \cite{md1433_noise_top_energy,md_noise_bbLHC} for colliding beams at the LHC.
+The general dynamics of a resonant excitation seen in simulations is a fast adjustment to a new equilibrium distribution.
+In the presence of only a resonant excitation, this distribution is stable with a constant emittance as shown for example in Figs.~\ref{fig:patternsim} and \ref{fig:10thsim}.
+If the resonant excitation interacts with the natural noise sources, the beam distribution still changes to a new equilibrium distribution but this change is also accompanied by a resonant excitation amplitude dependent emittance growth as seen for example in Figs.~\ref{fig:10thsim},\ref{fig:7thsim} and \ref{fig:8thsim}, and also observed experimentally, see for example Figs.~\ref{fig:10thexp} and ~\ref{fig:7thexp}.
+An excitation with random uniform noise on the other hand results in constant emittance growth without any initial adjustment phase.
+The interaction with the natural noise source then only manifests itself in an underestimation of loss rates and emittance growth.
+This is visible in simulations, see Fig.~\ref{fig:ransim} as well as experimentally, see Fig.~\ref{fig:ranexp}.
+The effect of a random excitation was also studied in more detail in \cite{md1433_noise_top_energy,md_noise_bbLHC} for colliding beams at the LHC.
 
 \subsubsection{Prediction of the most efficient pulsing pattern\label{sec:pattern}}
-\begin{figure}[b]
+\begin{figure}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
-The excitation was applied in both planes with an amplitude of 96~nrad.}
+The resonant excitation is applied in both planes with an amplitude of 96~nrad.
+No random noise component is added.}
 \end{figure}
-In order to predict the most efficient pulsing pattern, the loss rate and emittance growth from the Distribution Tracking are compared for the different pulsing patterns.
-The results for the 2016 optics and tune are shown in Fig.~\ref{fig:patternsim}.
-The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and the random excitation.
-As the bunch length decreases accordingly, the losses are mostly associated to off-momentum particles.
-Only for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as random excitation a considerable emittance growth is visible, where the random excitation has by far the strongest effect compared to any $k^{\mathrm{th}}$~turn pulsing pattern.
-
+The most efficient pulsing pattern can be predicted by the comparison of the loss rate and emittance growth from the Distribution Tracking for the different pulsing patterns.
+As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pulsing pattern.
+The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and a uniform random excitation.
+As the bunch length decreases in accordance with the losses, the losses can be mostly associated to off-momentum particles.
+Only for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as for a random excitation considerable emittance growth is visible.
+Compared to any $k^{\mathrm{th}}$~turn pulsing pattern, the random excitation has by far the strongest effect.
+The sensitivity to $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are the strong sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer}.
+This suggests that latter are the source of non-linearity responsible for this sensitivity.
+The driven resonances are revealed by the FMA analysis shown in Fig.~\ref{fig:patternfma}.
+The $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing for an excitation in both planes.
+As octupoles can only drive even resonances, the source of the $7Q_x$ resonances are the sextupoles while the octupoles only generate the tune footprint.
+The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect thus is based on the magnetic field errors assigned, explicitly also the seed chosen for the simulation.
 \begin{figure}[h]
@@
 The excitation is 120~nrad in the corresponding plane.
-The absence of a strong excitation of any resonance in case of $7^{\mathrm{th}}$~turn pulsing only in V and the strong excitation in case of pulsing only in H further confirms that indeed the $7Q_x$ resonance is driven.
-For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V.}
+The absence of a strong excitation of any resonance in case of $7^{\mathrm{th}}$~turn pulsing only in V and the strong excitation in case of pulsing only in H further confirms that the $7Q_x$ resonance is the main resonance driven.
+For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V (see Appendix~\ref{sec:fma:10}, Fig.~\ref{fig:fma:10}).}
 \end{figure}
-The sensitivity to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are the strong sextupoles and octupoles \cite{md_sim_hel_res_ex_fitterer}.
-The FMA analysis shown in Fig.~\ref{fig:patternfma} reveals that
-the $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing.
-As octupoles can only drive even resonances, the source of the $7Q_x$ resonances are the sextupoles while the octupoles only generate the tune footprint.
-The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect thus is based on the magnetic field errors assigned, explicitly also the seed chosen for the simulations.
 
@@
 	\end{minipage}	
-	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from simulation results (Distribution Tracking) based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
-The excitation was applied in both planes with an amplitude of 96~nrad.
+	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn pulsing (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from Distribution Tracking simulations based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ and standard errors.
+The excitation is applied in both planes with an amplitude of 96~nrad.
 The residual is defined as the final distribution (here $10^4$~turns) minus the initial distribution (here $10^2$~turns), and the ratio as the final distribution divided by the initial distribution.
-The red line in the residual shows the difference between the Gaussian fit of the distribution and the distribution itself.}
+The red line in the plot of the residual is the difference between the Gaussian fit of the distribution and the distribution itself.}
 \end{figure}
-For the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing the emittance also appears to start at an increased initial value and then stays almost constant.
-This behavior is due to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution.
-As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}.
-Besides an increased emittance, the new distribution also becomes non-Gaussian as can be seen from the residual in respect to the Gaussian distribution in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:patternhist}).
-For comparison, also the distribution in both planes for a random excitation is shown This initial fast change of the beam distribution was also to the knowledge of the authors for the first time experimentally observed (Sec.~\ref{sec:simex10} and Sec.~\ref{sec:simex7}).
-The timescale of this fast adjustment is experimentally however much slower than predicted in simulations - on the timescale of minutes compared to seconds.
+
+As already addressed earlier, the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time.
+This behavior is typical for any $n^{\mathrm{th}}$ turn pulsing and can be ascribed to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution.
+As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}, and also any other simulations over $10^6$ turns presented in this paper.
+Besides an increased emittance, the new distribution also becomes non-Gaussian.
+This can be seen by means of the residual in respect to the Gaussian distribution illustrated for example for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in Fig.~\ref{fig:patternhist}.
+For the purpose of comparison, also the distribution in both planes for a random excitation is shown.
+This initial fast change of the beam distribution is not only an artifact of the simulation, but has been, to the knowledge of the authors, also for the first time experimentally observed as will be presented later in this paper.
+The timescale of this fast adjustment is in experiments however much slower than predicted in simulations - on the timescale of minutes compared to seconds.
 
 \subsubsection{10th turn pulsing\label{sec:simex10}}
-The $10^{\mathrm{th}}$~turn pulsing was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
+The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
 \begin{enumerate}
-	\item increase of the loss rate with the excitation amplitude (Fig.~\ref{fig:10thexp}, \ref{fig:10thexpsim}),
-	\item increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane (Fig.~\ref{fig:10thexp}).
-	\item change of the beam distribution due to the resonant excitation (Fig.~\ref{fig:10thexp}, \ref{fig:10thexpprof})
+	\item Increase of the loss rate with the excitation amplitude as shown in Figs.~\ref{fig:10thexp} and \ref{fig:10thexpsim}.
+	\item Increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane shown in Fig.~\ref{fig:10thexp}.
+	\item Change of the beam distribution due to the resonant excitation illustrated in Fig.~\ref{fig:10thexp} and \ref{fig:10thexpprof}.
+	\item As loss rates, emittance growth and distribution stayed unchanged for the $4$~reference bunches, the above observations can be truly associated with the effect of the resonant excitation.	
 \end{enumerate}
-As loss rates, emittance growth and distribution stayed unchanged for the 4~reference bunches, the above observations can be truly associated with the effect of the resonant excitation.
 \begin{figure}[h]
@@
 
-Besides the determination of the scaling of the loss rate and emittance growth with the excitation amplitude as required for the specification of the hollow electron lens, the most novel observation is the direct measurement of the change of the beam distribution, here with the Beam Synchrotron Radiation Telescope (BSRT) \cite{bsrtprofinj}.
-As an example for the change in distribution, the vertical profile of a reference bunch and one bunch experiencing the maximum excitation are compared in Fig.~\ref{fig:10thexpprof}.
-The distribution of the excited bunch clearly changes and becomes non-Gaussian, while the reference bunch stays almost unchanged.
+Besides the determination of the scaling of the loss rate and emittance growth with the excitation amplitude as required for the specification of the hollow electron lens, the most novel observation is the direct measurement of the change of the beam distribution, here with the Beam Synchrotron Radiation Telescope (BSRT)~\cite{bsrtprofinj}.
+As an example for the change in distribution, the vertical profile of a reference bunch and one bunch experiencing the maximum excitation of $5\cdot\Delta A=A_{\mathrm{max}}$ are compared in Fig.~\ref{fig:10thexpprof}.
+The distribution of the excited bunch clearly changes and becomes non-Gaussian, while the reference bunch stays unchanged.
 \begin{figure}[h]
@@
 		\centering
+		reference bunch, no excitation
 		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_50.png}
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_1300.png}
+		excited bunch, $10^{\mathrm{th}}$~turn V		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_1300.png}
 	\end{minipage}
@@
 
-Comparing these experimental results with the simulations including only the $10^{\mathrm{th}}$~turn pulsing (solid line, Fig.~\ref{fig:10thsim}), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth seen experimentally.
+Comparing these experimental results with the simulations including only the excitation through the $10^{\mathrm{th}}$~turn pulsing and no random component and emittance growth only in the vertical plane (solid line in Fig.~\ref{fig:10thsim}), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth observed in experiments.
 \begin{figure}[h]
@@
 So obviously there is an essential ingredient missing.
-As it turns out, the constant emittance can be changed to an excitation amplitude dependent emittance growth by adding a random dipole noise component with uniform distribution (dotted line in Fig.~\ref{fig:10thsim}).
-As a first estimate the value obtained in \cite{md1433_noise_top_energy,md_noise_bbLHC} for 6.5~TeV was used and scaled by the magnetic rigidity to the injection energy of 450~GeV, as no noise estimate is available at injection.
+As it turns out, the constant emittance can be changed to an excitation amplitude dependent emittance growth by adding a random dipole noise component with uniform distribution representing the natural noise present in the LHC (dotted line in Fig.~\ref{fig:10thsim}).
+As no noise estimate is available at injection for the LHC, a first estimate is obtained by scaling the estimate for 6.5~TeV \cite{md1433_noise_top_energy,md_noise_bbLHC} by the magnetic rigidity to the injection energy of 450~GeV.
 This yields a maximum kick amplitude at the transverse damper of approximately
@@
 \end{equation}
-Adding this random noise component, the simulations then yield correctly an excitation amplitude dependent emittance growth in the vertical plane only after a first fast adjustment of the beam distribution visible as an increased initial value.
-
-Adding a random noise component has in addition the effect of increasing the loss rate.
-As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced, but a large difference still prevails (Fig.~\ref{fig:10thexpsim}).
+Adding this random noise component, the simulations then yield correctly an excitation amplitude dependent emittance growth in the vertical plane after a first fast adjustment of the beam distribution visible as an increased initial value.
+Adding a random noise component has also the effect of increasing the loss rate.
+As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced, but a large difference still prevails as illustrated in Fig.~\ref{fig:10thexpsim} by means of a qualitative comparison of the loss rates obtained in experiments and simulations.
 \begin{figure}[h]
@@
 	\caption{\label{fig:10thexpsim} Comparison of scaling of loss rates with the $10^{\mathrm{th}}$~turn pulsing excitation amplitude.
-The experimental results are shown in black (maximum excitation with 48~nrad) and blue (maximum excitation with 96~nrad) and the simulation results (2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad) in red.}
+The experimental results are shown in black (maximum excitation with 48~nrad) and blue (maximum excitation with 96~nrad).
+The excitation amplitude error is 50\% due to limited measurement accuracy and the error on the loss rate is the pure statistical error.
+The simulation results (2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad) are shown in red.}
 \end{figure}
@@
 
+\clearpage
+
 \appendix
+\section{FMA analysis for 10th turn pulsing and 2016 optics and tune}
+\label{sec:fma:10}
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		no excitation
+		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$10^{\mathrm{th}}$ H+V
+		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$10^{\mathrm{th}}$ H
+		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skh_dp0_ord10.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$10^{\mathrm{th}}$ V
+		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skv_dp0_ord10.png}
+	\end{minipage}
+	\caption{\label{fig:fma:10} FMA analysis for $10^{\mathrm{th}}$~turn pulsing based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
+The excitation is 120~nrad in the corresponding plane.
+There is no significant difference in the FMA analysis between pulsing only in H, only in V or in H+V.}
+\end{figure}
 
-\section{Appendixes}
 \section{Results of $7^{\mathrm{th}}$~turn pulsing during 2016 and 2017 experiments}

### 2018-01-27 resex.tex
finished section about 7th turn pulsing

@@
 		\centering
-		2016 experiment
+		2016 experiment \textcolor{red}{(update picture)}
 		\includegraphics[width=1.0\linewidth]{bunchfilling.png}	
@@
 \end{figure}
-The most efficient pulsing pattern can be predicted by the comparison of the loss rate and emittance growth from the Distribution Tracking for the different pulsing patterns.
-As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pulsing pattern.
-The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and a uniform random excitation.
-As the bunch length decreases in accordance with the losses, the losses can be mostly associated to off-momentum particles.
-Only for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as for a random excitation considerable emittance growth is visible.
-Compared to any $k^{\mathrm{th}}$~turn pulsing pattern, the random excitation has by far the strongest effect.
-The sensitivity to $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are the strong sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer}.
-This suggests that latter are the source of non-linearity responsible for this sensitivity.
-The driven resonances are revealed by the FMA analysis shown in Fig.~\ref{fig:patternfma}.
-The $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing for an excitation in both planes.
-As octupoles can only drive even resonances, the source of the $7Q_x$ resonances are the sextupoles while the octupoles only generate the tune footprint.
-The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect thus is based on the magnetic field errors assigned, explicitly also the seed chosen for the simulation.
 \begin{figure}[h]
@@
 The absence of a strong excitation of any resonance in case of $7^{\mathrm{th}}$~turn pulsing only in V and the strong excitation in case of pulsing only in H further confirms that the $7Q_x$ resonance is the main resonance driven.
-For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V (see Appendix~\ref{sec:fma:10}, Fig.~\ref{fig:fma:10}).}
+For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V (see Appendix~\ref{app:sec:fma:10}, Fig.~\ref{app:fig:fma:10}).}
 \end{figure}  
+The most efficient pulsing pattern can be predicted by the comparison of the loss rate and emittance growth from the Distribution Tracking for the different pulsing patterns.
+As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pulsing pattern.
+The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and a uniform random excitation.
+As the bunch length decreases in accordance with the losses, the losses can be mostly associated to off-momentum particles.
+Only for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as for a random excitation considerable emittance growth is visible.
+Compared to any $k^{\mathrm{th}}$~turn pulsing pattern, the random excitation has by far the strongest effect.
+The sensitivity to $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are the strong sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer}.
+This suggests that latter are the source of non-linearity responsible for this sensitivity.
+The driven resonances are revealed by the FMA analysis shown in Fig.~\ref{fig:patternfma}.
+The $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing for an excitation in both planes.
+As octupoles can only drive even resonances, the source of the $7Q_x$ resonances are the sextupoles while the octupoles only generate the tune footprint.
+The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect thus is based on the magnetic field errors assigned, explicitly also the seed chosen for the simulation.
 
@@
 
-\subsubsection{10th turn pulsing\label{sec:simex10}}
+\subsubsection{$10^{\mathrm{th}}$ turn pulsing\label{sec:simex10}}
 The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
@@
 		\centering
-		\includegraphics[width=0.49\linewidth]{bunch_intensity_v10th_no_damper_avg.png}
+		\includegraphics[width=0.49\linewidth]{2016_bunch_intensity_v10th_no_damper_avg.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{emith_avg_rel_v10th_no_damper.png}
+		\includegraphics[width=1.0\linewidth]{2016_emith_avg_rel_v10th_no_damper.png}
 	\end{minipage}	
@@
 		\centering
-		\includegraphics[width=1.0\linewidth]{emitv_avg_rel_v10th_no_damper.png}
+		\includegraphics[width=1.0\linewidth]{2016_emitv_avg_rel_v10th_no_damper.png}
 	\end{minipage}	
-	\caption{\label{fig:10thexp} Relative intensity (top) and emittance (bottom) obtained during the 2016 experiments.
+	\caption{\label{fig:10thexp} Relative intensity (top) and relative emittance (bottom) averaged over the bunches experiencing the same excitation amplitude during the 2016 experiments.
 For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Sec.~\ref{sec:adt}).
@@
 \end{figure}
-
-Besides the determination of the scaling of the loss rate and emittance growth with the excitation amplitude as required for the specification of the hollow electron lens, the most novel observation is the direct measurement of the change of the beam distribution, here with the Beam Synchrotron Radiation Telescope (BSRT)~\cite{bsrtprofinj}.
-As an example for the change in distribution, the vertical profile of a reference bunch and one bunch experiencing the maximum excitation of $5\cdot\Delta A=A_{\mathrm{max}}$ are compared in Fig.~\ref{fig:10thexpprof}.
-The distribution of the excited bunch clearly changes and becomes non-Gaussian, while the reference bunch stays unchanged.
 \begin{figure}[h]
@@
 		\centering
-		reference bunch, no excitation
+		reference bunch,\\ no excitation
 		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_50.png}
@@
 		\centering
-		excited bunch, $10^{\mathrm{th}}$~turn V		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_1300.png}
+		excited bunch,\\ $10^{\mathrm{th}}$~turn V		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_1300.png}
 	\end{minipage}
@@
 The profiles are taken at the end of the $10^{\mathrm{th}}$~turn pulsing in V.
-The black line in the plot of the residual is the current profile minus the initial profile and is a measure for the overall change of the distribution.
+The black line in the plot of the residual is the current profile minus the profile before applying the excitation and is a measure for the overall change of the distribution.
 The red line is the current profile minus its Gaussian fit, which gives and indication for the deviation of the distribution from a Gaussian distribution.
@@
 \end{figure}
-
-Comparing these experimental results with the simulations including only the excitation through the $10^{\mathrm{th}}$~turn pulsing and no random component and emittance growth only in the vertical plane (solid line in Fig.~\ref{fig:10thsim}), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth observed in experiments.
 \begin{figure}[h]
@@
 \end{figure}
+
+Besides the determination of the scaling of the loss rate and emittance growth with the excitation amplitude as required for the specification of the hollow electron lens, the most novel observation is the direct measurement of the change of the beam distribution, here with the Beam Synchrotron Radiation Telescope (BSRT)~\cite{bsrtprofinj}.
+As an example for the change in distribution, the vertical profile of a reference bunch and one bunch experiencing the maximum excitation of $5\cdot\Delta A=A_{\mathrm{max}}$ are compared in Fig.~\ref{fig:10thexpprof}.
+The distribution of the excited bunch clearly changes and becomes non-Gaussian, while the reference bunch stays unchanged.
+
+Comparing these experimental results with the simulations including only the excitation through the $10^{\mathrm{th}}$~turn pulsing and no random component and emittance growth only in the vertical plane (solid line in Fig.~\ref{fig:10thsim}), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth observed in experiments.
 Instead, the simulations yield a constant emittance after an initial increase during the first $10^4$~turns or first second.
@@
 Adding a random noise component has also the effect of increasing the loss rate.
-As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced, but a large difference still prevails as illustrated in Fig.~\ref{fig:10thexpsim} by means of a qualitative comparison of the loss rates obtained in experiments and simulations.
+As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced, but a large difference still prevails as illustrated in Fig.~\ref{fig:10thexploss} by means of a qualitative comparison of the loss rates obtained in experiments and simulations.
 \begin{figure}[h]
 		\centering
-		\includegraphics[width=1.0\linewidth]{scale_amp_10v_ran_lbllong_sim.png}
-	\caption{\label{fig:10thexpsim} Comparison of scaling of loss rates with the $10^{\mathrm{th}}$~turn pulsing excitation amplitude.
+		\includegraphics[width=1.0\linewidth]{2016_scale_amp_10v_ran_lbllong_sim.png}
+	\caption{\label{fig:10thexploss} Comparison of scaling of loss rates with excitation amplitude for $10^{\mathrm{th}}$~turn pulsing.
 The experimental results are shown in black (maximum excitation with 48~nrad) and blue (maximum excitation with 96~nrad).
@@
 
-\subsubsection{8th turn pulsing\label{sec:simex8}}
-\subsubsection{random excitation\label{sec:simexran}}
-
-\subsubsection{7th turn pulsing\label{sec:simex7}}
-The $7^{\mathrm{th}}$~turn pulsing pattern is the only pulsing pattern which has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the results.
+\subsubsection{$7^{\mathrm{th}}$ turn pulsing\label{sec:simex7}}
+\begin{table}[h]
+	\caption{\label{tab:7thexp}%
+		Overview of the effect of $7^{\mathrm{th}}$~turn pulsing during the two resonant excitation experiments in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}.
+The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V for horizontal and vertical at the same time.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lcc}
+			Parameter & Experiment 2017 & Experiment 2016  \\
+			\colrule
+			optics & 2017 injection optics & 2016 injection optics \\
+			tune $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\\hline
+			excitation plane  & H,V and H+V&  H\\
+			losses & similar loss rates for  & large losses for \\
+			& pulsing in H,V &  pulsing in H \\
+			&  and H+V & \\
+			hor. emit. growth & no & yes, very small \\
+			vert. emit. growth & yes, for pulsing in & no\\
+			&  V and H+V & \\
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
+The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the results.
 However, in 2017 the machine tune was changed from (64.28,59.31) to (62.27,60.295) in standard operation accompanied by a slight change in optics.
-The attempt to change back to the fractional tune of (.28,.31) as chosen in 2016 lead to large losses during the experiments in 2017 and therefore the 2017 working point of (62.27,60.295) was used during the experiment.
-This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016  (Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (Fig.~\ref{fig:7th2017fma}) and also a change in the affected amplitudes of the beam distribution.
+The attempt to change back to the fractional tune of (.28,.31) as chosen in 2016 lead to large losses during the experiment in 2017 and therefore the 2017 working point of (62.27,60.295) was used instead.
+This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016 (Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (Fig.~\ref{fig:7th2017fma}).
 \begin{figure}[h]
@@
 \end{figure}
-Consequently the effects on the beam also differ for both experiments (see Table~\ref{tab:7thexp} for a general overview and Fig.~\ref{} and Appendix~\ref{} for more details).
-\begin{table}[h]
-	\caption{\label{tab:7thexp}%
-		Overview of the effect of $7^{\mathrm{th}}$~turn pulsing during the two resonant excitation experiments in 2016 and 2017~\cite{resexmd2016,resexmd2017}.
-The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V for horizontal and vertical at the same time.
-	}
-	\begin{ruledtabular}
-		\begin{tabular}{lcc}
-			Parameter & Experiment 2017 & Experiment 2016  \\
-			\colrule
-			tune $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\\hline
-			excitation plane  & H,V and H+V&  H\\
-			losses & similar loss rates for  & losses for pulsing \\
-			& pulsing in H,V & in H \\
-			&  and H+V & \\
-			hor. emit. growth & no & yes, very small \\
-			vert. emit. growth & yes, for pulsing in V & no\\
-			&  and H+V & \\
-		\end{tabular}
-	\end{ruledtabular}
-\end{table}
-Simulations more or less fail completely (Fig.~\ref{fig:7thsim}). Higher amplitudes have been used for simulations to see noticable effect.
+Consequently, the effects on the beam also differ as can be illustrated by a comparison of the Distribution Tracking for both experiments~(Fig.~\ref{fig:7thsim}).
 \begin{figure}[h]
@@
 
-\begin{itemize}
-	\item Don't understand why there is emittance growth in H for 2016. There is no emittance growth in the experiment.
-	\item repeat simulations for 2017 for all three planes and with random noise => check if there is only emittance growth in V \textcolor{red}{done}.
-	\item experimentally: 2016 + 2017 losses in all cases, only emittance growth in V in 2017
-\end{itemize}
+\begin{figure}[h]
+	%	\begin{minipage}[t]{0.49\linewidth}
+	%		\centering
+	%		2016 experiment
+	%		\includegraphics[width=1.0\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
+	%	\end{minipage}	
+	%	\begin{minipage}[t]{0.49\linewidth}
+	%		\centering
+	%		2017 experiment
+	%		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
+	%	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		2016 experiment, $7^\mathrm{th}$ H	\includegraphics[width=1.0\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		2017 experiment, $7^\mathrm{th}$ H+V
+		\includegraphics[width=1.0\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
+	\end{minipage}	
+	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude.}
+\end{figure}
+Similar changes of the emittance are also experimentally observed as shown in Fig.~\ref{fig:7thexp}, where emittance growth was observed only in the horizontal plane for 2016 for pulsing in H, and only in the vertical plane for 2017 for pulsing in V and H+V (for pulsing in H no emittance growth is observed).
+Just as for the $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:10thexp}), the emittance growth can be divided also for the $7^{\mathrm{th}}$~turn pulsing in a phase of fast adjustment of the beam distribution and the reach of a new equilibrium state (Fig.~\ref{fig:7thexp}).
+Also in this case, the distribution changes are also directly visible on the beam profiles measured with the BSRT (Fig.~\ref{fig:7thexpprof}).
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch,\\ no excitation
+		\includegraphics[width=1.0\linewidth]{profile_hv7th_slot_2862.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ $7^{\mathrm{th}}$~turn h+V		\includegraphics[width=1.0\linewidth]{profile_hv7th_slot_2696.png}
+	\end{minipage}
+	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V.
+The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
+\end{figure}
+
+The loss rates measured during the experiment however differ from the simulations in terms of for which case the largest rate is observed as well as quantitatively for the specific cases.
+For example smaller losses are expected for the 2016 experiments compared to the 2017 experiments from simulations, while experimentally higher loss rates were measured in 2016 (Fig.~\ref{fig:7thexploss}).
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.0\linewidth]{2017_scale_amp_7h_lbllong.png}
+	\caption{\label{fig:7thexploss} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in H and data obtained during the 2016 experiment (black and blue) and 2017 experiment (red and yellow).
+The excitation amplitude error is 50\% in 2016 and could be reduced to 15\% in 2017.}
+\end{figure}
+
+\subsubsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
+Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation.
+With this motivation in mind the $8^{\mathrm{th}}$ turn pulsing pattern was studied, for which a small effect is expected in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing.
+In the Distribution Tracking, the excitation amplitudes had to be increased to 96~nrad to have a visible effect
+\subsubsection{random excitation\label{sec:simexran}}
+
 \subsubsection{Effect of the transverse damper\label{sec:damp}}
@@
 
-In general:
-\begin{itemize}
-	\item repeat simulations for each kth turn pattern with only b2 errors (\textcolor{green}{done})
-	\item show results when adding random noise -> amplitude dependent emittance growth + losses
-	\item FMA plot of 7th (2017) -> explain why in H and in V both losses, FMA plot of 10th (2016), FMA plot of 8th and random
-	\item beam distribution change for 10th
-\end{itemize}
-\begin{itemize}
-	\item 7th turn: small emittance but high losses
-	\item 8th turn: almost no effect
-	\item 10th turn: adjustment of distribution -> emittance growth + small losses
-	\item random: constant filamentation -> emittance growth + later strong losses when beam hits aperature
-
-\end{itemize}
-
 \section{Summary\label{sec:sum}}
+We now want to close the loop to the effects on the beam core for hollow electron lenses in HL-LHC.
+For DC operation, no influence on the beam core are are to be expected.
+To study the effect of pulsed operation beam experiments were conducted in 20216 and 2017 at injection.
+The experimental results showed that for the kick amplitudes expected for the e-lens, considerable effects are to be expected.
+The random pulsing mode is the most sensitive pulsing pattern and as it excites all beam frequencies results in emittance growth in the plane of excitation and losses once the beam size increased enough to hit the physical aperture.
+The resonant excitation or nth turn pulsing on the other hand features a rich non-linear dynamics as it excites only certain resonances and diversity in beam distribution and diffusion changes.
+This means, that each pulsing patterns results in different loss rates and emittance growth.
+Typical for this kind of excitation is also the adjustment of the beam distribution to a new equilibrium state.
 
@@
 \section{FMA analysis for 10th turn pulsing and 2016 optics and tune}
-\label{sec:fma:10}
+\label{app:sec:fma:10}
 \begin{figure}[h]
@@
 	\end{minipage}
-	\caption{\label{fig:fma:10} FMA analysis for $10^{\mathrm{th}}$~turn pulsing based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
+	\caption{\label{app:fig:fma:10} FMA analysis for $10^{\mathrm{th}}$~turn pulsing based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
 The excitation is 120~nrad in the corresponding plane.
@@
 
-\section{Results of $7^{\mathrm{th}}$~turn pulsing during 2016 and 2017 experiments}
+\section{Results of $7^{\mathrm{th}}$~turn pulsing during 2016 and 2017 experiments (additional figures)}
+\label{app:sec:simexp:7}
 

### 2018-01-27 scripts/lifetrac_sim_results/plot_loss_emit_7H_V_H+V_2017_7H_2016.py

@@ -32,6 +32,8 @@ optsk = "2017injnocolc15o+19_6erra2b2ut7skran/1/eps3.5/%s%se-8ran1_2e-3/p1e4t1e4
 t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
 t0_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0_16),plt_dir=plt_dir)
 
+grid_flag = True
+
 plt.close('all')
 # losses
 plt.figure("losses, 3.5 um emittance")
@@ -49,7 +51,7 @@ for amp,lamp in zip(['2_4','4_8'],['--',':']):
   t_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk_16%(amp)),plt_dir=plt_dir)
   pltlbl = 'H '+amp.replace('_','')+' nrad'
   plt.plot(t_16.data['time'],t_16.data['intensity'],linestyle=lamp,color='orange',label=pltlbl)
-#plt.grid(b=True)
+plt.grid(b=grid_flag)
 plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
 handles,labels = plt.gca().get_legend_handles_labels()
@@ -74,7 +76,7 @@ for amp,lamp in zip(['2_4','4_8'],['--',':']):
   t_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk_16%(amp)),plt_dir=plt_dir)
   pltlbl = 'H '+amp.replace('_','')+' nrad'
   plt.plot(t_16.data['time'],t_16.data['sigm'],linestyle=lamp,color='orange',label=pltlbl)
-#plt.grid(b=True)
+plt.grid(b=grid_flag)
 plt.xlabel('time [s]')
 plt.ylabel('bunch length [cm]')
 handles,labels = plt.gca().get_legend_handles_labels()
@@ -100,7 +102,7 @@ for phv,p12 in zip(['hor.','vert.'],'12'):
     t_16 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk_16%(amp)),plt_dir=plt_dir)
     pltlbl = 'H '+amp.replace('_','')+' nrad'
     plt.plot(t_16.data['time'],t_16.data['emit%s'%p12],linestyle=lamp,color='orange',label=pltlbl)
-#  plt.grid(b=True)
+  plt.grid(b=grid_flag)
[…]

### 2018-01-27 scripts/lifetrac_sim_results/plot_loss_emit_8H_V_H+V_2017.py

@@ -25,6 +25,8 @@ optsk = "2017injnocolc15o+19_6erra2b2ut8skran/1/eps3.5/%s%se-8ran1_2e-3/p1e4t1e4
 # no excitation
 t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
 
+grid_flag = False
+
 plt.close('all')
 # losses
 plt.figure("losses 2017, 3.5 um emittance")
@@ -33,12 +35,12 @@ plt.plot(t0.data['time'],t0.data['intensity'],'k-',label='0 nrad')
 for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
   for amp,lamp in zip(['9_6','19_2'],['--',':']):
     t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%(plane.replace('+',''),amp)),plt_dir=plt_dir)
-    print os.path.join(basedir,optsk%(plane.replace('+',''),amp))
     pltlbl = plane.upper()+' '+amp.replace('_','')+' nrad'
     plt.plot(t.data['time'],t.data['intensity'],linestyle=lamp,color=cplane,label=pltlbl)
-plt.grid(b=True)
+plt.grid(b=grid_flag)
 plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
+plt.ylim(0.99975,0.999825)
 leg = plt.legend(title='excitation\n8th turn',ncol=1,fontsize=12)
 plt.setp(leg.get_title(),fontsize=12)
 plt.savefig(os.path.join(plt_dir,'2017injerra2b2uran1_2e-3_8th_3_5um_intensity.png'),bbox_inches='tight')
@@ -51,9 +53,10 @@ for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
     t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%(plane.replace('+',''),amp)),plt_dir=plt_dir)
     pltlbl = plane.upper()+' '+amp.replace('_','')+' nrad'
     plt.plot(t.data['time'],t.data['sigm'],linestyle=lamp,color=cplane,label=pltlbl)
-plt.grid(b=True)
+plt.grid(b=grid_flag)
 plt.xlabel('time [s]')
 plt.ylabel('bunch length [cm]')
+plt.ylim(9.1905,9.1925)
[…]

### 2018-01-27 scripts/lifetrac_sim_results/plot_loss_emit_ranH_V_H+V_2017.py

@@ -18,13 +18,14 @@ plt_dir = "plots"
 
 # 2017 optics
 # only with random noise
-#opt0 = "2017injnocolc15o+19_6erra2b2u/1/eps3.5/p1e4t1e4s100weight"
-opt0 = "2017injnocolc15o+19_6erra2b2uran/1/eps3.5/1_2e-3/p1e4t1e4s100weight"
+opt0 = "2017injnocolc15o+19_6erra2b2u/1/eps3.5/p1e4t1e4s100weight"
 # random noise + 7th pulsing
-optsk = "2017injnocolc15o+19_6erra2b2uranadt/1/eps3.5/%s%se-8ran1_2e-3/p1e4t1e4s100weight"
+optsk = "2017injnocolc15o+19_6erra2b2uranadt/1/eps3.5/%s%se-8/p1e4t1e4s100weight"
 
 # no excitation
 t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
+
+grid_flag = True
 
 plt.close('all')
 # losses
@@ -36,9 +37,10 @@ for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
     t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%(plane.replace('+',''),amp)),plt_dir=plt_dir)
     pltlbl = plane.upper()+' '+amp.replace('_','')+' nrad'
     plt.plot(t.data['time'],t.data['intensity'],linestyle=lamp,color=cplane,label=pltlbl)
-plt.grid(b=True)
+plt.grid(b=grid_flag)
 plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
+#plt.ylim(0.99975,0.999825)
 leg = plt.legend(title='random\nexcitation',ncol=1,fontsize=12)
 plt.setp(leg.get_title(),fontsize=12)
 plt.savefig(os.path.join(plt_dir,'2017injerra2b2u_ranadt_3_5um_intensity.png'),bbox_inches='tight')
@@ -51,9 +53,10 @@ for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
     t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optsk%(plane.replace('+',''),amp)),plt_dir=plt_dir)
     pltlbl = plane.upper()+' '+amp.replace('_','')+' nrad'
[…]

### 2018-01-28 resex.tex
finished 8th turn pulsing section

@@
 
-\author{Giulia Papotti}
 \author{Stefano Redaelli}
@@
 		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_50.png}
+		\includegraphics[width=1.0\linewidth]{profile_v_10thv_slot_50.png}
 	\end{minipage}
@@
 		\centering
-		excited bunch,\\ $10^{\mathrm{th}}$~turn V		\includegraphics[width=1.0\linewidth]{profile_v10th_slot_1300.png}
+		excited bunch,\\ $10^{\mathrm{th}}$~turn V		\includegraphics[width=1.0\linewidth]{profile_v_10thv_slot_1300.png}
 	\end{minipage}
@@
 	\end{minipage}	
-	\caption{\label{fig:7th2017fma} FMA analysis for $7^{\mathrm{th}}$~turn pulsing based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
+	\caption{\label{fig:7th2017fma} FMA analysis in frequency space for $7^{\mathrm{th}}$~turn pulsing based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
 The excitation is 96~nrad in both planes.
@@
 		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_hv7th_slot_2862.png}
+		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
 	\end{minipage}
@@
 		\centering
-		excited bunch,\\ $7^{\mathrm{th}}$~turn h+V		\includegraphics[width=1.0\linewidth]{profile_hv7th_slot_2696.png}
+		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
 	\end{minipage}
@@
 \subsubsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
-Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation.
-With this motivation in mind the $8^{\mathrm{th}}$ turn pulsing pattern was studied, for which a small effect is expected in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing.
-In the Distribution Tracking, the excitation amplitudes had to be increased to 96~nrad to have a visible effect
+\begin{figure}[b]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_intensity.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_sigm.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit1.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit2.png}
+	\end{minipage}	
+	\caption{\label{fig:8thsim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $8^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
+\end{figure}
+Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they on the other hand have a large effect on the tails under consideration of the non-linear field generated by the electron lens itself.
+With this motivation in mind the $8^{\mathrm{th}}$ turn pulsing pattern was studied, for which a small effect is expected in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing as can be demonstrated in the Distribution Tracking, for which the excitation amplitudes had to be increased to 96~nrad to have a visible effect (Fig.~\ref{fig:8thsim}).
+The loss rates and also changes in bunch length are very small and any differences are here just due to statistical fluctuations.
+In the horizontal and vertical plane a small emittance growth due to the excitation is visible, but without any clear dependency on the amplitude nor plane of the $8^{\mathrm{th}}$ turn pulsing.
+Adducting the FMA analysis in amplitude space depicted in Fig.~\ref{fig:8th2017fmaamp}, the $8^{\mathrm{th}}$ turn pulsing drives the $16Q_y$ and \mbox{$8Q_x-4Q_y$} resonances and several even higher orders.
+These are in general very high order resonances whose effect is thus in general expected to be small.
+\begin{figure}[t]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		no excitation
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$8^{\mathrm{th}}$ H+V
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut8skhv_dp0_amp_annotate.png}
+	\end{minipage}	
+	\caption{\label{fig:8th2017fmaamp} FMA analysis in amplitude space for $8^{\mathrm{th}}$~turn pulsing based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
+The excitation is 96~nrad in both planes.
+The $16Q_y$ and \mbox{$8Q_x-4Q_y$} are both excited.}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.0\linewidth]{2017_scale_amp_8hv_lbllong.png}
+	\caption{\label{fig:8thexploss} Comparison of scaling of loss rates with excitation amplitude for $8^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars.}
+\end{figure}
+The only small effect on the beam expected from simulations could then be also confirmed experimentally during the 2017 experiments.
+For this purpose the excitation amplitude was increased to 96~nrad, which is the maximum kick strength reachable with the transverse damper at injection without risking saturation.
+As depicted in Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the excitation amplitude for an excitation in H+V.
+For an excitation only in V or only in H, no statistically relevant increase was observed.
+In addition, the beam distribution also changed in the horizontal plane for an excitation only in H as visible in the BSRT profiles Fig.~\ref{fig:8thexpprof} indicating an increase at around $2.0~\sigma$ simultaneously with a depletion of the beam core.
+For an excitation in H+V the emittances of the references bunches were much smaller, around $1.8~\mu$m normalized emittance on average, than the one of the excited bunches ranging from $2.3~\mu$m to $3.0~\mu$m \cite{resexmd2017}.
+The references bunches thus experienced a stronger emittance growth due to intra-beam scattering than the excited bunches due to the excitation on top of their smaller emittance growth due to intra-beam scattering.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch,\\ no excitation
+		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_804.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ $8^{\mathrm{th}}$~turn H	\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_632.png}
+	\end{minipage}
+	\caption{\label{fig:8thexpprof} Horizontal beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+The profiles are taken at the end of the $8^{\mathrm{th}}$~turn pulsing in H.
+The distribution changes in the reference bunch (left), here slot 804, are much less than for the bunch experiencing the maximum excitation (right), here slot 632.}
+\end{figure}
+
 \subsubsection{random excitation\label{sec:simexran}}
+In view of the electron lens, the random excitation presents the pulsing pattern leading to the fastest halo depletion rates.
+On the other hand it also represents the most dangerous pattern for the beam core.
+The aim of this experiment was to test the effect on the beam core in comparison to the resonant excitation represented in the 2017 experiment by the $7^{\mathrm{th}}$~turn pulsing as for $8^{\mathrm{th}}$~turn pulsing basically no effect was observed.
+The random excitation is so powerful in cleaning the halo as it excites all frequencies equally.
+This can be seen formidably with the FMA analysis in amplitude space depicted in Fig.~\ref{fig:ran2017fmaamp}.
+The diffusion in tune is increased equally independent of the amplitude.
+\begin{figure}[t]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		no excitation
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		uniform random H+V
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerruranadthv_1nrad_dp0_amp.png}
+	\end{minipage}	
+	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space for a uniform random excitation with 1~nrad amplitude in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
+The diffusion in tune is increased equally independent of the amplitude.}
+\end{figure}
+
+The results from the Distribution Tracking are shown in Fig.~\ref{fig:ransim}
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png}
+	\end{minipage}
+	\caption{\label{fig:ransim} Relative intensity (top left) and relative vert.
+emittance (top right), and hor.
+(bottom left) and vert.
+(bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$ and 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $7^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
+\end{figure}
 
@@
 \begin{acknowledgments}
-We wish to acknowledge Roderik Bruce and Gianluca Valentino for there very helpful suggestions in preparing the experiment at the LHC and we would like to thank all participants of the experiment for helping acquire the data presented in this paper.
+We wish to acknowledge Giulia Papotti with her help with the preparations for the 2016 experiments and her general support from the operations side, Roderik Bruce and Gianluca Valentino for there very helpful suggestions in preparing the experiment at the LHC and we would like to thank all participants of the experiment for helping acquire the data presented in this paper.
 We would like to acknowledge Dmitry Shatilov for his help with the Lifetrac code, and St\'{e}phane Fartoukh, Riccardo De Maria and Rogelio Tomás for their help with generating the appropiate optics model.

### 2018-01-29 resex.tex
version 2018-01-28

@@
 The diffusion in tune is increased equally independent of the amplitude.}
-\end{figure}
-
-The results from the Distribution Tracking are shown in Fig.~\ref{fig:ransim}
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png}
-	\end{minipage}
-	\caption{\label{fig:ransim} Relative intensity (top left) and relative vert.
-emittance (top right), and hor.
-(bottom left) and vert.
-(bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$ and 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
-The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $7^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
+%\end{figure}
+%\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_intensity.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_sigm.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_emit1.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_emit2.png}
+	\end{minipage}
+	\caption{\label{fig:ransim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+The solid line indicates the case with no excitation, and the dotted and dashed line the results for a uniform random excitation of 12~nrad and 24~nrad respectively.}
+\end{figure}
+
+The results from the Distribution Tracking are shown in Fig.~\ref{fig:ransim}.
+In contrast to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing, the effect of a random excitation is very symmetric:
+\begin{itemize}
+	\item Emittance growth is only caused in the plane of excitation.
+	\item The emittance growth is comparable between an excitation in H and V in the corresponding plane as well as the scaling with the excitation amplitude.
+Only a small increase or decrease is observed if the excitation is applied in both planes.
+The difference can be attributed to coupling.
+	\item The random excitation leads to an increased constant emittance growth rate which is reached without undergoing an initial adjustment phase.
+	\item Negligible losses are observed independent of the plane of excitation.
+\end{itemize}
+
+Qualitatively, this behavior could be also confirmed experimentally during the 2017 experiments and as example the emittances are depicted in Fig.~\ref{fig:ranexp} for pulsing in V and in H+V, showing no emittance growth in the horizontal for an excitation only in V and emittance growth in both planes for an excitation in H+V.
+All emittance growth rates are in a comparable range.
+Note that despite the large emittance growth, the growth rate stays constant and there is no fast adjustment phase succeeded by a new equilibrium state as was the case for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		random excitation V
+		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_vran_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		random excitation H+V
+		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hvran_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_emith_avg_rel_vran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_emith_avg_rel_hvran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_hvran_no_damper.png}
+	\end{minipage}	
+	\caption{\label{fig:ranexp} Relative intensity (top) and relative horizontal (center) and vertical (bottom) emittance for a random excitation only in V (left) and in H+V (right).
+Emittance growth is only caused in the plane of excitation.}
+\end{figure}
+These large changes of the emittance can of course also be seen directly in the BSRT profiles shown in Fig.~\ref{fig:10thexp} for a random excitation only in V.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch,\\ no excitation
+		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ random excitation V		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
+	\end{minipage}
+	\caption{\label{fig:ranexpprof}\textcolor{red}{update plot} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+The profiles are taken at the end of the random excitation in V.
+The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
+\end{figure}
+
+The experimental results also confirm the small loss predicted in simulations.
+The losses then suddenly increase for 12~nrad, which is most likely the point when, due to the large emittance increase, the physical aperture starts eating into the more central region of the distribution causing higher losses.
+Furthermore, the loss rates also add up, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only V.
+This is illustrated in  Fig.~\ref{fig:ranexploss}, which also shows the loss rates for the $7^{\mathrm{th}}$~turn pulsing in comparison.
+Furthermore the loss rate is considerably reduced if the transverse damper is active, which is not the case for the $7^{\mathrm{th}}$~turn pulsing.
+We will address this subject separately in the next Section, Sec.~\ref{sec:damp}.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+	\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranv_lbllong.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranhv_lbllong.png}
+	\end{minipage}	
+	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing in V (left) and H+V (right) measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars.}
 \end{figure}
@@
 \subsubsection{Effect of the transverse damper\label{sec:damp}}
-Compare random and nth turn -> no damping for nth, only for random.
+For nth turn pulsing no effect. Show example of emittance for 7th turn pulsing and random and refer to Fig.~\ref{fig:ranexploss} for the loss rate.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		random excitation H+V
+		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hvran_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$7^{\mathrm{th}}$~turn H+V
+		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hvran_with_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hv7th_with_damper_avg.png}
+	\end{minipage}	
+	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for $7^{\mathrm{th}}$~turn pulsing in V (left) and for a random excitation only in V (right).}
+\end{figure}
 
@@
 For DC operation, no influence on the beam core are are to be expected.
-To study the effect of pulsed operation beam experiments were conducted in 20216 and 2017 at injection.
+To study the effect of pulsed operation beam experiments were conducted in 2016 and 2017 at injection.
 The experimental results showed that for the kick amplitudes expected for the e-lens, considerable effects are to be expected.

### 2018-01-29 scripts/lifetrac_sim_results/plot_loss_emit_ranH_V_H+V_2017.py

@@ -40,7 +40,7 @@ for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
 plt.grid(b=grid_flag)
 plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
-#plt.ylim(0.99975,0.999825)
+plt.ylim(0.9997,0.999825)
 leg = plt.legend(title='random\nexcitation',ncol=1,fontsize=12)
 plt.setp(leg.get_title(),fontsize=12)
 plt.savefig(os.path.join(plt_dir,'2017injerra2b2u_ranadt_3_5um_intensity.png'),bbox_inches='tight')
@@ -56,7 +56,7 @@ for plane,cplane in zip(['h','v','h+v'],['b','r','g']):
 plt.grid(b=grid_flag)
 plt.xlabel('time [s]')
 plt.ylabel('bunch length [cm]')
-#plt.ylim(9.1905,9.1925)
+plt.ylim(9.1905,9.1927)
 leg = plt.legend(title='random\nexcitation',ncol=1,fontsize=12)
 plt.setp(leg.get_title(),fontsize=12)
 plt.savefig(os.path.join(plt_dir,'2017injerra2b2u_ranadt_3_5um_sigm.png'),bbox_inches='tight')
@@ -73,6 +73,7 @@ for phv,p12 in zip(['hor.','vert.'],'12'):
   plt.grid(b=grid_flag)
   plt.xlabel('time [s]')
   plt.ylabel('%s normalized emittance [$\mu$m]'%phv)
+  plt.ylim(3.48,4.35)
   leg = plt.legend(title='random\nexcitation',ncol=1,fontsize=12)
   plt.setp(leg.get_title(),fontsize=12)
   plt.savefig(os.path.join(plt_dir,'2017injerra2b2u_ranadt_3_5um_emit%s.png'%p12),bbox_inches='tight')

### 2018-01-29 resex.tex
finished random excitation and damper part

@@
 The profiles are taken at the end of the $10^{\mathrm{th}}$~turn pulsing in V.
+For both bunches the transverse damper is not active.
 The black line in the plot of the residual is the current profile minus the profile before applying the excitation and is a measure for the overall change of the distribution.
@@
 Just as for the $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:10thexp}), the emittance growth can be divided also for the $7^{\mathrm{th}}$~turn pulsing in a phase of fast adjustment of the beam distribution and the reach of a new equilibrium state (Fig.~\ref{fig:7thexp}).
-Also in this case, the distribution changes are also directly visible on the beam profiles measured with the BSRT (Fig.~\ref{fig:7thexpprof}).
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
-	\end{minipage}
-	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V.
-The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
-\end{figure}
+Also in this case, the distribution changes are also directly visible on the beam profiles measured with the BSRT (Appendix~\ref{app:sec:bsrtprof}, Fig.~\ref{fig:7thexpprof}).
 
@@
 For an excitation only in V or only in H, no statistically relevant increase was observed.
-In addition, the beam distribution also changed in the horizontal plane for an excitation only in H as visible in the BSRT profiles Fig.~\ref{fig:8thexpprof} indicating an increase at around $2.0~\sigma$ simultaneously with a depletion of the beam core.
+In addition, the beam distribution also changed in the horizontal plane for an excitation only in H as visible in the BSRT profiles (Appendix~\ref{app:sec:bsrtprof}, Fig.~\ref{fig:8thexpprof}) indicating an increase at around $2.0~\sigma$ simultaneously with a depletion of the beam core.
 For an excitation in H+V the emittances of the references bunches were much smaller, around $1.8~\mu$m normalized emittance on average, than the one of the excited bunches ranging from $2.3~\mu$m to $3.0~\mu$m \cite{resexmd2017}.
 The references bunches thus experienced a stronger emittance growth due to intra-beam scattering than the excited bunches due to the excitation on top of their smaller emittance growth due to intra-beam scattering.
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_804.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ $8^{\mathrm{th}}$~turn H	\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_632.png}
-	\end{minipage}
-	\caption{\label{fig:8thexpprof} Horizontal beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the $8^{\mathrm{th}}$~turn pulsing in H.
-The distribution changes in the reference bunch (left), here slot 804, are much less than for the bunch experiencing the maximum excitation (right), here slot 632.}
-\end{figure}
 
@@
 \end{figure}
-These large changes of the emittance can of course also be seen directly in the BSRT profiles shown in Fig.~\ref{fig:10thexp} for a random excitation only in V.
+These large changes of the emittance can of course also be seen directly in the BSRT profiles shown in Fig.~\ref{fig:ranexpprof} for a random excitation only in V.
+In comparison to for example the $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:10thexpprof}), the distribution also stays Gaussian in case of a random excitation while for the $10^{\mathrm{th}}$~turn pulsing it clearly adopts a non-Gaussian shape.
 \begin{figure}[h]
@@
 		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ random excitation V		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
-	\end{minipage}
-	\caption{\label{fig:ranexpprof}\textcolor{red}{update plot} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+		\includegraphics[width=1.0\linewidth]{profile_v_ranv_slot_1698.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ random excitation V
+		\includegraphics[width=1.0\linewidth]{profile_v_ranv_slot_1532.png}
+	\end{minipage}
+	\caption{\label{fig:ranexpprof}Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
 The profiles are taken at the end of the random excitation in V.
-The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
+For both bunches the transverse damper is not active.
+The distribution changes in the reference bunch (left), here slot 1698, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
 \end{figure}
@@
 	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing in V (left) and H+V (right) measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.}
+The excitation amplitude error is 15\% indicated by the horizontal error bars.
+Note that for $7^{\mathrm{th}}$~turn pulsing the transverse damper does not influence the loss rate, while for a random excitation the loss rate is considerably reduced for the bunches with transverse damper active.}
 \end{figure}
@@
 \subsubsection{Effect of the transverse damper\label{sec:damp}}
+For all $n^{\mathrm{th}}$~turn pulsing patterns, the transverse damper does not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation.
+On the other hand, the transverse damper considerably decreases loss rates, emittance growth and distribution changes for the random excitation.
+As an example, the vertical emittance growth is compared for the $7^{\mathrm{th}}$~turn pulsing in V and a random excitation in V in Fig.~\ref{fig:damp} and the loss rates in Fig.~\ref{fig:ranexploss} illustrating impressively the above behavior.
 For nth turn pulsing no effect. Show example of emittance for 7th turn pulsing and random and refer to Fig.~\ref{fig:ranexploss} for the loss rate.
 \begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		random excitation H+V
-		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hvran_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$7^{\mathrm{th}}$~turn H+V
-		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hvran_with_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hv7th_with_damper_avg.png}
+	\centering
+	random excitation H+V\\
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		damper not active
+		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		damper active
+		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_with_damper.png}
+	\end{minipage}	
+		\centering
+		$7^{\mathrm{th}}$~turn H+V\\
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		damper not active
+		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		damper active
+		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
 	\end{minipage}	
@@
 \end{figure}
+\textcolor{red}{Any idea why? Daniel, help!}
 
@@
 
-\section{Results of $7^{\mathrm{th}}$~turn pulsing during 2016 and 2017 experiments (additional figures)}
-\label{app:sec:simexp:7}
+\section{BSRT profiles for $7^{\mathrm{th}}$ and $8^{\mathrm{th}}$~turn pulsing during 2017 experiments}
+\label{app:sec:bsrtprof}
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch,\\ no excitation
+		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
+		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
+	\end{minipage}
+	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V.
+For both bunches the transverse damper is not active.
+The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
+\end{figure}
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch,\\ no excitation
+		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_804.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ $8^{\mathrm{th}}$~turn H
+		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_632.png}
+	\end{minipage}
+	\caption{\label{fig:8thexpprof} Horizontal beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+The profiles are taken at the end of the $8^{\mathrm{th}}$~turn pulsing in H.
+For both bunches the transverse damper is not active.
+The distribution changes in the reference bunch (left), here slot 804, are much less than for the bunch experiencing the maximum excitation (right), here slot 632.}
+\end{figure}
+
+\clearpage
 

### 2018-01-29 resex.tex
add hel residual field part from Giulio

@@
 	\end{minipage}	
-	\caption{\label{fig:active_halo_control} Sketch of passive halo control as with the collimation system (top) and active halo control using in addition for example a hollow electron lens (HEL) to control the diffusion speed in the region tail region without affected the beam core (bottom).
-\textcolor{red}{Any acknowledgement needed for the plots?}.}
-\end{figure}
-
-In view of the need of an active halo control system for LHC and HL-LHC in particular and in general for future high power accelerators, different methods have been studied in the last recent years \cite{helreview_bruce}, of which the HEL is considered the superior device at least in case of the HL-LHC \cite{helreview} and is also considered for other future high energy colliders like HE-LHC and FCC-hh \cite{helhcparam2011,fccparam2016}.
+	\caption{\label{fig:active_halo_control} Sketch of passive halo control as with the collimation system (top) and active halo control using in addition for example a hollow electron lens (HEL) to control the diffusion speed in the region tail region without affected the beam core (bottom).}
+\end{figure}
+
+In view of the need of an active halo control system for LHC and HL-LHC in particular and in general for future high power accelerators, different methods have been studied in the last recent years \cite{helreview_bruce}, of which the HEL is considered the superior device at least in case of the HL-LHC \cite{helreview} and is also considered for other future high energy colliders like HE-LHC and FCC-hh \cite{helhcparam2011,fcc_coll_ipac2017}.
 
@@
 Assuming HL-LHC and HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) the maximum kick of the HEL is:
-\begin{equation}
+\begin{equation}\label{eqn:helkick}
 \theta_{\rm max,B1} = 392~\rm{nrad}
@@
 Effects on the beam core thus only arise in case of imperfections, where possible sources are the bends of the HEL, as the electron beam crosses directly the proton beam, and distortions in the electron beam profile (see Section~\ref{sec:core}).
-Both sources result in non-linear kicks (see for example \cite{hel_bends_stancari,hel_model_polynomial_morozov} (\textcolor{red}{is there a better reference for the profile imperfections, updated note from Giulio???})).
-
-In DC operation of the HEL, the non-linear kicks from the HEL stay in the shadow of the non-linearities otherwise present in the machine.
-Tolerances on imperfections are therefore not particular stringent and not a concern.
-The picture however changes significantly in case of pulsed operation of the HEL, in which case the electron gun voltage is modulated using a white or colored noise spectrum.
+Both sources result in non-linear kicks \cite{hel_bends_stancari,hel_model_polynomial_morozov}.
+
+In DC operation of the HEL, these non-linear kicks stay in the shadow of the non-linearities otherwise present in the machine.
+Tolerances on imperfections are therefore not particularly stringent and of no concern.
+
+The picture changes significantly in case of pulsed operation of the HEL, in which case the electron gun voltage is modulated using a white or colored noise spectrum.
 If the electromagnetic field of the HEL does not vanish at the proton beam core, noise is transferred not only to the halo particles, as intended, but also to the beam core with naturally much smaller amplitude.
-Tolerances in this case rapidly become much more stringent than in DC operation and studies of the effect of the HEL on the beam core are therefore focusing on this mode of operation.
+Tolerances in this case rapidly become much more stringent than in the case of DC operation and studies of the effect of the HEL on the beam core are therefore focusing on this mode of operation which is also the main topic of this paper.
 \begin{table}[t]
@@
 
-The main benefit of pulsing the HEL is to increase the halo removal rate in case a fast depletion of the halo is desired.
+Despite the fact that pulsed operation is much more dangerous than DC operation for the proton beam core, pulsed operation is still strongly under consideration.
+The main benefit of pulsing the HEL is to increase the halo removal rate if a fast depletion of the halo is desired.
 The halo diffusion rate is increased in this case as in addition to the non-linear kick, noise is exerted on the halo particles of the proton beam.
-For the HL-LHC two different pulsing patterns are currently being considered:
+For the HL-LHC two different pulsing patterns are considered:
 \begin{itemize}
@@
 Currently, no HEL is installed in the LHC, and the kick on the proton beam core must therefore be approximated to first order by a dipole kick for the experiments at the LHC.
-Explicitly, the dipole kick with the corresponding excitation pattern is applied with the LHC transverse damper (see Section~\ref{sec:adt}).
-
-The expected kick to first order from the HEL bends~(Eqn.~\ref{eqn:kick_bends}) is
+During the experiments presented in this paper, the dipole kick with the corresponding excitation pattern was applied with the LHC transverse damper (see Section~\ref{sec:adt}).
+
+Leaping ahead of the more detailed derivations in Sec.~\ref{core:sec:1} and Sec.~\ref{core:sec:2}, the expected kick to first order from the HEL bends is estimated to be~(Eqn.~\ref{eqn:kick_bends}):
 \begin{equation}
@@
 \begin{equation}
-	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \leq \SI{15.0}{nrad}
+	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \leq \SI{19.6}{nrad}
 \end{equation}
-where the current HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) are used.
-% yielding the maximum kick, explicitly $B_{m}=\SI{5}{T}$, $I_e=\q{5.0}{A}$, $E_{e} = \SI{10.0}{keV}$, $L=\SI{3}{m}$ and $E_{p} = \SI{7.0}{TeV}$.
-The contribution from the central region is clearly dominating. A more detailed derivation of both values is given in Sec.~\ref{core:sec:1}--\ref{core:sec:2}.
+where the current HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) are used. The contribution from the central region is clearly dominating.
+
 \subsection{Uncompensated kicks from HEL bends}
@@
 \label{core:sec:2}
-For a perfectly uniform, annular and radially symmetric profile, the field in the region of the proton beam core vanishes (see Eqn.~\ref{eq:field_2}).
-In case of electron beam profile imperfections in particular the radial symmetry is broken, leading to a residual field at the beam core (see Fig.~\ref{core:fig:1}).
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1.0\linewidth]{e-beam_profile.png}
-	\caption{Calculated hollow electron beam field from measured profile of 9~kV, 2.49~A e-gun and different main solenoidal fields $Bm$.
-The field has been calculated using the code WARP \cite{warp}.
-\textcolor{red}{change $Bm$ to $B_m$ +substitute with new picture.}}
+For a perfectly uniform, annular and radially symmetric profile, the field in the region of the proton beam core vanishes (see Eqn.~\ref{eq:field_2} and Fig.~\ref{fig:hel_field}).
+Fields at the location of the proton beam core, thus only arise in case of profile imperfections of the electron beam, in particular if the radial symmetry is broken.
+
+Last year, new measurements could be performed at the Fermilab electron lens test stand \cite{hel_test_stand_fnal} of the newest hollow electron gun CHG1b.
+A typical measurement of the electron beam current density obtained during these measurements is shown in Fig.~\ref{core:fig:0}.
+\begin{figure}[t]
+		\centering
+		\includegraphics[width=0.8\linewidth]{e-field-measured.png}
+	\caption{Current density from measurements of the CHG1b hollow electron gun at the Fermilab electron lens test stand in 2017 \cite{hel_res_field_stancari_2017}.}
+	\label{core:fig:0}
+\end{figure}
+As at the test stand the solenoidal fields required for the HL-LHC electron lens can not be reached, an estimate of the electron beam profile for HL-LHC can only be obtained in terms of simulations.
+For this purpose, the measured distribution is compressed to the inner e-beam radius of $0.97$~mm and than an input distribution with $40\ 000$ particles generated.
+The electric field can in general be scaled with the electron beam current and energy, but can not easily be scaled with the inner and outer radius of the electron beam distribution.
+Therefore it is essential to track a distribution with the correct dimensions.
+The potential and fields can then be calculated with the particle in cell simulation code WARP \cite{warp} yielding the relative electric field strengths shown in Fig.~\ref{core:fig:1}.
+Further details on the measurements and WARP simulations can be found in~\cite{hel_res_field_stancari_2017}.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice.png}
+	\end{minipage}
+	\caption{Calculated relative electric field for the new CHG1b hollow electron gun in the x-y plane and as projection on the x- and y-axis respectively based on measurements at the Fermilab test stand combined with WARP simulations for the calculation of the electric potentials and fields.}
 	\label{core:fig:1}
 \end{figure}
-A first estimate of the residual kick for the HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) with a main solenoid field of $B_{m}=\SI{5}{T}$ can be obtained by scaling the electric field at the center from the measurements and simulations shown in Fig.~\ref{core:fig:1}, yielding:
-\begin{equation}\label{eqn:kick_central}
-\Delta x', \Delta y'=\q{15}{nrad}
+
+As a first estimate for the fields at the proton beam core, one can use the relative electric field strength along the x- and y-axis, yielding roughly (Fig.~\ref{core:fig:1}, right):
+\begin{equation}
+	E_{\mathrm{rel}}\approx 5\%
 \end{equation}
-\textcolor{red}{New gun measurements, can we maybe have a WARP simulation for the new design parameters instead of scaling?}
+For a maximum kick of $392$~nrad (Eqn.~\ref{eqn:helkick}), we then obtain as estimated dipole kick amplitude at the proton beam core:
+\begin{equation}
+\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \leq \SI{19.6}{nrad}
+\end{equation}
 
@@
 As already addressed earlier, the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time.
-This behavior is typical for any $n^{\mathrm{th}}$ turn pulsing and can be ascribed to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution.
+This behavior is typical for any $k^{\mathrm{th}}$ turn pulsing and can be ascribed to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution.
 As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}, and also any other simulations over $10^6$ turns presented in this paper.
@@
 \subsubsection{Effect of the transverse damper\label{sec:damp}}
-For all $n^{\mathrm{th}}$~turn pulsing patterns, the transverse damper does not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation.
+For all $k^{\mathrm{th}}$~turn pulsing patterns, the transverse damper does not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation.
 On the other hand, the transverse damper considerably decreases loss rates, emittance growth and distribution changes for the random excitation.

### 2018-01-29 resex.tex
remove grid lines

@@
 \begin{abstract}
-In the HL-LHC a considerable amount of energy is stored in the beam tails due to the high beam intensity and an overpopulation of the tails compared to a Gaussian distribution.
-To control and clean the tail population, the installation of two hollow electron lenses, one per beam, is considered.
-Beside the DC operation, also a pulsed operation of the hollow electron lens is considered, which would considerably increase the diffusion speed by putting noise on the halo particles.
-In the ideal case, that is in case of no field at the beam core, only the halo particles are excited while leaving the core unperturbed.
-The picture though changes, if a residual field is present also at the location of the beam core putting noise also on the beam core.
-In this paper we present for estimates of the residual field at the beam core expected from the HL-LHC hollow electron lens and first experimental results of the effect of this excitation on the beam core together with the supporting simulations.
+In case of the High Luminosity Large Hadron Collider (HL-LHC), a considerable amount of energy is stored in the beam tails due to the high beam intensity and, in addition, an overpopulation of the tails compared to a Gaussian distribution.
+To control and clean the tail population, the installation of two hollow electron lenses, one per beam, is foreseen.
+Beside the DC operation of the lenses, also pulsed operation is considered having the advantage to considerably increase the diffusion speed by exerting noise on the halo particles.
+In the ideal case, that is in case of no field at the proton beam core, only the halo particles are excited while leaving the core unperturbed.
+The picture though changes, if a residual field is present also at the location of the beam core exposing also the the particles in the center to noise.
+In this paper, we present first estimates of the residual field at the proton beam core expected from the HL-LHC hollow electron lens and summarize the results of the LHC experiments conducted in 2016 and 2017 to study the effects on the beam core in case of pulsed operation.
 \end{abstract}
@@
 
-All of the above reasons lead to the conclusion that a mechanism is needed to deplete the beam tails in a controlled manner (see for example \cite{helreview} in case of HL-LHC).
+All of the above reasons lead to the conclusion that a mechanism is needed to deplete the beam tails in a controlled manner \cite{helreview}.
 The most obvious idea is to decrease the collimator gaps or scrape the tails with a collimator type device.
-The minimum gap size of the collimators is however limited by the impedance of the machine and small enough gap sizes can therefore not be reached without risking an instability of the beam.
+The minimum gap size of the collimators is however limited by the impedance of the machine.
+Small enough gap sizes can therefore not be reached without risking severe beam instabilities.
 Other mechanisms must therefore be found to actively deplete the tails.
-
-Most promising are methods, which considerably increase the diffusion speed in the region of the halo particles while leaving the beam core unperturbed.
-The diffusing halo particles are then intercepted by the collimation system and removed (see Fig.~\ref{fig:active_halo_control} for illustration).
-This concept is also referred to as active halo control compared to a passive system, which only intercepts the halo particles without actively controlling the diffusion speed.
-Currently, the LHC collimation system is a passive halo control system.
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{passive_halo_control.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{active_halo_control.png}
+Most promising are methods, which increase the diffusion speed in the region of the halo particles while leaving the beam core unperturbed.
+The diffusing halo particles are then intercepted by the collimation system and removed.
+This concept is also referred to as active halo control compared to a passive system, which only intercepts the halo particles without actively controlling the diffusion speed (see Fig.~\ref{fig:active_halo_control} for illustration).
+\begin{figure}[h]
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.7\linewidth]{passive_halo_control.png}
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.7\linewidth]{active_halo_control.png}
 	\end{minipage}	
@@
 
-In view of the need of an active halo control system for LHC and HL-LHC in particular and in general for future high power accelerators, different methods have been studied in the last recent years \cite{helreview_bruce}, of which the HEL is considered the superior device at least in case of the HL-LHC \cite{helreview} and is also considered for other future high energy colliders like HE-LHC and FCC-hh \cite{helhcparam2011,fcc_coll_ipac2017}.
+In a recent review, the need for such an active halo control system for HL-LHC has been assessed with the conclusion that it would considerably increase the margin and reduce the risk for machine protection compared to the current LHC collimation system being a passive halo control system  \cite{helreview}.
+In view of the general need of active halo control for future high power accelerators and HL-LHC in particular, different methods have been studied in the last recent years \cite{helreview_bruce}, of which the HEL is considered the superior device at least in case of the HL-LHC \cite{helreview} and is also considered for other future high energy colliders like HE-LHC and FCC-hh \cite{helhcparam2011,fcc_coll_ipac2017}.
 
@@
 Section~\ref{sec:core} is dedicated to describing the sources of a residual field from the HEL in the core region.
-Section~\ref{sec:exp_sum}--\ref{sec:exp_sim} describes then in detail the two beam experiments conducted at the LHC to study the effect on the beam core in case of a pulsed HEL operation.
+Section~\ref{sec:exp} describes then in detail the two beam experiments conducted at the LHC to study the effect on the beam core in case of a pulsed HEL operation.
 This includes the detailed analysis of the observed losses, emittance and transverse beam distribution changes.

### 2018-01-29 scripts/lifetrac_sim_results/plot_loss_emit_ranH_V_H+V_2017.py

@@ -25,7 +25,7 @@ optsk = "2017injnocolc15o+19_6erra2b2uranadt/1/eps3.5/%s%se-8/p1e4t1e4s100weight
 # no excitation
 t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
 
-grid_flag = True
+grid_flag = False
 
 plt.close('all')
 # losses

### 2018-01-30 resex.tex
complete first revision

@@
 
-In a recent review, the need for such an active halo control system for HL-LHC has been assessed with the conclusion that it would considerably increase the margin and reduce the risk for machine protection compared to the current LHC collimation system being a passive halo control system  \cite{helreview}.
-In view of the general need of active halo control for future high power accelerators and HL-LHC in particular, different methods have been studied in the last recent years \cite{helreview_bruce}, of which the HEL is considered the superior device at least in case of the HL-LHC \cite{helreview} and is also considered for other future high energy colliders like HE-LHC and FCC-hh \cite{helhcparam2011,fcc_coll_ipac2017}.
-
-The beneficial effect of an active halo control system is however abrogated if also the beam core is affected as this would ultimately lead to a degradation of the performance due to losses of particles in the core region and emittance growth.
-In this paper, we will concentrate on this aspect in view of future HEL to be installed in the HL-LHC.
-We will summarize possible sources of perturbations of the beam core concentrating in particular on the case of pulsed operation and with the focus on the beam experiments at the LHC performed in this context.
-
-This paper is structured as follows: Section~\ref{sec:hel} gives an introduction to the concept of HELs and summarizes the design parameters of the HL-LHC HEL.
+In a recent review, the need for such an active halo control system for HL-LHC has been assessed with the conclusion that it would considerably increase the margin and reduce the risk for machine protection compared to the current LHC collimation system being a passive halo control system ~\cite{helreview}.
+In view of the need of active halo control for HL-LHC and for future high power accelerators like HE-LHC and FCC-hh~\cite{helhcparam2011,fcc_coll_ipac2017}, different active halo control methods have been studied in recent years~\cite{helreview_bruce}, of which the hollow electron lens (HEL) is considered the superior device for the HL-LHC~\cite{helreview}.
+The beneficial effect of a future HEL for HL-LHC in terms of machine protection would be however abrogated if the beam core was negatively affected leading ultimately to a degradation of the performance due to increased particle losses and emittance growth.
+In this paper, we will concentrate on this aspect.
+We will summarize possible sources of perturbations of the proton beam core focusing in particular on the case of pulsed operation and the beam experiments at the LHC conducted in this context.
+
+This paper is structured as follows: Section~\ref{sec:hel} gives an introduction to the concept of HELs and summarizes the design parameters of the HL-LHC HELs.
 Section~\ref{sec:core} is dedicated to describing the sources of a residual field from the HEL in the core region.
-Section~\ref{sec:exp} describes then in detail the two beam experiments conducted at the LHC to study the effect on the beam core in case of a pulsed HEL operation.
-This includes the detailed analysis of the observed losses, emittance and transverse beam distribution changes.
+Section~\ref{sec:exp} then describes in detail the two beam experiments conducted at the LHC studying the effect of pulsed operation on the proton beam core.
+
+%This includes the detailed analysis of the observed losses, emittance and transverse beam distribution changes.
 To the knowledge of the authors, the observed distribution changes presented in this paper have never been measured before in the context of a resonant excitation.
@@
 \subsection{General Overview\label{sec:hel:intro}}
-In general, electron lenses are DC or pulsed low energy electron beams.
-The electron beam is generated in an electron gun, then guided and confined by strong solenoids and finally dumped onto a collector.
-Exemplary for the conceptual design of all electron lenses, the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
+Electron lenses are in general DC or pulsed low energy electron beams, where the electron beam is generated in an electron gun, then guided and confined by strong solenoids and finally dumped onto a collector.
+As an example the conceptual design of the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
 \begin{figure}[h]
@@
 \end{figure}
+
+\begin{figure}[b]
+	\includegraphics[width=0.8\linewidth]{kick_hel_lhc_no_grid}% Here is how to import EPS art
+	\caption{\label{fig:hel_field} Illustration of the hollow electron beam distribution (blue), the kick experienced by the proton beam (red) and the collimators (gray).}
+\end{figure}
 The circulating beam, in case of the LHC the proton beam, is then affected by the electromagnetic field of the electron beam.
-For the application of active halo control, the electron beam needs to generate an electromagnetic field at the location of the halo particles while leaving the core untouched.
-This can be achieved by using a uniform hollow distribution in radius $r=\sqrt{x^2+y^2}$ with inner radius $R_1$ and outer radius $R_2$.
+For the application of active halo control, the electron beam needs to generate an electromagnetic field at the location of the halo particles while leaving the core unperturbed.
+This field distribution can be achieved by using a uniform hollow distribution in radius $r=\sqrt{x^2+y^2}$ with inner radius $R_1$ and outer radius $R_2$.
 In this case, the circulating proton beam experiences a radial kick $\theta(r)$
@@
 and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle given by
-\begin{equation}
+\begin{equation}\label{hel_kick_max}
 \theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{q}{p_0}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2},
 \end{equation}
-where $L$ is the length of the e-lens, $I_T$ the total electron beam current, $\beta_{e}$ and $\beta_{p}$ the relativistic $\beta$ of electron and proton beam, $\frac{q}{p_0}=\left(B\rho\right)_p$ the magnetic rigidity for the reference particle, $c$ the speed of light and $\epsilon_0$ the vacuum permittivity.
-\begin{figure}[b]
-	\includegraphics[width=0.8\linewidth]{kick_hel_lhc_no_grid}% Here is how to import EPS art
-	\caption{\label{fig:hel_field} Illustration of the hollow electron beam distribution (blue), the kick experienced by the proton beam (red) and the collimators (gray).}
-\end{figure}
-\begin{table}[t]
-	\caption{\label{tab:hel_param}%
-		HL-LHC hollow electron lens parameters as in~\cite{hel_cdr}.
-	}
-	\begin{ruledtabular}
-		\begin{tabular}{lcc}
-			Geometry & Value& Unit\\
-			\colrule
-		Length $L$    &  3 & m\\
-		Desired range of scraping positions\footnote{$\sigma$ denotes here the Beam sigma and not the collimation $\sigma$ which assumes a normalized emittance of 3.5~$\mu$m.} & 4-8 &$\sigma_p$\\
-		\colrule
-		Magnetic fields & & \\
-		\colrule
-		Gun solenoid, $B_g$ & 0.2-0.4 & T\\
-		Main solenoid, $B_m$ & 2-6 & T\\
-		Collector solenoid, $B_c$ & 0.2-0.4 & T\\
-		Compression factor ($k=\sqrt{B_m/B_g}$) & 2.2-5.5 & -\\
-		\colrule
-		Electron gun & & \\
-		\colrule
-		Peak yield $I_e$ at 10~keV & 5.0 & A\\
-		Gun perveance $P$ & 5 & $\mu$perv\\
-		Inner/outer cathode radius, $R_1/R_2$ & 6.75/12.7 & mm\\
-		\colrule
-		High voltage modulator & & \\
-		\colrule
-		Cathode-anode voltage & 10.0 & kV\\
-		Rise time (10\%-90\%) & 200 & ns \\
-		Repetition rate & 35 & kHz
-		\end{tabular}
-	\end{ruledtabular}
-\end{table}
-
+with $L$ the length of the HEL, $I_T$ the total electron beam current, $\beta_{e}$ and $\beta_{p}$ the relativistic $\beta$ of electron and proton beam, $\frac{q}{p_0}=\left(B\rho\right)_p$ the magnetic rigidity for the proton beam reference particle, $c$ the speed of light and $\epsilon_0$ the vacuum permittivity.
 The distribution together with the resulting kick are illustrated in Fig.~\ref{fig:hel_field}.
-The $\pm$-sign represents the two cases of the electron beam traveling in the direction of the proton beam ($v_e v_p>0$) leading to ``$-$" or in the opposite direction ($v_e v_p<0$) leading to ``$+$".
-For hollow electron beam collimation, electron and proton beam travel in opposite directions.
-Assuming HL-LHC and HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) the maximum kick of the HEL is:
+The $\pm$-sign in Eq.~\ref{hel_kick_max} represents the two cases of the electron beam traveling in the direction of the proton beam ($v_e v_p>0$) leading to ``$-$" or in the opposite direction ($v_e v_p<0$) leading to ``$+$".
+For hollow electron beam collimation, the beams are chosen to counterrotate.
+Assuming HL-LHC and HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) the maximum kick of the HEL is:
 \begin{equation}\label{eqn:helkick}
@@
 %\end{eqnarray}
-for an inner radius $R_1=4\sigma_p$, outer radius $R_2=7.5\sigma_p$ and a peak current of $I_e=$\SI{5.0}{A}. Similar values are obtained for Beam~2.
-
-\subsection{Operation modes and effects on the beam core\label{sec:hel:core}}
-As depicted in Fig.~\ref{fig:hel_field} and in evidence from Eq.~\ref{eq:field_1}--\ref{eq:field_2}, the field at the beam core vanishes in the ideal case.
-Effects on the beam core thus only arise in case of imperfections, where possible sources are the bends of the HEL, as the electron beam crosses directly the proton beam, and distortions in the electron beam profile (see Section~\ref{sec:core}).
-Both sources result in non-linear kicks \cite{hel_bends_stancari,hel_model_polynomial_morozov}.
-
-In DC operation of the HEL, these non-linear kicks stay in the shadow of the non-linearities otherwise present in the machine.
-Tolerances on imperfections are therefore not particularly stringent and of no concern.
-
-The picture changes significantly in case of pulsed operation of the HEL, in which case the electron gun voltage is modulated using a white or colored noise spectrum.
-If the electromagnetic field of the HEL does not vanish at the proton beam core, noise is transferred not only to the halo particles, as intended, but also to the beam core with naturally much smaller amplitude.
-Tolerances in this case rapidly become much more stringent than in the case of DC operation and studies of the effect of the HEL on the beam core are therefore focusing on this mode of operation which is also the main topic of this paper.
+for an inner radius of $R_1=4\sigma_p$, outer radius $R_2=7.5\sigma_p$, peak current of $I_e=$\SI{5.0}{A} and Beam~1 of the LHC. Similar values are obtained for Beam~2.
 \begin{table}[t]
@@
 \end{table}
-
-Despite the fact that pulsed operation is much more dangerous than DC operation for the proton beam core, pulsed operation is still strongly under consideration.
-The main benefit of pulsing the HEL is to increase the halo removal rate if a fast depletion of the halo is desired.
-The halo diffusion rate is increased in this case as in addition to the non-linear kick, noise is exerted on the halo particles of the proton beam.
-For the HL-LHC two different pulsing patterns are considered:
+\begin{table}[t]
+	\caption{\label{tab:hel_param}%
+		HL-LHC hollow electron lens parameters as in~\cite{hel_cdr}.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lcc}
+			Geometry & Value& Unit\\
+			\colrule
+			Length $L$    &  3 & m\\
+			Desired range of scraping positions\footnote{$\sigma$ denotes here the Beam sigma and not the collimation $\sigma$ which assumes a normalized emittance of 3.5~$\mu$m.} & 4-8 &$\sigma_p$\\
+			\colrule
+			Magnetic fields & & \\
+			\colrule
+			Gun solenoid, $B_g$ & 0.2-0.4 & T\\
+			Main solenoid, $B_m$ & 2-6 & T\\
+			Collector solenoid, $B_c$ & 0.2-0.4 & T\\
+			Compression factor ($k=\sqrt{B_m/B_g}$) & 2.2-5.5 & -\\
+			\colrule
+			Electron gun & & \\
+			\colrule
+			Peak yield $I_e$ at 10~keV & 5.0 & A\\
+			Gun perveance $P$ & 5 & $\mu$perv\\
+			Inner/outer cathode radius, $R_1/R_2$ & 6.75/12.7 & mm\\
+			\colrule
+			High voltage modulator & & \\
+			\colrule
+			Cathode-anode voltage & 10.0 & kV\\
+			Rise time (10\%-90\%) & 200 & ns \\
+			Repetition rate & 35 & kHz
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
+
+\subsection{Operation modes and effects on the beam core\label{sec:hel:core}}
+As depicted in Fig.~\ref{fig:hel_field} and in evidence from Eqs.~\ref{eq:field_1}--\ref{eq:field_2}, the field at the beam core vanishes in case of an ideal HEL.
+Effects on the beam core thus only arise from imperfections, where possible sources are the bends of the HEL (see Section~\ref{core:sec:1}), as here the electron beam crosses directly the proton beam, and distortions in the electron beam profile (see Section~\ref{core:sec:2}).
+Both sources result in non-linear kicks~\cite{hel_bends_stancari,hel_model_polynomial_morozov}.
+In DC operation, these non-linear kicks stay in the shadow of the non-linearities otherwise present in the machine.
+Tolerances on imperfections are therefore not particularly stringent and of no concern.
+The picture changes significantly in case of pulsed operation, in which case the electron gun voltage is modulated using a white or colored noise spectrum.
+If the electromagnetic field of the HEL does not vanish at the proton beam core, noise is transferred not only to the halo particles, as intended, but also to the beam core.
+Tolerances on the residual field in this case rapidly become much more stringent than in the case of DC operation.
+Studies of the effect of the HEL on the beam core are therefore focusing on this mode of operation which will also be the main subject of this paper.
+
+The main benefit of pulsed HEL operation is the increase in halo removal rate particularly relevant if a fast depletion of the halo is desired.
+The halo diffusion rate increases in this case as in addition to the non-linear kick, noise is exerted on the halo particles.
+For the HL-LHC two different pulsing patterns are considered for this purpose:
 \begin{itemize}
@@
 \end{itemize}
-In general, the random excitation mode is the most effective and operationally easiest pulsing mode, but also the most stringent in terms of tolerances on the residual field at the beam core as it induces white noise and thus excites all beam frequencies.
+In general, the random excitation mode is the most effective and operationally easiest pulsing mode, but also the most stringent in terms of tolerances on the residual field at the proton beam center as it induces white noise exciting all beam frequencies.
 As the resonant excitation acts on specific resonances, the pulsing pattern has to be chosen according to the machine and beam parameters.
-On the one hand, this makes the operation more difficult as one has to find the most efficient $k^{\mathrm{th}}$~turn pulsing for each operation mode, on the other hand, it also relaxes the tolerances on the residual field at the beam core as, depending on the machine and beam configuration, it might be possible to find a resonant excitation pattern, which mainly affects the halo and not the core.
+On the one hand, this makes the operation more difficult as one has to find the most efficient $k^{\mathrm{th}}$~turn pulsing for each operation mode, on the other hand, it also relaxes the tolerances on the residual field at the beam core as, depending on the machine and beam configuration, it might be possible to find a resonant excitation pattern, which mainly affects the halo and not the beam core.
 
 \section{Sources of residual field in the proton beam core region and first order estimates\label{sec:core}}
-With the current HEL layout (see Fig.~\ref{fig:hel_layout}), parasitic kicks on the proton beam core can arise in the central region (main solenoid) due to profile imperfections in the electron beam and at the entrance and exit of the HEL, where electron and proton beam overlap.
-Currently, no HEL is installed in the LHC, and the kick on the proton beam core must therefore be approximated to first order by a dipole kick for the experiments at the LHC.
-During the experiments presented in this paper, the dipole kick with the corresponding excitation pattern was applied with the LHC transverse damper (see Section~\ref{sec:adt}).
-
-Leaping ahead of the more detailed derivations in Sec.~\ref{core:sec:1} and Sec.~\ref{core:sec:2}, the expected kick to first order from the HEL bends is estimated to be~(Eqn.~\ref{eqn:kick_bends}):
+With the current HEL layout as shown in Fig.~\ref{fig:hel_layout}, parasitic kicks on the proton beam core can arise in the central region due to profile imperfections in the electron beam and at the entrance and exit of the HEL, where electron and proton beam overlap.
+Currently, no HEL is installed in the LHC, and the kick on the proton beam core must therefore be emulated with the help of other devices already installed in the LHC.
+For this purpose the kick is approximated to first order by a dipole kick.
+During the experiments presented in this paper, the dipole kick with the corresponding excitation pattern could then be applied with the LHC transverse damper as will be described in detail in Section~\ref{sec:adt}.
+Leaping ahead of the more detailed derivations in Sec.~\ref{core:sec:1} and \ref{core:sec:2}, the expected kick to first order from the HEL bends is estimated to be~(Eq.~\ref{eqn:kick_bends}):
 \begin{equation}
-	\Delta x'_{\rm{bends}},\Delta y'_{\rm{bends}}\leq \SI{0.5}{nrad},
+	\Delta x'_{\rm{bends}},\Delta y'_{\rm{bends}}\approx \SI{0.5}{nrad},
 \end{equation}
-and for the central region due to profile imperfections~(Eqn.~\ref{eqn:kick_central}):
+and for the central region due to profile imperfections~(Eq.~\ref{eqn:kick_central}):
 \begin{equation}
-	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \leq \SI{19.6}{nrad}
+	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \SI{19.6}{nrad}
 \end{equation}
-where the current HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) are used. The contribution from the central region is clearly dominating.
+where the current HEL design parameters are used (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}). The contribution from the central region is clearly dominating.
 
@@
 As reference for the estimate of the dipole component originating from the HEL bends, the symplectic map derived in \cite{hel_bends_stancari} is used.
-In this case the HEL bends are modeled as a cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons bend in the horizontal plane.
-In case of an U-shaped HEL, the transverse dipole kicks at entrance and exit add up, while for a S-shaped HEL the kicks compensated each other.
+In this case the HEL bends are modeled as a cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons deflected in the horizontal plane.
+In case of an U-shaped HEL, the transverse dipole kicks at entrance and exit add up, while for a S-shape the kicks compensate each other.
 For this reason a S-shape has been chosen for the HL-LHC HEL (see Fig.~\ref{fig:hel_layout}).
-In this case, uncompensated kicks therefore only arise from imperfections, which can originate from profile imperfections and current fluctuations.
-As a first rough estimate, we assume in this paper 10\% fluctuations between the entrance and exit and that the kicks from entrance and exit due to these imperfections add up.
-Using the electric field calculations in \cite{hel_bends_stancari}, the maximum values for the integrated electric field are then around
+Assuming and S-shaped HEL, uncompensated kicks therefore only arise from imperfections, which can originate from profile imperfections and current fluctuations.
+As a first estimate, we assume in this paper 10\% fluctuations between the entrance and exit and, in addition, that the kicks from entrance and exit due to these imperfections add up.
+Using the electric field calculations in \cite{hel_bends_stancari}, the maximum values for the integrated electric field are then
 \begin{equation}
@@
 \end{equation}
-assuming an \SI{1}{A}, \SI{5}{keV} electron beam.
+given an \SI{1}{A}, \SI{5}{keV} electron beam.
 %For \SI{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
@@
 %\end{equation}
-Neglecting magnetic effects and scaling to HL-LHC and HEL design parameters (Table~\ref{tab:hel_param}--\ref{tab:hllhc_param}) this yields a kick of:
+Neglecting magnetic effects and scaling to HL-LHC and HEL design parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}) yields a kick of:
 \begin{equation}
@@
 \begin{equation}\label{eqn:kick_bends}
-\Delta x', \Delta y'= \q{0.5}{nrad}.
+\Delta x', \Delta y'\approx \q{0.5}{nrad}.
 \end{equation}
 
-\subsection{Kicks in the central region (main solenoid)}
-\label{core:sec:2}
-For a perfectly uniform, annular and radially symmetric profile, the field in the region of the proton beam core vanishes (see Eqn.~\ref{eq:field_2} and Fig.~\ref{fig:hel_field}).
-Fields at the location of the proton beam core, thus only arise in case of profile imperfections of the electron beam, in particular if the radial symmetry is broken.
-
-Last year, new measurements could be performed at the Fermilab electron lens test stand \cite{hel_test_stand_fnal} of the newest hollow electron gun CHG1b.
-A typical measurement of the electron beam current density obtained during these measurements is shown in Fig.~\ref{core:fig:0}.
-\begin{figure}[t]
-		\centering
-		\includegraphics[width=0.8\linewidth]{e-field-measured.png}
-	\caption{Current density from measurements of the CHG1b hollow electron gun at the Fermilab electron lens test stand in 2017 \cite{hel_res_field_stancari_2017}.}
-	\label{core:fig:0}
-\end{figure}
-As at the test stand the solenoidal fields required for the HL-LHC electron lens can not be reached, an estimate of the electron beam profile for HL-LHC can only be obtained in terms of simulations.
-For this purpose, the measured distribution is compressed to the inner e-beam radius of $0.97$~mm and than an input distribution with $40\ 000$ particles generated.
-The electric field can in general be scaled with the electron beam current and energy, but can not easily be scaled with the inner and outer radius of the electron beam distribution.
-Therefore it is essential to track a distribution with the correct dimensions.
-The potential and fields can then be calculated with the particle in cell simulation code WARP \cite{warp} yielding the relative electric field strengths shown in Fig.~\ref{core:fig:1}.
-Further details on the measurements and WARP simulations can be found in~\cite{hel_res_field_stancari_2017}.
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice.png}
-	\end{minipage}
-	\caption{Calculated relative electric field for the new CHG1b hollow electron gun in the x-y plane and as projection on the x- and y-axis respectively based on measurements at the Fermilab test stand combined with WARP simulations for the calculation of the electric potentials and fields.}
-	\label{core:fig:1}
-\end{figure}
-
-As a first estimate for the fields at the proton beam core, one can use the relative electric field strength along the x- and y-axis, yielding roughly (Fig.~\ref{core:fig:1}, right):
-\begin{equation}
-	E_{\mathrm{rel}}\approx 5\%
-\end{equation}
-For a maximum kick of $392$~nrad (Eqn.~\ref{eqn:helkick}), we then obtain as estimated dipole kick amplitude at the proton beam core:
-\begin{equation}
-\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \leq \SI{19.6}{nrad}
-\end{equation}
-
-\section{LHC experiments\label{sec:exp}}
-\subsection{Overview of LHC experiments\label{sec:exp_sum}}
-The experiments at the LHC presented in this paper were aiming at deriving tolerances on the residual HEL field at the location of the beam core in case of pulsed operation.
-In total, two experiments were conducted, the first one in 2016 \cite{resexmd2016} and the second one in 2017 \cite{resexmd2017}.
-The beam and machine parameters for both experiments are summarized in Table~\ref{tab:md_param}.
+% insert table here to put it on top of page
 \begin{table*}
@@
 \end{table*}
-As there is currently no HEL installed in the LHC, the kick on the beam core is emulated by a dipole kick applied with the transverse damper of which more details are given in Section~\ref{sec:adt}.
-Explicitly, all orders higher than the first order are neglected.
-Preceding the beam experiments, simulations were performed to identify the most effective pulsing pattern for the $k^{\mathrm{th}}$~turn pulsing (see Sec.~\ref{sec:pattern}) and yield estimates for losses and emittance growth.
-The two $k^{\mathrm{th}}$~turn pulsing patterns with the largest effect, one $k^{\mathrm{th}}$~turn pulsing pattern with no effect and the random excitation could then be experimentally tested in the LHC limited obviously by the beam time available.
-To at least test the reproducibility of the results on the example of one pulsing pattern, the $7^{\mathrm{th}}$~turn turn pulsing was tested in the 2016 experiments and then repeated in the 2017.
-During the experiments the losses were measured with the fast beam current transformers (FBCT) and the emittance and transverse beam profiles with the beam synchrotron radiation telescope (BSRT).
-The analysis of the BSRT profiles is quite involved and we therefore concentrate in this paper only on the main results.
-A detailed description of the analysis can be found in \cite{bsrtprofinj} and for the individual experiments in \cite{resexmd2016,resexmd2017}.
+% insert table here to put it on top of page
+
+\subsection{Kicks in the central region (main solenoid)}
+\label{core:sec:2}
+For a perfectly uniform, annular and radially symmetric profile, the field in the region of the proton beam core vanishes as can be seen from Eq.~\ref{eq:field_2} and is illustrated in Fig.~\ref{fig:hel_field}.
+Fields at the location of the proton beam core thus only arise in case of profile imperfections of the electron beam, in particular if the radial symmetry is broken.
+Recently, new measurements of the newest hollow electron gun CHG1b could be performed at the Fermilab electron lens test stand \cite{hel_test_stand_fnal}.
+A typical measurement of the electron beam current density obtained during these measurements is shown in Fig.~\ref{core:fig:0}.
+\begin{figure}[t]
+		\centering
+		\includegraphics[width=0.8\linewidth]{e-field-measured.png}
+	\caption{Current density from measurements of the CHG1b hollow electron gun at the Fermilab electron lens test stand in 2017 \cite{hel_res_field_stancari_2017}.}
+	\label{core:fig:0}
+\end{figure}
+As at the test stand the solenoidal fields required for the HL-LHC electron lens can not be reached, an estimate of the electron beam profile for HL-LHC can only be obtained through simulations.
+For this purpose, the measured distribution is compressed to the inner electron beam radius of $0.97$~mm.
+Afterwards an input distribution with $40\ 000$ particles generated.
+The potential and fields can then be calculated with the particle in cell simulation code WARP~\cite{warp} yielding the relative electric field strengths shown in Fig.~\ref{core:fig:1}.
+The electric field can in general be scaled with the beam current and energy, but can not easily be scaled with the inner and outer radius of the electron beam distribution.
+Therefore it is essential to track a distribution with the correct dimensions.
+Further details on the measurements and WARP simulations can be found in~\cite{hel_res_field_stancari_2017}.
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice.png}
+	\end{minipage}
+	\caption{Calculated relative electric field for the new CHG1b hollow electron gun in the x-y plane and as projection on the x- and y-axis.
+The field calculations are based on measurements at the Fermilab test stand combined with WARP simulations for the calculation of the electric potentials and fields.}
+	\label{core:fig:1}
+\end{figure}
+
+Using these measurements and simulations of the relative electric field strength along the x- and y-axis (Fig.~\ref{core:fig:1}), we obtain
+\begin{equation}
+	E_{\mathrm{rel}}\approx 5\%
+\end{equation}
+For a maximum kick of $392$~nrad as derived in Eq.~\ref{eqn:helkick}, the estimated dipole kick amplitude at the proton beam core then is:
+\begin{equation}\label{eqn:kick_central}
+\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \SI{19.6}{nrad}
+\end{equation}
+
+\section{LHC experiments on the effect of a resonant and random excitation on the beam core\label{sec:exp}}
+%put picture on top
+\begin{figure*}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		2016 experiment \textcolor{red}{(update picture)}\\
+		\includegraphics[width=0.7\linewidth]{bunchfilling.png}	
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		$ $\\
+		2017 experiment\\
+		\includegraphics[width=0.7\linewidth]{bunchfilling.png}	
+	\end{minipage}
+	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).}
+\end{figure*}
+%put picture on top
+
+\subsection{Overview of LHC experiments\label{sec:exp_sum}}
+The experiments at the LHC which are presented in this paper were aiming at deriving tolerances on the residual HEL field at the location of the beam core in case of pulsed operation.
+In total, two experiments were conducted, the first one in 2016 \cite{resexmd2016} and the second one in 2017 \cite{resexmd2017}.
+Beam and machine parameters for both experiments are summarized in Table~\ref{tab:md_param}.
+As there is currently no HEL installed in the LHC, the kick on the beam core was emulated by a dipole kick applied with the transverse damper of which more details will be given in Section~\ref{sec:adt}.
+Preceding the beam experiments, simulations were performed to identify the most effective $k^{\mathrm{th}}$~turn pulsing pattern and yield estimates for losses and emittance growth.
+The two $k^{\mathrm{th}}$~turn pulsing patterns with the largest effect, one $k^{\mathrm{th}}$~turn pulsing pattern with no effect and the random excitation could then be experimentally tested in the LHC.
+In order to test the reproducibility of the results at least on the example of one pulsing pattern, the $7^{\mathrm{th}}$~turn pulsing was first tried in 2016 and then repeated in 2017.
+
+During the experiments the losses were measured with the fast beam current transformers (FBCTs) and the emittance and transverse beam profiles with the beam synchrotron radiation telescope (BSRT).
+The analysis of the BSRT profiles is quite involved and we will therefore focus in this paper only on the direct comparison of the profiles.
+A detailed description of the analysis with more details can be found in \cite{bsrtprofinj} and for the individual experiments in \cite{resexmd2016,resexmd2017}.
 
@@
 The primary function of the transverse feedback system in LHC, also referenced to as ADT, is to provide injection oscillation damping and actively damp the coupled bunch instabilities driven by the machine impedance.
-The main building blocks of the ADT system are stipline pickups at the position Q7 and Q9 at Point~4 of the LHC, connected to the beam position measurement modules at the surface, the digital signal processing modules (mDSPU) and a set of tetrode power amplifiers feeding electrostatic kickers in the RF zone of the LHC \cite{adt_sum_2008,adt_sum_2011}.
-
-Though originally designed to damp the undesired transverse activities only, the damper is being routinely used for sophisticated beam excitations as well (e.g.
-Abort gap and injection gap cleaning, single bunch excitation for tune and linear coupling measurements, machine development etc.) thanks to the state of the art digital signal processing hardware.
+The main building blocks of the ADT system are stipline pickups at the position Q7 and Q9 at Point~4 of the LHC, connected to the beam position measurement modules at the surface, the digital signal processing modules (mDSPU) and a set of tetrode power amplifiers feeding electrostatic kickers in the RF zone of the LHC~\cite{adt_sum_2008,adt_sum_2011}.
+Though originally designed to damp the undesired transverse activities only, the damper is being routinely used for sophisticated beam excitations as well thanks to the state of the art digital signal processing hardware.
+These include among others for example abort and injection gap cleaning, single bunch excitation for tune and linear coupling measurements and special modes of operation for dedicated experiments at the LHC.
 
@@
 All excitation signals referenced to in this paper, are digitally synthesized in the damper's digital signal processing units and therefore can be considered as ``noise-free''.
+This point is important as then the effect on the beam can be attributed entirely to the applied excitation and any effect due to unwanted residual noise can be excluded.
 
@@
 Both schemes are illustrated in Fig.~\ref{fig:fill}.
-\begin{figure*}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		2016 experiment \textcolor{red}{(update picture)}
-		\includegraphics[width=1.0\linewidth]{bunchfilling.png}	
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-	\centering
-	2017 experiment
-	\includegraphics[width=1.0\linewidth]{bunchfilling.png}	
-	\end{minipage}
-	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).}
-\end{figure*}
 
@@
 The measured values are therefore believed to be known with an error margin of 10-15\% for the experiments in 2017 and 50\% in 2016.
-The error for 2016 is much larger as no calibration of the ADT was performed beforehand, which was done in 2017.
-As an example Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor.
+The error in 2016 is much larger as no calibration of the ADT was performed beforehand, which was done in 2017.
+As an example Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor during the 2017 experiments and an excitation with 8th turn pulsing.
 The measured activity is well in line with the expected excitation pattern.
@@
 	\includegraphics[width=1.0\linewidth]{bunchfilling_measured.png}	
-	\caption{\label{fig:fill_meas} Bunch by bunch excitation detected by the real time transverse activity monitor during the 2017 experiment (8th turn pulsing}
+	\caption{\label{fig:fill_meas} Bunch by bunch excitation detected by the real time transverse activity monitor during the 2017 experiment and an excitation with 8th turn pulsing.}
 \end{figure}
@@
 For the preparation and interpretation of the experiments two different types of simulations were performed:
-\begin{table}[b]
+\begin{enumerate}
+	\item tracking of a Gaussian distribution, referred to as ``distribution tracking" in this paper, to obtain loss rates and emittance growth,
+	\item Frequency map analysis (FMA) for the visualization of resonances~\cite{fmalaskar}.
+\end{enumerate}
+For both simulation types, the tracking code Lifetrac \cite{lifetrac} was used.
+The simulation parameters are summarized in Table~\ref{tab:sim_param}.
+Further details on the simulations are given in \cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
+\begin{table}[h]
 	\caption{\label{tab:sim_param}%
-		Summary of simulation parameters for the Distribution Tracking and FMA analysis. For details see \cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
+		Summary of simulation parameters for the distribution tracking and FMA analysis. For details see \cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
 	}
@@
 		\begin{tabular}{lcc}
-			Parameter & Distribution Tracking & FMA \\
+			Parameter & distribution tracking & FMA \\
 			\colrule
@@
 \end{table}
-\begin{enumerate}
-	\item tracking of a Gaussian distribution (Distribution Tracking) to obtain loss rates and emittance growth,
-	\item Frequency map analysis (FMA)~\cite{fmalaskar} for the visualization of resonances.
-\end{enumerate}
-For both simulation types, the tracking code Lifetrac \cite{lifetrac} was used.
-The simulation parameters are summarized in Table~\ref{tab:sim_param}.
-Further details on the simulations can be found in \cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
 
 \subsection{Simulation and experimental results\label{sec:simex}}
-In the following we will summarize the simulation and experimental results.
-We will jump forth and back between both as the experimental results are best understood on the base of the simulations.
-In summary, the most efficient pulsing patterns could be successfully predicted in simulations (Sec.~\ref{sec:pattern}).
-Quantitative predictions of loss rates and emittance growth proved however challenging as both observables are influenced in addition by the natural noise present in the LHC originating for example from ground motion and magnetic field ripple.
+In the following we will jump forth and back between both experimental and simulation results as the experimental results are best understood on the base of the simulations.
+In summary, the most efficient pulsing patterns could be successfully predicted in simulations, which we will demonstrate on the example of the 2016 experiments (see Sec.~\ref{sec:pattern}).
+Afterwards we will present the simulation and experimental results of the tested excitation patterns, explicitly the $10^{\mathrm{th}}$~turn, $7^{\mathrm{th}}$~turn and $8^{\mathrm{th}}$~turn pulsing in Secs.~\ref{sec:simex10}--\ref{sec:simex8}) as well as the random excitation in Sec.~\ref{sec:simexran}.
+As a last chapter (Sec.~\ref{sec:damp}), we will address the influence of the transverse damper in the presence of an external excitation source on emittance, losses and beam distribution.
+
+Quantitative predictions of loss rates and emittance growth proved in general challenging as both observables are influenced in addition by the natural noise present in the LHC, originating for example from ground motion and magnetic field ripple.
 The observed effect is then the result of the interaction of the external excitation with the natural noise.
-At injection, this natural noise component is not known for the LHC.
+At injection, this natural noise component is not known and thus presents a very influential but undefined input to the simulations.
 Beside the natural noise, also collective effects like intra-beam scattering and electron cloud influence the time evolution of emittance and losses.
-Both are not taken into account in the simulations.
-
-The general dynamics of a resonant excitation seen in simulations is a fast adjustment to a new equilibrium distribution.
-In the presence of only a resonant excitation, this distribution is stable with a constant emittance as shown for example in Figs.~\ref{fig:patternsim} and \ref{fig:10thsim}.
-If the resonant excitation interacts with the natural noise sources, the beam distribution still changes to a new equilibrium distribution but this change is also accompanied by a resonant excitation amplitude dependent emittance growth as seen for example in Figs.~\ref{fig:10thsim},\ref{fig:7thsim} and \ref{fig:8thsim}, and also observed experimentally, see for example Figs.~\ref{fig:10thexp} and ~\ref{fig:7thexp}.
-An excitation with random uniform noise on the other hand results in constant emittance growth without any initial adjustment phase.
-The interaction with the natural noise source then only manifests itself in an underestimation of loss rates and emittance growth.
-This is visible in simulations, see Fig.~\ref{fig:ransim} as well as experimentally, see Fig.~\ref{fig:ranexp}.
-The effect of a random excitation was also studied in more detail in \cite{md1433_noise_top_energy,md_noise_bbLHC} for colliding beams at the LHC.
+Both are also not taken into account in the simulations.
 
 \subsubsection{Prediction of the most efficient pulsing pattern\label{sec:pattern}}
+The most efficient pulsing pattern can be predicted by the comparison of the loss rate and emittance growth from the distribution tracking for the different pulsing patterns.
+As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pattern.
 \begin{figure}[h]
@@
 	\end{minipage}	
-	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
+	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance for different pulsing patterns obtained from distribution tracking based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
 The resonant excitation is applied in both planes with an amplitude of 96~nrad.
@@
 \end{figure}
+
+The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and a uniform random excitation.
+As the bunch length decreases in accordance with the losses, the losses can be mostly associated to off-momentum particles.
+Only $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as random excitation exhibit considerable emittance growth.
+Compared to any resonant excitation, the random excitation has by far the strongest effect.
+
+The sensitivity to $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer} suggesting that latter are responsible for the observed sensitivity.
+The driven resonances are revealed by the FMA analysis shown in Fig.~\ref{fig:patternfma}.
+The $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing.
+As octupoles can only drive even resonances, the source of the $7Q_x$ resonances are the sextupoles while the octupoles only generate the tune footprint.
+The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect can be thus attributed to the magnetic field errors, implying also the seed chosen for the simulation.
 \begin{figure}[h]
@@
 	\end{minipage}
-	\caption{\label{fig:patternfma} FMA analysis for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
+	\caption{\label{fig:patternfma} FMA analysis without excitation (top left) for $10^{\mathrm{th}}$~turn pulsing in H+V (top left) and for $7^{\mathrm{th}}$~pulsing only in H (bottom left) and only in V (bottom right) and  based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
 The excitation is 120~nrad in the corresponding plane.
-The absence of a strong excitation of any resonance in case of $7^{\mathrm{th}}$~turn pulsing only in V and the strong excitation in case of pulsing only in H further confirms that the $7Q_x$ resonance is the main resonance driven.
-For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V (see Appendix~\ref{app:sec:fma:10}, Fig.~\ref{app:fig:fma:10}).}
-\end{figure}  
-The most efficient pulsing pattern can be predicted by the comparison of the loss rate and emittance growth from the Distribution Tracking for the different pulsing patterns.
-As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pulsing pattern.
-The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and a uniform random excitation.
-As the bunch length decreases in accordance with the losses, the losses can be mostly associated to off-momentum particles.
-Only for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as for a random excitation considerable emittance growth is visible.
-Compared to any $k^{\mathrm{th}}$~turn pulsing pattern, the random excitation has by far the strongest effect.
-The sensitivity to $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are the strong sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer}.
-This suggests that latter are the source of non-linearity responsible for this sensitivity.
-The driven resonances are revealed by the FMA analysis shown in Fig.~\ref{fig:patternfma}.
-The $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing for an excitation in both planes.
-As octupoles can only drive even resonances, the source of the $7Q_x$ resonances are the sextupoles while the octupoles only generate the tune footprint.
-The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect thus is based on the magnetic field errors assigned, explicitly also the seed chosen for the simulation.
-
+The absence of a strong excitation of any resonance in case of $7^{\mathrm{th}}$~turn pulsing only in V and the strong excitation in case of pulsing only in H confirms the excitation of the $7Q_x$ resonance.
+For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V (see Appendix~\ref{app:sec:10}, Fig.~\ref{app:fig:fma:10}).}
+\end{figure}
+
+The experienced or also inclined unexperienced reader might have noticed that the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time, which usually points to an error in the chosen simulation setup.
+This behavior is however typical for any resonant excitation and can be ascribed to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium.
+As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}, and also any other simulations over $10^6$ turns presented in this paper.
+Besides an increased emittance, the new distribution also becomes non-Gaussian.
+This is illustrated by means of the residual in respect to the Gaussian distribution for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in Fig.~\ref{fig:patternhist}.
+This initial fast change of the beam distribution is not only an artifact of the simulation, but has been, to the knowledge of the authors, also for the first time experimentally observed during the 2016 and 2017 LHC experiments presented in this paper.
+The timescale of this fast adjustment is in experiments however much slower than predicted in simulations - on the timescale of minutes compared to seconds.
 \begin{figure}[t]
@@
 	\end{minipage}	
-	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn pulsing (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from Distribution Tracking simulations based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ and standard errors.
+	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn pulsing (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from distribution tracking simulations based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ and standard errors.
 The excitation is applied in both planes with an amplitude of 96~nrad.
@@
 
-As already addressed earlier, the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time.
-This behavior is typical for any $k^{\mathrm{th}}$ turn pulsing and can be ascribed to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution.
-As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}, and also any other simulations over $10^6$ turns presented in this paper.
-Besides an increased emittance, the new distribution also becomes non-Gaussian.
-This can be seen by means of the residual in respect to the Gaussian distribution illustrated for example for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in Fig.~\ref{fig:patternhist}.
-For the purpose of comparison, also the distribution in both planes for a random excitation is shown.
-This initial fast change of the beam distribution is not only an artifact of the simulation, but has been, to the knowledge of the authors, also for the first time experimentally observed as will be presented later in this paper.
-The timescale of this fast adjustment is in experiments however much slower than predicted in simulations - on the timescale of minutes compared to seconds.
+In the presence of only a resonant excitation, this distribution is stable in simulations with a constant emittance.
+Experimentally however, an increase of the emittance growth with the applied excitation amplitude is observed, see for example Figs.~\ref{fig:10thexp} and ~\ref{fig:7thexp}, not matching the prediction by simulations.
+By adding a random noise component representative for the natural noise present in the LHC, the constant emittance is changed to an excitation amplitude dependent emittance growth as depicted for example in Figs.~\ref{fig:10thsim} and \ref{fig:7thsim}.
+An excitation with random uniform noise on the other hand results in constant emittance growth without any initial adjustment phase.
+The interaction with the natural noise source then only results in an underestimation of loss rates and emittance growth as it is basically equivalent to the application of random noise with an increased amplitude.
+In addition to the experiments presented in this paper, the effect of a random excitation was also studied in a separate experiment at the LHC for the case of colliding beams \cite{md1433_noise_top_energy,md_noise_bbLHC}.
 
@@
 \begin{enumerate}
-	\item Increase of the loss rate with the excitation amplitude as shown in Figs.~\ref{fig:10thexp} and \ref{fig:10thexpsim}.
-	\item Increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane shown in Fig.~\ref{fig:10thexp}.
-	\item Change of the beam distribution due to the resonant excitation illustrated in Fig.~\ref{fig:10thexp} and \ref{fig:10thexpprof}.
-	\item As loss rates, emittance growth and distribution stayed unchanged for the $4$~reference bunches, the above observations can be truly associated with the effect of the resonant excitation.	
+	\item Increase of the loss rate with the excitation amplitude (Figs.~\ref{fig:10thexp}.
+	\item Increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane (Fig.~\ref{fig:10thexp}).
+	\item Change of the beam distribution for the excited bunches (Fig.~\ref{fig:10thexpprof}).	
 \end{enumerate}
+As loss rates, emittance growth and beam distribution stayed unchanged for the $4$~reference bunches, the above observations can be also truly associated with the effect of the $10^{\mathrm{th}}$~turn pulsing.
 \begin{figure}[h]
@@
 	\end{minipage}	
-	\caption{\label{fig:10thexp} Relative intensity (top) and relative emittance (bottom) averaged over the bunches experiencing the same excitation amplitude during the 2016 experiments.
-For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Sec.~\ref{sec:adt}).
-The different colored solid lines labeled with $n\Delta A$ indicate the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation as their envelope.}
-\end{figure}
+	\caption{\label{fig:10thexp} Relative intensity (top) and relative horizontal (bottom left) and vertical (bottom right) emittance averaged over the bunches experiencing the same excitation amplitude during the 2016 experiments.
+For all bunches the damper was not active.
+Relative refers here to the normalization to the initial value.
+For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Fig.~\ref{fig:fill}).
+The different colored solid lines labeled with $n\Delta A$ indicate the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation indicated as envelope.}
+\end{figure}
+
+Besides the determination of the scaling of the loss rate and emittance growth with the excitation amplitude important for the actual specification of the hollow electron lens, the vertical emittance actually features the strange behavior predicted in simulations: a fast adjustment phase of the beam distribution visible as a rapid increase of the emittance followed by the reach of a new equilibrium distribution apparent by a slower continuous emittance growth.
+These two phases are indicated in Fig.~\ref{fig:10thexp} in blue and black.
+
 \begin{figure}[h]
@@
 The red line is the current profile minus its Gaussian fit, which gives and indication for the deviation of the distribution from a Gaussian distribution.
-For details on the analysis see \cite{bsrtprofinj}.
+For details on the analysis it is referred to \cite{bsrtprofinj}.
 The distribution of the reference bunch (left), here slot 50, stays unchanged, while the bunch experiencing the maximum excitation (right), here slot 1300, clearly shows a change to a non-Gaussian distribution.}
 \end{figure}
+Thanks to the good performance of the Beam Synchrotron Radiation Telescope (BSRT)~\cite{bsrtprofinj}, the distribution changes indirectly observed as changes of the emittance could also be for the first time directly measured.
+As example Fig.~\ref{fig:10thexpprof} shows the vertical profile of a reference bunch and one bunch experiencing the maximum excitation of $5\cdot\Delta A=A_{\mathrm{max}}$.
+The distribution of the excited bunch not only changes but becomes clearly non-Gaussian, while the reference bunch stays unchanged.
 \begin{figure}[h]
@@
 (bottom left) and vert.
-(bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$.
+(bottom right) emittance obtained from simulations (distribution tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$.
 The solid line indicates an excitation with only $10^{\mathrm{th}}$~turn pulsing in V and the dotted line with $10^{\mathrm{th}}$~turn pulsing in V plus a random dipole noise component in H+V of 6~nrad.}
 \end{figure}
-
-Besides the determination of the scaling of the loss rate and emittance growth with the excitation amplitude as required for the specification of the hollow electron lens, the most novel observation is the direct measurement of the change of the beam distribution, here with the Beam Synchrotron Radiation Telescope (BSRT)~\cite{bsrtprofinj}.
-As an example for the change in distribution, the vertical profile of a reference bunch and one bunch experiencing the maximum excitation of $5\cdot\Delta A=A_{\mathrm{max}}$ are compared in Fig.~\ref{fig:10thexpprof}.
-The distribution of the excited bunch clearly changes and becomes non-Gaussian, while the reference bunch stays unchanged.
-
-Comparing these experimental results with the simulations including only the excitation through the $10^{\mathrm{th}}$~turn pulsing and no random component and emittance growth only in the vertical plane (solid line in Fig.~\ref{fig:10thsim}), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth observed in experiments.
-Instead, the simulations yield a constant emittance after an initial increase during the first $10^4$~turns or first second.
-So obviously there is an essential ingredient missing.
-As it turns out, the constant emittance can be changed to an excitation amplitude dependent emittance growth by adding a random dipole noise component with uniform distribution representing the natural noise present in the LHC (dotted line in Fig.~\ref{fig:10thsim}).
-As no noise estimate is available at injection for the LHC, a first estimate is obtained by scaling the estimate for 6.5~TeV \cite{md1433_noise_top_energy,md_noise_bbLHC} by the magnetic rigidity to the injection energy of 450~GeV.
+The simulation results from the distribution tracking for $10^{\mathrm{th}}$~turn pulsing and without (solid line) and with (dotted line) additional random noise component to emulate the natural noise present in the LHC are shown in Fig.~\ref{fig:10thsim}.
+As no estimate of the natural noise is available at injection for the LHC, a first estimate is obtained by scaling the estimate for 6.5~TeV \cite{md1433_noise_top_energy,md_noise_bbLHC} by the magnetic rigidity to the injection energy of 450~GeV.
 This yields a maximum kick amplitude at the transverse damper of approximately
@@
 \end{equation}
-Adding this random noise component, the simulations then yield correctly an excitation amplitude dependent emittance growth in the vertical plane after a first fast adjustment of the beam distribution visible as an increased initial value.
-Adding a random noise component has also the effect of increasing the loss rate.
-As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced, but a large difference still prevails as illustrated in Fig.~\ref{fig:10thexploss} by means of a qualitative comparison of the loss rates obtained in experiments and simulations.
-\begin{figure}[h]
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016_scale_amp_10v_ran_lbllong_sim.png}
+As can be seen from simulations, adding a random noise component not only yields the correct qualitative behavior of the emittance but also increases the loss rates.
+As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment can be thus further reduced.
+A large difference however still prevails which is illustrated in Fig.~\ref{fig:10thexploss} by means of a qualitative comparison of the loss rates obtained in experiments and simulations.
+\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.8\linewidth]{2016_scale_amp_10v_ran_lbllong_sim.png}
 	\caption{\label{fig:10thexploss} Comparison of scaling of loss rates with excitation amplitude for $10^{\mathrm{th}}$~turn pulsing.
-The experimental results are shown in black (maximum excitation with 48~nrad) and blue (maximum excitation with 96~nrad).
-The excitation amplitude error is 50\% due to limited measurement accuracy and the error on the loss rate is the pure statistical error.
-The simulation results (2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad) are shown in red.}
+The experimental results for a maximum excitation with 48~nrad are shown in black and for 96~nrad in blue.
+The excitation amplitude error is 50\% and the error on the loss rate is the pure statistical error.
+The simulation results for 2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad are shown in red.}
 \end{figure}
@@
 \subsubsection{$7^{\mathrm{th}}$ turn pulsing\label{sec:simex7}}
-\begin{table}[h]
+The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the experimental results.
+Unfortunate for this experiment, the machine tune was changed in standard operation from (64.28,59.31) in 2016 to (62.27,60.295) in 2017 accompanied by a slight change in optics considered irrelevant in the context of the presented measurements.
+The attempt to change back to the fractional tunes of (.28,.31) as chosen in 2016 lead to large losses during the 2017 experiment and therefore the 2017 working point of (62.27,60.295) had to be used instead.
+This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016 (see Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (see Fig.~\ref{fig:7th2017fma}).
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		no excitation
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$7^{\mathrm{th}}$ H+V
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png}
+	\end{minipage}	
+	\caption{\label{fig:7th2017fma} FMA analysis in frequency space without excitation (left) and for $7^{\mathrm{th}}$~turn pulsing in H+V (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
+The excitation is 96~nrad in both planes.
+The $7Q_x$ and $7Q_y$ resonance are both excited and in addition the 14th order $7Q_x+7Q_y$.}
+\end{figure}
+
+\begin{table}[b]
 	\caption{\label{tab:7thexp}%
@@
 \end{table}
-The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the results.
-However, in 2017 the machine tune was changed from (64.28,59.31) to (62.27,60.295) in standard operation accompanied by a slight change in optics.
-The attempt to change back to the fractional tune of (.28,.31) as chosen in 2016 lead to large losses during the experiment in 2017 and therefore the 2017 working point of (62.27,60.295) was used instead.
-This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016 (Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (Fig.~\ref{fig:7th2017fma}).
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		no excitation
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$7^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png}
-	\end{minipage}	
-	\caption{\label{fig:7th2017fma} FMA analysis in frequency space for $7^{\mathrm{th}}$~turn pulsing based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
-The excitation is 96~nrad in both planes.
-The $7Q_x$ and $7Q_y$ resonance are both excited and in addition the 14th order $7Q_x+7Q_y$.}
-\end{figure}
-Consequently, the effects on the beam also differ as can be illustrated by a comparison of the Distribution Tracking for both experiments~(Fig.~\ref{fig:7thsim}).
+Consequently, the effect of the $7^{\mathrm{th}}$~turn pulsing on the beam losses, emittance and beam distribution also differ as depicted by the comparison of the distribution tracking for both experiments in Fig.~\ref{fig:7thsim}, the qualitative comparison of the experimental results in Table~\ref{tab:7thexp}, the comparison of the measured emittance in Fig.~\ref{fig:7thexp} and loss rate in Fig.~\ref{fig:7thexploss}.
 \begin{figure}[h]
@@
 	\end{minipage}
-	\caption{\label{fig:7thsim} Relative intensity (top left) and relative vert.
-emittance (top right), and hor.
-(bottom left) and vert.
-(bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$ and 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
-The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $7^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
-\end{figure}
-
+	\caption{\label{fig:7thsim} Relative intensity (top left), relative vertical emittance (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$ and 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $7^{\mathrm{th}}$~turn pulsing with two different excitation amplitudes plus a random dipole noise component in H+V of 6~nrad.}
+\end{figure}
+
+The simulation results from the distribution tracking shown in Fig.~\ref{fig:7thsim} in general predict in terms of emittance for the two experiments:
+\begin{description}
+	\item[2016 experiment] Excitation amplitude dependent emittance growth in the horizontal but not in the vertical plane.
+	\item[2017 experiment] Excitation amplitude dependent emittance growth in the vertical plane but not in the horizontal plane for an excitation in V and H+V.
+For an excitation only in H no emittance growth is observed.
+\end{description}
+Similar changes of the emittance as predicted in simulations were also experimentally observed as depicted in Fig.~\ref{fig:7thexp}, where emittance growth was observed only in the horizontal plane for pulsing in H in 2016, and only in the vertical plane for pulsing in V and H+V in 2017 with no emittance growth at all for pulsing in H.
+Just as for the $10^{\mathrm{th}}$~turn pulsing, the emittance growth can be divided also for the $7^{\mathrm{th}}$~turn pulsing in a phase of fast adjustment of the beam distribution and the reach of a new equilibrium state indicated in blue (fast adjustment) and black (equilibrium state) in Fig.~\ref{fig:7thexp}.
+Also in this case, the distribution changes are directly visible on the beam profiles taken with the BSRT (see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexpprof}).
 \begin{figure}[h]
@@
 	\end{minipage}	
-	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude.}
-\end{figure}
-Similar changes of the emittance are also experimentally observed as shown in Fig.~\ref{fig:7thexp}, where emittance growth was observed only in the horizontal plane for 2016 for pulsing in H, and only in the vertical plane for 2017 for pulsing in V and H+V (for pulsing in H no emittance growth is observed).
-Just as for the $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:10thexp}), the emittance growth can be divided also for the $7^{\mathrm{th}}$~turn pulsing in a phase of fast adjustment of the beam distribution and the reach of a new equilibrium state (Fig.~\ref{fig:7thexp}).
-Also in this case, the distribution changes are also directly visible on the beam profiles measured with the BSRT (Appendix~\ref{app:sec:bsrtprof}, Fig.~\ref{fig:7thexpprof}).
+	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude.
+For all bunches the damper was not active.}
+\end{figure}
 
 The loss rates measured during the experiment however differ from the simulations in terms of for which case the largest rate is observed as well as quantitatively for the specific cases.
-For example smaller losses are expected for the 2016 experiments compared to the 2017 experiments from simulations, while experimentally higher loss rates were measured in 2016 (Fig.~\ref{fig:7thexploss}).
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1.0\linewidth]{2017_scale_amp_7h_lbllong.png}
+For example smaller losses are expected for the 2016 experiments compared to the 2017 experiments from simulations, while experimentally higher loss rates were measured in 2016 (see Fig.~\ref{fig:7thexploss}).
+For the 2017 experiment, the highest loss rates are expected for pulsing in H+V, than in H and then with a considerably reduction in V.
+During the experiment comparable loss rates are however measured for all three planes (see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexploss2017})
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\linewidth]{2016+2017_scale_amp_7h_lbllong.png}
 	\caption{\label{fig:7thexploss} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in H and data obtained during the 2016 experiment (black and blue) and 2017 experiment (red and yellow).
@@
 \subsubsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
-\begin{figure}[b]
+Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they in addition have a large effect on the halo, which is in general feasible considering the highly nonlinear field generated at the location of the halo particles by the HEL.
+With this motivation in mind the $8^{\mathrm{th}}$~turn pulsing pattern was studied, for which a small effect in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing is expected from the distribution tracking summarized in Fig.~\ref{fig:8thsim}.
+\begin{figure}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 	\end{minipage}	
-	\caption{\label{fig:8thsim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+	\caption{\label{fig:8thsim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (distribution tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
 The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $8^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
-\end{figure}
-Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they on the other hand have a large effect on the tails under consideration of the non-linear field generated by the electron lens itself.
-With this motivation in mind the $8^{\mathrm{th}}$ turn pulsing pattern was studied, for which a small effect is expected in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing as can be demonstrated in the Distribution Tracking, for which the excitation amplitudes had to be increased to 96~nrad to have a visible effect (Fig.~\ref{fig:8thsim}).
-The loss rates and also changes in bunch length are very small and any differences are here just due to statistical fluctuations.
-In the horizontal and vertical plane a small emittance growth due to the excitation is visible, but without any clear dependency on the amplitude nor plane of the $8^{\mathrm{th}}$ turn pulsing.
-Adducting the FMA analysis in amplitude space depicted in Fig.~\ref{fig:8th2017fmaamp}, the $8^{\mathrm{th}}$ turn pulsing drives the $16Q_y$ and \mbox{$8Q_x-4Q_y$} resonances and several even higher orders.
-These are in general very high order resonances whose effect is thus in general expected to be small.
-\begin{figure}[t]
+%\end{figure}
+%\begin{figure}[t]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 	\end{minipage}	
-	\caption{\label{fig:8th2017fmaamp} FMA analysis in amplitude space for $8^{\mathrm{th}}$~turn pulsing based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
+	\caption{\label{fig:8th2017fmaamp} FMA analysis in amplitude space without excitation (left) and for $8^{\mathrm{th}}$~turn pulsing (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
 The excitation is 96~nrad in both planes.
@@
 
+For excitation amplitudes smaller than 96~nrad no visible effect is indeed observed.
+In the horizontal and vertical plane a small emittance growth due to the excitation is visible, but without any clear dependency on the amplitude nor plane.
+Adducting the FMA analysis in amplitude space shown in Fig.~\ref{fig:8th2017fmaamp}, the $8^{\mathrm{th}}$ turn pulsing drives the $16Q_y$ and \mbox{$8Q_x-4Q_y$} resonances and several even higher orders.
+These are in general very high order resonances whose effect is thus expected to be small.
+
+The only small effect on the beam expected from simulations could then be also confirmed experimentally during the 2017 experiment.
+For this purpose the excitation amplitude was increased to 96~nrad, which is the maximum kick strength reachable with the transverse damper at injection without risking saturation.
+As depicted in Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the excitation amplitude for an excitation in H+V.
+For an excitation only in V or only in H, no statistically relevant increase was observed.
 \begin{figure}[h]
@@
 \end{figure}
-The only small effect on the beam expected from simulations could then be also confirmed experimentally during the 2017 experiments.
-For this purpose the excitation amplitude was increased to 96~nrad, which is the maximum kick strength reachable with the transverse damper at injection without risking saturation.
-As depicted in Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the excitation amplitude for an excitation in H+V.
-For an excitation only in V or only in H, no statistically relevant increase was observed.
-In addition, the beam distribution also changed in the horizontal plane for an excitation only in H as visible in the BSRT profiles (Appendix~\ref{app:sec:bsrtprof}, Fig.~\ref{fig:8thexpprof}) indicating an increase at around $2.0~\sigma$ simultaneously with a depletion of the beam core.
-For an excitation in H+V the emittances of the references bunches were much smaller, around $1.8~\mu$m normalized emittance on average, than the one of the excited bunches ranging from $2.3~\mu$m to $3.0~\mu$m \cite{resexmd2017}.
+
+In the horizontal plane, a change in beam distribution was measured with the BSRT profiles (Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexpprof}) for an excitation only in H, explicitly an increase of the distribution at around $2.0~\sigma$ simultaneously with a depletion of the beam core.
+For an excitation in H+V the emittances of the references bunches were much smaller, around $1.8~\mu$m normalized emittance on average, than the ones of the excited bunches ranging from $2.3~\mu$m to $3.0~\mu$m as described in \cite{resexmd2017}.
 The references bunches thus experienced a stronger emittance growth due to intra-beam scattering than the excited bunches due to the excitation on top of their smaller emittance growth due to intra-beam scattering.
+Therefore no excitation amplitude dependent emittance growth was observed for an excitation in H+V albeit it was recorded for an excitation only in H.
+In the vertical plane, the beam distribution and thus emittance stayed unchanged.
+For an excitation only in V, no dependence on the excitation was measured.
 
 \subsubsection{random excitation\label{sec:simexran}}
-In view of the electron lens, the random excitation presents the pulsing pattern leading to the fastest halo depletion rates.
-On the other hand it also represents the most dangerous pattern for the beam core.
-The aim of this experiment was to test the effect on the beam core in comparison to the resonant excitation represented in the 2017 experiment by the $7^{\mathrm{th}}$~turn pulsing as for $8^{\mathrm{th}}$~turn pulsing basically no effect was observed.
-The random excitation is so powerful in cleaning the halo as it excites all frequencies equally.
-This can be seen formidably with the FMA analysis in amplitude space depicted in Fig.~\ref{fig:ran2017fmaamp}.
-The diffusion in tune is increased equally independent of the amplitude.
-\begin{figure}[t]
+The random excitation presents in general the pulsing pattern leading to the fastest halo depletion rates achievable with an HEL.
+On the other hand this pattern is also the most dangerous one for the proton beam core.
+The aim of this experiment was to test the effect on the core in comparison to the resonant excitation represented in the 2017 experiment by primarily the $7^{\mathrm{th}}$~turn pulsing as for $8^{\mathrm{th}}$~turn pulsing basically no effect was observed.
+The random excitation is in general so powerful as it excites all frequencies of the beam equally.
+\begin{figure}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 	\end{minipage}	
-	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space for a uniform random excitation with 1~nrad amplitude in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
-The diffusion in tune is increased equally independent of the amplitude.}
-%\end{figure}
-%\begin{figure}[h]
+	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space without excitation (left) and for a uniform random excitation with 1~nrad amplitude (right) in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).}
+\end{figure}
+This can be seen formidably on the example of the FMA analysis in amplitude space, depicted in Fig.~\ref{fig:ran2017fmaamp}, for a random excitation as applied during the 2017 experiment.
+In this case, the diffusion in tune is increased equally by the random excitation independent of the particle amplitude.
+
+\begin{figure}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 	\end{minipage}
-	\caption{\label{fig:ransim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+	\caption{\label{fig:ransim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
 The solid line indicates the case with no excitation, and the dotted and dashed line the results for a uniform random excitation of 12~nrad and 24~nrad respectively.}
 \end{figure}
-
-The results from the Distribution Tracking are shown in Fig.~\ref{fig:ransim}.
+The results from the distribution tracking are shown in Fig.~\ref{fig:ransim}.
 In contrast to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing, the effect of a random excitation is very symmetric:
@@
 Only a small increase or decrease is observed if the excitation is applied in both planes.
-The difference can be attributed to coupling.
+The difference could be due to coupling.
 	\item The random excitation leads to an increased constant emittance growth rate which is reached without undergoing an initial adjustment phase.
@@
 
-Qualitatively, this behavior could be also confirmed experimentally during the 2017 experiments and as example the emittances are depicted in Fig.~\ref{fig:ranexp} for pulsing in V and in H+V, showing no emittance growth in the horizontal for an excitation only in V and emittance growth in both planes for an excitation in H+V.
-All emittance growth rates are in a comparable range.
-Note that despite the large emittance growth, the growth rate stays constant and there is no fast adjustment phase succeeded by a new equilibrium state as was the case for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
-\begin{figure}[h]
+\begin{figure}[b]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 \end{figure}
-These large changes of the emittance can of course also be seen directly in the BSRT profiles shown in Fig.~\ref{fig:ranexpprof} for a random excitation only in V.
-In comparison to for example the $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:10thexpprof}), the distribution also stays Gaussian in case of a random excitation while for the $10^{\mathrm{th}}$~turn pulsing it clearly adopts a non-Gaussian shape.
+Qualitatively, this behavior could be also confirmed experimentally during the 2017 experiments, where excitation amplitude dependent emittance growth was only recorded in the plane of excitation with all excitation amplitude dependent emittance growth rates in a comparable range.
+As example the relative horizontal and vertical emittance for pulsing in V and in H+V are depicted in Fig.~\ref{fig:ranexp}.
+Note that as expected from simulations and despite the large excitation amplitude dependent emittance growth, the growth rate stays constant and there is explicitly no fast adjustment phase succeeded by a new equilibrium state as was observed for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
+
 \begin{figure}[h]
@@
 \end{figure}
-
-The experimental results also confirm the small loss predicted in simulations.
-The losses then suddenly increase for 12~nrad, which is most likely the point when, due to the large emittance increase, the physical aperture starts eating into the more central region of the distribution causing higher losses.
-Furthermore, the loss rates also add up, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only V.
-This is illustrated in  Fig.~\ref{fig:ranexploss}, which also shows the loss rates for the $7^{\mathrm{th}}$~turn pulsing in comparison.
-Furthermore the loss rate is considerably reduced if the transverse damper is active, which is not the case for the $7^{\mathrm{th}}$~turn pulsing.
-We will address this subject separately in the next Section, Sec.~\ref{sec:damp}.
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-	\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranv_lbllong.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranhv_lbllong.png}
-	\end{minipage}	
-	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing in V (left) and H+V (right) measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.
-Note that for $7^{\mathrm{th}}$~turn pulsing the transverse damper does not influence the loss rate, while for a random excitation the loss rate is considerably reduced for the bunches with transverse damper active.}
-\end{figure}
+The large changes of the emittance are of course also directly visible in the beam profiles measured with the BSRT shown in Fig.~\ref{fig:ranexpprof} for a random excitation only in V.
+In comparison to the $10^{\mathrm{th}}$~turn pulsing shown in Fig.~\ref{fig:10thexpprof}, the distribution also stays Gaussian in case of a random excitation while for the $10^{\mathrm{th}}$~turn pulsing it clearly adopts to a non-Gaussian shape.
+
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=0.8\linewidth]{2017_scale_amp_ranhv_lbllong.png}
+	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars.}
+\end{figure}
+The experimental results for a maximum excitation amplitude of $A_{\mathrm{max}}=1$~nrad and $A_{\mathrm{max}}=6$~nrad also confirm the in general small losses predicted in simulations.
+The losses then suddenly increase once the maximum excitation amplitude is increased to $A_{\mathrm{max}}=12$~nrad, which is most likely the point when, due to the large emittance increase, the physical aperture starts eating into the more central region of the distribution resulting in higher losses despite the larger emittance growth for a random excitation.
+In comparison to the $7^{\mathrm{th}}$~turn pulsing, the loss rate for a random excitation is smaller for a maximum excitation amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}).
+This indicates an increase of the diffusion mostly in the beam tails for $7^{\mathrm{th}}$~turn pulsing, while for a random excitation mostly the core is affected.
+Furthermore, the dependence of the loss rate on the excitation plane is additive, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only V (see Appendix~\ref{app:sec:ran}, Fig.~\ref{fig:ranexplossplane}).
+In view of the next section, Sec.~\ref{sec:damp}, not that the loss rate is also considerably reduced for a random excitation if the transverse damper is active, while in case of the $7^{\mathrm{th}}$~turn pulsing the loss rate does not change.
 
 \subsubsection{Effect of the transverse damper\label{sec:damp}}
-For all $k^{\mathrm{th}}$~turn pulsing patterns, the transverse damper does not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation.
-On the other hand, the transverse damper considerably decreases loss rates, emittance growth and distribution changes for the random excitation.
-As an example, the vertical emittance growth is compared for the $7^{\mathrm{th}}$~turn pulsing in V and a random excitation in V in Fig.~\ref{fig:damp} and the loss rates in Fig.~\ref{fig:ranexploss} illustrating impressively the above behavior.
-For nth turn pulsing no effect. Show example of emittance for 7th turn pulsing and random and refer to Fig.~\ref{fig:ranexploss} for the loss rate.
-\begin{figure}[h]
+\begin{figure}[b]
 	\centering
@@
 \end{figure}
+For all $k^{\mathrm{th}}$~turn pulsing patterns, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation.
+On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation.
+As an example, the vertical emittance growth for the $7^{\mathrm{th}}$~turn pulsing in V  is compared with a random excitation in V in Fig.~\ref{fig:damp} and the loss rates in Fig.~\ref{fig:ranexploss} illustrating impressively the aforesaid behavior.
 \textcolor{red}{Any idea why? Daniel, help!}
@@
 \section{Summary\label{sec:sum}}
-We now want to close the loop to the effects on the beam core for hollow electron lenses in HL-LHC.
-For DC operation, no influence on the beam core are are to be expected.
-To study the effect of pulsed operation beam experiments were conducted in 2016 and 2017 at injection.
-The experimental results showed that for the kick amplitudes expected for the e-lens, considerable effects are to be expected.
-The random pulsing mode is the most sensitive pulsing pattern and as it excites all beam frequencies results in emittance growth in the plane of excitation and losses once the beam size increased enough to hit the physical aperture.
-The resonant excitation or nth turn pulsing on the other hand features a rich non-linear dynamics as it excites only certain resonances and diversity in beam distribution and diffusion changes.
-This means, that each pulsing patterns results in different loss rates and emittance growth.
-Typical for this kind of excitation is also the adjustment of the beam distribution to a new equilibrium state.
+This paper intends to give an overview of the effects of the HEL on the proton beam core and the experiments conducted in this context at the LHC in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}.
+For a perfect HEL without any imperfections no effect on the core is to be expected as the field at the center vanishes by construction.
+Only in the case of imperfections in the guidance and profile of the electron beam, a residual kick on the proton beam core can occur.
+In this case, the contribution from the electron lens bends with approximately $0.5$~nrad to first order is negligible to the kick expected in the central region of the HEL of approximatley $20$~nrad to first order caused by profile imperfections in the electron beam assuming HL-LHC parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}).
+
+In DC operation these kicks, even if nonlinear in reality, are negligible compared to the nonlinearities present in the LHC and HL-LHC and are thus of no concern.
+The picture changes if the HEL is operated in pulsed mode, which introduces noise on the halo particles (intended) as well as on the beam core (unintended) in the presence of imperfections of the HEL entailing much more stringent tolerances.
+To study the effect on the beam core in case of the much more dangerous pulsed operation, beam experiments were conducted in 2016 and 2017 at injection.
+The experimental results in general suggest that for the kick amplitudes as expected from the HEL, considerable effects are to be expected for a random excitation and certain resonant excitations, explicitly $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing were tried during the experiments.
+In addition, also one resonant excitation pattern, the $8^{\mathrm{th}}$~turn pulsing, was tested featuring a small effect on the beam core.
+Assuming that this pattern has a large effect on the beam halo, it would be an optimal candidate for a feasible HEL excitation pattern.
+
+The resonant excitation in general features a rich nonlinear dynamics and diversity in beam distribution and diffusion changes as can be seen from the very diverse effects of the different $k^{\mathrm{th}}$~turn pulsing patterns.
+The origin of this diversity is the fact that only certain frequencies and thus resonances are excited.
+Typical for the resonant excitation is also the fast adjustment of the beam distribution to a new equilibrium state which could be for the first time measured in the experiments and predicted in the simulations presented in this paper.
+The random excitation on the other hand excites all beam frequencies equally, leading to the highest increase in halo removal rates for the HEL \cite{hel_halo_hllhc_fitterer} but also the strongest effects on the core, mainly in terms of strong emittance growth as predicted in simulations and also a multitude of earlier publications on the subject of random noise.
+
+In summary, the effects of the HEL on the beam core in case of pulsed operation are not negligible.
+Tolerances in case of a random excitation are extremely stringent, while a resonant excitation appears feasible if the order $k$ of the $k^{\mathrm{th}}$~turn pulsing is tuned to the machine and beam configuration, explicitly tune, tune spread and machine nonlinearities.
 
@@
 \appendix
-\section{FMA analysis for 10th turn pulsing and 2016 optics and tune}
-\label{app:sec:fma:10}
+\section{Additional plots for 10th turn pulsing during the 2016 experiments}
+\label{app:sec:10}
 \begin{figure}[h]
@@
 
-\section{BSRT profiles for $7^{\mathrm{th}}$ and $8^{\mathrm{th}}$~turn pulsing during 2017 experiments}
-\label{app:sec:bsrtprof}
+\newpage
+
+\section{Additional plots for $7^{\mathrm{th}}$~turn pulsing during 2017 experiments}
+\label{app:sec:7}
+\begin{figure}[h]
+	\begin{minipage}[t]{1.0\linewidth}
+	\centering
+	\includegraphics[width=0.8\linewidth]{2017_scale_amp_7h_lbllong.png}
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.8\linewidth]{2017_scale_amp_7v_lbllong.png}
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.8\linewidth]{2017_scale_amp_7hv_lbllong.png}
+	\end{minipage}
+	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in only in H (top), only in V (top) and in H+V during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by horizontal error bars.
+The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
+\end{figure}
 \begin{figure}[h]
@@
 \end{figure}
+
+\newpage
+
+\section{Additional plots for $8^{\mathrm{th}}$~turn pulsing during 2017 experiments}
+\label{app:sec:8}
 \begin{figure}[h]
@@
 
+\newpage
+
+\section{Additional plots for random excitation during 2017 experiments}
+\label{app:sec:ran}
+\begin{figure}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranh_lbllong.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranv_lbllong.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranhv_lbllong.png}
+	\end{minipage}	
+	\caption{\label{fig:ranexplossplane} Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars.
+Note that for $7^{\mathrm{th}}$~turn pulsing the transverse damper does not influence the loss rate, while for a random excitation the loss rate is considerably reduced for the bunches with transverse damper active.}
+\end{figure}
+
 \clearpage

### 2018-01-30 resex.tex
version 2018-01-29
first revised final version

@@
 Only in the case of imperfections in the guidance and profile of the electron beam, a residual kick on the proton beam core can occur.
-In this case, the contribution from the electron lens bends with approximately $0.5$~nrad to first order is negligible to the kick expected in the central region of the HEL of approximatley $20$~nrad to first order caused by profile imperfections in the electron beam assuming HL-LHC parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}).
+In this case and assuming HL-LHC parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}), the contribution from the electron lens bends with approximately $0.5$~nrad to first order is negligible with respect to the kick expected from the central region of the HEL of approximatley $20$~nrad to first order caused by profile imperfections in the electron beam.
 
 In DC operation these kicks, even if nonlinear in reality, are negligible compared to the nonlinearities present in the LHC and HL-LHC and are thus of no concern.
-The picture changes if the HEL is operated in pulsed mode, which introduces noise on the halo particles (intended) as well as on the beam core (unintended) in the presence of imperfections of the HEL entailing much more stringent tolerances.
+The picture changes if the HEL is operated in pulsed mode, in which case noise is introduced on the halo particles (intended) as well as on the beam core (unintended) in the presence of imperfections of the HEL entailing much more stringent tolerances.
+
 To study the effect on the beam core in case of the much more dangerous pulsed operation, beam experiments were conducted in 2016 and 2017 at injection.
-The experimental results in general suggest that for the kick amplitudes as expected from the HEL, considerable effects are to be expected for a random excitation and certain resonant excitations, explicitly $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing were tried during the experiments.
+The experimental results in general suggest that for the kick amplitudes as expected from the HEL, considerable effects are to be expected for a random excitation and certain resonant excitations, explicitly $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing during the conducted experiments.
 In addition, also one resonant excitation pattern, the $8^{\mathrm{th}}$~turn pulsing, was tested featuring a small effect on the beam core.
@@
 
-The resonant excitation in general features a rich nonlinear dynamics and diversity in beam distribution and diffusion changes as can be seen from the very diverse effects of the different $k^{\mathrm{th}}$~turn pulsing patterns.
-The origin of this diversity is the fact that only certain frequencies and thus resonances are excited.
+The resonant excitation in general features a rich nonlinear dynamics leading to a diversity in beam distribution and diffusion changes as can be seen for example from the very diverse effects of the different $k^{\mathrm{th}}$~turn pulsing patterns observed in simulations and experiments.
+The origin of this diversity is due to the fact that only certain frequencies and thus resonances are excited.
 Typical for the resonant excitation is also the fast adjustment of the beam distribution to a new equilibrium state which could be for the first time measured in the experiments and predicted in the simulations presented in this paper.
-The random excitation on the other hand excites all beam frequencies equally, leading to the highest increase in halo removal rates for the HEL \cite{hel_halo_hllhc_fitterer} but also the strongest effects on the core, mainly in terms of strong emittance growth as predicted in simulations and also a multitude of earlier publications on the subject of random noise.
+The random excitation on the other hand excites all beam frequencies equally, leading to the highest increase in halo removal rates for the HEL \cite{hel_halo_hllhc_fitterer} but also the strongest effects on the core, mainly in terms of strong emittance growth as predicted in simulations, experiments and also a multitude of earlier publications on the subject of random noise.
 
@@
 
-\clearpage
+\newpage
 

### 2018-02-06 resex.tex
Changes Stefano, 2018-02-06

@@
 \begin{abstract}
-In case of the High Luminosity Large Hadron Collider (HL-LHC), a considerable amount of energy is stored in the beam tails due to the high beam intensity and, in addition, an overpopulation of the tails compared to a Gaussian distribution.
-To control and clean the tail population, the installation of two hollow electron lenses, one per beam, is foreseen.
-Beside the DC operation of the lenses, also pulsed operation is considered having the advantage to considerably increase the diffusion speed by exerting noise on the halo particles.
-In the ideal case, that is in case of no field at the proton beam core, only the halo particles are excited while leaving the core unperturbed.
+At the High Luminosity Large Hadron Collider (HL-LHC), a considerable amount of energy is expected to be stored in the beam tails due to the high beam intensity and, in addition, an overpopulation of the tails compared to a Gaussian distribution.
+To control and clean the tail population, the installation of two hollow electron lenses, one per beam, is presently under consideration.
+Beside the DC operation of the lenses, also pulsed operation is considered having the advantage to considerably increase the diffusion speed by exerting white or colored noise on the halo particles.
+In an ideal electron beam with azimuthal symmetry and no perturbations at the incoming and outgoing electron beam locations, only the halo particles are excited while leaving the core unperturbed.
 The picture though changes, if a residual field is present also at the location of the beam core exposing also the the particles in the center to noise.
@@
 In case of the LHC explicitly around 5\% of the beam population is stored in the tails above 3.5~$\sigma$ compared to 0.22\% in case of a Gaussian distribution leading to 19~MJ of stored energy in the tails in case of nominal LHC parameters and 34~MJ in case of HL-LHC~\cite{helreview_valentino}.
-
-All of the above reasons lead to the conclusion that a mechanism is needed to deplete the beam tails in a controlled manner \cite{helreview}.
+This leads to the conclusion that a mechanism is needed to deplete the beam tails in a controlled manner \cite{helreview}.
+
 The most obvious idea is to decrease the collimator gaps or scrape the tails with a collimator type device.
-The minimum gap size of the collimators is however limited by the impedance of the machine.
-Small enough gap sizes can therefore not be reached without risking severe beam instabilities.
-Other mechanisms must therefore be found to actively deplete the tails.
-Most promising are methods, which increase the diffusion speed in the region of the halo particles while leaving the beam core unperturbed.
+This approach is however not feasible as it would generate unacceptably large loss spikes due to the high intensity halo population in the LHC.
+Most promising are methods, which increase the diffusion speed in the region of the halo particles resulting in a smooth and continuous removal of the high amplitude tails while leaving the core of the beam unperturbed.
 The diffusing halo particles are then intercepted by the collimation system and removed.
@@
 In view of the need of active halo control for HL-LHC and for future high power accelerators like HE-LHC and FCC-hh~\cite{helhcparam2011,fcc_coll_ipac2017}, different active halo control methods have been studied in recent years~\cite{helreview_bruce}, of which the hollow electron lens (HEL) is considered the superior device for the HL-LHC~\cite{helreview}.
-The beneficial effect of a future HEL for HL-LHC in terms of machine protection would be however abrogated if the beam core was negatively affected leading ultimately to a degradation of the performance due to increased particle losses and emittance growth.
-In this paper, we will concentrate on this aspect.
+The beneficial effect of a future HEL for HL-LHC in terms of machine protection and beam collimation can not be, however, deployed at the expenses of the beam core as latter would ultimately lead to a degradation of the performance due to increased particle losses and emittance growth.
+In this paper, we will concentrate on the assessment in simulations and measurements of possible detrimental effets on the beam core.
 We will summarize possible sources of perturbations of the proton beam core focusing in particular on the case of pulsed operation and the beam experiments at the LHC conducted in this context.
@@
 
-%This includes the detailed analysis of the observed losses, emittance and transverse beam distribution changes.
-To the knowledge of the authors, the observed distribution changes presented in this paper have never been measured before in the context of a resonant excitation.
-A resonant excitation has been previously studied at the Tevatron in the context of the HEL experiments \cite{hel_tevatron_stancari} and the abort gap cleaning used in operation \cite{hel_tevatron_abortgap_zhang}.
-Both studies however only concentrated on the losses and emittance changes without measuring the detailed changes of the distribution itself.
-In addition, the presented experiments also provide scalings of the losses and emittance growth with the excitation amplitude.
-This allows to provide first tolerances for the residual field of the HL-LHC HEL at the beam core and also to compare the obtained results with simulations and thus benchmark the same.
-
 \section{Hollow electron lens for HL-LHC\label{sec:hel}}
 \subsection{General Overview\label{sec:hel:intro}}
-Electron lenses are in general DC or pulsed low energy electron beams, where the electron beam is generated in an electron gun, then guided and confined by strong solenoids and finally dumped onto a collector.
+Electron lenses produce DC or pulsed low energy electron beams, where the electron beam is generated in an electron gun, then guided and confined along a straight section by strong solenoids and finally dumped onto a collector.
 As an example the conceptual design of the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
@@
 	\includegraphics[width=1.0\linewidth]{hel_layout_epbeam}% Here is how to import EPS art
-	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL.}
+	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL. Courtesy of CERN EN-MME group.}
 \end{figure}
@@
 			Length $L$    &  3 & m\\
-			Desired range of scraping positions\footnote{$\sigma$ denotes here the Beam sigma and not the collimation $\sigma$ which assumes a normalized emittance of 3.5~$\mu$m.} & 4-8 &$\sigma_p$\\
+			% 3-8 sigmap with 3.5 um -> 3.55-9.47 for 2.5 um emittance
+			Desired range of scraping positions & 3.5-9.5 &$\sigma_p$\\
 			\colrule
@@
 \subsection{Operation modes and effects on the beam core\label{sec:hel:core}}
-As depicted in Fig.~\ref{fig:hel_field} and in evidence from Eqs.~\ref{eq:field_1}--\ref{eq:field_2}, the field at the beam core vanishes in case of an ideal HEL.
-Effects on the beam core thus only arise from imperfections, where possible sources are the bends of the HEL (see Section~\ref{core:sec:1}), as here the electron beam crosses directly the proton beam, and distortions in the electron beam profile (see Section~\ref{core:sec:2}).
-Both sources result in non-linear kicks~\cite{hel_bends_stancari,hel_model_polynomial_morozov}.
-In DC operation, these non-linear kicks stay in the shadow of the non-linearities otherwise present in the machine.
-Tolerances on imperfections are therefore not particularly stringent and of no concern.
-The picture changes significantly in case of pulsed operation, in which case the electron gun voltage is modulated using a white or colored noise spectrum.
-If the electromagnetic field of the HEL does not vanish at the proton beam core, noise is transferred not only to the halo particles, as intended, but also to the beam core.
-Tolerances on the residual field in this case rapidly become much more stringent than in the case of DC operation.
-Studies of the effect of the HEL on the beam core are therefore focusing on this mode of operation which will also be the main subject of this paper.
-
-The main benefit of pulsed HEL operation is the increase in halo removal rate particularly relevant if a fast depletion of the halo is desired.
-The halo diffusion rate increases in this case as in addition to the non-linear kick, noise is exerted on the halo particles.
-For the HL-LHC two different pulsing patterns are considered for this purpose:
+For the HEL two modes of operation are currently under consideration, the DC mode as standard operation mode and the pulsed mode.
+The main benefit of pulsed HEL operation is the increase in halo removal rates which might become essential as in ideal operation conditions featuring low non-linearities, in particular low chromaticity and octupole current, the DC mode is not effective~\cite{hel_halo_hllhc_fitterer,hl_halo_ipac2017}.
+For this purpose, two different pulsing patterns are thus considered for the HL-LHC:
 \begin{itemize}
@@
 \end{itemize}
-In general, the random excitation mode is the most effective and operationally easiest pulsing mode, but also the most stringent in terms of tolerances on the residual field at the proton beam center as it induces white noise exciting all beam frequencies.
+
+As depicted in Fig.~\ref{fig:hel_field} and in evidence from Eqs.~\ref{eq:field_1}--\ref{eq:field_2}, the field at the beam core vanishes in case of an ideal HEL.
+Effects on the beam core thus only arise from imperfections, where possible sources are the bends of the HEL (see Section~\ref{core:sec:1}), as here the electron beam crosses directly the proton beam, and distortions in the electron beam profile (see Section~\ref{core:sec:2}).
+Both sources result in non-linear kicks~\cite{hel_bends_stancari,hel_model_polynomial_morozov}.
+In DC operation, these non-linear kicks stay in the shadow of the non-linearities otherwise present in the machine if the electron beam is kept at a decent quality.
+Tolerances on imperfections are therefore not particularly stringent and of no concern.
+The picture changes significantly in case of pulsed operation, in which case the electron gun voltage is modulated using a white or colored noise spectrum.
+If the electromagnetic field of the HEL does not vanish at the proton beam core, noise is transferred not only to the halo particles, as intended, but also to the beam core.
+Tolerances on the residual field in this case rapidly become much more stringent than in the case of DC operation.
+Studies of the effect of the HEL on the beam core therefore focus on this mode of operation which will also be the main subject of this paper.
+
+%In comparison, the random excitation mode is the most effective and operationally easiest pulsing mode, but also the most stringent in terms of tolerances on the residual field at the proton beam center as it induces white noise exciting all beam frequencies.
 As the resonant excitation acts on specific resonances, the pulsing pattern has to be chosen according to the machine and beam parameters.
@@
 As reference for the estimate of the dipole component originating from the HEL bends, the symplectic map derived in \cite{hel_bends_stancari} is used.
-In this case the HEL bends are modeled as a cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons deflected in the horizontal plane.
+In this case the HEL bends are modeled as a cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons deflected in the horizontal plane, where the parameters were her still based on the old gun design for the Tevatron electron lens \textcolor{red}{Giulio, is this correct?}.
 In case of an U-shaped HEL, the transverse dipole kicks at entrance and exit add up, while for a S-shape the kicks compensate each other.
@@
 \begin{equation}
-\int_{z_1}^{z_2} E_{x,y} dz= 10 \ \mathrm{kV}
+\int_{z_1}^{z_2} E_{x,y} dz= 10 \ \mathrm{kV}.
 \end{equation}
-given an \SI{1}{A}, \SI{5}{keV} electron beam.
+%given an \SI{1}{A}, \SI{5}{keV} electron beam.
 %For \SI{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
@@
 %\end{equation}
-Neglecting magnetic effects and scaling to HL-LHC and HEL design parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}) yields a kick of:
+Neglecting magnetic effects and scaling to HL-LHC and HEL design parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}) yields a kick of \cite{md_sim_hel_res_ex_fitterer}:
 \begin{equation}
@@
 		\includegraphics[width=0.8\linewidth]{e-field-measured.png}
-	\caption{Current density from measurements of the CHG1b hollow electron gun at the Fermilab electron lens test stand in 2017 \cite{hel_res_field_stancari_2017}.}
+	\caption{Current density from measurements of the CHG1b hollow electron gun at the Fermilab electron lens test stand in 2017 \cite{hel_res_field_stancari_2017}.
+\textcolor{red}{Replace by 2D simulations - Giulio.}}
 	\label{core:fig:0}
@@
 Beam and machine parameters for both experiments are summarized in Table~\ref{tab:md_param}.
-As there is currently no HEL installed in the LHC, the kick on the beam core was emulated by a dipole kick applied with the transverse damper of which more details will be given in Section~\ref{sec:adt}.
 Preceding the beam experiments, simulations were performed to identify the most effective $k^{\mathrm{th}}$~turn pulsing pattern and yield estimates for losses and emittance growth.
@@
 
-During the experiments the losses were measured with the fast beam current transformers (FBCTs) and the emittance and transverse beam profiles with the beam synchrotron radiation telescope (BSRT).
+During the experiments the losses were measured with the fast beam current transformers (FBCTs) and the emittance and transverse beam profiles with the beam synchrotron radiation telescope (BSRT), where all instruments are capable of delivering bunch-by-bunch data.
 The analysis of the BSRT profiles is quite involved and we will therefore focus in this paper only on the direct comparison of the profiles.
@@
 Though originally designed to damp the undesired transverse activities only, the damper is being routinely used for sophisticated beam excitations as well thanks to the state of the art digital signal processing hardware.
-These include among others for example abort and injection gap cleaning, single bunch excitation for tune and linear coupling measurements and special modes of operation for dedicated experiments at the LHC.
+These include among others abort and injection gap cleaning, single bunch excitation for tune and linear coupling measurements and special modes of operation for dedicated experiments at the LHC.
 
 The transverse feedback is in general active during the whole LHC cycle, therefore it is of very high importance, that the noise introduced by the system does not contribute to any emittance growth.
-The typical machine cycle requires a short damping time of ~10-20 turns (high feedback gain) for injection oscillation damping, which is increased during the ramp and stable beams to ~50-100 turns (lower feedback gain).
+The typical machine cycle requires a short damping time of ~10-20 turns (high feedback gain) for injection oscillation damping, which is increased during the ramp and collisions to ~50-100 turns (lower feedback gain).
 It has been demonstrated that noise introduced by the transverse feedback system itself is sufficiently low to introduce any measurable emittance growth \cite{adt_noise_emit_2017}.
 All excitation signals referenced to in this paper, are digitally synthesized in the damper's digital signal processing units and therefore can be considered as ``noise-free''.
-This point is important as then the effect on the beam can be attributed entirely to the applied excitation and any effect due to unwanted residual noise can be excluded.
+This point is important as then the effect on the beam can be attributed entirely to the applied excitation and any effect due to unwanted residual noise can be neglected.
 
@@
 The error in 2016 is much larger as no calibration of the ADT was performed beforehand, which was done in 2017.
-As an example Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor during the 2017 experiments and an excitation with 8th turn pulsing.
+As an example Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor during the 2017 experiments and an excitation with $8^{\mathrm{th}}$ turn pulsing.
 The measured activity is well in line with the expected excitation pattern.
@@
 			optics & \multicolumn{2}{c}{2016 or 2017 injection optics,}\\
-			& \multicolumn{2}{c}{$\beta^*=11$~m}\\
+			& \multicolumn{2}{c}{$\beta_{\mathrm{IP1/5}}^*=11$~m}\\
 			machine  & standard errors & no errors \\
@@
 
-Quantitative predictions of loss rates and emittance growth proved in general challenging as both observables are influenced in addition by the natural noise present in the LHC, originating for example from ground motion and magnetic field ripple.
+Quantitative predictions of loss rates and emittance growth proved in general challenging as both observables are influenced in addition by the natural noise present in the LHC, originating for example from vibrations and magnetic field ripple.
 The observed effect is then the result of the interaction of the external excitation with the natural noise.
@@
 Beside the natural noise, also collective effects like intra-beam scattering and electron cloud influence the time evolution of emittance and losses.
-Both are also not taken into account in the simulations.
-
-\subsubsection{Prediction of the most efficient pulsing pattern\label{sec:pattern}}
-The most efficient pulsing pattern can be predicted by the comparison of the loss rate and emittance growth from the distribution tracking for the different pulsing patterns.
+Both are also not taken into account in the simulations, but the effect was minimized during the experiments by choosing an as small as possible bunch intensity of only $0.7\times10^{11}$ protons per bunch.
+
+\subsubsection{Dependence on the pulsing pattern\label{sec:pattern}}
+The effect of each pulsing pattern in comparison to the other patterns can be identified by the means of the loss rate and emittance growth from the distribution tracking.
 As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pattern.
@@
 The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and a uniform random excitation.
-As the bunch length decreases in accordance with the losses, the losses can be mostly associated to off-momentum particles.
+As the bunch length decreases in accordance with the losses, the losses can be mostly associated to off-momentum particles hitting the transverse aperture.
 Only $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as random excitation exhibit considerable emittance growth.
@@
 
-The experienced or also inclined unexperienced reader might have noticed that the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time, which usually points to an error in the chosen simulation setup.
+The experienced reader might have noticed that the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time, which usually points to an error in the chosen simulation setup.
 This behavior is however typical for any resonant excitation and can be ascribed to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium.
@@
 In the presence of only a resonant excitation, this distribution is stable in simulations with a constant emittance.
-Experimentally however, an increase of the emittance growth with the applied excitation amplitude is observed, see for example Figs.~\ref{fig:10thexp} and ~\ref{fig:7thexp}, not matching the prediction by simulations.
-By adding a random noise component representative for the natural noise present in the LHC, the constant emittance is changed to an excitation amplitude dependent emittance growth as depicted for example in Figs.~\ref{fig:10thsim} and \ref{fig:7thsim}.
+Experimentally however, an increase of the emittance growth with the applied excitation amplitude is observed, see for example Figs.~\ref{fig:10thexp},~\ref{fig:7thexp}, not matching the prediction by simulations.
+By adding a random noise component representative for the natural noise present in the LHC, the constant emittance is changed to an excitation amplitude dependent emittance growth as depicted for example in Figs.~\ref{fig:10thsim},~\ref{fig:7thsim}.
 An excitation with random uniform noise on the other hand results in constant emittance growth without any initial adjustment phase.

### 2018-02-06 resex.tex
change abstract

@@
 To control and clean the tail population, the installation of two hollow electron lenses, one per beam, is presently under consideration.
-Beside the DC operation of the lenses, also pulsed operation is considered having the advantage to considerably increase the diffusion speed by exerting white or colored noise on the halo particles.
-In an ideal electron beam with azimuthal symmetry and no perturbations at the incoming and outgoing electron beam locations, only the halo particles are excited while leaving the core unperturbed.
-The picture though changes, if a residual field is present also at the location of the beam core exposing also the the particles in the center to noise.
-In this paper, we present first estimates of the residual field at the proton beam core expected from the HL-LHC hollow electron lens and summarize the results of the LHC experiments conducted in 2016 and 2017 to study the effects on the beam core in case of pulsed operation.
+Beside the DC operation of the lenses, also pulsed operation is considered having the advantage to increase the diffusion speed by exerting white or colored noise on the halo particles.
+For an ideal electron lens, only the halo particles are excited while the core stays untouched.
+The picture though changes, if a residual field is present also at the location of the beam core exposing also the particles in the center to noise.
+In this paper, we present
+a summary of the LHC experiments conducted in 2016 and 2017 to study the effects on the beam core in case of pulsed operation, where uniform random noise and a resonant excitation were studied.
+In particular the resonant excitation features rich non-linear dynamics inducing unusual changes of the beam distribution and with it emittance and losses.
+These changes will be one of the main topics of this paper as they have not been observed and studied before in this detail.
+In addition to the presentation of the experimental and supportive simulation results, we also give first estimates of the residual field at the proton beam core expected from the HL-LHC hollow electron lens for comparison to the noise tolerances obtained in the experiments.
 \end{abstract}

### 2018-02-12 resex.tex
add 2016 filling scheme

@@
 		\centering
-		2016 experiment \textcolor{red}{(update picture)}\\
-		\includegraphics[width=0.7\linewidth]{bunchfilling.png}	
+		2016 experiment\\
+		\includegraphics[width=0.7\linewidth]{bunchfilling_2016.png}	
 	\end{minipage}
@@
 		2017 experiment\\
-		\includegraphics[width=0.7\linewidth]{bunchfilling.png}	
-	\end{minipage}
-	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).}
+		\includegraphics[width=0.7\linewidth]{bunchfilling_2017.png}	
+	\end{minipage}
+	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).
+In 2016 the same excitation amplitude was applied on a group of 4 bunch while in 2016 the number was increased to 6 bunches.
+In 2016 the excitation was also only applied in one plane leading to $2\times24=48$ bunches in total in 2016 and $3\times2\times36=216$ bunches in 2017.}
 \end{figure*}

### 2018-02-13 resex.tex
Giulios comments until section IV - LHC experiments ...

@@
 
-\title{Effects of pulsed hollow electron lens operation on the beam core in HL-LHC: \\First experimental studies and simulations.}% Force line breaks with \\
+%\title{Effects of pulsed hollow electron lens operation on the beam core in HL-LHC: \\First experimental studies and simulations.}% Force line breaks with \\
+\title{Effect of resonant and random excitation on the LHC beam in the context of pulsed hollow electron lens operation.}% Force line breaks with \\
 \thanks{Fermilab is operated by Fermi Research Alliance, LLC under
@@
 For an ideal electron lens, only the halo particles are excited while the core stays untouched.
-The picture though changes, if a residual field is present also at the location of the beam core exposing also the particles in the center to noise.
+If a residual field is present at the location of the beam core, also the particles in the center are being exposed to noise.
 In this paper, we present
 a summary of the LHC experiments conducted in 2016 and 2017 to study the effects on the beam core in case of pulsed operation, where uniform random noise and a resonant excitation were studied.
+In both cases the excitation was generated with the transverse damper, which is capable of producing almost arbitrary colored noise spectra.
 In particular the resonant excitation features rich non-linear dynamics inducing unusual changes of the beam distribution and with it emittance and losses.
@@
 \begin{equation}\label{hel_kick_max}
-\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{q}{p_0}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2},
+\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{p_0}{q}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2},
 \end{equation}
@@
 			\colrule
-			\multicolumn{4}{l}{Optics paramters at HEL (Beam~1) \footnote{As the twiss parameters at IP4 do not change during the entire squeeze, and IP4 and the HEL are only separated by a drift space, the twiss parameters stay constant also at the HEL during the entire squeeze.}} \\
+			\multicolumn{4}{l}{Optics paramters at HEL (Beam~1) \footnote{As the Twiss parameters at IP4 do not change during the entire squeeze, and IP4 and the HEL are only separated by a drift space, the Twiss parameters stay constant also at the HEL during the entire squeeze.}} \\
 			\colrule
@@
 	\end{eqnarray}
-	with $U_{\mathrm{max}}$ the equivalent voltage in DC operation, $a$ the modulation strength with $a\in[0,1]$, and $\mathrm{ran}(0,1)$ a uniformly distributed random number between~[0,1].
+	with $U_{\mathrm{max}}$ the maximum voltage in DC operation, $a$ the modulation strength with $a\in[0,1]$, and $\mathrm{ran}(0,1)$ a uniformly distributed random number between~[0,1].
 	\item \textbf{resonant excitation:} The HEL is switched on only every $k^{\mathrm{th}}$ turn. The excitation can then be represented by:
 	\begin{eqnarray}\label{intro:eqn:1}
-	f(t)&=&\sum_{p=-\infty}^{+\infty}\delta(t-n\cdot(kT)),
+	f(t)&=&\sum_{n=-\infty}^{+\infty}\delta(t-n\cdot(kT)),
 	\end{eqnarray}
-	where $p$ is the turn number and $T$ the revolution time, and its Fourier series by:
+	where $n$ is the turn number and $T$ the revolution time, and its Fourier series by:
 	\begin{eqnarray}\label{intro:eqn:2}
@@
 During the experiments presented in this paper, the dipole kick with the corresponding excitation pattern could then be applied with the LHC transverse damper as will be described in detail in Section~\ref{sec:adt}.
-Leaping ahead of the more detailed derivations in Sec.~\ref{core:sec:1} and \ref{core:sec:2}, the expected kick to first order from the HEL bends is estimated to be~(Eq.~\ref{eqn:kick_bends}):
+As will be shown in Sec.~\ref{core:sec:1} and \ref{core:sec:2}, the expected kick to first order from the HEL bends is estimated to be~(Eq.~\ref{eqn:kick_bends}):
 \begin{equation}
@@
 \begin{equation}
-	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \SI{19.6}{nrad}
+	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \SI{19.6}{nrad} % this should be rounded to the same number of digits as the 0.5nrad
 \end{equation}
@@
 As reference for the estimate of the dipole component originating from the HEL bends, the symplectic map derived in \cite{hel_bends_stancari} is used.
-In this case the HEL bends are modeled as a cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons deflected in the horizontal plane, where the parameters were her still based on the old gun design for the Tevatron electron lens \textcolor{red}{Giulio, is this correct?}.
+In this case the HEL bends are modeled as a bent cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons deflected in the horizontal plane, where the parameters were her still based on the old gun design for the Tevatron electron lens \textcolor{red}{Giulio, is this correct?}.
 In case of an U-shaped HEL, the transverse dipole kicks at entrance and exit add up, while for a S-shape the kicks compensate each other.
@@
 \begin{equation}
-\int_{z_1}^{z_2} E_{x} dz= 36 \ \mathrm{kV} \Rightarrow \Delta x'= \q{5.1}{nrad}
+\int_{z_1}^{z_2} E_{y} dz= 36 \ \mathrm{kV} \Rightarrow \Delta y'= \q{5.1}{nrad}
 \end{equation}
@@
 
-% insert table here to put it on top of page
+%\begin{table*}
+%	\caption{\label{tab:md_param}%
+%		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and 2017 \cite{resexmd2016,resexmd2017}.
+The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V horizontal and vertical at the same time.
+%	}
+%	\begin{ruledtabular}
+%		\begin{tabular}{lcc}
+%			Parameter & Experiment 2017 & Experiment 2016  \\
+%			\colrule
+%			beam &\multicolumn{2}{c}{Beam~1} \\
+%			beam energy &\multicolumn{2}{c}{injection energy, 450 GeV} \\\hline
+%			single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
+%			normalized emittance &\multicolumn{2}{c}{$2.5-3.5~\mathrm{\mu m}$} \\
+%			$4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
+%			$1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
+%			number of bunches & $3\times72=216$ bunches & $12\times4=48$ single bunches\\
+%			& (+ 1 pilot + 12 nominal) & \\\hline
+%			injection optics, $\beta^*=11$~m & standard optics 2017 & standard optics 2016\\
+%			Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=\pm19.6~\mathrm{A}$, explicitly +19.6 A for MOF circuit and}\\
+%			& \multicolumn{2}{c}{-19.6~A for MOD circuit (standard 2016 settings)}\\\hline
+%			working point $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\
+%			chromaticity $(Q'_x,Q'_y)$ & \multicolumn{2}{c}{(+15,+15)}\\\hline
+%			pulsing patterns  & $7^{\mathrm{th}}$~turn H,V,H+V;& $7^{\mathrm{th}}$~turn H;\\
+%			& $8^{\mathrm{th}}$~turn H,V,H+V;& $10^{\mathrm{th}}$~turn V;\\
+%			& random  H,V,H+V;& \\
+%		\end{tabular}
+%	\end{ruledtabular}
+%\end{table*}
+
 \begin{table*}
 	\caption{\label{tab:md_param}%
-		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and 2017 \cite{resexmd2016,resexmd2017}.
+		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and~2017~\cite{resexmd2016,resexmd2017}.
 The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V horizontal and vertical at the same time.
@@
 		\begin{tabular}{lcc}
-			Parameter & Experiment 2017 & Experiment 2016  \\
+			Parameter & Experiment 2016 & Experiment 2017  \\
 			\colrule
@@
 			$1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
-			number of bunches & $3\times72=216$ bunches & $12\times4=48$ single bunches\\
-			& (+ 1 pilot + 12 nominal) & \\\hline
-			injection optics, $\beta^*=11$~m & standard optics 2017 & standard optics 2016\\
-			Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=\pm19.6~\mathrm{A}$, explicitly +19.6 A for MOF circuit and}\\
-			& \multicolumn{2}{c}{-19.6~A for MOD circuit (standard 2016 settings)}\\\hline
-			working point $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\
+			number of bunches & $12\times4=48$ single bunches & $3\times72=216$ bunches\\
+			&  & (+ 1 pilot + 12 nominal) \\\hline
+			injection optics, $\beta^*=11$~m & standard optics 2016 & standard optics 2017\\
+			Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=+19.6~\mathrm{A}$ for MOF circuit and}\\
+			& \multicolumn{2}{c}{$I_{\mathrm{MO}}=-19.6~\mathrm{A}$ for MOD circuit (standard 2016 settings)}\\\hline
+			working point $(Q_x,Q_y)$ & (64.28,59.31) & (62.27,60.295) \\
 			chromaticity $(Q'_x,Q'_y)$ & \multicolumn{2}{c}{(+15,+15)}\\\hline
-			pulsing patterns  & $7^{\mathrm{th}}$~turn H,V,H+V;& $7^{\mathrm{th}}$~turn H;\\
-			& $8^{\mathrm{th}}$~turn H,V,H+V;& $10^{\mathrm{th}}$~turn V;\\
-			& random  H,V,H+V;& \\
+			pulsing patterns  &$7^{\mathrm{th}}$~turn H; &$7^{\mathrm{th}}$~turn H,V,H+V; \\
+			& $10^{\mathrm{th}}$~turn V; & $8^{\mathrm{th}}$~turn H,V,H+V; \\
+			& &  random  H,V,H+V;\\
 		\end{tabular}
@@
 \end{table*}
-% insert table here to put it on top of page
 
@@
 \label{core:sec:2}
-For a perfectly uniform, annular and radially symmetric profile, the field in the region of the proton beam core vanishes as can be seen from Eq.~\ref{eq:field_2} and is illustrated in Fig.~\ref{fig:hel_field}.
-Fields at the location of the proton beam core thus only arise in case of profile imperfections of the electron beam, in particular if the radial symmetry is broken.
-Recently, new measurements of the newest hollow electron gun CHG1b could be performed at the Fermilab electron lens test stand \cite{hel_test_stand_fnal}.
-A typical measurement of the electron beam current density obtained during these measurements is shown in Fig.~\ref{core:fig:0}.
+For a perfectly annular and axially symmetric profile, the field in the region of the proton beam core vanishes as can be seen from Eq.~\ref{eq:field_2} and as illustrated in Fig.~\ref{fig:hel_field}.
+Fields at the location of the proton beam core thus only arise in case of profile imperfections of the electron beam, in particular if the axial symmetry is broken.
+Recently, the latest LHC hollow electron gun prototype was characterized at the Fermilab electron lens test stand \cite{hel_test_stand_fnal}.
+A typical measurement of the electron beam current density is shown in Fig.~\ref{core:fig:0}.
 \begin{figure}[t]
@@
 For this purpose, the measured distribution is compressed to the inner electron beam radius of $0.97$~mm.
-Afterwards an input distribution with $40\ 000$ particles generated.
+Afterwards an input distribution with $40\ 000$ particles is generated.
+As boundary condition, the LHC inner beam pipe radius $b=30$~mm is used.
 The potential and fields can then be calculated with the particle in cell simulation code WARP~\cite{warp} yielding the relative electric field strengths shown in Fig.~\ref{core:fig:1}.
@@
 	\end{minipage}
-	\caption{Calculated relative electric field for the new CHG1b hollow electron gun in the x-y plane and as projection on the x- and y-axis.
-The field calculations are based on measurements at the Fermilab test stand combined with WARP simulations for the calculation of the electric potentials and fields.}
+	\caption{Calculated relative electric field for the new CHG1b hollow electron gun in the x-y plane (left) and as 1-dimensional cuts though the x- and y-axes (right).
+The field calculations are based on measurements at the Fermilab test stand combined with WARP  calculations of the electric potentials and fields in a cylindrical beam pipe.}
 	\label{core:fig:1}
@@
 
-Using these measurements and simulations of the relative electric field strength along the x- and y-axis (Fig.~\ref{core:fig:1}), we obtain
+Using these measurements and calculations of the relative electric field strengths in the transverse plane (Fig.~\ref{core:fig:1}), we obtain an average field in the hole of
 \begin{equation}
-	E_{\mathrm{rel}}\approx 5\%
+	E_{\mathrm{rel}}= 5\%
 \end{equation}
@@
 \begin{equation}\label{eqn:kick_central}
-\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \SI{19.6}{nrad}
+\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \SI{19.6}{nrad}
 \end{equation}

### 2018-02-13 resex.tex
split section IV in setup and results

@@
 
-\section{LHC experiments on the effect of a resonant and random excitation on the beam core\label{sec:exp}}
+\section{Experimental and simulation setup\label{sec:exp}}
 %put picture on top
@@
 
-\subsection{Simulation and experimental results\label{sec:simex}}
+\section{Experimental and simulation results\label{sec:simex}}
 In the following we will jump forth and back between both experimental and simulation results as the experimental results are best understood on the base of the simulations.
@@
 
-\subsubsection{Dependence on the pulsing pattern\label{sec:pattern}}
+\subsection{Dependence on the pulsing pattern\label{sec:pattern}}
 The effect of each pulsing pattern in comparison to the other patterns can be identified by the means of the loss rate and emittance growth from the distribution tracking.
@@
 
-\subsubsection{$10^{\mathrm{th}}$ turn pulsing\label{sec:simex10}}
+\subsection{$10^{\mathrm{th}}$ turn pulsing\label{sec:simex10}}
 The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
@@
 
-\subsubsection{$7^{\mathrm{th}}$ turn pulsing\label{sec:simex7}}
+\subsection{$7^{\mathrm{th}}$ turn pulsing\label{sec:simex7}}
 The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the experimental results.
@@
 
-\subsubsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
+\subsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
 Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they in addition have a large effect on the halo, which is in general feasible considering the highly nonlinear field generated at the location of the halo particles by the HEL.
@@
 
-\subsubsection{random excitation\label{sec:simexran}}
+\subsection{random excitation\label{sec:simexran}}
 The random excitation presents in general the pulsing pattern leading to the fastest halo depletion rates achievable with an HEL.
@@
 
-\subsubsection{Effect of the transverse damper\label{sec:damp}}
+\subsection{Effect of the transverse damper\label{sec:damp}}
 \begin{figure}[b]

### 2018-02-14 resex.tex
comments from Giulio up to 7th turn pulsing section

@@
 		2016 experiment\\
-		\includegraphics[width=0.7\linewidth]{bunchfilling_2016.png}	
+		\includegraphics[width=0.9\linewidth]{bunchfilling_2016.png}	
 	\end{minipage}
@@
 		2017 experiment\\
-		\includegraphics[width=0.7\linewidth]{bunchfilling_2017.png}	
+		\includegraphics[width=0.9\linewidth]{bunchfilling_2017.png}	
 	\end{minipage}
@@
 
-The transverse feedback is in general active during the whole LHC cycle, therefore it is of very high importance, that the noise introduced by the system does not contribute to any emittance growth.
+The transverse feedback is in general active during the whole LHC cycle, therefore it is critical that the noise introduced by the system does not contribute to any emittance growth.
 The typical machine cycle requires a short damping time of ~10-20 turns (high feedback gain) for injection oscillation damping, which is increased during the ramp and collisions to ~50-100 turns (lower feedback gain).
-It has been demonstrated that noise introduced by the transverse feedback system itself is sufficiently low to introduce any measurable emittance growth \cite{adt_noise_emit_2017}.
-All excitation signals referenced to in this paper, are digitally synthesized in the damper's digital signal processing units and therefore can be considered as ``noise-free''.
+It has been demonstrated that noise introduced by the transverse feedback system itself is sufficiently low not to introduce any measurable emittance growth \cite{adt_noise_emit_2017}.
+All excitation signals referenced to in this paper are digitally synthesized in the damper's digital signal processing units and therefore can be considered as ``noise-free''.
 This point is important as then the effect on the beam can be attributed entirely to the applied excitation and any effect due to unwanted residual noise can be neglected.
@@
 The resonant excitation experiment in 2017 involved simultaneous measurements on three groups of 72 bunches with a dedicated excitation pattern and transverse feedback configuration on each sub-group of 6 bunches.
-In 2016 a similar configuration was used, but with only 48 bunches in total.
+In 2016 a similar configuration was used, but with only 48 bunches in total and sub-groups of 4 bunches.
 Both schemes are illustrated in Fig.~\ref{fig:fill}.
+In order to reduce the error of the measurements, the observables are averaged over each sub-group of bunches experiencing the same excitation amplitude, yielding in total 5~different amplitudes $n\Delta A$ with $n=1,\cdots, 5$ plus the references bunch without excitation for each transverse feedback configuration and excitation plane.
+Furthermore, the experiments were usually split in 3 different parts:
+\begin{description}
+	\item[Part 1] No excitation is applied in order to allow the distribution to reach an equilibrium state after injection.
+	\item[Part 2] The excitation is applied with a first maximum excitation amplitude of $A_{\mathrm{max,1}}=5\Delta A_1$.
+	\item[Part 3] The maximum excitation amplitude is further increased to $A_{\mathrm{max,1}} < A_{\mathrm{max,2}}=5\Delta A_2$.
+\end{description}
+Each part lasted approximately 10-15~minutes, which is considered to be a long enough time window to allow the beam distribution to reach its new equilibrium state.
+In the plots of the experimental results shown later in the paper, the maximum excitation amplitude is indicated with black horizontal arrows, while the average of each sub-group of bunches with the same excitation amplitude $n\Delta A$ is distinguished with different colors.
 
@@
 The error in 2016 is much larger as no calibration of the ADT was performed beforehand, which was done in 2017.
+The maximum kick strength reachable with the ADT at injection without risking saturation is 96~nrad.
 As an example Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor during the 2017 experiments and an excitation with $8^{\mathrm{th}}$ turn pulsing.
@@
 The simulation parameters are summarized in Table~\ref{tab:sim_param}.
-Further details on the simulations are given in \cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
+Further details on the simulations are given in Refs.~\cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
 \begin{table}[h]
@@
 In summary, the most efficient pulsing patterns could be successfully predicted in simulations, which we will demonstrate on the example of the 2016 experiments (see Sec.~\ref{sec:pattern}).
-Afterwards we will present the simulation and experimental results of the tested excitation patterns, explicitly the $10^{\mathrm{th}}$~turn, $7^{\mathrm{th}}$~turn and $8^{\mathrm{th}}$~turn pulsing in Secs.~\ref{sec:simex10}--\ref{sec:simex8}) as well as the random excitation in Sec.~\ref{sec:simexran}.
-As a last chapter (Sec.~\ref{sec:damp}), we will address the influence of the transverse damper in the presence of an external excitation source on emittance, losses and beam distribution.
+Afterwards we will present the simulation and experimental results of the tested excitation patterns, explicitly the $10^{\mathrm{th}}$~turn, $7^{\mathrm{th}}$~turn and $8^{\mathrm{th}}$~turn pulsing in Secs.~\ref{sec:simex10}--\ref{sec:simex8} as well as the random excitation in Sec.~\ref{sec:simexran}.
+As a last chapter, Sec.~\ref{sec:damp}, we will address the influence of the transverse damper in the presence of an external excitation source on emittance, losses and beam distribution.
 
@@
 Beside the natural noise, also collective effects like intra-beam scattering and electron cloud influence the time evolution of emittance and losses.
-Both are also not taken into account in the simulations, but the effect was minimized during the experiments by choosing an as small as possible bunch intensity of only $0.7\times10^{11}$ protons per bunch.
+Both are also not taken into account in these simulations, but the effect was minimized during the experiments by choosing an as small as possible bunch intensity of only $0.7\times10^{11}$ protons per bunch.
 
@@
 As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pattern.
-\begin{figure}[h]
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance for different pulsing patterns obtained from distribution tracking based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
-The resonant excitation is applied in both planes with an amplitude of 96~nrad.
-No random noise component is added.}
-\end{figure}
+The resonant and random excitation respectively is applied in both planes with an amplitude of 96~nrad.
+No random noise component is added in addition.}
+\end{figure*}
 
@@
 
-The experienced reader might have noticed that the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time, which usually points to an error in the chosen simulation setup.
+One may have noticed that the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time, which usually points to an error in the chosen simulation setup.
 This behavior is however typical for any resonant excitation and can be ascribed to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium.
@@
 This is illustrated by means of the residual in respect to the Gaussian distribution for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in Fig.~\ref{fig:patternhist}.
-This initial fast change of the beam distribution is not only an artifact of the simulation, but has been, to the knowledge of the authors, also for the first time experimentally observed during the 2016 and 2017 LHC experiments presented in this paper.
-The timescale of this fast adjustment is in experiments however much slower than predicted in simulations - on the timescale of minutes compared to seconds.
+This initial fast change of the beam distribution is not an artifact of the simulation, but has been, to the knowledge of the authors, also for the first time experimentally observed during the 2016 and 2017 LHC experiments presented in this paper.
+The timescale of this fast adjustment is in experiments however much slower than predicted in simulations --- on the timescale of minutes compared to seconds.
 \begin{figure}[t]
@@
 The excitation is applied in both planes with an amplitude of 96~nrad.
-The residual is defined as the final distribution (here $10^4$~turns) minus the initial distribution (here $10^2$~turns), and the ratio as the final distribution divided by the initial distribution.
+The residual is defined as the final distribution (here after $10^4$~turns) minus the initial distribution (here after $10^2$~turns) as \%, and the ratio as the final distribution divided by the initial distribution.
 The red line in the plot of the residual is the difference between the Gaussian fit of the distribution and the distribution itself.}
@@
 
-In the presence of only a resonant excitation, this distribution is stable in simulations with a constant emittance.
-Experimentally however, an increase of the emittance growth with the applied excitation amplitude is observed, see for example Figs.~\ref{fig:10thexp},~\ref{fig:7thexp}, not matching the prediction by simulations.
-By adding a random noise component representative for the natural noise present in the LHC, the constant emittance is changed to an excitation amplitude dependent emittance growth as depicted for example in Figs.~\ref{fig:10thsim},~\ref{fig:7thsim}.
-An excitation with random uniform noise on the other hand results in constant emittance growth without any initial adjustment phase.
+In the presence of only a resonant excitation, this modified distribution is stable in simulations with a constant emittance.
+As we will see later, an increase of the emittance with the applied excitation amplitude is however observed experimentally not matching the prediction by simulations.
+By adding a random noise component representative for the natural noise present in the LHC, the constant emittance is changed in simulations to an excitation amplitude dependent emittance growth after an initial adjustment phase.
+In case of random uniform noise, the excitation leads to constant emittance growth without any initial adjustment phase.
 The interaction with the natural noise source then only results in an underestimation of loss rates and emittance growth as it is basically equivalent to the application of random noise with an increased amplitude.
@@
 
+\clearpage
+
 \subsection{$10^{\mathrm{th}}$ turn pulsing\label{sec:simex10}}
-The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
+The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an excitation in the vertical plane only.
+As this was the first time a resonant excitation was tested in the LHC, only in total 48~bunches could be injected in order to still guarantee the protection of the machine.
+To test different excitation amplitudes during one fill and have enough statistics, the excitation was therefore only applied in one plane.
+As described earlier, the experiment started with a period of 12~minutes without excitation in order to let the beam distribution fully adjust to its equilibrium state after injection.
+Afterwards the excitation was applied with the excitation scheme shown in Fig.~\ref{fig:fill} and a maximum excitation amplitude of $A_{\mathrm{max}}=5\Delta A = 48~$nrad for a total of 11~minutes.
+The excitation was then further increased to the maximum value reachable without saturation of $A_{\mathrm{max}}=5\Delta A = 96~$nrad and kept for 11~minutes.
+In both cases, the excitation induced the following changes:
 \begin{enumerate}
@@
 As loss rates, emittance growth and beam distribution stayed unchanged for the $4$~reference bunches, the above observations can be also truly associated with the effect of the $10^{\mathrm{th}}$~turn pulsing.
-\begin{figure}[h]
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.49\linewidth]{2016_bunch_intensity_v10th_no_damper_avg.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016_emith_avg_rel_v10th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016_emitv_avg_rel_v10th_no_damper.png}
+\begin{figure*}[h]
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_v10th_no_damper_avg.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_v10th_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_no_damper.png}
 	\end{minipage}	
@@
 Relative refers here to the normalization to the initial value.
-For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Fig.~\ref{fig:fill}).
-The different colored solid lines labeled with $n\Delta A$ indicate the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation indicated as envelope.}
-\end{figure}
-
-Besides the determination of the scaling of the loss rate and emittance growth with the excitation amplitude important for the actual specification of the hollow electron lens, the vertical emittance actually features the strange behavior predicted in simulations: a fast adjustment phase of the beam distribution visible as a rapid increase of the emittance followed by the reach of a new equilibrium distribution apparent by a slower continuous emittance growth.
+For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad indicated with black arrows, the excitation amplitude was linearly increased for each group of 4~bunches (see Fig.~\ref{fig:fill}).
+The different colored solid lines labeled with $n\Delta A$ are the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation shown as envelope.
+The area with a blue background indicates the fast adjustment period of the beam distribution transitioning into a new equilibrium state highlighted with a gray background.}
+\end{figure*}
+These experimentally obtained scalings of the loss rate and emittance growth with the excitation amplitude can be then used for the actual specification of the hollow electron lens.
+
+Besides the numerical determination of tolerances, the vertical emittance actually features indeed the strange and interesting behavior predicted in simulations: a fast adjustment phase of the beam distribution visible as a rapid increase of the emittance followed by the reach of a new equilibrium distribution apparent by a slower continuous emittance growth.
 These two phases are indicated in Fig.~\ref{fig:10thexp} in blue and black.
@@
 Thanks to the good performance of the Beam Synchrotron Radiation Telescope (BSRT)~\cite{bsrtprofinj}, the distribution changes indirectly observed as changes of the emittance could also be for the first time directly measured.
-As example Fig.~\ref{fig:10thexpprof} shows the vertical profile of a reference bunch and one bunch experiencing the maximum excitation of $5\cdot\Delta A=A_{\mathrm{max}}$.
+As example Fig.~\ref{fig:10thexpprof} shows the vertical profile of a reference bunch and one bunch experiencing the maximum excitation of $5\cdot\Delta A=A_{\mathrm{max}}=48$~nrad and afterwards 96~nrad.
 The distribution of the excited bunch not only changes but becomes clearly non-Gaussian, while the reference bunch stays unchanged.
-\begin{figure}[h]
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 The solid line indicates an excitation with only $10^{\mathrm{th}}$~turn pulsing in V and the dotted line with $10^{\mathrm{th}}$~turn pulsing in V plus a random dipole noise component in H+V of 6~nrad.}
-\end{figure}
+\end{figure*}
 The simulation results from the distribution tracking for $10^{\mathrm{th}}$~turn pulsing and without (solid line) and with (dotted line) additional random noise component to emulate the natural noise present in the LHC are shown in Fig.~\ref{fig:10thsim}.
@@
 \begin{equation}
-\theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = 6~\mathrm{nrad},
+\theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = 6~\mathrm{nrad}.
 \end{equation}
-As can be seen from simulations, adding a random noise component not only yields the correct qualitative behavior of the emittance but also increases the loss rates.
-As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment can be thus further reduced.
-A large difference however still prevails which is illustrated in Fig.~\ref{fig:10thexploss} by means of a qualitative comparison of the loss rates obtained in experiments and simulations.
-\begin{figure}[h]
-		\centering
-		\includegraphics[width=0.8\linewidth]{2016_scale_amp_10v_ran_lbllong_sim.png}
-	\caption{\label{fig:10thexploss} Comparison of scaling of loss rates with excitation amplitude for $10^{\mathrm{th}}$~turn pulsing.
-The experimental results for a maximum excitation with 48~nrad are shown in black and for 96~nrad in blue.
-The excitation amplitude error is 50\% and the error on the loss rate is the pure statistical error.
-The simulation results for 2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad are shown in red.}
-\end{figure}
+As can be seen from simulations, adding a random noise component not only yields the correct qualitative behavior of the emittance but also increases the losses.
+As the losses are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment can be thus reduced.
+To better compare the experimental results for the two different maximum excitation amplitudes applied for slightly different time intervals and also the simulations conducted for a much shorter interval of only 90~s, we define the relative loss rate $R$:
+\begin{equation}\label{eqn:lossrate}
+R = \frac{I_{\mathrm{start}}-I_{\mathrm{end}}}{I_{\mathrm{start}}\cdot \Delta t}
+\end{equation}
+where $I$ is the intensity and $\Delta t$ the time interval during which the excitation with fixed $A_{\mathrm{max}}$ is applied during the experiments or the duration of the simulation respectively.
+A qualitative comparison of the loss rates obtained in experiments and simulations is shown in Fig.~\ref{fig:10thexploss} featuring a surprisingly good agreement.
+It should be also kept in mind that the additional random noise component considerably influences the simulation results and as long as it is unknown should be seen more as a fit parameter than a real experimentally confirmed input parameter.
+The measured loss rates for $A_{\mathrm{max}}=48$~nrad and $A_{\mathrm{max}}=96$~nrad should furthermore follow the same scaling.
+Instead the loss rates for the same excitation amplitude are smaller for $A_{\mathrm{max}}=96$~nrad compared to $A_{\mathrm{max}}=48$~nrad.
+That losses and also emittance growth for a fixed excitation amplitude do not exactly agree for both experiments, is a general tendency observed in all experiments.
+As the second excitation however starts from an already by the previous excitation changed beam distribution, the difference is however not surprising and can be explained by this difference in initial conditions.
+For checking the reproducibility of the results in a possible future experiment, it would be interesting to see if the same loss rates are measured if for each $A_{\mathrm{max}}$ a new fill is used.
+\begin{figure}[h]
+		\centering
+%		\includegraphics[width=0.8\linewidth]{2016_scale_amp_10v_ran_lbllong_sim.png}
+\includegraphics[width=1.0\linewidth]{2016_scale_amp_10v_ran_lblshort_sim_prstab.png}
+	\caption{\label{fig:10thexploss} Comparison of scaling of loss rates $R$ as defined in Eqn.~\ref{eqn:lossrate} with the excitation amplitude for $10^{\mathrm{th}}$~turn pulsing in~V.
+The experimental results for a maximum excitation of $A_{\mathrm{max}}=48$~nrad are shown in black and for $A_{\mathrm{max}}=96$~nrad in green.
+The excitation amplitude error is 50\% and the error on the loss rate is the pure statistical error on the mean.
+The simulation results (see Fig.~\ref{fig:10thsim}) for 2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad are shown in red together with their statistical error.
+The loss rates follow approximately a quadratic behavior illustrated by a second order polynomial fit to the data shown as solid (for bunches with transverse damper off) or dashed (for bunches with transverse damper on) line.}
+\end{figure}
+
+\clearpage
 

### 2018-02-14 scripts/lifetrac_sim_results/get_loss_scaling_10V.py

@@ -1,0 +1,1 @@
+/home/mfittere/work/MDs/mdelens/MD23_08_2016/get_loss_scaling_10V.py

### 2018-02-14 scripts/lifetrac_sim_results/plot_loss_emit_10V_2016.py

@@ -95,7 +95,7 @@ for phv,p12 in zip(['hor.','vert.'],'12'):
     plt.plot(t.data['time'],t.data['emit%s'%p12]/t.data['emit%s'%p12][0],linestyle='-',color=camp,label=pltlbl)
     plt.plot(tran.data['time'],tran.data['emit%s'%p12]/tran.data['emit%s'%p12][0],linestyle=':',color=camp,label=pltlbl)
   plt.xlabel('time [s]')
-  plt.ylabel('rel. %s normalized emittance [$\mu$m]'%phv)
+  plt.ylabel('rel. %s normalized emittance'%phv)
 #  plt.grid(b=True)
 
 # now do the legends for all plots

### 2018-02-14 scripts/lifetrac_sim_results/plot_loss_emit_pulsing.py

@@ -22,8 +22,8 @@ opt0 = "2016injnocolc15o+19_6erra2b2u/1/eps3.5/p1e4t1e4s100weight"
 # with pulsing
 optsk = "2016injnocolc15o+19_6erra2b2ut%ssk/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
 # random
-optran8 = "2016injnocolc15o+19_6erra2b2uran/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
-optran9 = "2016injnocolc15o+19_6erra2b2uran/1/eps3.5/hv9_6e-9/p1e4t1e4s100weight"
+optran = "2016injnocolc15o+19_6erra2b2uran/1/eps3.5/hv9_6e-8/p1e4t1e4s100weight"
+#optran = "2016injnocolc15o+19_6erra2b2uran/1/eps3.5/hv9_6e-9/p1e4t1e4s100weight"
 
 # no excitation
 t0 = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,opt0),plt_dir=plt_dir)
@@ -46,7 +46,7 @@ plt.xlabel('time [s]')
 plt.ylabel('relative intensity')
 ax1=plt.gca()
 # second axis for random
-t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optran8),plt_dir=plt_dir)
+t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optran),plt_dir=plt_dir)
 ax2 = plt.gca().twinx()
 ax2.plot(t.data['time'],t.data['intensity'],linestyle='-',color='grey',label='random')
 ax2.set_ylabel('relative intensity random',color='grey')
@@ -74,7 +74,7 @@ plt.xlabel('time [s]')
 plt.ylabel('bunch length [cm]')
 ax1 = plt.gca()
 # second axis for random
-t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optran8),plt_dir=plt_dir)
+t = LifeDeskDB.getdata(ltr_dir=os.path.join(basedir,optran),plt_dir=plt_dir)
 ax2 = plt.gca().twinx()
 ax2.plot(t.data['time'],t.data['sigm'],linestyle='-',color='grey',label='random')
 ax2.set_ylabel('bunch length random [cm]',color='grey')
@@ -103,7 +103,7 @@ for phv,p12 in zip(['hor.','vert.'],'12'):
 #  plt.grid(b=True)
[…]

### 2018-02-14 resex.tex
make font bold for FMA plot annotations

@@
 \subsection{$7^{\mathrm{th}}$ turn pulsing\label{sec:simex7}}
-The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017 and was intended to serve as a proof for the reproducibility of the experimental results.
-Unfortunate for this experiment, the machine tune was changed in standard operation from (64.28,59.31) in 2016 to (62.27,60.295) in 2017 accompanied by a slight change in optics considered irrelevant in the context of the presented measurements.
-The attempt to change back to the fractional tunes of (.28,.31) as chosen in 2016 lead to large losses during the 2017 experiment and therefore the 2017 working point of (62.27,60.295) had to be used instead.
-This change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016 (see Fig.~\ref{fig:patternfma}) to the $7Q_x$, $7Q_y$ and $7Q_x + 7Q_y$ resonance in 2017 (see Fig.~\ref{fig:7th2017fma}).
+The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017.
+In 2016 the resonant excitation was tested for the first time in the LHC and the experiments were therefore still in the exploratory state.
+Based on the experience gained in 2016, the experiments could then be repeated more systematically with more bunches allowing to also test in addition the dependence on the excitation plane.
+Also in case of the $7^{\mathrm{th}}$~turn pulsing, the experiment in 2016 as well as in 2017 can be divided in three parts: the first part without excitation, the second part with a resonant excitation with a maximum amplitude of $A_{\mathrm{max}}=6$~nrad, and a third part in which the excitation amplitude is further increased to $A_{\mathrm{max}}=12$~nrad.
+As the machine tune was changed in standard operation from (64.28,59.31) in 2016 to (62.27,60.295) in 2017 accompanied by a slight change in optics, considered irrelevant in the context of the presented measurements, a direct comparison of the two experiments and thus check of reproducibility is however not possible.
+
+The described change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016 to mainly the $7Q_x$ and $7Q_y$ resonance in 2017.
+This can be seen from the FMA analysis for the 2016 optics (see Fig.~\ref{fig:patternfma} bottom left and bottom right), which shows a strong increase in diffusion for $7^{\mathrm{th}}$~turn pulsing in H around the $7Q_x$ resonances and only small changes for pulsing in V.
+In case of the 2017 optics and tune, the $7Q_x$ as well as the $7Q_y$ resonances cross the tune footprint (see Fig.~\ref{fig:7th2017fma} top left) and are also excited by the $7^{\mathrm{th}}$~turn pulsing visible as an increase of the diffusion in tune around the corresponding resonance lines for pulsing only in H or only in V and thus also for H+V (see Fig.~\ref{fig:7th2017fma}).
+An increase in the diffusion in tune however much weaker than for the $7Q_x$ and $7Q_y$ resonances is also observed around the $7Q_x + 7Q_y$ resonances for pulsing in V and H+V suggesting also an excitation of this $14^{\mathrm{th}}$ order resonance in addition.
+Why it is not visible for pulsing only in H is not understood.
 \begin{figure}[h]
@@
 		no excitation
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate.png}
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate_7th.png}
 	\end{minipage}
@@
 	\end{minipage}	
-	\caption{\label{fig:7th2017fma} FMA analysis in frequency space without excitation (left) and for $7^{\mathrm{th}}$~turn pulsing in H+V (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$7^{\mathrm{th}}$ H
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skh_dp0_ord7.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		$7^{\mathrm{th}}$ V
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skv_dp0_ord7.png}
+	\end{minipage}	
+	\caption{\label{fig:7th2017fma} FMA analysis in frequency space without excitation (top left) and for $7^{\mathrm{th}}$~turn pulsing in H+V (top right), only in H (bottom left) and only in V (bottom right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
 The excitation is 96~nrad in both planes.
-The $7Q_x$ and $7Q_y$ resonance are both excited and in addition the 14th order $7Q_x+7Q_y$.}
+The $7Q_x$ and $7Q_y$ resonance are both excited and in addition to the 14th order $7Q_x+7Q_y$.}
 \end{figure}
@@
 \end{table}
-Consequently, the effect of the $7^{\mathrm{th}}$~turn pulsing on the beam losses, emittance and beam distribution also differ as depicted by the comparison of the distribution tracking for both experiments in Fig.~\ref{fig:7thsim}, the qualitative comparison of the experimental results in Table~\ref{tab:7thexp}, the comparison of the measured emittance in Fig.~\ref{fig:7thexp} and loss rate in Fig.~\ref{fig:7thexploss}.
+
 \begin{figure}[h]
@@
 
-The loss rates measured during the experiment however differ from the simulations in terms of for which case the largest rate is observed as well as quantitatively for the specific cases.
+The losses measured during the experiment however differ from the simulations in terms of for which case the largest rate is observed as well as quantitatively for the specific cases.
 For example smaller losses are expected for the 2016 experiments compared to the 2017 experiments from simulations, while experimentally higher loss rates were measured in 2016 (see Fig.~\ref{fig:7thexploss}).
 For the 2017 experiment, the highest loss rates are expected for pulsing in H+V, than in H and then with a considerably reduction in V.
-During the experiment comparable loss rates are however measured for all three planes (see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexploss2017})
+During the experiment comparable loss rates are however measured for all three planes (see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexploss2017}).
+Last but not least the loss rates as defined in Eqn.~\ref{eqn:lossrate} for the 2016 (red and yellow) and 2017 (black and gree) experiments are compared in Fig.~\ref{fig:7thexploss}.
+In both cases and as also observed for the $10^{\mathrm{th}}$~turn pulsing, the loss rate follows a quadratic scaling with the excitation amplitude.
+Loss rates were in general higher during the 2016 experiments.
+The loss rates from simulations are not shown in Fig.~\ref{fig:7thexploss} as losses were only observed for much higher excitation amplitudes than tried in the experiments.
 \begin{figure}[h]

### 2018-02-14 resex.tex
remove um from rel emittance label for 7th turn pulsing simulation plots

@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{2016+2017_scale_amp_7h_lbllong.png}
+	\includegraphics[width=0.8\linewidth]{2016+2017_scale_amp_7h_lblshort.png}
 	\caption{\label{fig:7thexploss} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in H and data obtained during the 2016 experiment (black and blue) and 2017 experiment (red and yellow).
-The excitation amplitude error is 50\% in 2016 and could be reduced to 15\% in 2017.}
+The excitation amplitude error is 50\% in 2016 and was reduced to 15\% in 2017.
+The solid and dashed line indicate the second order polynomial fit to the loss rate.}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{2017_scale_amp_7h_lbllong.png}
+	\includegraphics[width=0.8\linewidth]{2017_scale_amp_7h_lblshort.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=0.8\linewidth]{2017_scale_amp_7v_lbllong.png}
+		\includegraphics[width=0.8\linewidth]{2017_scale_amp_7v_lblshort.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=0.8\linewidth]{2017_scale_amp_7hv_lbllong.png}
-	\end{minipage}
-	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in only in H (top), only in V (top) and in H+V during the 2017 experiment.
+		\includegraphics[width=0.8\linewidth]{2017_scale_amp_7hv_lblshort.png}
+	\end{minipage}
+	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
 The excitation amplitude error is 15\% indicated by horizontal error bars.

### 2018-02-14 scripts/lifetrac_sim_results/plot_loss_emit_7H_V_H+V_2017_7H_2016.py

@@ -130,7 +130,7 @@ for phv,p12 in zip(['hor.','vert.'],'12'):
     plt.plot(t_16.data['time'],t_16.data['emit%s'%p12]/t_16.data['emit%s'%p12][0],linestyle=lamp,color='orange',label=pltlbl)
   plt.grid(b=grid_flag)
   plt.xlabel('time [s]')
-  plt.ylabel('rel. %s normalized emittance [$\mu$m]'%phv)
+  plt.ylabel('rel. %s normalized emittance'%phv)
   handles,labels = plt.gca().get_legend_handles_labels()
   leg1 = plt.legend(title='(2017)\n7th turn pulsing',ncol=1,fontsize=12,handles=handles[:-2],labels=labels[:-2],loc='upper left')
   leg2 = plt.legend(title='(2016)\n7th turn pulsing',ncol=1,fontsize=12,handles=handles[-3:],labels=labels[-3:],loc='upper center')

### 2018-02-14 resex.tex
update emittance plots for 7th turn pulsing

@@
 Based on the experience gained in 2016, the experiments could then be repeated more systematically with more bunches allowing to also test in addition the dependence on the excitation plane.
-Also in case of the $7^{\mathrm{th}}$~turn pulsing, the experiment in 2016 as well as in 2017 can be divided in three parts: the first part without excitation, the second part with a resonant excitation with a maximum amplitude of $A_{\mathrm{max}}=6$~nrad, and a third part in which the excitation amplitude is further increased to $A_{\mathrm{max}}=12$~nrad.
+Also in case of the $7^{\mathrm{th}}$~turn pulsing, the experiment in 2016 as well as in 2017 can be divided in different parts: the first part without excitation, the second part with a resonant excitation with a maximum amplitude of $A_{\mathrm{max}}=6$~nrad, a third part in which the excitation amplitude is further increased to $A_{\mathrm{max}}=12$~nrad and in case of the 2016 experiments even further to $A_{\mathrm{max}}=24$~nrad.
 As the machine tune was changed in standard operation from (64.28,59.31) in 2016 to (62.27,60.295) in 2017 accompanied by a slight change in optics, considered irrelevant in the context of the presented measurements, a direct comparison of the two experiments and thus check of reproducibility is however not possible.
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $7^{\mathrm{th}}$~turn pulsing with two different excitation amplitudes plus a random dipole noise component in H+V of 6~nrad.}
-\end{figure}
+\end{figure*}
 
@@
 Also in this case, the distribution changes are directly visible on the beam profiles taken with the BSRT (see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexpprof}).
-\begin{figure}[h]
+\begin{figure*}[h]
 	%	\begin{minipage}[t]{0.49\linewidth}
@@
 For all bunches the damper was not active.}
-\end{figure}
+\end{figure*}
 
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{2016+2017_scale_amp_7h_lblshort.png}
+	\includegraphics[width=1.0\linewidth]{2016+2017_scale_amp_7h_lblshort.png}
 	\caption{\label{fig:7thexploss} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in H and data obtained during the 2016 experiment (black and blue) and 2017 experiment (red and yellow).
@@
 \end{figure}
-
+\clearpage
 \subsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}

### 2018-02-14 resex.tex
comments from Giulio including 8th turn pulsing

@@
 	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).
-In 2016 the same excitation amplitude was applied on a group of 4 bunch while in 2016 the number was increased to 6 bunches.
-In 2016 the excitation was also only applied in one plane leading to $2\times24=48$ bunches in total in 2016 and $3\times2\times36=216$ bunches in 2017.}
+In 2016 a total of 48~bunches was used and in 2017 a total of 216~bunches.
+Each bunch is indicated with a cyan vertical bar.
+The bunches were grouped in sub-groups of 4~for 2016 and 6~for 2017 bunches, where each sub-group experienced the same excitation pattern and amplitude.
+The excitation amplitude is shown in blue and red.
+In 2016 the excitation was only applied in one plane.
+In 2017 more bunches could be injected without compromising machine protection making it possible to test all three planes in one fill.}
 \end{figure*}
@@
 The measured values are therefore believed to be known with an error margin of 10-15\% for the experiments in 2017 and 50\% in 2016.
-The error in 2016 is much larger as no calibration of the ADT was performed beforehand, which was done in 2017.
+The estimate for 2017 is based on only a single measurement and in the future and with more time available the determination of the error would definitely have to be improved.
+The error in 2016 is an educated guess as no calibration was performed beforehand.
 The maximum kick strength reachable with the ADT at injection without risking saturation is 96~nrad.
@@
 \end{figure}
+
 \clearpage
+
 \subsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
-Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they in addition have a large effect on the halo, which is in general feasible considering the highly nonlinear field generated at the location of the halo particles by the HEL.
-With this motivation in mind the $8^{\mathrm{th}}$~turn pulsing pattern was studied, for which a small effect in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing is expected from the distribution tracking summarized in Fig.~\ref{fig:8thsim}.
-\begin{figure}[h]
+Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they in addition have a large effect on the halo.
+This is in general feasible considering the highly nonlinear field generated at the location of the halo particles by the HEL compared to no field or only a very small field at the location of the beam core.
+With this motivation in mind the $8^{\mathrm{th}}$~turn pulsing pattern was studied, for which a small effect in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing is expected (see Fig.~\ref{fig:8thsim}).
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $8^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
-%\end{figure}
-%\begin{figure}[t]
+\end{figure*}
+\begin{figure}[t]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 \end{figure}
-
-For excitation amplitudes smaller than 96~nrad no visible effect is indeed observed.
-In the horizontal and vertical plane a small emittance growth due to the excitation is visible, but without any clear dependency on the amplitude nor plane.
-Adducting the FMA analysis in amplitude space shown in Fig.~\ref{fig:8th2017fmaamp}, the $8^{\mathrm{th}}$ turn pulsing drives the $16Q_y$ and \mbox{$8Q_x-4Q_y$} resonances and several even higher orders.
+In simulations no visible effect is indeed observed for excitation amplitudes smaller than 96~nrad.
+In the horizontal and vertical plane a small emittance growth due to the excitation is visible, but without any clear dependence on the amplitude or plane.
+The FMA analysis in amplitude space reveals also the driven resonances (Fig.~\ref{fig:8th2017fmaamp}), which are the $16Q_y$ and \mbox{$8Q_x-4Q_y$} resonances and several even higher orders.
 These are in general very high order resonances whose effect is thus expected to be small.
 
-The only small effect on the beam expected from simulations could then be also confirmed experimentally during the 2017 experiment.
-For this purpose the excitation amplitude was increased to 96~nrad, which is the maximum kick strength reachable with the transverse damper at injection without risking saturation.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.0\linewidth]{2017_scale_amp_8hv_lblshort.png}
+	\caption{\label{fig:8thexploss} Comparison of scaling of loss rates with excitation amplitude for $8^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars, the error on the loss rate is the error on the mean obtained through the average over the 6~bunches with the same excitation amplitude.}
+\end{figure}
+The small effect on the beam expected from simulations could then be also confirmed experimentally during the 2017 experiment.
+For this purpose the excitation amplitude was increased to the maximum reachable amplitude of 96~nrad.
 As depicted in Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the excitation amplitude for an excitation in H+V.
-For an excitation only in V or only in H, no statistically relevant increase was observed.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1.0\linewidth]{2017_scale_amp_8hv_lbllong.png}
-	\caption{\label{fig:8thexploss} Comparison of scaling of loss rates with excitation amplitude for $8^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.}
-\end{figure}
-
-In the horizontal plane, a change in beam distribution was measured with the BSRT profiles (Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexpprof}) for an excitation only in H, explicitly an increase of the distribution at around $2.0~\sigma$ simultaneously with a depletion of the beam core.
-For an excitation in H+V the emittances of the references bunches were much smaller, around $1.8~\mu$m normalized emittance on average, than the ones of the excited bunches ranging from $2.3~\mu$m to $3.0~\mu$m as described in \cite{resexmd2017}.
-The references bunches thus experienced a stronger emittance growth due to intra-beam scattering than the excited bunches due to the excitation on top of their smaller emittance growth due to intra-beam scattering.
-Therefore no excitation amplitude dependent emittance growth was observed for an excitation in H+V albeit it was recorded for an excitation only in H.
-In the vertical plane, the beam distribution and thus emittance stayed unchanged.
-For an excitation only in V, no dependence on the excitation was measured.
+For an excitation only in V or only in H, no statistically relevant increase was observed (see Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexploss2017}).
+In the horizontal plane, a change in beam distribution was measured with the BSRT profiles (Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexpprof}) for an excitation only in~H.
+The distribution increased around $2.0~\sigma$ simultaneously with a depletion of the beam core.
+For an excitation in H+V one would expect to see a similar distribution change.
+However, this change is not visible as the emittance of the references bunches was much smaller, around $1.8~\mu$m normalized emittance, compared to around $2.6~\mu$m for the excited bunches \cite{resexmd2017}.
+This difference in emittance then resulted in a stronger emittance growth due to intra-beam scattering for the reference bunches in comparison to the excited bunches making the additional emittance growth and distribution changes due to the excitation undetectable.
+In the vertical plane, the beam distribution and emittance were not affected by the resonant excitation.
+
+\clearpage
 
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{2017_scale_amp_7h_lblshort.png}
+	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=0.8\linewidth]{2017_scale_amp_7v_lblshort.png}
+		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
 	\end{minipage}
@@
 		\centering
-		\includegraphics[width=0.8\linewidth]{2017_scale_amp_7hv_lblshort.png}
+		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
 	\end{minipage}
@@
 \begin{figure}[h]
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8h_lblshort.png}
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8v_lblshort.png}
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8hv_lblshort.png}
+	\end{minipage}
+	\caption{\label{fig:8thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $8^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by horizontal error bars.
+The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
+\end{figure}
+\begin{figure}[h]
 	\begin{minipage}[t]{0.49\linewidth}

### 2018-02-14 resex.tex
Comments from Giulio for random excitation

@@
 The estimate for 2017 is based on only a single measurement and in the future and with more time available the determination of the error would definitely have to be improved.
-The error in 2016 is an educated guess as no calibration was performed beforehand.
+The error in 2016 is estimated from the transverse damper operational parameters.
+Due to lack of the dedicated machine time no precision calibration was performed.
 The maximum kick strength reachable with the ADT at injection without risking saturation is 96~nrad.
@@
 \begin{figure*}[h]
-	%	\begin{minipage}[t]{0.49\linewidth}
-	%		\centering
-	%		2016 experiment
-	%		\includegraphics[width=1.0\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
-	%	\end{minipage}	
-	%	\begin{minipage}[t]{0.49\linewidth}
-	%		\centering
-	%		2017 experiment
-	%		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
-	%	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		2016 experiment, $7^\mathrm{th}$ H	\includegraphics[width=1.0\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		2017 experiment, $7^\mathrm{th}$ H+V
-		\includegraphics[width=1.0\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
+		2016 experiment, $7^\mathrm{th}$ H	\\	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
+	\end{minipage}	
+		2017 experiment, $7^\mathrm{th}$ H+V\\
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
 	\end{minipage}	
@@
 \end{figure}
-This can be seen formidably on the example of the FMA analysis in amplitude space, depicted in Fig.~\ref{fig:ran2017fmaamp}, for a random excitation as applied during the 2017 experiment.
-In this case, the diffusion in tune is increased equally by the random excitation independent of the particle amplitude.
-
-\begin{figure}[h]
+This can be seen clearly on the example of the FMA analysis in amplitude space depicted in Fig.~\ref{fig:ran2017fmaamp}.
+In the case of a random excitation, the diffusion in tune is increased in comparison to the $k^{\mathrm{th}}$~turn pulsing rather equally for all particle amplitudes.
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 The solid line indicates the case with no excitation, and the dotted and dashed line the results for a uniform random excitation of 12~nrad and 24~nrad respectively.}
-\end{figure}
-The results from the distribution tracking are shown in Fig.~\ref{fig:ransim}.
-In contrast to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing, the effect of a random excitation is very symmetric:
+\end{figure*}
+The distribution tracking shown in Fig.~\ref{fig:ransim} also reveals a quite symmetric evolution of the beam parameters for a random excitation:
 \begin{itemize}
-	\item Emittance growth is only caused in the plane of excitation.
-	\item The emittance growth is comparable between an excitation in H and V in the corresponding plane as well as the scaling with the excitation amplitude.
+	\item Emittance growth is only caused by an excitation in the same plane.
+	\item The emittance growth is quantitatively comparable for an excitation in only the corresponding plane.
 Only a small increase or decrease is observed if the excitation is applied in both planes.
 The difference could be due to coupling.
-	\item The random excitation leads to an increased constant emittance growth rate which is reached without undergoing an initial adjustment phase.
+	\item The random excitation leads to an increase of the emittance growth rate. Compared to the $k^{\mathrm{th}}$~turn pulsing there is explicitly no initial adjustment phase of the distribution.
 	\item Negligible losses are observed independent of the plane of excitation.
 \end{itemize}
-
-\begin{figure}[b]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		random excitation V
-		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_vran_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		random excitation H+V
-		\includegraphics[width=1.0\linewidth]{2017_bunch_intensity_hvran_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_emith_avg_rel_vran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_emith_avg_rel_hvran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_hvran_no_damper.png}
-	\end{minipage}	
-	\caption{\label{fig:ranexp} Relative intensity (top) and relative horizontal (center) and vertical (bottom) emittance for a random excitation only in V (left) and in H+V (right).
+\begin{figure*}[b]
+	random excitation V\\
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_vran_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_vran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
+	\end{minipage}	
+		random excitation H+V\\
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hvran_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hvran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hvran_no_damper.png}
+	\end{minipage}	
+	\caption{\label{fig:ranexp} Relative intensity (left) and relative horizontal (center) and vertical (right) emittance for a random excitation only in V (top) and in H+V (bottom).
 Emittance growth is only caused in the plane of excitation.}
-\end{figure}
-Qualitatively, this behavior could be also confirmed experimentally during the 2017 experiments, where excitation amplitude dependent emittance growth was only recorded in the plane of excitation with all excitation amplitude dependent emittance growth rates in a comparable range.
+\end{figure*}
+
+Qualitatively, this behavior could be also confirmed experimentally during the 2017 experiments.
+Here excitation amplitude dependent emittance growth was only recorded in the plane of excitation with all excitation amplitude dependent emittance growth rates in a comparable range.
 As example the relative horizontal and vertical emittance for pulsing in V and in H+V are depicted in Fig.~\ref{fig:ranexp}.
-Note that as expected from simulations and despite the large excitation amplitude dependent emittance growth, the growth rate stays constant and there is explicitly no fast adjustment phase succeeded by a new equilibrium state as was observed for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
-
+Note that as expected from simulations and despite the large excitation amplitude dependent emittance growth, the growth rates stays constant.
+There is explicitly no fast adjustment phase succeeded by a new equilibrium state as observed for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
 \begin{figure}[h]
@@
 \end{figure}
-The large changes of the emittance are of course also directly visible in the beam profiles measured with the BSRT shown in Fig.~\ref{fig:ranexpprof} for a random excitation only in V.
-In comparison to the $10^{\mathrm{th}}$~turn pulsing shown in Fig.~\ref{fig:10thexpprof}, the distribution also stays Gaussian in case of a random excitation while for the $10^{\mathrm{th}}$~turn pulsing it clearly adopts to a non-Gaussian shape.
-
-\begin{figure}[b]
-	\centering
-	\includegraphics[width=0.8\linewidth]{2017_scale_amp_ranhv_lbllong.png}
-	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.}
-\end{figure}
+The changes in emittance are also directly visible in the beam profiles measured with the BSRT.
+As an example Fig.~\ref{fig:ranexpprof} shows the profile for a random excitation only in V in comparison to a reference bunch.
+The distribution also stays Gaussian in case of a random excitation in contrast to the $10^{\mathrm{th}}$~turn pulsing in which case it clearly assumes a non-Gaussian shape (see Fig.~\ref{fig:10thexpprof}).
+
+\begin{figure*}[h]
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranh_lblshort.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranv_lblshort.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranhv_lblshort.png}
+	\end{minipage}	
+	\caption{\label{fig:ranexplossplane} Comparison of scaling of loss rates with excitation amplitude for a random excitation only in H (top), only in V (center) and in H+V (bottom) measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars.
+The loss rate is considerably reduced for the bunches with transverse damper active and is approximately additive with the excitation plane, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only in V.}
+\end{figure*}
 The experimental results for a maximum excitation amplitude of $A_{\mathrm{max}}=1$~nrad and $A_{\mathrm{max}}=6$~nrad also confirm the in general small losses predicted in simulations.
-The losses then suddenly increase once the maximum excitation amplitude is increased to $A_{\mathrm{max}}=12$~nrad, which is most likely the point when, due to the large emittance increase, the physical aperture starts eating into the more central region of the distribution resulting in higher losses despite the larger emittance growth for a random excitation.
-In comparison to the $7^{\mathrm{th}}$~turn pulsing, the loss rate for a random excitation is smaller for a maximum excitation amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}).
-This indicates an increase of the diffusion mostly in the beam tails for $7^{\mathrm{th}}$~turn pulsing, while for a random excitation mostly the core is affected.
-Furthermore, the dependence of the loss rate on the excitation plane is additive, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only V (see Appendix~\ref{app:sec:ran}, Fig.~\ref{fig:ranexplossplane}).
-In view of the next section, Sec.~\ref{sec:damp}, not that the loss rate is also considerably reduced for a random excitation if the transverse damper is active, while in case of the $7^{\mathrm{th}}$~turn pulsing the loss rate does not change.
+The losses then increase more dramatically around the time when the maximum excitation amplitude is increased to $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexpprof}).
+This is most likely the point when, due to the large emittance growth, the physical aperture starts cutting into the more central region of the distribution.
+Furthermore, the dependence of the loss rate on the excitation plane is additive, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only V.
+The loss rate is also considerably reduced for the bunches with transverse damper active independent of the plane the excitation is applied in (see Fig.~\ref{fig:ranexplossplane}).
+
+\clearpage
 
@@
 \end{figure}
-For all $k^{\mathrm{th}}$~turn pulsing patterns, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation.
+For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation.
 On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation.
@@
 \section{Summary\label{sec:sum}}
+In comparison to the $7^{\mathrm{th}}$~turn pulsing, the loss rate for a random excitation is smaller for a maximum excitation amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}).
+This indicates an increase of the diffusion mostly in the beam tails for $7^{\mathrm{th}}$~turn pulsing, while for a random excitation mostly the core is affected.
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=0.8\linewidth]{2017_scale_amp_7th_ranhv_lblshort.png}
+	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars.}
+\end{figure}
+
 This paper intends to give an overview of the effects of the HEL on the proton beam core and the experiments conducted in this context at the LHC in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}.
@@
 
-\section{Additional plots for random excitation during 2017 experiments}
-\label{app:sec:ran}
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranh_lbllong.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranv_lbllong.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranhv_lbllong.png}
-	\end{minipage}	
-	\caption{\label{fig:ranexplossplane} Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.
-Note that for $7^{\mathrm{th}}$~turn pulsing the transverse damper does not influence the loss rate, while for a random excitation the loss rate is considerably reduced for the bunches with transverse damper active.}
-\end{figure}
+%\section{Additional plots for random excitation during 2017 experiments}
+%\label{app:sec:ran}
 

### 2018-02-15 resex.tex
Changes from Giulio completed

@@
 \begin{equation}
-	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \SI{19.6}{nrad} % this should be rounded to the same number of digits as the 0.5nrad
+	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \SI{20}{nrad} % this should be rounded to the same number of digits as the 0.5nrad
 \end{equation}
@@
 \begin{equation}\label{eqn:kick_central}
-\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \SI{19.6}{nrad}
+\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \SI{20}{nrad}
 \end{equation}
@@
 The measured activity is well in line with the expected excitation pattern.
-\begin{figure}[h]
+\begin{figure*}[h]
 	\centering
@@
 	\caption{\label{fig:fill_meas} Bunch by bunch excitation detected by the real time transverse activity monitor during the 2017 experiment and an excitation with 8th turn pulsing.}
-\end{figure}
+\end{figure*}
 
@@
 The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect can be thus attributed to the magnetic field errors, implying also the seed chosen for the simulation.
-\begin{figure}[h]
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V (see Appendix~\ref{app:sec:10}, Fig.~\ref{app:fig:fma:10}).}
-\end{figure}
+\end{figure*}
 
@@
 The timescale of this fast adjustment is in experiments however much slower than predicted in simulations --- on the timescale of minutes compared to seconds.
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 The red line in the plot of the residual is the difference between the Gaussian fit of the distribution and the distribution itself.}
-\end{figure}
+\end{figure*}
 
@@
 	\end{minipage}	
-	\caption{\label{fig:10thexp} Relative intensity (top) and relative horizontal (bottom left) and vertical (bottom right) emittance averaged over the bunches experiencing the same excitation amplitude during the 2016 experiments.
+	\caption{\label{fig:10thexp} Relative intensity (top) and relative horizontal (bottom left) and vertical (bottom right) emittance averaged over the bunches experiencing the same excitation amplitude during the 2016 experiments and for $10^{\mathrm{th}}$~turn pulsing in~V.
 For all bunches the damper was not active.
@@
 
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
+\begin{figure*}[h]
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch, no excitation
 		\includegraphics[width=1.0\linewidth]{profile_v_10thv_slot_50.png}
@@
 		\centering
-		excited bunch,\\ $10^{\mathrm{th}}$~turn V		\includegraphics[width=1.0\linewidth]{profile_v_10thv_slot_1300.png}
+		excited bunch, $10^{\mathrm{th}}$~turn V		\includegraphics[width=1.0\linewidth]{profile_v_10thv_slot_1300.png}
 	\end{minipage}
@@
 The distribution of the reference bunch (left), here slot 50, stays unchanged, while the bunch experiencing the maximum excitation (right), here slot 1300, clearly shows a change to a non-Gaussian distribution.}
-\end{figure}
+\end{figure*}
 Thanks to the good performance of the Beam Synchrotron Radiation Telescope (BSRT)~\cite{bsrtprofinj}, the distribution changes indirectly observed as changes of the emittance could also be for the first time directly measured.
@@
 Why it is not visible for pulsing only in H is not understood.
-\begin{figure}[h]
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 The $7Q_x$ and $7Q_y$ resonance are both excited and in addition to the 14th order $7Q_x+7Q_y$.}
-\end{figure}
+\end{figure*}
 
@@
 \end{figure*}
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 The $16Q_y$ and \mbox{$8Q_x-4Q_y$} are both excited.}
-\end{figure}
+\end{figure*}
 In simulations no visible effect is indeed observed for excitation amplitudes smaller than 96~nrad.
@@
 The random excitation is in general so powerful as it excites all frequencies of the beam equally.
-\begin{figure}[h]
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space without excitation (left) and for a uniform random excitation with 1~nrad amplitude (right) in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).}
-\end{figure}
+\end{figure*}
 This can be seen clearly on the example of the FMA analysis in amplitude space depicted in Fig.~\ref{fig:ran2017fmaamp}.
@@
 There is explicitly no fast adjustment phase succeeded by a new equilibrium state as observed for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
-\begin{figure}[h]
+\begin{figure*}[h]
 	\begin{minipage}[t]{0.49\linewidth}
@@
 The distribution changes in the reference bunch (left), here slot 1698, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
-\end{figure}
+\end{figure*}
 The changes in emittance are also directly visible in the beam profiles measured with the BSRT.
@@
 The loss rate is also considerably reduced for the bunches with transverse damper active independent of the plane the excitation is applied in (see Fig.~\ref{fig:ranexplossplane}).
-
-\clearpage
-
-\subsection{Effect of the transverse damper\label{sec:damp}}
-\begin{figure}[b]
-	\centering
-	random excitation H+V\\
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		damper not active
-		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		damper active
-		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_vran_with_damper.png}
-	\end{minipage}	
-	\centering
-	$7^{\mathrm{th}}$~turn H+V\\
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		damper not active
-		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		damper active
-		\includegraphics[width=1.0\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
-	\end{minipage}	
-	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for $7^{\mathrm{th}}$~turn pulsing in V (left) and for a random excitation only in V (right).}
-\end{figure}
-For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not capable of damping any resonant excitation.
-On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation.
-As an example, the vertical emittance growth for the $7^{\mathrm{th}}$~turn pulsing in V  is compared with a random excitation in V in Fig.~\ref{fig:damp} and the loss rates in Fig.~\ref{fig:ranexploss} illustrating impressively the aforesaid behavior.
-\textcolor{red}{Any idea why? Daniel, help!}
-
-\section{Summary\label{sec:sum}}
 In comparison to the $7^{\mathrm{th}}$~turn pulsing, the loss rate for a random excitation is smaller for a maximum excitation amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}).
 This indicates an increase of the diffusion mostly in the beam tails for $7^{\mathrm{th}}$~turn pulsing, while for a random excitation mostly the core is affected.
-\begin{figure}[b]
-	\centering
-	\includegraphics[width=0.8\linewidth]{2017_scale_amp_7th_ranhv_lblshort.png}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1.0\linewidth]{2017_scale_amp_7th_ranhv_lblshort.png}
 	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
@@
 
-This paper intends to give an overview of the effects of the HEL on the proton beam core and the experiments conducted in this context at the LHC in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}.
+\clearpage
+
+\subsection{Effect of the transverse damper\label{sec:damp}}
+\begin{figure*}[b]
+	\centering
+	damper not active\\
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+	random excitation V, 2017\\
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		$7^{\mathrm{th}}$~turn V, 2017 exp.\\
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		$10^{\mathrm{th}}$~turn V, 2016 exp.\\
+		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_no_damper_no_text.png}
+	\end{minipage}	
+	damper active\\
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		random excitation V, 2017 exp.\\
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_with_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		$7^{\mathrm{th}}$~turn V, 2017 exp.\\
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		$10^{\mathrm{th}}$~turn V, 2016 exp.\\
+		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_with_damper_no_text.png}
+	\end{minipage}	
+	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for a random excitation only in V (left) and for $7^{\mathrm{th}}$ (center) and $10^{\mathrm{th}}$~turn pulsing only in V (right).}
+\end{figure*}
+For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not damping any resonant excitation.
+On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation.
+As an example, the vertical emittance growth for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in V is compared with a random excitation in V in Fig.~\ref{fig:damp}.
+The same observation is also made in case of losses as shown in Fig.~\ref{fig:ranexploss} comparing the $7^{\mathrm{th}}$~turn pulsing in H+V with a random excitation in H+V.
+For the $7^{\mathrm{th}}$~turn pulsing the two curves for the transverse damper active and not active almost coincide while for the random excitation a significant difference is visible.
+Until present, it is unknown why the transverse damper (ADT) appears to be incapable of damping any resonant excitation as it fulfills all requirements to detect as well as damp the caused oscillation:
+\begin{itemize}
+	\item The closed orbit distortion caused by the resonant excitation during the experiments is large enough to be detectable with the ADT pickups.
+	\item The ADT compares the position of each bunch with its position in the previous turn in order to detect an orbit distortion to be damped.
+Therefore an oscillation due to any $k^{\mathrm{th}}$~turn pulsing with $k>1$ would be noticed and damped.
+	\item The ADT is capable of damping the oscillation of each bunch individually.
+The ADT should therefore be capable of applying the correct corrective kick for each group of bunches seeing the same excitation amplitude.
+\end{itemize}
+Further studies are therefore needed in order to obtain a deeper understanding of the interaction of the transverse damper with a resonant excitation.
+\clearpage
+
+\section{Summary\label{sec:sum}}
+This paper intends to give an overview of the effects of the hollow electron lens (HEL) on the proton beam core and the experiments conducted in this context at the LHC in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}.
 For a perfect HEL without any imperfections no effect on the core is to be expected as the field at the center vanishes by construction.
@@
 In this case and assuming HL-LHC parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}), the contribution from the electron lens bends with approximately $0.5$~nrad to first order is negligible with respect to the kick expected from the central region of the HEL of approximatley $20$~nrad to first order caused by profile imperfections in the electron beam.
-
 In DC operation these kicks, even if nonlinear in reality, are negligible compared to the nonlinearities present in the LHC and HL-LHC and are thus of no concern.
@@
 To study the effect on the beam core in case of the much more dangerous pulsed operation, beam experiments were conducted in 2016 and 2017 at injection.
+As no hollow electron lens is currently installed in the LHC, the in reality non-linear kick was emulated to first order by a dipole kick with the corresponding excitation pattern.
 The experimental results in general suggest that for the kick amplitudes as expected from the HEL, considerable effects are to be expected for a random excitation and certain resonant excitations, explicitly $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing during the conducted experiments.
@@
 The random excitation on the other hand excites all beam frequencies equally, leading to the highest increase in halo removal rates for the HEL \cite{hel_halo_hllhc_fitterer} but also the strongest effects on the core, mainly in terms of strong emittance growth as predicted in simulations, experiments and also a multitude of earlier publications on the subject of random noise.
-
-In summary, the effects of the HEL on the beam core in case of pulsed operation are not negligible.
-Tolerances in case of a random excitation are extremely stringent, while a resonant excitation appears feasible if the order $k$ of the $k^{\mathrm{th}}$~turn pulsing is tuned to the machine and beam configuration, explicitly tune, tune spread and machine nonlinearities.
+From the experiments the tolerances summarized in Table~\ref{tab:tolexp} could be obtained, where for each excitation pattern and plane the smallest kick amplitude is listed for which an increase in emittance or losses was observed.
+\begin{table}
+	\caption{\label{tab:tolexp}%
+		Minimum kick amplitudes $\Delta z'$ for which either an increase in emittance or losses was observed during the LHC experiments.
+The tolerances are given for the case of the transverse damper not active.
+The same tolerances are obtained for the case with the damper active in case of all resonant excitation.
+In case of a random excitation the tolerance can be increased by roughly a factor~2.
+	}
+	\begin{ruledtabular}
+		\begin{tabular}{lccc}
+		pulsing pattern & exp. year & plane & $\Delta z' \ [\mathrm{nrad}]$  \\
+			\colrule
+			random & 2017 & H or V & 2.4 \\
+			 & & H+V & 1.2 \\\hline
+			$7^{\mathrm{th}}$~turn & 2016 & H & 2.4 \\
+			 & 2017 & H & 2.4 \\
+			 & 2017 & V & 1.2 \\
+			 & 2017 & H+V & 1.2 \\\hline
+			$8^{\mathrm{th}}$~turn & 2017 & H & 14.4 \\
+			& 2017 & V & $>96$ \\
+			& 2017 & H+V & 12 \\\hline
+			$10^{\mathrm{th}}$~turn & 2016 & V & 9.6 \\
+		\end{tabular}
+	\end{ruledtabular}
+\end{table}
+Tolerances in case of a random excitation are extremely stringent.
+For a resonant excitation the effect depends strongly on the order $k$ of the $k^{\mathrm{th}}$~turn pulsing.
+For a future HEL, the pattern could therefore be tuned to the machine and beam configuration in order to yield a strong effect on the halo particles and small effect on the beam core.
+Compared to the estimated kick amplitude of the HEL on the beam core of
+\begin{equation}
+\Delta x', \Delta y' = 20 \ \mathrm{nrad}
+\end{equation}
+the tolerances for all experimentally tried pulsing pattern except for $8^{\mathrm{th}}$~turn pulsing in V lie below this value.
+As a next step, the experiment should be conducted at collision energy where the HEL is most needed.
+At collision energy the inherent noise level of the LHC is smaller which could lead to more relaxed tolerances than the ones obtained at injection energy.
 
@@
 We wish to acknowledge Giulia Papotti with her help with the preparations for the 2016 experiments and her general support from the operations side, Roderik Bruce and Gianluca Valentino for there very helpful suggestions in preparing the experiment at the LHC and we would like to thank all participants of the experiment for helping acquire the data presented in this paper.
+We wish to thank Sergei Nagaitsev for the very useful discussion and profound question to find out why the transverse damper is not damping any resonant excitation.
 We would like to acknowledge Dmitry Shatilov for his help with the Lifetrac code, and St\'{e}phane Fartoukh, Riccardo De Maria and Rogelio Tomás for their help with generating the appropiate optics model.

### 2018-02-16 resex.tex
Updated figures of residual fields with 1.24-mm inner diameter.

@@
 \end{figure}
-As at the test stand the solenoidal fields required for the HL-LHC electron lens can not be reached, an estimate of the electron beam profile for HL-LHC can only be obtained through simulations.
-For this purpose, the measured distribution is compressed to the inner electron beam radius of $0.97$~mm.
-Afterwards an input distribution with $40\ 000$ particles is generated.
-As boundary condition, the LHC inner beam pipe radius $b=30$~mm is used.
-The potential and fields can then be calculated with the particle in cell simulation code WARP~\cite{warp} yielding the relative electric field strengths shown in Fig.~\ref{core:fig:1}.
-The electric field can in general be scaled with the beam current and energy, but can not easily be scaled with the inner and outer radius of the electron beam distribution.
-Therefore it is essential to track a distribution with the correct dimensions.
-Further details on the measurements and WARP simulations can be found in~\cite{hel_res_field_stancari_2017}.
+
+As at the test stand the solenoidal fields required for the HL-LHC
+electron lens can not be reached, an estimate of the electron beam
+profile for HL-LHC can only be obtained through simulations. For this
+purpose, the measured distribution is compressed to the inner electron
+beam radius of 1.24~mm. Afterwards an input distribution with
+$40\ 000$ particles is generated. As boundary condition, the LHC inner
+beam pipe radius $b=30$~mm is used. The potential and fields can then
+be calculated with the particle in cell simulation code
+Warp~\cite{warp} yielding the relative electric field strengths shown
+in Fig.~\ref{core:fig:1}. The electric field can in general be scaled
+with the beam current and energy, but can not easily be scaled with
+the inner and outer radius of the electron beam
+distribution. Therefore it is essential to track a distribution with
+the correct dimensions. Further details on the measurements and Warp
+simulations can be found in~\cite{hel_res_field_stancari_2017}.
+
 \begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice.png}
-	\end{minipage}
-	\caption{Calculated relative electric field for the new CHG1b hollow electron gun in the x-y plane (left) and as 1-dimensional cuts though the x- and y-axes (right).
-The field calculations are based on measurements at the Fermilab test stand combined with WARP  calculations of the electric potentials and fields in a cylindrical beam pipe.}
+\includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap}
+\includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
+\caption{Calculated relative electric field for the hollow electron
+  gun in the transverse $x$-$y$ plane (left) and as 1-dimensional cuts
+  through the $x$ and $y$ axes (right). The field calculations are
+  based on measurements at the Fermilab electron-lens test stand
+  combined with Warp calculations of the electric potentials and
+  fields in a cylindrical beam pipe.}
 \label{core:fig:1}

### 2018-02-16 resex.tex
Updated figures of measured current-density profiles

@@
 \label{core:sec:2}
-For a perfectly annular and axially symmetric profile, the field in the region of the proton beam core vanishes as can be seen from Eq.~\ref{eq:field_2} and as illustrated in Fig.~\ref{fig:hel_field}.
-Fields at the location of the proton beam core thus only arise in case of profile imperfections of the electron beam, in particular if the axial symmetry is broken.
-Recently, the latest LHC hollow electron gun prototype was characterized at the Fermilab electron lens test stand \cite{hel_test_stand_fnal}.
-A typical measurement of the electron beam current density is shown in Fig.~\ref{core:fig:0}.
-\begin{figure}[t]
-		\centering
-		\includegraphics[width=0.8\linewidth]{e-field-measured.png}
-	\caption{Current density from measurements of the CHG1b hollow electron gun at the Fermilab electron lens test stand in 2017 \cite{hel_res_field_stancari_2017}.
-\textcolor{red}{Replace by 2D simulations - Giulio.}}
+
+\begin{figure*}
+  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} \hfill
+  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}
+  \caption{Example of current-density distribution measurements for
+    the hollow electron gun prototype CHG1b, taken at the Fermilab
+    electron lens test stand in
+    2017~\cite{hel_res_field_stancari_2017}: 2-dimensional transverse
+    profile measurement (left) and calculated 1-dimensional radial
+    projection (right).}
   \label{core:fig:0}
-\end{figure}
+\end{figure*}
+
+\begin{figure*}
+  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} \hfill
+  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
+  \caption{Calculated relative electric field for the hollow electron
+    gun CHG1b in the transverse $x$-$y$ plane (left) and as
+    1-dimensional cuts through the $x$ and $y$ axes (right). The field
+    calculations are based on measurements at the Fermilab
+    electron-lens test stand combined with Warp calculations of the
+    electric potentials and fields in a cylindrical beam pipe.}
+\label{core:fig:1}
+\end{figure*}
+
+For a perfectly annular and axially symmetric profile, the field in
+the region of the proton beam core vanishes as can be seen from
+Eq.~\ref{eq:field_2} and as illustrated in
+Fig.~\ref{fig:hel_field}. Fields at the location of the proton beam
+core thus only arise in case of profile imperfections of the electron
+beam, in particular if the axial symmetry is broken. Recently, the
+latest LHC hollow electron gun prototype was characterized at the
+Fermilab electron lens test stand \cite{hel_test_stand_fnal}. A
+typical measurement of the electron beam current density is shown in
+Fig.~\ref{core:fig:0}.
 
@@
 simulations can be found in~\cite{hel_res_field_stancari_2017}.
-
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap}
-\includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
-\caption{Calculated relative electric field for the hollow electron
-  gun in the transverse $x$-$y$ plane (left) and as 1-dimensional cuts
-  through the $x$ and $y$ axes (right). The field calculations are
-  based on measurements at the Fermilab electron-lens test stand
-  combined with Warp calculations of the electric potentials and
-  fields in a cylindrical beam pipe.}
-\label{core:fig:1}
-\end{figure}
 

### 2018-02-16 resex.tex
Revised section on fields in overlap region

@@
 For a perfectly annular and axially symmetric profile, the field in
-the region of the proton beam core vanishes as can be seen from
-Eq.~\ref{eq:field_2} and as illustrated in
+the region of the proton beam core vanishes. This is expressed by
+Eq.~\ref{eq:field_2} and is illustrated in
 Fig.~\ref{fig:hel_field}. Fields at the location of the proton beam
-core thus only arise in case of profile imperfections of the electron
-beam, in particular if the axial symmetry is broken. Recently, the
-latest LHC hollow electron gun prototype was characterized at the
-Fermilab electron lens test stand \cite{hel_test_stand_fnal}. A
-typical measurement of the electron beam current density is shown in
-Fig.~\ref{core:fig:0}.
-
-As at the test stand the solenoidal fields required for the HL-LHC
-electron lens can not be reached, an estimate of the electron beam
-profile for HL-LHC can only be obtained through simulations. For this
-purpose, the measured distribution is compressed to the inner electron
-beam radius of 1.24~mm. Afterwards an input distribution with
-$40\ 000$ particles is generated. As boundary condition, the LHC inner
-beam pipe radius $b=30$~mm is used. The potential and fields can then
-be calculated with the particle in cell simulation code
-Warp~\cite{warp} yielding the relative electric field strengths shown
-in Fig.~\ref{core:fig:1}. The electric field can in general be scaled
-with the beam current and energy, but can not easily be scaled with
-the inner and outer radius of the electron beam
-distribution. Therefore it is essential to track a distribution with
-the correct dimensions. Further details on the measurements and Warp
-simulations can be found in~\cite{hel_res_field_stancari_2017}.
-
-Using these measurements and calculations of the relative electric field strengths in the transverse plane (Fig.~\ref{core:fig:1}), we obtain an average field in the hole of
+core can arise in case of profile imperfections of the electron
+beam, in particular if the axial symmetry is broken.
+
+Recently, a hollow electron gun prototype for the LHC (called CHG1b)
+was characterized at the Fermilab electron lens test
+stand~\cite{hel_test_stand_fnal}. An example of a measurement of the
+electron beam current density is shown in Fig.~\ref{core:fig:0}.
+
+In the test stand, only resistive solenoids are available. Through
+simulations, one can estimate the fields generated by the compressed
+electron beam profile needed for HL-LHC. The measured distribution is
+compressed to the inner electron beam radius of 1.24~mm, corresponding
+to $4\sigma_p$, as described in Table~\ref{tab:hllhc_param} and
+Table~\ref{tab:hel_param}. A distribution with about $65\,000$
+particles is generated according to the measured current-density
+profile. As boundary condition, the LHC inner beam pipe radius
+$b=30$~mm is used. The potential and fields are then calculated with
+the particle-in-cell code Warp~\cite{warp}. The resulting relative
+electric field strengths are shown in Fig.~\ref{core:fig:1}. The
+electric field is obviously proportional to the charge density of
+electrons. Its relative strength is rather insensitive to the hollow
+beam radii, as long as these remain small compared to the beam pipe
+radius. Further details on the measurements and Warp simulations can
+be found in Ref.~\cite{hel_res_field_stancari_2017}.
+
+Using these measurements and calculations of the relative electric
+field strengths in the transverse plane (Fig.~\ref{core:fig:1}), we
+obtain an average field in the hole of
+
 \begin{equation}
-	E_{\mathrm{rel}}= 5\%
+  E_{\mathrm{rel}}= 5\%.
 \end{equation}
-For a maximum kick of $392$~nrad as derived in Eq.~\ref{eqn:helkick}, the estimated dipole kick amplitude at the proton beam core then is:
-\begin{equation}\label{eqn:kick_central}
-\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \SI{20}{nrad}
+
+For a maximum kick of 392~nrad, as derived in Eq.~\ref{eqn:helkick},
+the estimated dipole kick amplitude at the proton beam core is then
+
+\begin{equation}
+  \label{eqn:kick_central}
+  \Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \SI{20}{nrad}.
 \end{equation}
+
+For the scope of this paper, only the approximate magnitude of the
+residual dipole kick is considered. In general, the field map,
+including its multipolar components, can be parameterized in a
+symplectic form, as described for instance in
+Ref.~\cite{hel_bends_stancari}, and used in tracking codes to estimate
+the effects on the circulating beam.
 

### 2018-02-16 scripts/diffusion_illustration/diffusion_steady-state_1D.R
Added first version of R script that generates illustration of diffusion enhancement

@@ -1,0 +1,109 @@
+# Cartoon of steady-state beam density
+# given the diffusion coefficient
+
+for(HEBC in c(FALSE, TRUE)) {
+
+# aperture [beam sigmas]
+Jm <- 0; JM <- 9
+
+# collimator
+Jc <- 6; Jc2 <- JM
+
+# hollow electron beam
+Jh <- 4; Jh2 <- 7
+
+# diffusion coefficient
+hat <- function(x, x1=0, x2=1, s=0.1) {
+  return(pnorm(x, mean=x1, sd=s) - pnorm(x, mean=x2, sd=s)) }
+
+D.pow <- function(J, D0=0.01, Dc=0.5, P=2) {
+ D <- D0 + (Dc-D0)*(J/Jc)^P
+ if(HEBC) D <- D + Dc*hat(J, Jh, Jh2)
+ sub <- Jc < J
+ D[sub] <- NA
+ return(D)
+}
+
+D.test <- function(J, Nc=1, Nh=1/3, P=2) {
+ D <- J / (Nc*J*exp(-0.5*J^2)/sqrt(2*pi) + P*(P+1)*Nh*(2*Jc-J)^(P-1)/Jc^(P+1))
+ if(HEBC) D <- D*(1 + 2*hat(J, Jh, Jh2))
+ sub <- Jc < J
+ D[sub] <- NA
+ return(D)
+}
+
+D <- D.test
+
+# calculation of gradient and distribution
+df <- function(J, k=1){ k*J/D(J) }
+f.calc <- function(J){
+  I <- 0*J
+  for(i in 1:length(J)){
+    if(J[i]<Jc) I[i] <- integrate(df, J[i], Jc)$value }
+  return(I)}
+
+f.test <- function(J){
+    I <- 0*J
+    subset <- abs(J)<Jc
+    Nc <- 1; Nh <- 1; P <- 1
+    if(HEBC){ Nc <- 1.7; Nh <- 0.3 }
+    I[subset] <- Nc*exp(-0.5*J[subset]^2)/sqrt(2*pi) +
+        Nh*(P+1)*(Jc-J[subset])^P/Jc^(P+1)
+    return(I) }
+
+f <- f.calc
+
+# plot setup
+wi <- 3.25; he <- 3.25; pt <- 10; fo <- "Times"
+dev.new(width=wi, height=he, pointsize=pt, family=fo)
[…]

### 2018-02-17 scripts/diffusion_illustration/diffusion_steady-state_1D.R
Added script to generate diffusion illustration

@@ -1,109 +1,118 @@
-# Cartoon of steady-state beam density
-# given the diffusion coefficient
+# Cartoon of steady-state beam density given the diffusion coefficient
+# illustrating the effect of the hollow electron beam collimator
 
-for(HEBC in c(FALSE, TRUE)) {
-
+# amplitudes of various elements
 # aperture [beam sigmas]
-Jm <- 0; JM <- 9
-
+Jm = 0; JM = 9
 # collimator
-Jc <- 6; Jc2 <- JM
-
+Jc = 6; Jc2 = JM
 # hollow electron beam
-Jh <- 4; Jh2 <- 7
+Jh = 4; Jh2 = 7
 
-# diffusion coefficient
-hat <- function(x, x1=0, x2=1, s=0.1) {
+# smooth "hat" function, 0-1-0
+hat = function(x
+               , x1=0 # starting point
+               , x2=1 # end point
+               , s=0.05 # width of transition
+               ) {
   return(pnorm(x, mean=x1, sd=s) - pnorm(x, mean=x2, sd=s)) }
 
-D.pow <- function(J, D0=0.01, Dc=0.5, P=2) {
- D <- D0 + (Dc-D0)*(J/Jc)^P
- if(HEBC) D <- D + Dc*hat(J, Jh, Jh2)
- sub <- Jc < J
- D[sub] <- NA
- return(D)
+# shapes of diffusion coefficient vs. amplitude
+# power law
+D.pow = function(J # amplitude
+                 , D0=1e-2 # diffusion coeff. at zero amplitude
+                 , Dc=1 # diffusion coeff. at collimator
+                 , P=4 # power
+                 , Jlim=1 # amplitude at which diffusion coeff. starts to grow
+                 , HEBC=FALSE # hollow electron beam on?
+                 , hbs=3 # diffusion amplification due to hollow beam
+                 ) {
+  D = D0+0*J
+  sub = Jlim < J & J <= Jc
+  D[sub] = D0 + (Dc-D0)*((J[sub]-Jlim)/(Jc-Jlim))^P
+  if(HEBC) D = D + Dc*hat(J, Jh, Jh2)*hbs
+  sub = Jc < J
+  D[sub] = NA
+  return(D)
 }
-
[…]

### 2018-02-17 resex.tex
Added diffusion illustration and caption

@@
 
-\section{Introduction\label{sec:intro}}%force line break with \protect\\
-Considering past, current and future high energy and intensity colliders, each new machine has represented a considerable leap in stored beam energy with rising values for future accelerators and colliders (see Table~\ref{tab:stored_energy}).
+\section{Introduction}
+\label{sec:intro}
+
 \begin{table*}
@@
 \end{table*}
-Recent measurements at the LHC furthermore show that the tails of the transverse beam distribution are overpopulated compared to a Gaussian distribution resulting in a considerable amount of energy being stored in the beam tails alone.
-In case of the LHC explicitly around 5\% of the beam population is stored in the tails above 3.5~$\sigma$ compared to 0.22\% in case of a Gaussian distribution leading to 19~MJ of stored energy in the tails in case of nominal LHC parameters and 34~MJ in case of HL-LHC~\cite{helreview_valentino}.
-This leads to the conclusion that a mechanism is needed to deplete the beam tails in a controlled manner \cite{helreview}.
-
-The most obvious idea is to decrease the collimator gaps or scrape the tails with a collimator type device.
-This approach is however not feasible as it would generate unacceptably large loss spikes due to the high intensity halo population in the LHC.
-Most promising are methods, which increase the diffusion speed in the region of the halo particles resulting in a smooth and continuous removal of the high amplitude tails while leaving the core of the beam unperturbed.
-The diffusing halo particles are then intercepted by the collimation system and removed.
-This concept is also referred to as active halo control compared to a passive system, which only intercepts the halo particles without actively controlling the diffusion speed (see Fig.~\ref{fig:active_halo_control} for illustration).
-\begin{figure}[h]
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.7\linewidth]{passive_halo_control.png}
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.7\linewidth]{active_halo_control.png}
-	\end{minipage}	
-	\caption{\label{fig:active_halo_control} Sketch of passive halo control as with the collimation system (top) and active halo control using in addition for example a hollow electron lens (HEL) to control the diffusion speed in the region tail region without affected the beam core (bottom).}
-\end{figure}
+
+\begin{figure*}
+  \begin{minipage}[c]{\columnwidth}
+    \includegraphics[width=\columnwidth]{passive_halo_control.png}
+    \includegraphics[width=\columnwidth]{active_halo_control.png}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[c]{\columnwidth}
+  \includegraphics[width=\columnwidth]{diffusion_illustration}
+  \end{minipage}
+  \caption{Left: Sketch of passive halo control with a conventional
+    collimation system (top) and active halo control with the addition
+    of a hollow electron lens (bottom). Right: Illustration of a
+    simplified model of active diffusion enhancement in the transverse
+    plane. The diffusion coefficient as a function of amplitude
+    (orange) is enhanced in a specific amplitude region when the
+    hollow beam is turned on (from solid to dashed line). A
+    corresponding reduction in beam tail population (black) is created
+    (from solid to dashed line).}
+  \label{fig:active_halo_control} 
+\end{figure*}
+
+Considering past, current and future high energy and intensity
+colliders, each new machine has represented a considerable leap in
+stored beam energy with rising values for future accelerators and
+colliders (see Table~\ref{tab:stored_energy}).
+
+Recent measurements at the LHC furthermore show that the tails of the
+transverse beam distribution are overpopulated compared to a Gaussian
+distribution resulting in a considerable amount of energy being stored
+in the beam tails alone. In case of the LHC explicitly around 5\% of
+the beam population is stored in the tails above 3.5~$\sigma$ compared
+to 0.22\% in case of a Gaussian distribution leading to 19~MJ of
+stored energy in the tails in case of nominal LHC parameters and 34~MJ
+in case of HL-LHC~\cite{helreview_valentino}. This leads to the
+conclusion that a mechanism is needed to deplete the beam tails in a
+controlled manner \cite{helreview}.
+
+The most obvious idea is to decrease the collimator gaps or scrape the
+tails with a collimator type device. This approach is however not
+feasible as it would generate unacceptably large loss spikes due to
+the high intensity halo population in the LHC. Most promising are
+methods, which increase the diffusion speed in the region of the halo
+particles resulting in a smooth and continuous removal of the high
+amplitude tails while leaving the core of the beam unperturbed. The
+diffusing halo particles are then intercepted by the collimation
+system and removed. This concept is also referred to as active halo
+control compared to a passive system, which only intercepts the halo
+particles without actively controlling the diffusion speed (see
+Fig.~\ref{fig:active_halo_control} for illustration).
 

### 2018-02-18 resex.tex
Revised title and abstract

@@
-% ****** Start of file apssamp.tex ******
-%
-%   This file is part of the APS files in the REVTeX 4.1 distribution.
-%   Version 4.1r of REVTeX, August 2010
-%
-%   Copyright (c) 2009, 2010 The American Physical Society.
-%
-%   See the REVTeX 4 README file for restrictions and more information.
-%
-% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
-% as well as the rest of the prerequisites for REVTeX 4.1
-%
-% See the REVTeX 4 README file
-% It also requires running BibTeX. The commands are as follows:
-%
-%  1)  latex apssamp.tex
-%  2)  bibtex apssamp
-%  3)  latex apssamp.tex
-%  4)  latex apssamp.tex
-%
 \documentclass[%
@@
 
-%\title{Effects of pulsed hollow electron lens operation on the beam core in HL-LHC: \\First experimental studies and simulations.}% Force line breaks with \\
-\title{Effect of resonant and random excitation on the LHC beam in the context of pulsed hollow electron lens operation.}% Force line breaks with \\
+\title{Experimental and numerical studies of the effects of resonant
+  and random excitations of the proton beam in the LHC, with
+  applications to the design of a pulsed hollow electron lens for
+  active halo control}
+
 \thanks{Fermilab is operated by Fermi Research Alliance, LLC under
@@
  \email{mfittere@fnal.gov}
-\author{Giulio Stancari}%
-\author{Alexander Valishev}%
-\affiliation{Fermi National Accelerator Laboratory, Batavia, Illinois, USA
-% This line break forced with \textbackslash\textbackslash
-}%
+\author{Giulio Stancari}
+\author{Alexander Valishev}
+\affiliation{Fermi National Accelerator Laboratory, Batavia, Illinois, USA}
 
@@
 \author{Daniel Valuch}
-\affiliation{CERN, Geneva, Switzerland}%
-
-\date{\today}% It is always \today, today,
-             %  but any date may be explicitly specified
+\affiliation{CERN, Geneva, Switzerland}
+
+\date{\today}
 
 \begin{abstract}
-At the High Luminosity Large Hadron Collider (HL-LHC), a considerable amount of energy is expected to be stored in the beam tails due to the high beam intensity and, in addition, an overpopulation of the tails compared to a Gaussian distribution.
-To control and clean the tail population, the installation of two hollow electron lenses, one per beam, is presently under consideration.
-Beside the DC operation of the lenses, also pulsed operation is considered having the advantage to increase the diffusion speed by exerting white or colored noise on the halo particles.
-For an ideal electron lens, only the halo particles are excited while the core stays untouched.
-If a residual field is present at the location of the beam core, also the particles in the center are being exposed to noise.
-In this paper, we present
-a summary of the LHC experiments conducted in 2016 and 2017 to study the effects on the beam core in case of pulsed operation, where uniform random noise and a resonant excitation were studied.
-In both cases the excitation was generated with the transverse damper, which is capable of producing almost arbitrary colored noise spectra.
-In particular the resonant excitation features rich non-linear dynamics inducing unusual changes of the beam distribution and with it emittance and losses.
-These changes will be one of the main topics of this paper as they have not been observed and studied before in this detail.
-In addition to the presentation of the experimental and supportive simulation results, we also give first estimates of the residual field at the proton beam core expected from the HL-LHC hollow electron lens for comparison to the noise tolerances obtained in the experiments.
+  In the High Luminosity Large Hadron Collider (HL-LHC), a
+  considerable amount of energy will be stored in the beam tails, due
+  to the high beam intensity and to an overpopulation of the tails
+  compared to a Gaussian distribution. To control and clean the beam
+  halo, the installation of two hollow electron lenses (HEL), one per
+  beam, is being considered. In standard electron-lens operation, a
+  proton bunch sees the same electron current at every
+  revolution. Pulsed operation (i.e., different current for different
+  turns) is also considered, because it can widen the range of
+  achievable halo removal rates. For an axially symmetric electron
+  lens, only the halo particles are excited, leaving the core
+  unperturbed. If a residual field is present at the location of the
+  beam core, these particles are exposed to transverse kicks and to
+  noise. In this paper, we present the results of numerical
+  simulations and of the LHC experiments conducted in~2016 and~2017 to
+  study the effects of random noise and of resonant excitations on the
+  beam core. The excitation patterns were generated by the transverse
+  damper system, which is a very flexible source of transverse dipole
+  kicks. The main observables were losses, emittances, and transverse
+  distributions as a function of excitation patterns and
+  strengths. The resonant excitations showed rich nonlinear dynamics
+  and nontrivial changes of the beam distribution, which, to our
+  knowledge, have not previously been observed and studied in this
+  detail. We also give estimates, based upon existing measurements of
+  prototype electron guns, of the residual fields that may be expected
+  in the HL-LHC hollow electron lens.
 \end{abstract}

### 2018-02-18 resex.tex
Revised introduction. Added to-do list in README file.

@@
 \begin{table*}
-	\caption{\label{tab:stored_energy}%
-		Stored beam energy for different past, present and future colliders. Each new machine represents a leap in stored beam energy.
-	}
+  \caption{ Stored beam energy for different past, present and future
+    colliders. Each new machine represents a leap in stored beam
+    energy.}
+  \label{tab:stored_energy}
   \begin{ruledtabular}
     \begin{tabular}{lccccc}
-			Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{chamonix2017param}
-			& LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{fcc_param_2017} \\
+      Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{chamonix2017param}	& LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{fcc_param_2017} \\
       \colrule
@@
   \caption{Left: Sketch of passive halo control with a conventional
-    collimation system (top) and active halo control with the addition
-    of a hollow electron lens (bottom). Right: Illustration of a
-    simplified model of active diffusion enhancement in the transverse
-    plane. The diffusion coefficient as a function of amplitude
-    (orange) is enhanced in a specific amplitude region when the
-    hollow beam is turned on (from solid to dashed line). A
+    collimation system (top) and active halo control, with the
+    addition of a hollow electron lens (bottom). Right: Illustration
+    of a simplified model of active diffusion enhancement in the
+    transverse plane. The diffusion coefficient as a function of
+    amplitude (orange) is enhanced in a specific amplitude region when
+    the hollow beam is turned on (from solid to dashed line). A
     corresponding reduction in beam tail population (black) is created
@@
 
-Considering past, current and future high energy and intensity
-colliders, each new machine has represented a considerable leap in
-stored beam energy with rising values for future accelerators and
-colliders (see Table~\ref{tab:stored_energy}).
-
-Recent measurements at the LHC furthermore show that the tails of the
+Considering past, current and future high energy colliders, each new
+machine has represented a considerable leap in stored beam energy
+(Table~\ref{tab:stored_energy}).
+
+Furthermore, recent measurements at the LHC show that the tails of the
 transverse beam distribution are overpopulated compared to a Gaussian
-distribution resulting in a considerable amount of energy being stored
-in the beam tails alone. In case of the LHC explicitly around 5\% of
-the beam population is stored in the tails above 3.5~$\sigma$ compared
-to 0.22\% in case of a Gaussian distribution leading to 19~MJ of
-stored energy in the tails in case of nominal LHC parameters and 34~MJ
-in case of HL-LHC~\cite{helreview_valentino}. This leads to the
-conclusion that a mechanism is needed to deplete the beam tails in a
-controlled manner \cite{helreview}.
-
-The most obvious idea is to decrease the collimator gaps or scrape the
-tails with a collimator type device. This approach is however not
-feasible as it would generate unacceptably large loss spikes due to
-the high intensity halo population in the LHC. Most promising are
-methods, which increase the diffusion speed in the region of the halo
-particles resulting in a smooth and continuous removal of the high
-amplitude tails while leaving the core of the beam unperturbed. The
-diffusing halo particles are then intercepted by the collimation
-system and removed. This concept is also referred to as active halo
-control compared to a passive system, which only intercepts the halo
-particles without actively controlling the diffusion speed (see
-Fig.~\ref{fig:active_halo_control} for illustration).
-
-In a recent review, the need for such an active halo control system for HL-LHC has been assessed with the conclusion that it would considerably increase the margin and reduce the risk for machine protection compared to the current LHC collimation system being a passive halo control system ~\cite{helreview}.
-In view of the need of active halo control for HL-LHC and for future high power accelerators like HE-LHC and FCC-hh~\cite{helhcparam2011,fcc_coll_ipac2017}, different active halo control methods have been studied in recent years~\cite{helreview_bruce}, of which the hollow electron lens (HEL) is considered the superior device for the HL-LHC~\cite{helreview}.
-The beneficial effect of a future HEL for HL-LHC in terms of machine protection and beam collimation can not be, however, deployed at the expenses of the beam core as latter would ultimately lead to a degradation of the performance due to increased particle losses and emittance growth.
-In this paper, we will concentrate on the assessment in simulations and measurements of possible detrimental effets on the beam core.
-We will summarize possible sources of perturbations of the proton beam core focusing in particular on the case of pulsed operation and the beam experiments at the LHC conducted in this context.
-
-This paper is structured as follows: Section~\ref{sec:hel} gives an introduction to the concept of HELs and summarizes the design parameters of the HL-LHC HELs.
-Section~\ref{sec:core} is dedicated to describing the sources of a residual field from the HEL in the core region.
-Section~\ref{sec:exp} then describes in detail the two beam experiments conducted at the LHC studying the effect of pulsed operation on the proton beam core.
-
-\section{Hollow electron lens for HL-LHC\label{sec:hel}}
+distribution. This results in a considerable amount of energy being
+stored in the beam tails. In particular, in the case of the LHC, about
+5\% of the beam population is stored in the tails (i.e., above
+3.5$\sigma$, where $\sigma$ is the Gaussian rms core beam size),
+compared to 0.22\% in an ideal Gaussian distribution, leading to 19~MJ
+of stored energy for nominal LHC parameters and 34~MJ in the case of
+HL-LHC~\cite{helreview_valentino}. This leads to the conclusion that a
+mechanism is needed to deplete the beam tails in a controlled
+manner. Further information on the needs for halo control in LHC can
+be found in Ref.~\cite{helreview}.
+
+The most direct approach is to decrease the collimator gaps or to
+periodically scrape the tails. However, this is not feasible, as it
+would generate unacceptably large loss spikes and possibly component
+damage. Most promising are methods which increase the diffusion speed
+in the region of the halo particles, resulting in a smooth and
+continuous removal of the high amplitude tails, while leaving the core
+of the beam unperturbed. The diffusing halo particles are then
+intercepted by the collimation system and removed. This concept is
+also referred to as active halo control, designed to enhance a
+conventional passive system, which is still needed to robustly
+intercept the halo particles. An illustration of the concept is shown
+in Fig.~\ref{fig:active_halo_control}.
+
+In a recent review, the need for such an active halo control system
+for HL-LHC has been assessed, with the conclusion that it would
+considerably increase the operational margins and reduce the risks for
+machine protection~\cite{helreview}. In view of the need of active
+halo control for HL-LHC and for future high power accelerators, like
+HE-LHC and FCC-hh~\cite{helhcparam2011, fcc_coll_ipac2017}, different
+active halo control methods have been
+studied~\cite{helreview_bruce}. The hollow electron lens (HEL) is
+considered the most established, flexible and suitable technology for
+the HL-LHC~\cite{helreview}.
+
+However, the beneficial effects of an HL-LHC HEL for machine
+protection and for collimation cannot come at the expense of
+performance degradation due to losses or emittance growth in the beam
+core. In standard electron-lens operation, a proton bunch sees the
+same electron current at every revolution. It is also possible to have
+different currents for different groups of bunches. Under these
+conditions, the imperfections of the hollow beam have a negligible
+effect. On the other hand, in order to extend the range of achievable
+removal rates, pulsed operation is also being considered. In this
+case, different currents can be set to act on the same bunch at each
+turn. If a residual field is present at the location of the beam core,
+core particles can be exposed to resonant transverse kicks and to
+noise.
+
+In this paper, we concentrate on the experimental and numerical
+assessment of possible detrimental effects on the beam core of a
+pulsed electron lens. Section~\ref{sec:hel} gives an introduction to
+the concept of HELs and summarizes the design parameters of the HL-LHC
+HELs. Section~\ref{sec:core} is dedicated to the sources of residual
+fields from the HEL in the core region. Section~\ref{sec:exp}
+describes the setup of simulations and experimental
+conditions. Results and comparisons between simulations and
+measurements are given in Section~\ref{sec:simex}, with discussion and
+summary in Section~\ref{sec:sum}. Auxiliary data and plots are
+collected in the Appendices.
+
+\section{Hollow electron lens for HL-LHC}
+\label{sec:hel}
+
 \subsection{General Overview\label{sec:hel:intro}}

### 2018-02-19 resex.tex
Started revision of HEL section and of bibliography. Renamed bibliography (now managed with BibDesk).

@@
 \usepackage{bm}% bold math
-%\usepackage{hyperref}% add hypertext capabilities
+\usepackage{hyperref}% add hypertext capabilities
 %\usepackage[mathlines]{lineno}% Enable numbering of text and display math
@@
 considered the most established, flexible and suitable technology for
-the HL-LHC~\cite{helreview}.
+the HL-LHC~\cite{hel_tevatron_stancari, helreview}.
 
@@
 
-\subsection{General Overview\label{sec:hel:intro}}
-Electron lenses produce DC or pulsed low energy electron beams, where the electron beam is generated in an electron gun, then guided and confined along a straight section by strong solenoids and finally dumped onto a collector.
-As an example the conceptual design of the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
-\begin{figure}[h]
-	\includegraphics[width=1.0\linewidth]{hel_layout_epbeam}% Here is how to import EPS art
-	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL. Courtesy of CERN EN-MME group.}
-\end{figure}
+\subsection{General Overview}
+\label{sec:hel:intro}
+
+Electron lenses are based upon continuous or pulsed low energy,
+magnetically confined electron
+beams~\cite{Shiltsev:elens-book:2016}. The electron beam is generated
+in an electron gun, guided and confined by strong solenoids and
+finally dumped in a collector. As an example, the conceptual design of
+the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
+
+\begin{figure*}
+  \includegraphics[width=0.75\textwidth]{hel_layout_epbeam}
+  \caption{Layout of the hollow electron lens for HL-LHC. (Courtesy of
+    CERN EN-MME mechanical engineering group.)}
+  \label{fig:hel_layout}
+\end{figure*}
 
@@
 
-\subsection{Operation modes and effects on the beam core\label{sec:hel:core}}
-For the HEL two modes of operation are currently under consideration, the DC mode as standard operation mode and the pulsed mode.
-The main benefit of pulsed HEL operation is the increase in halo removal rates which might become essential as in ideal operation conditions featuring low non-linearities, in particular low chromaticity and octupole current, the DC mode is not effective~\cite{hel_halo_hllhc_fitterer,hl_halo_ipac2017}.
-For this purpose, two different pulsing patterns are thus considered for the HL-LHC:
+\subsection{Operation modes and effects on the beam core}
+\label{sec:hel:core}
+
+For the HEL two modes of operation are currently under consideration,
+the continuous (also referred to as `DC', or direct current, in this
+paper) mode as standard operation mode and the pulsed mode. The main
+benefit of pulsed HEL operation is the increase in halo removal rates
+which might become essential as in ideal operation conditions
+featuring low non-linearities, in particular low chromaticity and
+octupole current, the DC mode is not
+effective~\cite{hel_halo_hllhc_fitterer,hl_halo_ipac2017}.
+
+For this purpose, two different pulsing patterns are thus considered
+for the HL-LHC:
+
 \begin{itemize}
@@
 
-\newpage
-
-%\section{Additional plots for random excitation during 2017 experiments}
-%\label{app:sec:ran}
-
 \clearpage
 
-\bibliography{bibliography}% Produces the bibliography via BibTeX.
+\bibliography{resex-bibdesk}
 
 \end{document}
-%
-% ****** End of file apssamp.tex ******

### 2018-02-19 resex.tex
Intermediate step in revising Section II on HEL

@@
 
-\subsection{General Overview}
-\label{sec:hel:intro}
-
-Electron lenses are based upon continuous or pulsed low energy,
-magnetically confined electron
-beams~\cite{Shiltsev:elens-book:2016}. The electron beam is generated
-in an electron gun, guided and confined by strong solenoids and
-finally dumped in a collector. As an example, the conceptual design of
-the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
-
 \begin{figure*}
@@
 
-\begin{figure}[b]
-	\includegraphics[width=0.8\linewidth]{kick_hel_lhc_no_grid}% Here is how to import EPS art
-	\caption{\label{fig:hel_field} Illustration of the hollow electron beam distribution (blue), the kick experienced by the proton beam (red) and the collimators (gray).}
+\begin{figure}
+  \includegraphics[width=0.8\columnwidth]{kick_hel_lhc_no_grid}
+  \caption{Illustration of the hollow electron beam charge
+    distribution (blue), of the magnitude of the transverse kick
+    experienced by the proton beam (red), and of the position of the
+    primary collimators (gray).}
+  \label{fig:hel_field}
 \end{figure}
-The circulating beam, in case of the LHC the proton beam, is then affected by the electromagnetic field of the electron beam.
-For the application of active halo control, the electron beam needs to generate an electromagnetic field at the location of the halo particles while leaving the core unperturbed.
-This field distribution can be achieved by using a uniform hollow distribution in radius $r=\sqrt{x^2+y^2}$ with inner radius $R_1$ and outer radius $R_2$.
-In this case, the circulating proton beam experiences a radial kick $\theta(r)$
-\begin{equation}\label{eq:field_1}
-\theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max},
-\end{equation}
-where $f(r)$ is a shape function with
-\begin{equation}\label{eq:field_2}
-f(r) =
-\begin{cases} 0 &,\quad r< R_1,\\
-\frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2,\\
-1 &,\quad R_2 \leq r
-\end{cases}
-\end{equation}
-and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle given by
-\begin{equation}\label{hel_kick_max}
-\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{p_0}{q}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2},
-\end{equation}
-with $L$ the length of the HEL, $I_T$ the total electron beam current, $\beta_{e}$ and $\beta_{p}$ the relativistic $\beta$ of electron and proton beam, $\frac{q}{p_0}=\left(B\rho\right)_p$ the magnetic rigidity for the proton beam reference particle, $c$ the speed of light and $\epsilon_0$ the vacuum permittivity.
-The distribution together with the resulting kick are illustrated in Fig.~\ref{fig:hel_field}.
-The $\pm$-sign in Eq.~\ref{hel_kick_max} represents the two cases of the electron beam traveling in the direction of the proton beam ($v_e v_p>0$) leading to ``$-$" or in the opposite direction ($v_e v_p<0$) leading to ``$+$".
-For hollow electron beam collimation, the beams are chosen to counterrotate.
-Assuming HL-LHC and HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) the maximum kick of the HEL is:
-\begin{equation}\label{eqn:helkick}
-\theta_{\rm max,B1} = 392~\rm{nrad}
-\end{equation}
-%\begin{eqnarray}
-%\theta_{\rm max,B1} = 392~\rm{nrad},\\
-%\theta_{\rm max,B2} = 341~\rm{nrad}.
-%\end{eqnarray}
-for an inner radius of $R_1=4\sigma_p$, outer radius $R_2=7.5\sigma_p$, peak current of $I_e=$\SI{5.0}{A} and Beam~1 of the LHC. Similar values are obtained for Beam~2.
-\begin{table}[t]
-	\caption{\label{tab:hllhc_param}%
-		HL-LHC design parameters at top energy \cite{hlcdr} and parameters relevant in connection with the HEL.
-Optics parameters at HEL are based on a position of the HEL of $-40$~m for Beam~1 (B1) and $+40$~m for Beam~2 (B2) from IP4 using HL-LHC optics V1.3 with $\beta^{*}=0.15$~m \cite{hlv13}.
-	}
+
+\begin{table}
+  \caption{HL-LHC design parameters at top energy~\cite{hlcdr} and
+    parameters relevant to the HEL. Optics parameters at the HEL are
+    based on a position of $-$40~m for Beam~1 (B1) and $+$40~m for
+    Beam~2 (B2) from the interaction point IP4, using HL-LHC optics
+    V1.3 with $\beta^{*} = \q{0.15}{m}$~\cite{hlv13}.}
+  \label{tab:hllhc_param}
   \begin{ruledtabular}
@@
 \end{table}
-\begin{table}[t]
-	\caption{\label{tab:hel_param}%
-		HL-LHC hollow electron lens parameters as in~\cite{hel_cdr}.
-	}
+
+\begin{table}
+  \caption{HL-LHC hollow electron lens parameters as defined in
+    Ref.~\cite{hel_cdr}.}
+  \label{tab:hel_param}
   \begin{ruledtabular}
@@
 \end{table}
+
+\subsection{General Overview}
+\label{sec:hel:intro}
+
+Electron lenses are based upon continuous or pulsed low-energy,
+magnetically confined electron beams~\cite{Shiltsev:PRSTAB:2008,
+  Fischer:PRL:2015, Shiltsev:elens-book:2016}. The electron beam is
+generated in an electron gun, guided and confined by strong solenoids
+and finally dumped in a collector. As an example, the conceptual
+design of the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
+
+The circulating beam (protons in the LHC case) is affected by the
+electromagnetic field of the electron beam. For the application of
+active halo control, the electron beam needs to generate an
+electromagnetic field only at the location of the halo particles. This
+field distribution can be achieved, for instance, by using a hollow
+charge distribution in radius $r=\sqrt{x^2+y^2}$, uniformly
+distributed between inner radius $R_1$ and outer radius $R_2$
+(Fig.~\ref{hel_field}). In this case, the circulating proton beam
+experiences the following radial kick $\theta(r)$:
+%
+\begin{equation}
+  \label{eq:field_1}
+  \theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max},
+\end{equation}
+%
+where $f(r)$ is a shape function with
+%
+\begin{equation}
+  \label{eq:field_2}
+  f(r) =
+  \begin{cases}
+    0 &,\quad r< R_1,\\
+    \frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2,\\
+    1 &,\quad R_2 \leq r
+  \end{cases}
+\end{equation}
+%
+and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle given by
+%
+\begin{equation}
+  \label{hel_kick_max}
+  \theta_{\rm max} = \theta(R_2) =
+  \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot
+    \left(p_0/q\right)_p \cdot \beta_e\beta_p
+    c^2}\cdot\frac{1}{R_2},
+\end{equation}
+%
+with $L$ the length of the HEL, $I_T$ the total electron beam current,
+$\beta_{e}$ and $\beta_{p}$ the relativistic velocity parameters of
+electrons and protons, $\left(p_0/q\right)_p = \left(B\rho\right)_p$
+the magnetic rigidity for the proton beam reference particle, $c$ the
+speed of light and $\epsilon_0$ the vacuum permittivity. The
+$\pm$-sign in Eq.~\ref{hel_kick_max} represents the two cases of the
+electron beam traveling in the direction of the proton beam
+($v_e v_p>0$) leading to ``$-$" or in the opposite direction
+($v_e v_p<0$) leading to ``$+$". For hollow electron beam collimation,
+electrons and protons are chosen to counterrotate, so that the
+magnetic and electric kicks add up. (For simplicity, the dependence of
+the electron axial velocity on radius is neglected in
+Eq.~\ref{hel_kick_max}.)
+
+In the case of HL-LHC HEL design parameters
+(Table~\ref{tab:hllhc_param} and Table~\ref{tab:hel_param}), the
+maximum kick is:
+%
+\begin{equation}
+  \label{eqn:helkick}
+  \theta_{\rm max,B1} = 392~\rm{nrad}
+\end{equation}
+%
+for an inner radius of $R_1 = 4\sigma_p$, outer radius $R_2 =
+7.5\sigma_p$, peak current of $I_e = \q{5.0}{A}$, using the Beam~1
+lattice of the LHC. Similar values are obtained for Beam~2.
 

### 2018-02-19 resex.tex
Revised Section II A on HEL

@@
     \begin{tabular}{lccc}
-      Beam parameters & Value(B1) & Value(B2) & Unit\\
+      Beam parameters & \multicolumn{2}{c}{Value} & Unit\\
+                       & B1 & B2 & \\
       \colrule
-      Beam energy  $E_{p}$  &  \multicolumn{2}{c}{7} & TeV\\
-      Number of bunches $n_b$ & \multicolumn{2}{c}{2748} & - \\
-      Number of particles per bunch $N_b$ & \multicolumn{2}{c}{$2.2\times 10^{11}$} & -\\
-      Normalized emittance $\epsilon_{N,x/y}$ & \multicolumn{2}{c}{2.5} & $\mu$m\\
+      Beam energy,  $E_{p}$  &  \multicolumn{2}{c}{7} & TeV\\
+      Number of bunches, $n_b$ & \multicolumn{2}{c}{2748} &  \\
+      Bunch population, $N_b$ & \multicolumn{2}{c}{$2.2\times 10^{11}$} & \\
+      Normalized emittance, $\epsilon_{N,x/y}$ & \multicolumn{2}{c}{2.5} & $\mu$m\\
       Bunch spacing & \multicolumn{2}{c}{25} & ns\\
       \colrule
-      \multicolumn{4}{l}{Optics paramters at HEL (Beam~1) \footnote{As the Twiss parameters at IP4 do not change during the entire squeeze, and IP4 and the HEL are only separated by a drift space, the Twiss parameters stay constant also at the HEL during the entire squeeze.}} \\
+      \multicolumn{4}{l}{Optics parameters at HEL (B1)\footnote{As the
+      Twiss parameters at IP4 do not change during the entire squeeze
+      of the optical functions, and IP4 and the HEL are only separated
+      by a drift space, the Twiss parameters stay constant also at the
+      HEL during the squeeze.}} \\
       \colrule
       $\beta_{x}$ at HEL  & 197.5 & 280.6 & m\\
-      $\beta_{y}$ at HEL & 211.9 & 262.6 m\\
+      $\beta_{y}$ at HEL & 211.9 & 262.6 & m\\
       Dispersion $D_{x}$ at HEL & 0.0& 0.0 & m\\
@@
       Proton beam size $\sigma_{p,y}$ at HEL & 0.27 & 0.30 &mm \\
-      \multicolumn{4}{l}{scraping position}\\ \hspace{1cm}$\sigma_{p}=\max(\sigma_{p,x},\sigma_{p,y})$ & 0.27& 0.31 & mm\\
+      \multicolumn{4}{l}{Scale of scraping positions}\\ \hspace{6ex}$\sigma_{p}=\max(\sigma_{p,x},\sigma_{p,y})$ & 0.27& 0.31 & mm\\
     \end{tabular}
@@
 \begin{table}
-  \caption{HL-LHC hollow electron lens parameters as defined in
+  \caption{HL-LHC hollow electron lens parameters, as defined in
     Ref.~\cite{hel_cdr}.}
@@
       \colrule
-      Length $L$    &  3 & m\\
+      Length, $L$    &  3 & m\\
       % 3-8 sigmap with 3.5 um -> 3.55-9.47 for 2.5 um emittance
-      Desired range of scraping positions & 3.5-9.5 &$\sigma_p$\\
+      Desired range of scraping positions & 3.5--9.5 &$\sigma_p$\\
       \colrule
@@
       \colrule
-      Gun solenoid, $B_g$ & 0.2-0.4 & T\\
-      Main solenoid, $B_m$ & 2-6 & T\\
-      Collector solenoid, $B_c$ & 0.2-0.4 & T\\
-      Compression factor ($k=\sqrt{B_m/B_g}$) & 2.2-5.5 & -\\
+      Gun solenoid, $B_g$ & 0.2--0.4 & T\\
+      Main solenoid, $B_m$ & 2--6 & T\\
+      Collector solenoid, $B_c$ & 0.2--0.4 & T\\
+      Compression factor, $k \equiv \sqrt{B_m/B_g}$ & 2.2--5.5 &  \\
       \colrule
@@
       Peak yield $I_e$ at 10~keV & 5.0 & A\\
-      Gun perveance $P$ & 5 & $\mu$perv\\
-      Inner/outer cathode radius, $R_1/R_2$ & 6.75/12.7 & mm\\
+      Gun perveance, $P$ & 5 & $\mu$perv\\
+      Inner/outer cathode radii, $R_1/R_2$ & 6.75/12.7 & mm\\
       \colrule
-      High voltage modulator & & \\
+      High-voltage modulator & & \\
       \colrule

### 2018-02-20 resex.tex
Revised Section II on HEL

@@
 
-\subsection{General Overview}
+\subsection{General overview}
 \label{sec:hel:intro}
@@
 distributed between inner radius $R_1$ and outer radius $R_2$
-(Fig.~\ref{hel_field}). In this case, the circulating proton beam
+(Fig.~\ref{fig:hel_field}). In this case, the circulating proton beam
 experiences the following radial kick $\theta(r)$:
@@
 
-For the HEL two modes of operation are currently under consideration,
-the continuous (also referred to as `DC', or direct current, in this
-paper) mode as standard operation mode and the pulsed mode. The main
-benefit of pulsed HEL operation is the increase in halo removal rates
-which might become essential as in ideal operation conditions
-featuring low non-linearities, in particular low chromaticity and
-octupole current, the DC mode is not
-effective~\cite{hel_halo_hllhc_fitterer,hl_halo_ipac2017}.
-
-For this purpose, two different pulsing patterns are thus considered
-for the HL-LHC:
+For the HEL, two modes of operation are currently under consideration:
+the continuous mode (also referred to as `DC', or direct current, in
+this paper) as standard operation mode, described above; and the
+pulsed mode. The main benefit of pulsed HEL operation is the increase
+in halo removal rates. A wider range of removal rates may become
+important under operating conditions with small nonlinearities, in
+particular low chromaticity and octupole current, when the DC mode may
+be too slow~\cite{hel_halo_hllhc_fitterer, hl_halo_ipac2017}.
+
+Two different pulsing patterns are considered for the HL-LHC. In both
+cases, at each passage, a given bunch sees a different electron-lens
+current and, therefore, experiences a different transverse kick. The
+patterns are defined as follows:
 
 \begin{itemize}
-	\item \textbf{random excitation:} The electron gun voltage is modulated between:
+
+\item \textbf{random excitation:} The extraction voltage in the
+  electron gun is modulated according to the
+  following expression:
+%
+  \begin{equation}
+    U_{\mathrm{e-gun}} = (1-a) \cdot U_{\mathrm{max}} +
+    a \cdot \eta \cdot U_{\mathrm{max}},
+  \end{equation}
+%
+  where $U_{\mathrm{max}}$ is the maximum voltage, $a$ is the
+  modulation strength, with $a \in [0,1]$, and $\eta$ is a uniformly
+  distributed random number in the interval~$[0,1]$. Simulations and
+  experiments, discussed below, were usually conducted with $a = 1$.
+
+\item \textbf{resonant excitation:} The electron beam is switched on
+  only every $k^{\mathrm{th}}$ turn. The excitation can be represented
+  by the following expression:
   \begin{eqnarray}
-	U_{\mathrm{e-gun}}&=&a\cdot U_{\mathrm{max}}\\
-	& &+(1-a)\cdot \mathrm{ran}(0,1)\cdot U_{\mathrm{max}}
+    \label{intro:eqn:1}
+    f(t) & = & \sum_{n=-\infty}^{+\infty}\delta\left[t-n\cdot(kT)\right],
   \end{eqnarray}
-	with $U_{\mathrm{max}}$ the maximum voltage in DC operation, $a$ the modulation strength with $a\in[0,1]$, and $\mathrm{ran}(0,1)$ a uniformly distributed random number between~[0,1].
-	\item \textbf{resonant excitation:} The HEL is switched on only every $k^{\mathrm{th}}$ turn. The excitation can then be represented by:
-	\begin{eqnarray}\label{intro:eqn:1}
-	f(t)&=&\sum_{n=-\infty}^{+\infty}\delta(t-n\cdot(kT)),
-	\end{eqnarray}
-	where $n$ is the turn number and $T$ the revolution time, and its Fourier series by:
-	\begin{eqnarray}\label{intro:eqn:2}
-	f(t)=\Sh_{kT}(t)
-	&=&\frac{1}{kT}\sum_{n=-\infty}^{+\infty}e^{2\pi i f_nt} \\
-	& & \quad \text{with} \ f_n=\frac{n}{k}f_{\rm rev},
-	\end{eqnarray}
-	where $\Sh_{kT}$ is the Dirac comb.
-As can be seen from the Fourier series, $k^{\mathrm{th}}$~turn pulsing in general drives $k^{\mathrm{th}}$ order resonances \cite{md_sim_hel_res_ex_fitterer}.
-Historically, the $k^{\mathrm{th}}$~turn pulsing was used in regular operation in the Tevatron for abort cleaning \cite{hel_tevatron_abortgap_zhang}.
+  where $n$ is the turn number and $T$ is the revolution period. Its
+  Fourier representation is
+  \begin{equation}
+    \label{intro:eqn:2}
+    f(t) = \Sh_{kT}(t) = \frac{1}{kT}\sum_{n=-\infty}^{+\infty}e^{2\pi i f_nt},
+  \end{equation}
+  with $f_n = n \cdot f_{\rm rev} / k$ and
+  where $\Sh_{kT}$ is the Dirac comb. In general,
+  $k^{\mathrm{th}}$~turn pulsing drives $k^{\mathrm{th}}$-order
+  resonances~\cite{md_sim_hel_res_ex_fitterer}. This type of pulsing
+  pattern was used in the Tevatron during regular collider operations
+  for abort-gap cleaning~\cite{hel_tevatron_abortgap_zhang}.
 \end{itemize}
 
-As depicted in Fig.~\ref{fig:hel_field} and in evidence from Eqs.~\ref{eq:field_1}--\ref{eq:field_2}, the field at the beam core vanishes in case of an ideal HEL.
-Effects on the beam core thus only arise from imperfections, where possible sources are the bends of the HEL (see Section~\ref{core:sec:1}), as here the electron beam crosses directly the proton beam, and distortions in the electron beam profile (see Section~\ref{core:sec:2}).
-Both sources result in non-linear kicks~\cite{hel_bends_stancari,hel_model_polynomial_morozov}.
-In DC operation, these non-linear kicks stay in the shadow of the non-linearities otherwise present in the machine if the electron beam is kept at a decent quality.
-Tolerances on imperfections are therefore not particularly stringent and of no concern.
-The picture changes significantly in case of pulsed operation, in which case the electron gun voltage is modulated using a white or colored noise spectrum.
-If the electromagnetic field of the HEL does not vanish at the proton beam core, noise is transferred not only to the halo particles, as intended, but also to the beam core.
-Tolerances on the residual field in this case rapidly become much more stringent than in the case of DC operation.
-Studies of the effect of the HEL on the beam core therefore focus on this mode of operation which will also be the main subject of this paper.
-
-%In comparison, the random excitation mode is the most effective and operationally easiest pulsing mode, but also the most stringent in terms of tolerances on the residual field at the proton beam center as it induces white noise exciting all beam frequencies.
-As the resonant excitation acts on specific resonances, the pulsing pattern has to be chosen according to the machine and beam parameters.
-On the one hand, this makes the operation more difficult as one has to find the most efficient $k^{\mathrm{th}}$~turn pulsing for each operation mode, on the other hand, it also relaxes the tolerances on the residual field at the beam core as, depending on the machine and beam configuration, it might be possible to find a resonant excitation pattern, which mainly affects the halo and not the beam core.
+For an axially symmetric electron lens, the field at the beam core
+vanishes. Effects on the proton core arise from imperfections, with
+two main sources: the injection and extraction bends of the HEL
+(discussed below in Section~\ref{core:sec:1}), where the electron beam
+crosses the proton beam; and distortions of the electron beam profile
+during its propagation under magnetic focusing and space charge
+(Section~\ref{core:sec:2}). Both sources result in nonlinear
+kicks~\cite{hel_bends_stancari, hel_model_polynomial_morozov}. In
+continuous operation, these nonlinear kicks are usually much smaller
+than the machine nonlinearities. Tolerances on imperfections are
+therefore not particularly stringent. The picture changes
+significantly in case of pulsed operation. If the electromagnetic
+field does not vanish at the proton beam core, noise or resonant kicks
+are transferred not only to the halo particles, as intended, but also
+to the beam core. Tolerances on the residual fields in this case
+become much more stringent. Studies of the effects of the HEL on the
+beam core therefore focus on this mode of operation, which is also the
+main subject of this paper.
 

### 2018-02-21 resex.tex
bibliography is fixed
Removed package siunitx. Started revision of Section III on residual kicks.

@@
 \usepackage{color}
-\usepackage{siunitx}
+%\usepackage{siunitx}
 \usepackage{url}
@@
 
-\section{Sources of residual field in the proton beam core region and first order estimates\label{sec:core}}
-With the current HEL layout as shown in Fig.~\ref{fig:hel_layout}, parasitic kicks on the proton beam core can arise in the central region due to profile imperfections in the electron beam and at the entrance and exit of the HEL, where electron and proton beam overlap.
+\section{Sources and estimates of residual fields}
+\label{sec:core}
+
+Parasitic kicks on the proton beam core can be due to electron-beam
+profile imperfections in the overlap region, or to the injection and
+extraction toroidal bends, where electrons and protons overlap, as
+shown in the layout of Fig.~\ref{fig:hel_layout}.
+
 Currently, no HEL is installed in the LHC, and the kick on the proton beam core must therefore be emulated with the help of other devices already installed in the LHC.
@@
 \begin{equation}
-	\Delta x'_{\rm{bends}},\Delta y'_{\rm{bends}}\approx \SI{0.5}{nrad},
+	\Delta x'_{\rm{bends}},\Delta y'_{\rm{bends}}\approx \q{0.5}{nrad},
 \end{equation}
@@
 \begin{equation}
-	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \SI{20}{nrad} % this should be rounded to the same number of digits as the 0.5nrad
+	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \q{20}{nrad} % this should be rounded to the same number of digits as the 0.5nrad
 \end{equation}
@@
 \end{equation}
-%given an \SI{1}{A}, \SI{5}{keV} electron beam.
-%For \SI{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
+%given an \q{1}{A}, \q{5}{keV} electron beam.
+%For \q{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
 %\begin{equation}
-%\Delta x'=\Delta y'= \SI{1.4}{nrad},
+%\Delta x'=\Delta y'= \q{1.4}{nrad},
 %\end{equation}
-%Scaling to the HEL design parameters of a \SI{5}{A} and \SI{10}{keV} electron beam (Table~\ref{tab:hel_param}), one obtains:
+%Scaling to the HEL design parameters of a \q{5}{A} and \q{10}{keV} electron beam (Table~\ref{tab:hel_param}), one obtains:
 %\begin{equation}
@@
   \label{eqn:kick_central}
-  \Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \SI{20}{nrad}.
+  \Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \q{20}{nrad}.
 \end{equation}

### 2018-02-22 resex.tex
fixed merge errors for bibliography
Revised section III A on residual kicks.

@@
 
-Currently, no HEL is installed in the LHC, and the kick on the proton beam core must therefore be emulated with the help of other devices already installed in the LHC.
-For this purpose the kick is approximated to first order by a dipole kick.
-During the experiments presented in this paper, the dipole kick with the corresponding excitation pattern could then be applied with the LHC transverse damper as will be described in detail in Section~\ref{sec:adt}.
-As will be shown in Sec.~\ref{core:sec:1} and \ref{core:sec:2}, the expected kick to first order from the HEL bends is estimated to be~(Eq.~\ref{eqn:kick_bends}):
+Because no HEL is currently installed in the LHC, the kicks on the
+proton beam core must be emulated by other devices to determine their
+effects experimentally in a given machine and to guide design and
+tolerances. In particular, during the experiments presented in this
+paper, the LHC transverse damper system (described in detail in
+Section~\ref{sec:adt}) was used for this purpose. This system can
+generate transverse dipole kicks with a wide range of excitation
+patterns.
+
+For comparison with experiments and simulations, here we estimate to
+first order the magnitude of the dipole kicks that may be expected
+from the HEL. As one can see below, the contribution from the central
+region is in general dominating.
+
+\subsection{Uncompensated kicks from injection and extraction bends}
+\label{core:sec:1}
+
+To estimate the dipole component of the kicks that originate from the
+injection and extraction bends, we used the approach described in
+Ref.~\cite{hel_bends_stancari}. The bends are modeled as a bent
+cylindrical pipe with a static charge distribution of electrons. The
+resulting electric field is calculated, with the vacuum pipe as
+boundary, using the solvers of the Warp particle-in-cell
+code~\cite{warp}. (The contribution of the magnetic field generated by
+the electron current was neglected in this study.) The field,
+integrated over the trajectory of the protons, is then translated into
+a symplectic kick map.
+
+In case of a U-shaped electron lens, where electron gun and collector
+are on the same side, the transverse (and possibly pulsed) dipole
+kicks generated by the electron charge at the entrance and exit add
+up. For an S-shaped electron lens, on the other hand, where gun and
+collector are on opposite sides, these kicks compensate each
+other. For this reason, an S-shape was chosen for the HL-LHC HEL
+(Fig.~\ref{fig:hel_layout}). A disadvantage of the S-shape is that the
+static magnetic kicks due to the toroidal sections do add up, but they
+can be compensated by conventional dipole correctors, especially in a
+high-energy machine.
+
+In the case under study of an S-shaped electron lens, residual
+uncompensated kicks arise from differences in electron charge
+distribution between the injection and extraction bends. Here we
+conservatively assume 10\% fluctuations between the entrance and exit
+and, furthermore, that these differences add up.
+
+The maximum values of the integrated electric fields calculated in
+Ref.~\cite{hel_bends_stancari}, based upon an electron beam of 1~A at
+5~keV, are
+%
 \begin{equation}
-	\Delta x'_{\rm{bends}},\Delta y'_{\rm{bends}}\approx \q{0.5}{nrad},
+  \int_{z_1}^{z_2} E_{x,y} \, dz= \q{10}{kV}.
 \end{equation}
-and for the central region due to profile imperfections~(Eq.~\ref{eqn:kick_central}):
+%
+Scaling to HL-LHC and HEL design parameters
+(Table~\ref{tab:hllhc_param} and Table~\ref{tab:hel_param}) yields the
+following integrated field and corresponding kick:
+%
 \begin{equation}
-	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \q{20}{nrad} % this should be rounded to the same number of digits as the 0.5nrad
+  \int_{z_1}^{z_2} E_{y} \, dz = \q{36}{kV} \Rightarrow \Delta y' = \q{5.1}{nrad},
 \end{equation}
-where the current HEL design parameters are used (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}). The contribution from the central region is clearly dominating.
-
-\subsection{Uncompensated kicks from HEL bends}
-\label{core:sec:1}
-As reference for the estimate of the dipole component originating from the HEL bends, the symplectic map derived in \cite{hel_bends_stancari} is used.
-In this case the HEL bends are modeled as a bent cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons deflected in the horizontal plane, where the parameters were her still based on the old gun design for the Tevatron electron lens \textcolor{red}{Giulio, is this correct?}.
-In case of an U-shaped HEL, the transverse dipole kicks at entrance and exit add up, while for a S-shape the kicks compensate each other.
-For this reason a S-shape has been chosen for the HL-LHC HEL (see Fig.~\ref{fig:hel_layout}).
-Assuming and S-shaped HEL, uncompensated kicks therefore only arise from imperfections, which can originate from profile imperfections and current fluctuations.
-As a first estimate, we assume in this paper 10\% fluctuations between the entrance and exit and, in addition, that the kicks from entrance and exit due to these imperfections add up.
-Using the electric field calculations in \cite{hel_bends_stancari}, the maximum values for the integrated electric field are then
+%
+as described in detail in
+Ref.~\cite{md_sim_hel_res_ex_fitterer}. Assuming a residual difference
+of 10\% between entrance and exit, the expected kick is approximately
+%
 \begin{equation}
-\int_{z_1}^{z_2} E_{x,y} dz= 10 \ \mathrm{kV}.
+  \label{eqn:kick_bends}
+  \Delta x', \Delta y' = \q{0.5}{nrad}.
 \end{equation}
-%given an \q{1}{A}, \q{5}{keV} electron beam.
-%For \q{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
-%\begin{equation}
-%\Delta x'=\Delta y'= \q{1.4}{nrad},
-%\end{equation}
-%Scaling to the HEL design parameters of a \q{5}{A} and \q{10}{keV} electron beam (Table~\ref{tab:hel_param}), one obtains:
-%\begin{equation}
-%\int_{z_1}^{z_2} E_{x} dz= 36 \ \mathrm{kV} \Rightarrow \Delta x'= \q{5.1}{nrad}
-%\end{equation}
-Neglecting magnetic effects and scaling to HL-LHC and HEL design parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}) yields a kick of \cite{md_sim_hel_res_ex_fitterer}:
+
+\subsection{Kicks in the central region (main solenoid)}
+\label{core:sec:2}
+
+\begin{figure*}
+  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} \hfill
+  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}
+  \caption{Example of current-density distribution measurements for
+    the hollow electron gun prototype CHG1b, taken at the Fermilab
+    electron lens test stand in
+    2017~\cite{hel_res_field_stancari_2017}: 2-dimensional transverse
+    profile measurement (left) and calculated 1-dimensional radial
+    projection (right).}
+  \label{core:fig:0}
+\end{figure*}
+
+\begin{figure*}
+  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} \hfill
+  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
+  \caption{Calculated relative electric field for the hollow electron
+    gun CHG1b in the transverse $x$-$y$ plane (left) and as
+    1-dimensional cuts through the $x$ and $y$ axes (right). The field
+    calculations are based on measurements at the Fermilab
+    electron-lens test stand combined with Warp calculations of the
+    electric potentials and fields in a cylindrical beam pipe.}
+\label{core:fig:1}
+\end{figure*}
+
+For a perfectly annular and axially symmetric profile, the field in
+the region of the proton beam core vanishes. This is expressed by
+Eq.~\ref{eq:field_2} and is illustrated in
+Fig.~\ref{fig:hel_field}. Fields at the location of the proton beam
+core can arise in case of profile imperfections of the electron
+beam, in particular if the axial symmetry is broken.
+
+Recently, a hollow electron gun prototype for the LHC (called CHG1b)
+was characterized at the Fermilab electron lens test
+stand~\cite{hel_test_stand_fnal}. An example of a measurement of the
+electron beam current density is shown in Fig.~\ref{core:fig:0}.
+
+In the test stand, only resistive solenoids are available. Through
+simulations, one can estimate the fields generated by the compressed
+electron beam profile needed for HL-LHC. The measured distribution is
+compressed to the inner electron beam radius of 1.24~mm, corresponding
+to $4\sigma_p$, as described in Table~\ref{tab:hllhc_param} and
+Table~\ref{tab:hel_param}. A distribution with about $65\,000$
+particles is generated according to the measured current-density
+profile. As boundary condition, the LHC inner beam pipe radius
+$b=30$~mm is used. The potential and fields are then calculated with
+the particle-in-cell code Warp~\cite{warp}. The resulting relative
+electric field strengths are shown in Fig.~\ref{core:fig:1}. The
+electric field is obviously proportional to the charge density of
+electrons. Its relative strength is rather insensitive to the hollow
+beam radii, as long as these remain small compared to the beam pipe
+radius. Further details on the measurements and Warp simulations can
+be found in Ref.~\cite{hel_res_field_stancari_2017}.
+
+Using these measurements and calculations of the relative electric
+field strengths in the transverse plane (Fig.~\ref{core:fig:1}), we
+obtain an average field in the hole of
+
 \begin{equation}
-\int_{z_1}^{z_2} E_{y} dz= 36 \ \mathrm{kV} \Rightarrow \Delta y'= \q{5.1}{nrad}
+  E_{\mathrm{rel}}= 5\%.
 \end{equation}
-Assuming now 10\% fluctuation between entrance and exit kick, the expected kick from the HEL bends is
-\begin{equation}\label{eqn:kick_bends}
-\Delta x', \Delta y'\approx \q{0.5}{nrad}.
+
+For a maximum kick of 392~nrad, as derived in Eq.~\ref{eqn:helkick},
+the estimated dipole kick amplitude at the proton beam core is then
+
+\begin{equation}
+  \label{eqn:kick_central}
+  \Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \q{20}{nrad}.
 \end{equation}
 
-%\begin{table*}
-%	\caption{\label{tab:md_param}%
-%		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and 2017 \cite{resexmd2016,resexmd2017}.
-The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V horizontal and vertical at the same time.
-%	}
-%	\begin{ruledtabular}
-%		\begin{tabular}{lcc}
-%			Parameter & Experiment 2017 & Experiment 2016  \\
-%			\colrule
-%			beam &\multicolumn{2}{c}{Beam~1} \\
-%			beam energy &\multicolumn{2}{c}{injection energy, 450 GeV} \\\hline
-%			single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
-%			normalized emittance &\multicolumn{2}{c}{$2.5-3.5~\mathrm{\mu m}$} \\
-%			$4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
-%			$1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
-%			number of bunches & $3\times72=216$ bunches & $12\times4=48$ single bunches\\
-%			& (+ 1 pilot + 12 nominal) & \\\hline
-%			injection optics, $\beta^*=11$~m & standard optics 2017 & standard optics 2016\\
-%			Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=\pm19.6~\mathrm{A}$, explicitly +19.6 A for MOF circuit and}\\
-%			& \multicolumn{2}{c}{-19.6~A for MOD circuit (standard 2016 settings)}\\\hline
-%			working point $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\
-%			chromaticity $(Q'_x,Q'_y)$ & \multicolumn{2}{c}{(+15,+15)}\\\hline
-%			pulsing patterns  & $7^{\mathrm{th}}$~turn H,V,H+V;& $7^{\mathrm{th}}$~turn H;\\
-%			& $8^{\mathrm{th}}$~turn H,V,H+V;& $10^{\mathrm{th}}$~turn V;\\
-%			& random  H,V,H+V;& \\
-%		\end{tabular}
-%	\end{ruledtabular}
-%\end{table*}
+For the scope of this paper, only the approximate magnitude of the
+residual dipole kick is considered. In general, the field map,
+including its multipolar components, can be parameterized in a
+symplectic form, as described for instance in
+Ref.~\cite{hel_bends_stancari}, and used in tracking codes to estimate
+the effects on the circulating beam.
+
+\section{Experimental and simulation setup\label{sec:exp}}
+% put picture on top
 
@@
 
-\subsection{Kicks in the central region (main solenoid)}
-\label{core:sec:2}
-
-\begin{figure*}
-  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} \hfill
-  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}
-  \caption{Example of current-density distribution measurements for
-    the hollow electron gun prototype CHG1b, taken at the Fermilab
-    electron lens test stand in
-    2017~\cite{hel_res_field_stancari_2017}: 2-dimensional transverse
-    profile measurement (left) and calculated 1-dimensional radial
-    projection (right).}
-  \label{core:fig:0}
-\end{figure*}
-
-\begin{figure*}
-  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} \hfill
-  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
-  \caption{Calculated relative electric field for the hollow electron
-    gun CHG1b in the transverse $x$-$y$ plane (left) and as
-    1-dimensional cuts through the $x$ and $y$ axes (right). The field
-    calculations are based on measurements at the Fermilab
-    electron-lens test stand combined with Warp calculations of the
-    electric potentials and fields in a cylindrical beam pipe.}
-\label{core:fig:1}
-\end{figure*}
-
-For a perfectly annular and axially symmetric profile, the field in
-the region of the proton beam core vanishes. This is expressed by
-Eq.~\ref{eq:field_2} and is illustrated in
-Fig.~\ref{fig:hel_field}. Fields at the location of the proton beam
-core can arise in case of profile imperfections of the electron
-beam, in particular if the axial symmetry is broken.
-
-Recently, a hollow electron gun prototype for the LHC (called CHG1b)
-was characterized at the Fermilab electron lens test
-stand~\cite{hel_test_stand_fnal}. An example of a measurement of the
-electron beam current density is shown in Fig.~\ref{core:fig:0}.
-
-In the test stand, only resistive solenoids are available. Through
-simulations, one can estimate the fields generated by the compressed
-electron beam profile needed for HL-LHC. The measured distribution is
-compressed to the inner electron beam radius of 1.24~mm, corresponding
-to $4\sigma_p$, as described in Table~\ref{tab:hllhc_param} and
-Table~\ref{tab:hel_param}. A distribution with about $65\,000$
-particles is generated according to the measured current-density
-profile. As boundary condition, the LHC inner beam pipe radius
-$b=30$~mm is used. The potential and fields are then calculated with
-the particle-in-cell code Warp~\cite{warp}. The resulting relative
-electric field strengths are shown in Fig.~\ref{core:fig:1}. The
-electric field is obviously proportional to the charge density of
-electrons. Its relative strength is rather insensitive to the hollow
-beam radii, as long as these remain small compared to the beam pipe
-radius. Further details on the measurements and Warp simulations can
-be found in Ref.~\cite{hel_res_field_stancari_2017}.
-
-Using these measurements and calculations of the relative electric
-field strengths in the transverse plane (Fig.~\ref{core:fig:1}), we
-obtain an average field in the hole of
-
-\begin{equation}
-  E_{\mathrm{rel}}= 5\%.
-\end{equation}
-
-For a maximum kick of 392~nrad, as derived in Eq.~\ref{eqn:helkick},
-the estimated dipole kick amplitude at the proton beam core is then
-
-\begin{equation}
-  \label{eqn:kick_central}
-  \Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \q{20}{nrad}.
-\end{equation}
-
-For the scope of this paper, only the approximate magnitude of the
-residual dipole kick is considered. In general, the field map,
-including its multipolar components, can be parameterized in a
-symplectic form, as described for instance in
-Ref.~\cite{hel_bends_stancari}, and used in tracking codes to estimate
-the effects on the circulating beam.
-
-\section{Experimental and simulation setup\label{sec:exp}}
-%put picture on top
 \begin{figure*}

### 2018-02-22 resex.tex
Revised Section III on residual kicks. Added 2 references to resex-bibtex.bib. Including PDF paper in this commit.

@@
 
+\begin{figure*}
+  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} \hfill
+  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}
+  \caption{Example of current-density distribution measurements for
+    the hollow electron gun prototype CHG1b, taken at the Fermilab
+    electron lens test stand in
+    2017~\cite{hel_res_field_stancari_2017}: 2-dimensional transverse
+    profile measurement (left) and calculated 1-dimensional radial
+    projection (right).}
+  \label{core:fig:0}
+\end{figure*}
+
+\begin{figure*}
+  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} \hfill
+  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
+  \caption{Calculated relative electric field for the hollow electron
+    gun CHG1b in the transverse $x$-$y$ plane (left) and as
+    1-dimensional cuts through the $x$ and $y$ axes (right). The field
+    calculations are based on measurements at the Fermilab
+    electron-lens test stand combined with Warp calculations of the
+    electric potentials and fields in a cylindrical beam pipe.}
+\label{core:fig:1}
+\end{figure*}
+
 Parasitic kicks on the proton beam core can be due to electron-beam
@@
 
-\subsection{Uncompensated kicks from injection and extraction bends}
+\subsection{Kicks from injection and extraction bends}
 \label{core:sec:1}
@@
 
-\subsection{Kicks in the central region (main solenoid)}
+\subsection{Kicks in the central overlap region}
 \label{core:sec:2}
 
-\begin{figure*}
-  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} \hfill
-  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}
-  \caption{Example of current-density distribution measurements for
-    the hollow electron gun prototype CHG1b, taken at the Fermilab
-    electron lens test stand in
-    2017~\cite{hel_res_field_stancari_2017}: 2-dimensional transverse
-    profile measurement (left) and calculated 1-dimensional radial
-    projection (right).}
-  \label{core:fig:0}
-\end{figure*}
-
-\begin{figure*}
-  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} \hfill
-  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
-  \caption{Calculated relative electric field for the hollow electron
-    gun CHG1b in the transverse $x$-$y$ plane (left) and as
-    1-dimensional cuts through the $x$ and $y$ axes (right). The field
-    calculations are based on measurements at the Fermilab
-    electron-lens test stand combined with Warp calculations of the
-    electric potentials and fields in a cylindrical beam pipe.}
-\label{core:fig:1}
-\end{figure*}
-
-For a perfectly annular and axially symmetric profile, the field in
-the region of the proton beam core vanishes. This is expressed by
-Eq.~\ref{eq:field_2} and is illustrated in
-Fig.~\ref{fig:hel_field}. Fields at the location of the proton beam
-core can arise in case of profile imperfections of the electron
-beam, in particular if the axial symmetry is broken.
+For a perfectly annular and axially symmetric electron beam profile,
+the electromagnetic field in the region of the proton beam core
+vanishes. This is expressed by Eq.~\ref{eq:field_2} and is illustrated
+in Fig.~\ref{fig:hel_field}. Fields at the location of the proton beam
+core can arise if the axial symmetry is broken.
 
@@
 
-In the test stand, only resistive solenoids are available. Through
-simulations, one can estimate the fields generated by the compressed
-electron beam profile needed for HL-LHC. The measured distribution is
+In the test stand, only resistive solenoids are available. One can
+estimate the fields generated by the compressed electron beam profile
+in the superconducting solenoids of the HL-LHC HEL using a combination
+of experimental measurements and calculations.
+
+Experimentally, it was verified that the current-density profiles
+scale with electron beam current and confining axial field according
+to space-charge evolution~\cite{Jo:PoP:2018,
+  hel_res_field_stancari_2017}. Specifically, the same profile is
+obtained for a given family of experimental conditions with constant
+ratio $\sqrt{V} / B$, where $V$ is the accelerating voltage and $B$ is
+the axial field. This ratio is proportional to the space-charge
+evolution number $g = \omega_D \cdot \tau \propto \sqrt{V}/B$,
+representing the number of $\vec{E} \times \vec{B}$ rotations in the
+propagation time~$\tau$, with
+$\omega_D \equiv \omega_p^2 / (2 \omega_c)$ the diocotron frequency,
+$\omega_p$ the plasma frequency, and $\omega_c$ the cyclotron
+frequency of the magnetically confined electrons in the
+solenoid~\cite{Davidson:nonneutral-plasmas:2001}.
+
+For a given HL-LHC HEL configuration, the corresponding
+current-density profile measured in the Fermilab test stand is used as
+input to calculate the electromagnetic fields. For the purpose of
+estimating the residual fields, the measured distribution is
 compressed to the inner electron beam radius of 1.24~mm, corresponding
@@
 profile. As boundary condition, the LHC inner beam pipe radius
-$b=30$~mm is used. The potential and fields are then calculated with
-the particle-in-cell code Warp~\cite{warp}. The resulting relative
-electric field strengths are shown in Fig.~\ref{core:fig:1}. The
-electric field is obviously proportional to the charge density of
-electrons. Its relative strength is rather insensitive to the hollow
-beam radii, as long as these remain small compared to the beam pipe
-radius. Further details on the measurements and Warp simulations can
-be found in Ref.~\cite{hel_res_field_stancari_2017}.
+$b = \q{30}{mm}$ is used. The potential and fields are then calculated
+with Warp~\cite{warp}. The resulting relative electric field strengths
+are shown in Fig.~\ref{core:fig:1}. The electric field is obviously
+proportional to the charge density of electrons. Its relative strength
+is rather insensitive to the hollow beam radii, as long as these
+remain small compared to the beam pipe radius. Further details on the
+measurements and Warp simulations can be found in
+Ref.~\cite{hel_res_field_stancari_2017}.
 
@@
 field strengths in the transverse plane (Fig.~\ref{core:fig:1}), we
-obtain an average field in the hole of
-
+obtain a median field in the hole of
+%
 \begin{equation}
-  E_{\mathrm{rel}}= 5\%.
+  \langle E_\mathrm{hole} \rangle / E_\mathrm{max} =  5\%.
 \end{equation}
-
+%
 For a maximum kick of 392~nrad, as derived in Eq.~\ref{eqn:helkick},
-the estimated dipole kick amplitude at the proton beam core is then
-
+the estimated dipole kick amplitude at the proton beam core is therefore
+approximately
+%
 \begin{equation}
   \label{eqn:kick_central}
-  \Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} = \q{20}{nrad}.
+  \Delta x' , \Delta y' = \q{20}{nrad}.
 \end{equation}
-
-For the scope of this paper, only the approximate magnitude of the
+%
+The relative magnitude of the residual field depends on several
+factors, including cathode quality, electron gun geometry, solenoid
+field configuration, space-charge evolution, etc. It can be improved,
+if needed.
+
+For the purposes of this paper, only the approximate magnitude of the
 residual dipole kick is considered. In general, the field map,
-including its multipolar components, can be parameterized in a
+including its multipolar components, can be parameterized in
 symplectic form, as described for instance in

### 2018-02-22 resex.tex

@@
 %floatfix,
+longbibliography,
 ]{revtex4-1}
@@
 \author{Miriam Fitterer}
- \email{mfittere@fnal.gov}
+ \email{Miriam.Fitterer@cern.ch}
 \author{Giulio Stancari}
@@
 (Section~\ref{core:sec:2}). Both sources result in nonlinear
-kicks~\cite{hel_bends_stancari, hel_model_polynomial_morozov}. In
+kicks~\cite{hel_bends_stancari,hel_model_polynomial_morozov}. In
 continuous operation, these nonlinear kicks are usually much smaller
@@
 
-\section{Experimental and simulation setup\label{sec:exp}}
-% put picture on top
+\section{Experimental and simulation setup}
+\label{sec:exp}
 
 \begin{table*}
-	\caption{\label{tab:md_param}%
-		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and~2017~\cite{resexmd2016,resexmd2017}.
-The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V horizontal and vertical at the same time.
-	}
+  \caption{ Beam parameters and machine configuration as used in the
+    two resonant excitation experiments in 2016
+    and~2017~\cite{resexmd2016,resexmd2017}. The plane of the
+    excitation is abbreviated with H for horizontal, V for vertical
+    and H+V horizontal and vertical at the same time.}
+  \label{tab:md_param}
   \begin{ruledtabular}
@@
   \end{minipage}
-	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).
-In 2016 a total of 48~bunches was used and in 2017 a total of 216~bunches.
-Each bunch is indicated with a cyan vertical bar.
-The bunches were grouped in sub-groups of 4~for 2016 and 6~for 2017 bunches, where each sub-group experienced the same excitation pattern and amplitude.
-The excitation amplitude is shown in blue and red.
-In 2016 the excitation was only applied in one plane.
-In 2017 more bunches could be injected without compromising machine protection making it possible to test all three planes in one fill.}
-\end{figure*}
-%put picture on top
+  \caption{Filling and excitation scheme for LHC experiments in 2016
+    (top) and 2017 (bottom). In 2016 a total of 48~bunches was used
+    and in 2017 a total of 216~bunches. Each bunch is indicated with a
+    cyan vertical bar. The bunches were grouped in sub-groups of 4~for
+    2016 and 6~for 2017 bunches, where each sub-group experienced the
+    same excitation pattern and amplitude. The excitation amplitude is
+    shown in blue and red. In 2016 the excitation was only applied in
+    one plane. In 2017 more bunches could be injected without
+    compromising machine protection making it possible to test all
+    three planes in one fill.}
+  \label{fig:fill}
+\end{figure*}
 
@@
 
-\bibliography{resex-bibdesk}
+%\bibliography{resex-bibdesk}
+\bibliography{resex-mf}
 

### 2018-02-27 resex.tex
Restored resex.tex and .gitignore
Intermediate step in including Miriam's bibliography; resex-bibdesk.bib and resex-mf.bib are still separate.
Revised figure and table in experimental setup (Section IV). Split setup section into experimental (Section IV) and simulations (Section V).

--- 2018-02-22 resex.tex
@@
 %floatfix,
+longbibliography
 ]{revtex4-1}
@@
 HELs. Section~\ref{sec:core} is dedicated to the sources of residual
-fields from the HEL in the core region. Section~\ref{sec:exp}
-describes the setup of simulations and experimental
-conditions. Results and comparisons between simulations and
-measurements are given in Section~\ref{sec:simex}, with discussion and
-summary in Section~\ref{sec:sum}. Auxiliary data and plots are
-collected in the Appendices.
+fields from the HEL in the core region. Sections~\ref{sec:exp}
+and~\ref{sec:sim} describe the experimental conditions and the setup
+of the numerical tracking simulations. Results and comparisons between
+simulations and measurements are given in Section~\ref{sec:simex},
+with discussion and summary in Section~\ref{sec:sum}. Auxiliary data
+and plots are collected in the Appendices.
 
@@
 
-\section{Experimental and simulation setup\label{sec:exp}}
-% put picture on top
+\section{Experimental setup}
+\label{sec:exp}
 
 \begin{table*}
-	\caption{\label{tab:md_param}%
-		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and~2017~\cite{resexmd2016,resexmd2017}.
-The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V horizontal and vertical at the same time.
-	}
+  \caption{Beam parameters and machine configuration for the two
+    resonant excitation experiments of~2016
+    and~2017~\cite{resexmd2016, resexmd2017}. The plane of the
+    excitation is abbreviated as H for horizontal, V for vertical and
+    H+V for horizontal and vertical at the same time.}
+  \label{tab:md_param}
   \begin{ruledtabular}
@@
       single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
-			normalized emittance &\multicolumn{2}{c}{$2.5-3.5~\mathrm{\mu m}$} \\
+      normalized emittance &\multicolumn{2}{c}{$2.5$--\q{3.5}{\mu m}} \\
       $4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
       $1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
-			number of bunches & $12\times4=48$ single bunches & $3\times72=216$ bunches\\
+      number of bunches & $12 \times 4 = 48$ & $3 \times 72 = 216$ \\
                 &  & (+ 1 pilot + 12 nominal) \\\hline
-			injection optics, $\beta^*=11$~m & standard optics 2016 & standard optics 2017\\
-			Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=+19.6~\mathrm{A}$ for MOF circuit and}\\
-			& \multicolumn{2}{c}{$I_{\mathrm{MO}}=-19.6~\mathrm{A}$ for MOD circuit (standard 2016 settings)}\\\hline
-			working point $(Q_x,Q_y)$ & (64.28,59.31) & (62.27,60.295) \\
-			chromaticity $(Q'_x,Q'_y)$ & \multicolumn{2}{c}{(+15,+15)}\\\hline
-			pulsing patterns  &$7^{\mathrm{th}}$~turn H; &$7^{\mathrm{th}}$~turn H,V,H+V; \\
-			& $10^{\mathrm{th}}$~turn V; & $8^{\mathrm{th}}$~turn H,V,H+V; \\
-			& &  random  H,V,H+V;\\
+      injection optics, $\beta^* = \q{11}{m}$ & standard optics 2016 & standard optics 2017\\
+      Landau-damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}
+                                  = \q{+19.6}{A}$ for MOF circuit and}\\
+                & \multicolumn{2}{c}{$I_{\mathrm{MO}} = \q{-19.6}{A}$
+                  for MOD circuit (standard 2016 settings)} \\ \hline
+      working point $(Q_x, Q_y)$ & (64.28, 59.31) & (62.27, 60.295) \\
+      chromaticity $(Q'_x, Q'_y)$ & \multicolumn{2}{c}{(+15, +15)}\\ \hline
+      pulsing patterns  & $7^{\mathrm{th}}$~turn H
+                                  &$7^{\mathrm{th}}$~turn H, V, H+V \\
+                & $10^{\mathrm{th}}$~turn V & $8^{\mathrm{th}}$~turn
+                                              H, V, H+V \\
+                & &  random  H, V, H+V\\
     \end{tabular}
@@
 \begin{figure*}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		2016 experiment\\
-		\includegraphics[width=0.9\linewidth]{bunchfilling_2016.png}	
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		$ $\\
-		2017 experiment\\
-		\includegraphics[width=0.9\linewidth]{bunchfilling_2017.png}	
-	\end{minipage}
-	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom).
-In 2016 a total of 48~bunches was used and in 2017 a total of 216~bunches.
-Each bunch is indicated with a cyan vertical bar.
-The bunches were grouped in sub-groups of 4~for 2016 and 6~for 2017 bunches, where each sub-group experienced the same excitation pattern and amplitude.
-The excitation amplitude is shown in blue and red.
-In 2016 the excitation was only applied in one plane.
-In 2017 more bunches could be injected without compromising machine protection making it possible to test all three planes in one fill.}
-\end{figure*}
-%put picture on top
+  \begin{tabular}{c}
+    \multicolumn{1}{l}{\emph{2016 experiment}} \\
+    \includegraphics[width=0.95\textwidth]{bunchfilling_2016.png} \\
+    \\
+    \multicolumn{1}{l}{\emph{2017 experiment}} \\
+    \includegraphics[width=0.95\textwidth]{bunchfilling_2017.png}
+  \end{tabular}
+  \caption{Bunch filling scheme and excitation patterns for the
+    2016~(top) and 2017~(bottom) LHC experiments. In 2016, a total of
+    48~bunches was used, whereas in 2017 there were 216~bunches. Each
+    bunch is represented by a vertical cyan bar. The bunches were
+    grouped in subsets of~4 in 2016 and in subsets of~6 in 2017. Each
+    subset experienced the same excitation pattern and amplitude. The
+    excitation amplitudes and relative phases are shown in blue or
+    red. In 2016, the excitation was only applied in one plane. In
+    2017, more injected bunches were allowed without compromising
+    machine protection, so it was possible to test all~3 excitation
+    planes in the same fill. The transverse damper was active on half
+    of the bunches, indicated by the black lines.}
+  \label{fig:fill}
+\end{figure*}
 
@@
 
-The resonant excitation experiment in 2017 involved simultaneous measurements on three groups of 72 bunches with a dedicated excitation pattern and transverse feedback configuration on each sub-group of 6 bunches.
-In 2016 a similar configuration was used, but with only 48 bunches in total and sub-groups of 4 bunches.
+The resonant excitation experiment in 2017 involved simultaneous measurements on three groups of 72 bunches with a dedicated excitation pattern and transverse feedback configuration on each subset of 6 bunches.
+In 2016 a similar configuration was used, but with only 48 bunches in total and subsets of 4 bunches.
 Both schemes are illustrated in Fig.~\ref{fig:fill}.
-In order to reduce the error of the measurements, the observables are averaged over each sub-group of bunches experiencing the same excitation amplitude, yielding in total 5~different amplitudes $n\Delta A$ with $n=1,\cdots, 5$ plus the references bunch without excitation for each transverse feedback configuration and excitation plane.
+In order to reduce the error of the measurements, the observables are averaged over each subset of bunches experiencing the same excitation amplitude, yielding in total 5~different amplitudes $n\Delta A$ with $n=1,\cdots, 5$ plus the references bunch without excitation for each transverse feedback configuration and excitation plane.
 Furthermore, the experiments were usually split in 3 different parts:
@@
 Each part lasted approximately 10-15~minutes, which is considered to be a long enough time window to allow the beam distribution to reach its new equilibrium state.
-In the plots of the experimental results shown later in the paper, the maximum excitation amplitude is indicated with black horizontal arrows, while the average of each sub-group of bunches with the same excitation amplitude $n\Delta A$ is distinguished with different colors.
+In the plots of the experimental results shown later in the paper, the maximum excitation amplitude is indicated with black horizontal arrows, while the average of each subset of bunches with the same excitation amplitude $n\Delta A$ is distinguished with different colors.
 
@@
 
-\subsection{Overview of simulations\label{sec:sim}}
+\section{Numerical tracking simulations}
+\label{sec:sim}
+
 For the preparation and interpretation of the experiments two different types of simulations were performed:
@@
 
-\section{Experimental and simulation results\label{sec:simex}}
+\section{Results}
+\label{sec:simex}
+
 In the following we will jump forth and back between both experimental and simulation results as the experimental results are best understood on the base of the simulations.

### 2018-02-27 resex.tex
Intermediate step in revision of experimental section.

@@
 
-\subsection{Overview of LHC experiments\label{sec:exp_sum}}
-The experiments at the LHC which are presented in this paper were aiming at deriving tolerances on the residual HEL field at the location of the beam core in case of pulsed operation.
-In total, two experiments were conducted, the first one in 2016 \cite{resexmd2016} and the second one in 2017 \cite{resexmd2017}.
-Beam and machine parameters for both experiments are summarized in Table~\ref{tab:md_param}.
-Preceding the beam experiments, simulations were performed to identify the most effective $k^{\mathrm{th}}$~turn pulsing pattern and yield estimates for losses and emittance growth.
-The two $k^{\mathrm{th}}$~turn pulsing patterns with the largest effect, one $k^{\mathrm{th}}$~turn pulsing pattern with no effect and the random excitation could then be experimentally tested in the LHC.
-In order to test the reproducibility of the results at least on the example of one pulsing pattern, the $7^{\mathrm{th}}$~turn pulsing was first tried in 2016 and then repeated in 2017.
-
-During the experiments the losses were measured with the fast beam current transformers (FBCTs) and the emittance and transverse beam profiles with the beam synchrotron radiation telescope (BSRT), where all instruments are capable of delivering bunch-by-bunch data.
-The analysis of the BSRT profiles is quite involved and we will therefore focus in this paper only on the direct comparison of the profiles.
-A detailed description of the analysis with more details can be found in \cite{bsrtprofinj} and for the individual experiments in \cite{resexmd2016,resexmd2017}.
-
-\subsection{Excitation with transverse damper and filling scheme\label{sec:adt}}
-The primary function of the transverse feedback system in LHC, also referenced to as ADT, is to provide injection oscillation damping and actively damp the coupled bunch instabilities driven by the machine impedance.
-The main building blocks of the ADT system are stipline pickups at the position Q7 and Q9 at Point~4 of the LHC, connected to the beam position measurement modules at the surface, the digital signal processing modules (mDSPU) and a set of tetrode power amplifiers feeding electrostatic kickers in the RF zone of the LHC~\cite{adt_sum_2008,adt_sum_2011}.
-Though originally designed to damp the undesired transverse activities only, the damper is being routinely used for sophisticated beam excitations as well thanks to the state of the art digital signal processing hardware.
-These include among others abort and injection gap cleaning, single bunch excitation for tune and linear coupling measurements and special modes of operation for dedicated experiments at the LHC.
-
-The transverse feedback is in general active during the whole LHC cycle, therefore it is critical that the noise introduced by the system does not contribute to any emittance growth.
-The typical machine cycle requires a short damping time of ~10-20 turns (high feedback gain) for injection oscillation damping, which is increased during the ramp and collisions to ~50-100 turns (lower feedback gain).
-It has been demonstrated that noise introduced by the transverse feedback system itself is sufficiently low not to introduce any measurable emittance growth \cite{adt_noise_emit_2017}.
-All excitation signals referenced to in this paper are digitally synthesized in the damper's digital signal processing units and therefore can be considered as ``noise-free''.
-This point is important as then the effect on the beam can be attributed entirely to the applied excitation and any effect due to unwanted residual noise can be neglected.
-
-The resonant excitation experiment in 2017 involved simultaneous measurements on three groups of 72 bunches with a dedicated excitation pattern and transverse feedback configuration on each subset of 6 bunches.
-In 2016 a similar configuration was used, but with only 48 bunches in total and subsets of 4 bunches.
-Both schemes are illustrated in Fig.~\ref{fig:fill}.
-In order to reduce the error of the measurements, the observables are averaged over each subset of bunches experiencing the same excitation amplitude, yielding in total 5~different amplitudes $n\Delta A$ with $n=1,\cdots, 5$ plus the references bunch without excitation for each transverse feedback configuration and excitation plane.
-Furthermore, the experiments were usually split in 3 different parts:
+\begin{figure*}
+  \includegraphics[width=\textwidth]{bunchfilling_measured.png}	
+  \caption{Example of bunch-by-bunch excitation detected by the
+    real-time transverse activity monitor during the 2017 experiment,
+    when the beam was excited every $8^\mathrm{th}$~turn.}
+  \label{fig:fill_meas} 
+\end{figure*}
+
+\subsection{Overview}
+\label{sec:exp_sum}
+
+The purpose of the experiments at the LHC is to quantify the effects
+of a pulsed excitation on the proton beam core. In addition, these
+measurements provide an experimental basis to guide the design and
+tolerances on the residual HEL fields at the location of the beam
+core, in case resonant excitation is needed for HL-LHC.
+
+Two experiments were conducted, one in 2016~\cite{resexmd2016} and one
+in 2017~\cite{resexmd2017}. Beam and machine parameters are summarized
+in Table~\ref{tab:md_param}.
+
+During the experiments, losses were measured with the fast beam
+current transformers (FBCTs). Transverse beam profiles and emittances
+were provided by the beam synchrotron radiation telescope (BSRT). All
+instruments were capable of delivering bunch-by-bunch data. The data
+analysis of the BSRT profiles is quite involved. In this paper, we
+focus on the direct comparison of the resulting profiles. A detailed
+description of the profile analysis can be found in
+Ref.~\cite{bsrtprofinj}, with individual experiments reported in
+Refs.~\cite{resexmd2016, resexmd2017}.
+
+The choice of excitation patterns for the experiments was guided by
+losses and emittance growths calculated in numerical tracking
+simulations, described below. It was chosen to study experimentally
+the two pulsing patterns with the largest calculated effects on the
+beam ($7^{\mathrm{th}}$- and $10^{\mathrm{th}}$-turn pulsing), one
+pattern with no effect ($8^{\mathrm{th}}$-turn pulsing), and the
+random excitation. In order to quantify the reproducibility of the
+results under different machine configurations, one pulsing pattern
+($7^{\mathrm{th}}$-turn pulsing) was tried first in~2016 and then
+repeated in~2017.
+
+\subsection{Excitations with the transverse damper and filling schemes}
+\label{sec:adt}
+
+The primary function of the transverse feedback system in LHC, also
+referenced to as ADT, is to provide injection oscillation damping and
+actively damp the coupled bunch instabilities driven by the machine
+impedance. The main building blocks of the ADT system are stipline
+pickups at the position Q7 and Q9 at Point~4 of the LHC, connected to
+the beam position measurement modules at the surface, the digital
+signal processing modules (mDSPU) and a set of tetrode power
+amplifiers feeding electrostatic kickers in the RF zone of the
+LHC~\cite{adt_sum_2008,adt_sum_2011}. Though originally designed to
+damp the undesired transverse activities only, the damper is being
+routinely used for sophisticated beam excitations as well thanks to
+the state of the art digital signal processing hardware. These include
+among others abort and injection gap cleaning, single bunch excitation
+for tune and linear coupling measurements and special modes of
+operation for dedicated experiments at the LHC.
+
+The transverse feedback is in general active during the whole LHC
+cycle, therefore it is critical that the noise introduced by the
+system does not contribute to any emittance growth. The typical
+machine cycle requires a short damping time of ~10-20 turns (high
+feedback gain) for injection oscillation damping, which is increased
+during the ramp and collisions to ~50-100 turns (lower feedback
+gain). It has been demonstrated that noise introduced by the
+transverse feedback system itself is sufficiently low not to introduce
+any measurable emittance growth \cite{adt_noise_emit_2017}. All
+excitation signals referenced to in this paper are digitally
+synthesized in the damper's digital signal processing units and
+therefore can be considered as ``noise-free''. This point is important
+as then the effect on the beam can be attributed entirely to the
+applied excitation and any effect due to unwanted residual noise can
+be neglected.
+
+The resonant excitation experiment in 2017 involved simultaneous
+measurements on three groups of 72 bunches with a dedicated excitation
+pattern and transverse feedback configuration on each subset of 6
+bunches. In 2016 a similar configuration was used, but with only 48
+bunches in total and subsets of 4 bunches. Both schemes are
+illustrated in Fig.~\ref{fig:fill}. In order to reduce the error of
+the measurements, the observables are averaged over each subset of
+bunches experiencing the same excitation amplitude, yielding in total
+5~different amplitudes $n\Delta A$ with $n=1,\cdots, 5$ plus the
+references bunch without excitation for each transverse feedback
+configuration and excitation plane. Furthermore, the experiments were
+usually split in 3 different parts:
 \begin{description}
@@
 \end{description}
-Each part lasted approximately 10-15~minutes, which is considered to be a long enough time window to allow the beam distribution to reach its new equilibrium state.
-In the plots of the experimental results shown later in the paper, the maximum excitation amplitude is indicated with black horizontal arrows, while the average of each subset of bunches with the same excitation amplitude $n\Delta A$ is distinguished with different colors.
-
-The excitation in the horizontal and the vertical plane is generated by a different set of signal processing hardware, however during this experiment the signals were synchronized at a turn level.
-This means the bunches of the third group (of 72 bunches) experienced the kick in both horizontal and vertical plane during the same turn.
-
-The ADT kicker deflection angle is calculated from the kicker geometry and the excitation voltage on the deflection plates.
-The excitation voltage used in the experiments is a complicated waveform, generated by a very complex hardware (digital signal processor, chain of transmission lines, low and high power amplifiers).
-The peak excitation voltage (and subsequently the kick) seen by the beam can be only indirectly measured using the probes on each kicker.
-The measured values are therefore believed to be known with an error margin of 10-15\% for the experiments in 2017 and 50\% in 2016.
-The estimate for 2017 is based on only a single measurement and in the future and with more time available the determination of the error would definitely have to be improved.
-The error in 2016 is estimated from the transverse damper operational parameters.
-Due to lack of the dedicated machine time no precision calibration was performed.
-The maximum kick strength reachable with the ADT at injection without risking saturation is 96~nrad.
-As an example Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor during the 2017 experiments and an excitation with $8^{\mathrm{th}}$ turn pulsing.
-The measured activity is well in line with the expected excitation pattern.
-\begin{figure*}[h]
-	\centering
-	\includegraphics[width=1.0\linewidth]{bunchfilling_measured.png}	
-	\caption{\label{fig:fill_meas} Bunch by bunch excitation detected by the real time transverse activity monitor during the 2017 experiment and an excitation with 8th turn pulsing.}
-\end{figure*}
+Each part lasted approximately 10-15~minutes, which is considered to
+be a long enough time window to allow the beam distribution to reach
+its new equilibrium state. In the plots of the experimental results
+shown later in the paper, the maximum excitation amplitude is
+indicated with black horizontal arrows, while the average of each
+subset of bunches with the same excitation amplitude $n\Delta A$ is
+distinguished with different colors.
+
+The excitation in the horizontal and the vertical plane is generated
+by a different set of signal processing hardware, however during this
+experiment the signals were synchronized at a turn level. This means
+the bunches of the third group (of 72 bunches) experienced the kick in
+both horizontal and vertical plane during the same turn.
+
+The ADT kicker deflection angle is calculated from the kicker geometry
+and the excitation voltage on the deflection plates. The excitation
+voltage used in the experiments is a complicated waveform, generated
+by a very complex hardware (digital signal processor, chain of
+transmission lines, low and high power amplifiers). The peak
+excitation voltage (and subsequently the kick) seen by the beam can be
+only indirectly measured using the probes on each kicker. The measured
+values are therefore believed to be known with an error margin of
+10-15\% for the experiments in 2017 and 50\% in 2016. The estimate for
+2017 is based on only a single measurement and in the future and with
+more time available the determination of the error would definitely
+have to be improved. The error in 2016 is estimated from the
+transverse damper operational parameters. Due to lack of the dedicated
+machine time no precision calibration was performed. The maximum kick
+strength reachable with the ADT at injection without risking
+saturation is 96~nrad. As an example Fig.~\ref{fig:fill_meas} shows
+the oscillation amplitude of each bunch in both H and V planes
+captured by the LHC real time transverse activity monitor during the
+2017 experiments and an excitation with $8^{\mathrm{th}}$ turn
+pulsing. The measured activity is well in line with the expected
+excitation pattern.
 

### 2018-02-27 resex.tex
Revised Section IV on experimental setup.

@@
   \includegraphics[width=\textwidth]{bunchfilling_measured.png}	
-  \caption{Example of bunch-by-bunch excitation detected by the
-    real-time transverse activity monitor during the 2017 experiment,
-    when the beam was excited every $8^\mathrm{th}$~turn.}
+  \caption{Example of bunch-by-bunch beam centroid motion for Beam~1
+    in the horizontal (B1H, left) and vertical (B1V, right) planes,
+    detected by the real-time transverse activity monitor during the
+    2017 experiment, when the beam was excited every
+    $8^\mathrm{th}$~turn.}
   \label{fig:fill_meas} 
@@
 
-\subsection{Excitations with the transverse damper and filling schemes}
+\subsection{Excitations with the transverse damper and bunch filling schemes}
 \label{sec:adt}
 
-The primary function of the transverse feedback system in LHC, also
-referenced to as ADT, is to provide injection oscillation damping and
-actively damp the coupled bunch instabilities driven by the machine
-impedance. The main building blocks of the ADT system are stipline
-pickups at the position Q7 and Q9 at Point~4 of the LHC, connected to
-the beam position measurement modules at the surface, the digital
-signal processing modules (mDSPU) and a set of tetrode power
-amplifiers feeding electrostatic kickers in the RF zone of the
-LHC~\cite{adt_sum_2008,adt_sum_2011}. Though originally designed to
-damp the undesired transverse activities only, the damper is being
-routinely used for sophisticated beam excitations as well thanks to
-the state of the art digital signal processing hardware. These include
-among others abort and injection gap cleaning, single bunch excitation
-for tune and linear coupling measurements and special modes of
-operation for dedicated experiments at the LHC.
-
-The transverse feedback is in general active during the whole LHC
-cycle, therefore it is critical that the noise introduced by the
-system does not contribute to any emittance growth. The typical
-machine cycle requires a short damping time of ~10-20 turns (high
-feedback gain) for injection oscillation damping, which is increased
-during the ramp and collisions to ~50-100 turns (lower feedback
-gain). It has been demonstrated that noise introduced by the
-transverse feedback system itself is sufficiently low not to introduce
-any measurable emittance growth \cite{adt_noise_emit_2017}. All
-excitation signals referenced to in this paper are digitally
-synthesized in the damper's digital signal processing units and
-therefore can be considered as ``noise-free''. This point is important
-as then the effect on the beam can be attributed entirely to the
-applied excitation and any effect due to unwanted residual noise can
-be neglected.
-
-The resonant excitation experiment in 2017 involved simultaneous
-measurements on three groups of 72 bunches with a dedicated excitation
-pattern and transverse feedback configuration on each subset of 6
-bunches. In 2016 a similar configuration was used, but with only 48
-bunches in total and subsets of 4 bunches. Both schemes are
-illustrated in Fig.~\ref{fig:fill}. In order to reduce the error of
-the measurements, the observables are averaged over each subset of
-bunches experiencing the same excitation amplitude, yielding in total
-5~different amplitudes $n\Delta A$ with $n=1,\cdots, 5$ plus the
-references bunch without excitation for each transverse feedback
-configuration and excitation plane. Furthermore, the experiments were
-usually split in 3 different parts:
+The primary function of the LHC transverse damping and feedback
+system, also known as ADT, is to mitigate injection oscillations and
+to actively damp the coupled-bunch instabilities driven by machine
+impedance~\cite{adt_sum_2008, adt_sum_2011}. The main building blocks
+of the system are the following: strip-line pickups at positions Q7
+and Q9 near Interaction Point~4 (IP4) of the LHC, which are connected
+to the beam position measurement modules at the surface; the digital
+signal processing modules (mDSPU); and a set of tetrode power
+amplifiers feeding electrostatic kickers in the same radio-frequency
+sector of the LHC (IP4).
+
+Because of its flexibility and state-of-the-art hardware, the system
+is being routinely used for sophisticated beam excitations. These
+include abort- and injection-gap cleaning, excitation of individual
+bunches for tune and linear-coupling measurements, and other special
+modes of operation for dedicated experiments in the LHC.
+
+The transverse feedback is in general active during all phases of LHC
+operation. The typical machine cycle requires a short damping time of
+10--20 turns (high feedback gain) for injection oscillation
+damping. During the acceleration ramp and during collisions, the
+damping time is increased to 50--100 turns (lower feedback gain).
+
+Because the damper is always active in operations, it is critical that
+the noise introduced by the system does not cause any measurable
+emittance growth, and this fact was verified
+experimentally~\cite{adt_noise_emit_2017}. All excitation signals
+described in this paper were digitally synthesized in the ADT's
+digital signal processing units and are therefore assumed to be
+`noise-free.' The observed effects on the beam are attributed to the
+applied excitations and the effects of unwanted residual noise are
+assumed to be negligible.
+
+The resonant excitation experiment in~2017 involved simultaneous
+measurements on 3~groups of 72~bunches, with dedicated excitation
+patterns and transverse feedback configurations (i.e., damper active
+or damper off) on each subset of 6~bunches. In~2016, a similar
+configuration was used, with 48~bunches in total and subsets of
+4~bunches. Both schemes are illustrated in Fig.~\ref{fig:fill}. A
+total of 5 different amplitudes could be applied simultaneously to
+different subsets of bunches, denoted as $n \, \Delta A$, with
+$n = 1, \ldots, 5$. In addition, there were reference bunches without
+excitation for each transverse feedback configuration and excitation
+plane. Observables (losses, emittances, etc.) were averaged over each
+subset of bunches experiencing the same excitation and damping
+conditions.
+
+In~2017, the excitations in the horizontal and vertical planes were
+generated by a different set of signal-processing devices, but
+synchronized at a turn-by-turn level. Therefore, the bunches of the
+third group of 72 bunches (see Fig.~\ref{fig:fill}) were affected by
+the kicks in both horizontal and vertical planes during the same turn.
+
+The experiments were usually split in time into 3 different periods:
+%
 \begin{description}
-	\item[Part 1] No excitation is applied in order to allow the distribution to reach an equilibrium state after injection.
-	\item[Part 2] The excitation is applied with a first maximum excitation amplitude of $A_{\mathrm{max,1}}=5\Delta A_1$.
-	\item[Part 3] The maximum excitation amplitude is further increased to $A_{\mathrm{max,1}} < A_{\mathrm{max,2}}=5\Delta A_2$.
+\item[Period 1] No excitation was applied, to allow the beam to reach
+  an equilibrium state after injection.
+\item[Period 2] The excitation was applied with a first maximum
+  excitation amplitude of $A_{\mathrm{max,1}} = 5 \, \Delta A_1$.
+\item[Period 3] The maximum excitation amplitude was further increased
+  to $A_{\mathrm{max,2}} = 5 \, \Delta A_2 > A_{\mathrm{max,1}}$.
 \end{description}
-Each part lasted approximately 10-15~minutes, which is considered to
-be a long enough time window to allow the beam distribution to reach
-its new equilibrium state. In the plots of the experimental results
-shown later in the paper, the maximum excitation amplitude is
-indicated with black horizontal arrows, while the average of each
-subset of bunches with the same excitation amplitude $n\Delta A$ is
-distinguished with different colors.
-
-The excitation in the horizontal and the vertical plane is generated
-by a different set of signal processing hardware, however during this
-experiment the signals were synchronized at a turn level. This means
-the bunches of the third group (of 72 bunches) experienced the kick in
-both horizontal and vertical plane during the same turn.
-
-The ADT kicker deflection angle is calculated from the kicker geometry
-and the excitation voltage on the deflection plates. The excitation
-voltage used in the experiments is a complicated waveform, generated
-by a very complex hardware (digital signal processor, chain of
-transmission lines, low and high power amplifiers). The peak
-excitation voltage (and subsequently the kick) seen by the beam can be
-only indirectly measured using the probes on each kicker. The measured
-values are therefore believed to be known with an error margin of
-10-15\% for the experiments in 2017 and 50\% in 2016. The estimate for
-2017 is based on only a single measurement and in the future and with
-more time available the determination of the error would definitely
-have to be improved. The error in 2016 is estimated from the
-transverse damper operational parameters. Due to lack of the dedicated
-machine time no precision calibration was performed. The maximum kick
-strength reachable with the ADT at injection without risking
-saturation is 96~nrad. As an example Fig.~\ref{fig:fill_meas} shows
-the oscillation amplitude of each bunch in both H and V planes
-captured by the LHC real time transverse activity monitor during the
-2017 experiments and an excitation with $8^{\mathrm{th}}$ turn
-pulsing. The measured activity is well in line with the expected
+%
+Each period lasted approximately 10--15~minutes, which was considered
+long enough to allow the beam to reach its new equilibrium state. In
+the discussion of the experimental results (Section~\ref{sec:simex}
+below), the 3~periods are labeled according to the maximum excitation
+amplitude~$A_\mathrm{max}$, and subsets of bunches with the same
+excitation amplitude $n \, \Delta A$ ($n = 1, \ldots, 5$) are grouped
+by color.
+
+The proton deflection angle generated by the ADT kicker was calculated
+from the kicker geometry and from the excitation voltage on the
+deflection plates. The voltage depends on a complex chain of hardware
+(digital signal processor, transmission lines, low- and high-power
+amplifiers, etc.). In~2016, the kick was estimated from the
+operational system parameters and it was assigned an uncertainty of
+50\%. In~2017, the excitation voltage could be indirectly measured
+once, using the probes mounted on the kickers, with an estimated
+uncertainty of 10--15\%. A precise in-situ calibration was not
+possible due to the limited machine availability for dedicated
+studies.
+
+The maximum ADT kick strength that could be obtained at injection
+without risking saturation was approximately 100~nrad. For the
+experiments, a maximum nominal kick strength of 96~nrad was chosen.
+
+Figure~\ref{fig:fill_meas} shows an example of the oscillation
+amplitude of each bunch centroid in both horizontal and vertical
+planes, captured by the LHC real-time transverse activity monitor
+during the 2017 experiments, when the beam was excited every
+$8^{\mathrm{th}}$~turn. The measured activity shows the typical
+amplitude of bunch centroid excursions and reflects the expected
 excitation pattern.

### 2018-02-27 resex.tex
Revised Section V on numerical simulations

@@
 
-For the preparation and interpretation of the experiments two different types of simulations were performed:
-\begin{enumerate}
-	\item tracking of a Gaussian distribution, referred to as ``distribution tracking" in this paper, to obtain loss rates and emittance growth,
-	\item Frequency map analysis (FMA) for the visualization of resonances~\cite{fmalaskar}.
-\end{enumerate}
-For both simulation types, the tracking code Lifetrac \cite{lifetrac} was used.
-The simulation parameters are summarized in Table~\ref{tab:sim_param}.
-Further details on the simulations are given in Refs.~\cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
-\begin{table}[h]
-	\caption{\label{tab:sim_param}%
-		Summary of simulation parameters for the distribution tracking and FMA analysis. For details see \cite{md_sim_hel_res_ex_fitterer,resexmd2017}.
-	}
+\begin{table*}
+  \caption{Summary of parameters used in numerical simulations of
+    distribution tracking and frequency-map analysis (FMA). Further
+    details can be found in Refs.~\cite{md_sim_hel_res_ex_fitterer,
+      resexmd2017}.}
+  \label{tab:sim_param}
   \begin{ruledtabular}
     \begin{tabular}{lcc}
-			Parameter & distribution tracking & FMA \\
+      Parameter & distribution & FMA \\
       \colrule
@@
       beam energy &\multicolumn{2}{c}{450 GeV} \\
-			emittance & $3.5~\mu$m& $2.5~\mu$m\\
-			$4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
-			$1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
-			particle & 6D Gaussian & equally spaced grid\\
-			distribution &  distribution & in $x,y$ up to $10~\sigma$\\
-			&  ($10^4$ particles) & with $\frac{\Delta p}{p_0}=0$\\
-			turns tracked & $10^6$ turns & $10^4$ turns \\\hline
-			optics & \multicolumn{2}{c}{2016 or 2017 injection optics,}\\
-			& \multicolumn{2}{c}{$\beta_{\mathrm{IP1/5}}^*=11$~m}\\
-			machine  & standard errors & no errors \\
-			imperfections &with $a_1=b_1=0$\footnote{Orbit errors are disabled due to different implementation of the $a_1,b_1$ errors in Lifetrac and MAD-X.
-$b_2$ errors are adjusted to yield an average (over 60 seeds) peak $\beta$-beat of 15\% as expected from optics measurements in the LHC.} &  \\
-			octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MOF}}=+19.6~\mathrm{A}$, $I_{\mathrm{MOD}}=-19.6~\mathrm{A}$}\\
-			tune $(Q_x,Q_y)$ & \multicolumn{2}{c}{(64.28,59.31) for 2016,}\\
-			& \multicolumn{2}{c}{(62.27,60.295) for 2017}\\
-			chromaticity & \multicolumn{2}{c}{(+15,+15)}\\
-			$(Q'_x,Q'_y)$ & & \\\hline
-			transv. aperture & $5.7~\sigma$ & - \\
-			long. aperture & $10~\sigma$ & - \\
+      normalized emittance & \q{3.5}{\mu m} & \q{2.5}{\mu m} \\
+      $4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns} \\
+      $1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm} \\
+      particle distribution & \parbox{0.4\textwidth}{6D Gaussian
+                              distribution with $10^4$ particles}
+                               & \parbox{0.4\textwidth}{equally spaced
+                                 grid in $x, y$ up to $10\sigma$, with
+                                 $(\Delta p / p_0) = 0$} \\
+      turns tracked & $10^6$ & $10^4$ \\
+      \hline
+      optics & \multicolumn{2}{c}{2016 or 2017 injection optics, with
+               $\beta^* = \q{11}{m}$ at IP1 and IP5} \\
+      machine imperfections & standard errors, with $a_1 = b_1 =
+                              0$\footnote{Orbit
+                              errors are disabled due to
+                              different implementation of the
+                              $a_1, b_1$ coefficients in Lifetrac and
+                              MAD-X. $b_2$ errors are adjusted to
+                              yield an average
+                              peak $\beta$-beat of 15\% over 60~seeds, as
+                              observed in optics measurements
+                              in the LHC.} & no errors \\
+      octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MOF}} =
+                   \q{+19.6}{A}$, $I_{\mathrm{MOD}} = \q{-19.6}{A}$} \\
+      tunes $(Q_x, Q_y)$ & \multicolumn{2}{c}{(64.28, 59.31) for 2016,
+                           (62.27, 60.295) for 2017} \\
+      chromaticities $(Q'_x, Q'_y)$ & \multicolumn{2}{c}{(+15, +15)} \\
+      \hline
+      transverse aperture & $5.7\sigma$ &  \\
+      longitudinal aperture & $10\sigma$ &  \\
     \end{tabular}
   \end{ruledtabular}
-\end{table}
+\end{table*}
+
+For the preparation and interpretation of the experiments, two
+different types of simulations were performed:
+%
+\begin{itemize}
+\item tracking of a Gaussian distribution, referred to as
+  `distribution tracking' in this paper, to obtain particle loss rates
+  and emittance evolution;
+\item frequency-map analysis (FMA), to visualize the location and
+  intensity of resonances~\cite{fmalaskar}.
+\end{itemize}
+%
+For both simulation types, the tracking code Lifetrac~\cite{lifetrac}
+was used. The simulation parameters are summarized in
+Table~\ref{tab:sim_param}. Further details are given in
+Refs.~\cite{md_sim_hel_res_ex_fitterer, resexmd2017}.
+
+Simulation results are discussed below in Section~\ref{sec:simex}
+together with the experimental results to enhance the understanding
+and interpretation of the models and observations.
 

### 2018-03-01 resex.tex
Started revision of Section VI on results. Moved discussion of simulation systematics to Section V on simulations.

@@
 
-Simulation results are discussed below in Section~\ref{sec:simex}
-together with the experimental results to enhance the understanding
-and interpretation of the models and observations.
+Quantitative predictions of loss rates and emittance growth are in
+general challenging, as both observables are influenced by several
+factors. For instance, the natural noise present in the machine
+(originating from mechanical vibrations, current ripple in the magnet
+power supplies, etc.) may have complex interactions with the external
+excitations. In the LHC, noise sources at top energy are well
+characterized, whereas they are not well known at injection,
+presenting an influential but undefined input for simulations.
+
+Collective effects, such as intra-beam scattering and electron cloud,
+influence the time evolution of losses and emittance as well. To
+minimize their effects, experiments were done at low bunch intensities
+($0.7 \times 10^{11}$ protons per bunch). The presence of collective
+effects was neglected in these simulations.
+
+The results of tracking simulations are discussed below in
+Section~\ref{sec:simex} together with the experimental results, to
+enhance the understanding and interpretation of the models and
+observations.
 
@@
 
-In the following we will jump forth and back between both experimental and simulation results as the experimental results are best understood on the base of the simulations.
-In summary, the most efficient pulsing patterns could be successfully predicted in simulations, which we will demonstrate on the example of the 2016 experiments (see Sec.~\ref{sec:pattern}).
-Afterwards we will present the simulation and experimental results of the tested excitation patterns, explicitly the $10^{\mathrm{th}}$~turn, $7^{\mathrm{th}}$~turn and $8^{\mathrm{th}}$~turn pulsing in Secs.~\ref{sec:simex10}--\ref{sec:simex8} as well as the random excitation in Sec.~\ref{sec:simexran}.
-As a last chapter, Sec.~\ref{sec:damp}, we will address the influence of the transverse damper in the presence of an external excitation source on emittance, losses and beam distribution.
-
-Quantitative predictions of loss rates and emittance growth proved in general challenging as both observables are influenced in addition by the natural noise present in the LHC, originating for example from vibrations and magnetic field ripple.
-The observed effect is then the result of the interaction of the external excitation with the natural noise.
-At injection, this natural noise component is not known and thus presents a very influential but undefined input to the simulations.
-Beside the natural noise, also collective effects like intra-beam scattering and electron cloud influence the time evolution of emittance and losses.
-Both are also not taken into account in these simulations, but the effect was minimized during the experiments by choosing an as small as possible bunch intensity of only $0.7\times10^{11}$ protons per bunch.
+In this Section, we present the experimental observations and compare
+them with the results of numerical simulations.  In
+Section~\ref{sec:pattern}, we show which pulsing patterns are most
+efficient in the LHC, using the 2016 experiments as an example to show
+that these patterns can be predicted.  Next, we present results for
+each of the specific excitation patterns that could be tested
+experimentally, namely pulsing every $10^{\mathrm{th}}$~turn
+(Section~\ref{sec:simex10}), $7^{\mathrm{th}}$~turn
+(Section~\ref{sec:simex7}), $8^{\mathrm{th}}$~turn
+(Section~\ref{sec:simex8}), and the random excitation
+(Section~\ref{sec:simexran}).  In Section~\ref{sec:damp}, we describe
+how the transverse damping system influenced the effects of the
+external excitation sources on losses, beam distributions, and
+emittances.
 

### 2018-03-01 resex.tex
Intermediate step in revision of results section on pulsing patterns.

@@
 them with the results of numerical simulations.  In
-Section~\ref{sec:pattern}, we show which pulsing patterns are most
-efficient in the LHC, using the 2016 experiments as an example to show
-that these patterns can be predicted.  Next, we present results for
-each of the specific excitation patterns that could be tested
-experimentally, namely pulsing every $10^{\mathrm{th}}$~turn
+Section~\ref{sec:pattern}, we show which pulsing patterns are
+predicted to be the most efficient in the LHC.  Next, we present
+results for each of the specific excitation patterns that could be
+tested experimentally, namely pulsing every $10^{\mathrm{th}}$~turn
 (Section~\ref{sec:simex10}), $7^{\mathrm{th}}$~turn
@@
 
-\subsection{Dependence on the pulsing pattern\label{sec:pattern}}
-The effect of each pulsing pattern in comparison to the other patterns can be identified by the means of the loss rate and emittance growth from the distribution tracking.
-As an example, the simulation results for the 2016 experiment are shown in Fig.~\ref{fig:patternsim} featuring a clear dependence of both parameters on the pattern.
-\begin{figure*}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_intensity.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_sigm.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit1.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit2.png}
-	\end{minipage}	
-	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance for different pulsing patterns obtained from distribution tracking based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors.
-The resonant and random excitation respectively is applied in both planes with an amplitude of 96~nrad.
-No random noise component is added in addition.}
-\end{figure*}
-
-The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and a uniform random excitation.
-As the bunch length decreases in accordance with the losses, the losses can be mostly associated to off-momentum particles hitting the transverse aperture.
-Only $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as random excitation exhibit considerable emittance growth.
-Compared to any resonant excitation, the random excitation has by far the strongest effect.
-
-The sensitivity to $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing is also observed in absence of machine errors in which case the only source of non-linearities are sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer} suggesting that latter are responsible for the observed sensitivity.
-The driven resonances are revealed by the FMA analysis shown in Fig.~\ref{fig:patternfma}.
-The $7Q_x$ resonance is excited in case of the $7^{\mathrm{th}}$~turn pulsing and the $10Q_x$ and $10Q_y$ resonance in case of the $10^{\mathrm{th}}$~turn pulsing.
-As octupoles can only drive even resonances, the source of the $7Q_x$ resonances are the sextupoles while the octupoles only generate the tune footprint.
-The other pulsing patterns do not exhibit any increase in losses or emittance growth without magnetic errors and their effect can be thus attributed to the magnetic field errors, implying also the seed chosen for the simulation.
-\begin{figure*}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		no excitation
-		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$10^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$7^{\mathrm{th}}$ H
-		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut7skh_dp0_ord7.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$7^{\mathrm{th}}$ V
-		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut7skv_dp0_ord7.png}
-	\end{minipage}
-	\caption{\label{fig:patternfma} FMA analysis without excitation (top left) for $10^{\mathrm{th}}$~turn pulsing in H+V (top left) and for $7^{\mathrm{th}}$~pulsing only in H (bottom left) and only in V (bottom right) and  based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
-The excitation is 120~nrad in the corresponding plane.
-The absence of a strong excitation of any resonance in case of $7^{\mathrm{th}}$~turn pulsing only in V and the strong excitation in case of pulsing only in H confirms the excitation of the $7Q_x$ resonance.
-For $10^{\mathrm{th}}$~turn pulsing there is in contrast no significant difference between pulsing only in H, only in V or in H+V (see Appendix~\ref{app:sec:10}, Fig.~\ref{app:fig:fma:10}).}
-\end{figure*}
-
-One may have noticed that the emittance in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an increased initial value and then stays almost constant during the entire simulation time, which usually points to an error in the chosen simulation setup.
-This behavior is however typical for any resonant excitation and can be ascribed to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium.
-As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}, and also any other simulations over $10^6$ turns presented in this paper.
-Besides an increased emittance, the new distribution also becomes non-Gaussian.
-This is illustrated by means of the residual in respect to the Gaussian distribution for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in Fig.~\ref{fig:patternhist}.
-This initial fast change of the beam distribution is not an artifact of the simulation, but has been, to the knowledge of the authors, also for the first time experimentally observed during the 2016 and 2017 LHC experiments presented in this paper.
-The timescale of this fast adjustment is in experiments however much slower than predicted in simulations --- on the timescale of minutes compared to seconds.
-\begin{figure*}[t]
+\subsection{Dependence on the pulsing pattern}
+\label{sec:pattern}
+
+\begin{figure*}
+\begin{tabular}{cc}
+\includegraphics[width=0.49\textwidth]{2016injerra2b2u_pattern_3_5um_intensity.png}
+  &
+\includegraphics[width=0.49\textwidth]{2016injerra2b2u_pattern_3_5um_sigm.png}
+  \\
+\includegraphics[width=0.49\textwidth]{2016injerra2b2u_pattern_3_5um_emit1.png} &
+\includegraphics[width=0.49\textwidth]{2016injerra2b2u_pattern_3_5um_emit2.png}
+  \\
+\end{tabular}
+\caption{Relative intensity (top left), bunch length (top right) and
+  horizontal (bottom left) and vertical (bottom right) emittances for
+  different pulsing patterns, calculated by distribution tracking
+  based on the 2016 injection optics with
+  $(Q_x, Q_y) = (64.28, 59.31)$ and standard lattice errors. The
+  resonant and random excitations are applied in both planes, with an
+  amplitude of 96~nrad. No random noise component is added. Because of
+  its much larger effects, the random excitation is shown with
+  separate vertical axes.}
+\label{fig:patternsim}
+\end{figure*}
+
+\begin{figure*}
+  \begin{tabular}{cc}
+    no excitation & $10^{\mathrm{th}}$, H+V \\
+    \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png} &
+    \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
+    \\
+    $7^{\mathrm{th}}$, H & $7^{\mathrm{th}}$, V \\
+    \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
+    \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
+  \end{tabular}
+  \caption{Frequency-map analysis in betatron tune space based on the
+    2016 injection optics with no machine errors and tunes (64.28,
+    59.31): without excitation (top left); $10^{\mathrm{th}}$~turn
+    pulsing in both horizontal and vertical planes (top left);
+    $7^{\mathrm{th}}$~turn pulsing in the horizontal (bottom left) and
+    vertical (bottom right) planes. The excitation amplitude is
+    120~nrad in the corresponding plane. The colors (blue to red)
+    represent the tune jitter of tracked particles starting at each
+    given location in tune space~\cite{fmalaskar}. The absence of a
+    strong excitation of any resonance in case of vertical
+    $7^{\mathrm{th}}$~turn pulsing and the strong excitation in case
+    of horizontal pulsing confirms the excitation of the $7 Q_x$
+    resonance. For $10^{\mathrm{th}}$~turn pulsing, there is in
+    contrast no significant difference between H, V, or H+V (see
+    Appendix~\ref{app:sec:10}, Fig.~\ref{app:fig:fma:10}).}
+  \label{fig:patternfma}
+\end{figure*}
+
+\begin{figure*}
 	\begin{minipage}[t]{0.49\linewidth}
@@
 
-In the presence of only a resonant excitation, this modified distribution is stable in simulations with a constant emittance.
-As we will see later, an increase of the emittance with the applied excitation amplitude is however observed experimentally not matching the prediction by simulations.
-By adding a random noise component representative for the natural noise present in the LHC, the constant emittance is changed in simulations to an excitation amplitude dependent emittance growth after an initial adjustment phase.
-In case of random uniform noise, the excitation leads to constant emittance growth without any initial adjustment phase.
-The interaction with the natural noise source then only results in an underestimation of loss rates and emittance growth as it is basically equivalent to the application of random noise with an increased amplitude.
-In addition to the experiments presented in this paper, the effect of a random excitation was also studied in a separate experiment at the LHC for the case of colliding beams \cite{md1433_noise_top_energy,md_noise_bbLHC}.
-
-\clearpage
+The effect of each pulsing pattern is characterized in terms of the
+resulting losses and emittance growth, calculated in
+distribution-tracking simulations.
+
+As an example, the simulation results for the 2016 experiment are
+shown in Fig.~\ref{fig:patternsim}.  A clear dependence of both losses
+and emittances on the pattern can be seen.  The largest losses are
+predicted for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and
+$10^{\mathrm{th}}$~turn pulsing, and for a uniform random
+excitation. Because the bunch length decreases with the number of lost
+particles, losses are attributed to off-momentum particles hitting the
+transverse aperture. Significant emittance growth is visible only for
+$7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and for the
+random excitation. Compared to any resonant excitation, the random
+excitation shows by far the strongest effect.
+
+The effects of $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing
+are also observed in the absence of machine lattice errors. In this
+case, the only sources of nonlinearity are sextupoles and
+octupoles~\cite{md_sim_hel_res_ex_fitterer}, suggesting that these
+nonlinearities are responsible for the pronounced beam
+sensitivity. The driven resonances are revealed by the frequency-map
+analysis, shown in Fig.~\ref{fig:patternfma}. The $7 Q_x$ resonance is
+excited by the $7^{\mathrm{th}}$~turn pulsing, and the $10 Q_x$ and
+$10 Q_y$ resonances by the $10^{\mathrm{th}}$~turn pulsing. As
+octupoles can only drive even resonances, the sources of the $7 Q_x$
+resonances are the sextupoles, while the octupoles generate the tune
+footprint. The other pulsing patterns do not exhibit any increase in
+losses or emittance growth without magnetic errors. Their effect can
+be thus attributed to an interaction of the excitation with the
+lattice errors, implying also a sensitivity to the chosen random seed
+in simulations and to the specific lattice in experiments.
+
+One may have noticed that the emittance in case of the
+$7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an
+increased initial value and then stays almost constant during the
+entire simulation time, which usually points to an error in the chosen
+simulation setup. This behavior is however typical for any resonant
+excitation and can be ascribed to the adjustment of the beam
+distribution during the first $10^4$ turns to a new equilibrium. As
+the beam distribution is only dumped every $10^4$ turns, this initial
+fast adjustment manifests itself in an increased initial emittance
+value in Fig.~\ref{fig:patternsim}, and also any other simulations
+over $10^6$ turns presented in this paper. Besides an increased
+emittance, the new distribution also becomes non-Gaussian. This is
+illustrated by means of the residual in respect to the Gaussian
+distribution for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing
+in Fig.~\ref{fig:patternhist}. This initial fast change of the beam
+distribution is not an artifact of the simulation, but has been, to
+the knowledge of the authors, also for the first time experimentally
+observed during the 2016 and 2017 LHC experiments presented in this
+paper. The timescale of this fast adjustment is in experiments however
+much slower than predicted in simulations --- on the timescale of
+minutes compared to seconds.
+
+In the presence of only a resonant excitation, this modified
+distribution is stable in simulations with a constant emittance. As we
+will see later, an increase of the emittance with the applied
+excitation amplitude is however observed experimentally not matching
+the prediction by simulations. By adding a random noise component
+representative for the natural noise present in the LHC, the constant
+emittance is changed in simulations to an excitation amplitude
+dependent emittance growth after an initial adjustment phase. In case
+of random uniform noise, the excitation leads to constant emittance
+growth without any initial adjustment phase. The interaction with the
+natural noise source then only results in an underestimation of loss
+rates and emittance growth as it is basically equivalent to the
+application of random noise with an increased amplitude. In addition
+to the experiments presented in this paper, the effect of a random
+excitation was also studied in a separate experiment at the LHC for
+the case of colliding beams
+\cite{md1433_noise_top_energy,md_noise_bbLHC}.
 

### 2018-03-02 resex.tex
Another intermediate step in the revision of the results section: discussion and plot of calculated beam re-distribution.

@@
 \begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$7^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$10^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		random H+V
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		no excitation
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_3_5um_hist_y.png}
-	\end{minipage}	
-	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn pulsing (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from distribution tracking simulations based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ and standard errors.
-The excitation is applied in both planes with an amplitude of 96~nrad.
-The residual is defined as the final distribution (here after $10^4$~turns) minus the initial distribution (here after $10^2$~turns) as \%, and the ratio as the final distribution divided by the initial distribution.
-The red line in the plot of the residual is the difference between the Gaussian fit of the distribution and the distribution itself.}
+  \begin{tabular}{cc}
+    no excitation & random, H+V \\
+    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_3_5um_hist_y.png} &
+    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
+    $7^{\mathrm{th}}$, H+V & $10^{\mathrm{th}}$, H+V \\
+    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
+                  &
+    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png} \\
+  \end{tabular}
+  \caption{Calculated beam distributions as a function of vertical
+    position from distribution-tracking simulations based on the 2016
+    injection optics with $(Q_x, Q_y) = (64.28, 59.31)$ and standard
+    lattice errors: no excitation (top left), random excitation (top
+    right), $7^{\mathrm{th}}$~turn pulsing (bottom left), and
+    $10^{\mathrm{th}}$~turn pulsing (bottom right). The excitations
+    are applied in both planes with an amplitude of 96~nrad. For each
+    of the 4~cases, 3~plots are shown. The top plot shows the
+    normalized transverse distributions: `initial' (after
+    $10^2$~turns, in gray), `final' (after $10^4$~turns, in black),
+    and their Gaussian fits (light and dark red, respectively). The
+    middle plots show the relative residuals (i.e., differences, in
+    percent of the peak value) between final and initial distributions
+    (in black) and between each distribution and its Gaussian fit (in
+    light and dark red). The ratios between final and initial
+    distributions are drawn in black in the bottom plots. The
+    residuals emphasize changes near the core of the distributions,
+    whereas ratios (when statistically significant), reveal variations
+    in the tails.}
+  \label{fig:patternhist}
 \end{figure*}
@@
 
-One may have noticed that the emittance in case of the
-$7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing starts at an
-increased initial value and then stays almost constant during the
-entire simulation time, which usually points to an error in the chosen
-simulation setup. This behavior is however typical for any resonant
-excitation and can be ascribed to the adjustment of the beam
-distribution during the first $10^4$ turns to a new equilibrium. As
-the beam distribution is only dumped every $10^4$ turns, this initial
-fast adjustment manifests itself in an increased initial emittance
-value in Fig.~\ref{fig:patternsim}, and also any other simulations
-over $10^6$ turns presented in this paper. Besides an increased
-emittance, the new distribution also becomes non-Gaussian. This is
-illustrated by means of the residual in respect to the Gaussian
-distribution for $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing
-in Fig.~\ref{fig:patternhist}. This initial fast change of the beam
-distribution is not an artifact of the simulation, but has been, to
-the knowledge of the authors, also for the first time experimentally
-observed during the 2016 and 2017 LHC experiments presented in this
-paper. The timescale of this fast adjustment is in experiments however
-much slower than predicted in simulations --- on the timescale of
-minutes compared to seconds.
+In Fig.~\ref{fig:patternsim}, one can see that, in the cases of
+$7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing, the emittances
+start from an increased initial value and then stay almost constant
+for the duration of the simulation. This behavior is typical of the
+resonant excitation and it is not an artifact of the simulation. It
+was verified that it is due to the adjustment of the input beam
+distribution to a new equilibrium during the first $10^4$~turns. In
+all long-term simulations ($10^6$~turns, corresponding to 89~s)
+presented in this paper, the beam distribution is saved every
+$10^4$~turns. Therefore, this initial fast adjustment manifests itself
+in an increased initial emittance value.
+
+Besides having a larger emittance, the new distribution also differs
+from a Gaussian distribution.  This is illustrated in
+Fig.~\ref{fig:patternhist}, which shows, for different excitation
+patterns, two snapshots of the normalized transverse beam
+distributions, the `initial' one at $10^2$~turns and the `final' one
+at $10^4$~turns, and how they differ from each other and from a
+Gaussian distribution. For instance, for $7^{\mathrm{th}}$ turn
+pulsing (bottom left plot in Fig.~\ref{fig:patternhist}), simulations
+indicate a clear shift in beam population from the core ($< 1\sigma$)
+towards the regions at $\pm 1.5\sigma$ and towards the tails at
+$\pm 4\sigma$.
+
+Although on a different time scale (minutes instead of seconds), a
+similar re-adjustment of the beam distribution in response to an
+external excitation was directly observed experimentally for the first
+time, to our knowledge, during the 2016 and 2017 LHC experiments
+presented in this paper, as discussed below.
 

### 2018-03-02 resex.tex
Revised results section on pattern dependence (VI A).

@@
 
-In the presence of only a resonant excitation, this modified
-distribution is stable in simulations with a constant emittance. As we
-will see later, an increase of the emittance with the applied
-excitation amplitude is however observed experimentally not matching
-the prediction by simulations. By adding a random noise component
-representative for the natural noise present in the LHC, the constant
-emittance is changed in simulations to an excitation amplitude
-dependent emittance growth after an initial adjustment phase. In case
-of random uniform noise, the excitation leads to constant emittance
-growth without any initial adjustment phase. The interaction with the
-natural noise source then only results in an underestimation of loss
-rates and emittance growth as it is basically equivalent to the
-application of random noise with an increased amplitude. In addition
-to the experiments presented in this paper, the effect of a random
-excitation was also studied in a separate experiment at the LHC for
-the case of colliding beams
-\cite{md1433_noise_top_energy,md_noise_bbLHC}.
+If only the resonant excitation is present, without noise, simulations
+show that the modified distribution is stable, with constant
+emittance. By adding a random noise component, representative of the
+natural noise present in the LHC, one calculates, after the initial
+adjustment phase, emittance-growth rates that increase with the
+amplitude of the excitation. As discussed below, an increase of the
+emittance-growth rates with the applied excitation amplitude is what
+is observed experimentally.
+
+In the case of random excitation, there is constant emittance growth,
+without any initial adjustment phase. The interaction with the natural
+noise sources in the machine results in increased losses and emittance
+growth, as it is basically equivalent to the application of a random
+excitation with an increased amplitude. In addition to the experiments
+presented in this paper, the effect of a random excitation was also
+studied in a separate experiment at the LHC for the case of colliding
+beams~\cite{md1433_noise_top_energy, md_noise_bbLHC}.
 

### 2018-03-03 resex.tex
Revised text of results on 10th-turn pulsing. Moved estimate of random noise to simulation section.

@@
 characterized, whereas they are not well known at injection,
-presenting an influential but undefined input for simulations.
+presenting an influential but undefined input for simulations.  An
+estimate is obtained by scaling the value at
+6.5~TeV~\cite{md1433_noise_top_energy, md_noise_bbLHC} with proton
+magnetic rigidity to the injection energy of 450~GeV. This yields a
+maximum kick amplitude at the transverse damper of approximately
+%
+\begin{equation}
+  \label{eq:noise-kick}
+  \theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = \q{6}{nrad}.
+\end{equation}
 
@@
 
-\subsection{$10^{\mathrm{th}}$ turn pulsing\label{sec:simex10}}
-The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an excitation in the vertical plane only.
-As this was the first time a resonant excitation was tested in the LHC, only in total 48~bunches could be injected in order to still guarantee the protection of the machine.
-To test different excitation amplitudes during one fill and have enough statistics, the excitation was therefore only applied in one plane.
-As described earlier, the experiment started with a period of 12~minutes without excitation in order to let the beam distribution fully adjust to its equilibrium state after injection.
-Afterwards the excitation was applied with the excitation scheme shown in Fig.~\ref{fig:fill} and a maximum excitation amplitude of $A_{\mathrm{max}}=5\Delta A = 48~$nrad for a total of 11~minutes.
-The excitation was then further increased to the maximum value reachable without saturation of $A_{\mathrm{max}}=5\Delta A = 96~$nrad and kept for 11~minutes.
-In both cases, the excitation induced the following changes:
-\begin{enumerate}
-	\item Increase of the loss rate with the excitation amplitude (Figs.~\ref{fig:10thexp}.
-	\item Increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane (Fig.~\ref{fig:10thexp}).
-	\item Change of the beam distribution for the excited bunches (Fig.~\ref{fig:10thexpprof}).	
-\end{enumerate}
-As loss rates, emittance growth and beam distribution stayed unchanged for the $4$~reference bunches, the above observations can be also truly associated with the effect of the $10^{\mathrm{th}}$~turn pulsing.
-\begin{figure*}[h]
+\subsection{Pulsing every $10^{\mathrm{th}}$ turn}
+\label{sec:simex10}
+
+\begin{figure*}
 	\begin{minipage}[t]{0.32\linewidth}
@@
 \end{figure*}
-These experimentally obtained scalings of the loss rate and emittance growth with the excitation amplitude can be then used for the actual specification of the hollow electron lens.
-
-Besides the numerical determination of tolerances, the vertical emittance actually features indeed the strange and interesting behavior predicted in simulations: a fast adjustment phase of the beam distribution visible as a rapid increase of the emittance followed by the reach of a new equilibrium distribution apparent by a slower continuous emittance growth.
-These two phases are indicated in Fig.~\ref{fig:10thexp} in blue and black.
-
-\begin{figure*}[h]
+
+\begin{figure*}
 	\begin{minipage}[t]{0.49\linewidth}
@@
 \end{figure*}
-Thanks to the good performance of the Beam Synchrotron Radiation Telescope (BSRT)~\cite{bsrtprofinj}, the distribution changes indirectly observed as changes of the emittance could also be for the first time directly measured.
-As example Fig.~\ref{fig:10thexpprof} shows the vertical profile of a reference bunch and one bunch experiencing the maximum excitation of $5\cdot\Delta A=A_{\mathrm{max}}=48$~nrad and afterwards 96~nrad.
-The distribution of the excited bunch not only changes but becomes clearly non-Gaussian, while the reference bunch stays unchanged.
-\begin{figure*}[h]
+
+\begin{figure*}
 	\begin{minipage}[t]{0.49\linewidth}
@@
 \end{figure*}
-The simulation results from the distribution tracking for $10^{\mathrm{th}}$~turn pulsing and without (solid line) and with (dotted line) additional random noise component to emulate the natural noise present in the LHC are shown in Fig.~\ref{fig:10thsim}.
-As no estimate of the natural noise is available at injection for the LHC, a first estimate is obtained by scaling the estimate for 6.5~TeV \cite{md1433_noise_top_energy,md_noise_bbLHC} by the magnetic rigidity to the injection energy of 450~GeV.
-This yields a maximum kick amplitude at the transverse damper of approximately
-\begin{equation}
-\theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = 6~\mathrm{nrad}.
-\end{equation}
-As can be seen from simulations, adding a random noise component not only yields the correct qualitative behavior of the emittance but also increases the losses.
-As the losses are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment can be thus reduced.
-To better compare the experimental results for the two different maximum excitation amplitudes applied for slightly different time intervals and also the simulations conducted for a much shorter interval of only 90~s, we define the relative loss rate $R$:
-\begin{equation}\label{eqn:lossrate}
-R = \frac{I_{\mathrm{start}}-I_{\mathrm{end}}}{I_{\mathrm{start}}\cdot \Delta t}
-\end{equation}
-where $I$ is the intensity and $\Delta t$ the time interval during which the excitation with fixed $A_{\mathrm{max}}$ is applied during the experiments or the duration of the simulation respectively.
-A qualitative comparison of the loss rates obtained in experiments and simulations is shown in Fig.~\ref{fig:10thexploss} featuring a surprisingly good agreement.
-It should be also kept in mind that the additional random noise component considerably influences the simulation results and as long as it is unknown should be seen more as a fit parameter than a real experimentally confirmed input parameter.
-The measured loss rates for $A_{\mathrm{max}}=48$~nrad and $A_{\mathrm{max}}=96$~nrad should furthermore follow the same scaling.
-Instead the loss rates for the same excitation amplitude are smaller for $A_{\mathrm{max}}=96$~nrad compared to $A_{\mathrm{max}}=48$~nrad.
-That losses and also emittance growth for a fixed excitation amplitude do not exactly agree for both experiments, is a general tendency observed in all experiments.
-As the second excitation however starts from an already by the previous excitation changed beam distribution, the difference is however not surprising and can be explained by this difference in initial conditions.
-For checking the reproducibility of the results in a possible future experiment, it would be interesting to see if the same loss rates are measured if for each $A_{\mathrm{max}}$ a new fill is used.
-\begin{figure}[h]
+
+\begin{figure}
 		\centering
@@
 
-\clearpage
+The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an
+excitation in the vertical plane only. As this was the first time a
+resonant excitation was tested in the LHC, only 48~bunches were
+injected, in order to guarantee safe operation of the machine. To test
+different excitation amplitudes during one fill and have enough
+statistics, the excitation was only applied in one plane. The
+experiment started with a period of 12~minutes without excitation, to
+let the beam distribution fully adjust to its equilibrium state after
+injection. The excitation was then applied for 11~minutes, with the
+excitation scheme shown in Fig.~\ref{fig:fill} and a maximum
+excitation amplitude of
+$A_{\mathrm{max}} = 5 \, \Delta A = \q{48}{nrad}$. The excitation was
+then further increased to the maximum value of
+$A_{\mathrm{max}} = 5 \, \Delta A = \q{96}{nrad}$ and kept for another
+period of 11~minutes.
+
+The main observations are collected in Figures~\ref{fig:10thexp}
+and~\ref{fig:10thexpprof}. In Fig.~\ref{fig:10thexp}, one can see the
+evolution in time of losses and emittances for the control bunches and
+for the affected bunches, as a function of excitation amplitude. In
+brief, during both phases of the experiment, the excitation induced
+the following changes:
+%
+\begin{itemize}
+  \item loss rates increased with excitation amplitude
+    (Fig.~\ref{fig:10thexp}, left);
+  \item emittance growth increased with amplitude in the vertical
+    plane, but not in the horizontal plane (Fig.~\ref{fig:10thexp},
+    center and right);
+  \item a change of the beam distribution for the excited bunches was
+    directly observed (Fig.~\ref{fig:10thexpprof}).
+\end{itemize}
+
+As loss rates, emittance-growth rates and beam distributions of the 4
+reference bunches were unchanged, the above observations can be
+directly attributed to the pulsing pattern. The quantitative
+dependence of the loss rates and emittance-growth rates on excitation
+amplitude measured in these experiments provides an estimate of the
+effects of external resonant excitations, such as the residual fields
+of a pulsed hollow electron lens.
+
+The vertical emittance features a behavior similar to that predicted
+in simulations: a fast adjustment phase of the beam distribution,
+which manifests itself as a rapid increase of the emittance, followed
+by a new equilibrium, inferred from a slower and continuous emittance
+growth. In Fig.~\ref{fig:10thexp}, these two phases are indicated in
+blue and black.
+
+Distribution changes could also be directly observed, thanks to the
+performance of the Beam Synchrotron Radiation Telescope
+(BSRT)~\cite{bsrtprofinj}. As an example, Fig.~\ref{fig:10thexpprof}
+shows the vertical profile of a reference bunch and of a bunch
+experiencing the maximum excitation
+$A_{\mathrm{max}} = 5 \, \Delta A = \q{48}{nrad}$ and later
+96~nrad. While the reference bunch stays unchanged, the distribution
+of the excited bunch clearly changes and re-adjusts to a non-Gaussian
+shape, with particles shifting from the core towards the regions at
+approximately $\pm 2\sigma$. This behavior is similar to what was
+calculated in simulations, as discussed in Section~\ref{sec:pattern}
+and in Fig.~\ref{fig:patternhist}.
+
+Numerical simulations of intensities and emittances for this
+experiment are shown in Fig.~\ref{fig:10thsim}. They are the results
+of distribution tracking for $10^{\mathrm{th}}$~turn pulsing without
+additional noise (solid lines) and with an additional noise component
+(dotted lines), to emulate the natural noise present in the LHC
+(Eq.~\ref{eq:noise-kick}). The presence of noise significantly changes
+the calculated effect of the resonant excitation on both losses and
+emittances.
+
+A direct comparison of measured and predicted loss rates is presented
+in Fig.~\ref{fig:10thexploss}. As experiments and simulations were
+conducted on different time intervals (11~minutes vs.\ 90~s), we
+compare the relative average loss rate~$R$:
+%
+\begin{equation}
+  \label{eqn:lossrate}
+  R = \frac{I_{\mathrm{start}} - I_{\mathrm{end}}}{I_{\mathrm{start}}}
+  \cdot \frac{1}{\Delta t},
+\end{equation}
+%
+where $I$ is the beam intensity and
+$\Delta t = t_\mathrm{end} - t_\mathrm{start}$ is the time interval
+during which the excitation with given amplitude is
+applied. Uncertainties on the excitation amplitude arise from the
+calibration of the transverse damping system
+(Section~\ref{sec:adt}). For both experiments and simulations, the
+error bars on the loss rate are statistical. The magnitude of the
+random noise component introduces a systematic uncertainty in the
+calculations, which was not evaluated for the scope of this paper. The
+systematic uncertainty on the measurements can be inferred from the
+difference between the 2~periods of the experiment, which should in
+principle be the same. Part of this systematic effect could be due to
+the fact that the second set of excitations with double amplitude
+started with a modified beam distribution. The magnitude of this
+effect could be verified in future experiments in which the
+excitations are reversed or where a new beam fill is used for each set
+of measurements. Overall, we conclude that the effects of this
+resonant excitation could be measured in the LHC within a factor~2 and
+could be predicted with similar accuracy.
 

### 2018-03-05 resex.tex
Revised Fig. 11 on 10th-turn measurements.

@@
 \begin{figure*}
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_v10th_no_damper_avg.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_v10th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_no_damper.png}
-	\end{minipage}	
-	\caption{\label{fig:10thexp} Relative intensity (top) and relative horizontal (bottom left) and vertical (bottom right) emittance averaged over the bunches experiencing the same excitation amplitude during the 2016 experiments and for $10^{\mathrm{th}}$~turn pulsing in~V.
-For all bunches the damper was not active.
-Relative refers here to the normalization to the initial value.
-For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad indicated with black arrows, the excitation amplitude was linearly increased for each group of 4~bunches (see Fig.~\ref{fig:fill}).
-The different colored solid lines labeled with $n\Delta A$ are the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation shown as envelope.
-The area with a blue background indicates the fast adjustment period of the beam distribution transitioning into a new equilibrium state highlighted with a gray background.}
+  \begin{tabular}{ccc}
+    \includegraphics[width=0.32\textwidth]{2016_bunch_intensity_v10th_no_damper_avg.png}&
+    \includegraphics[width=0.32\textwidth]{2016_emith_avg_rel_v10th_no_damper.png}&
+    \includegraphics[width=0.32\textwidth]{2016_emitv_avg_rel_v10th_no_damper.png}\\
+  \end{tabular}
+  \caption{Summary of the 2016 experiments on $10^{\mathrm{th}}$~turn
+    pulsing in the vertical plane: losses (left), horizontal
+    emittances (middle) and vertical emittances (right), relative to
+    their initial values. The transverse damping system was not active
+    in this case. The 3 excitation periods are labeled in black
+    according to the value of the maximum excitation amplitude
+    $A_{\mathrm{max}} = 5 \, \Delta A$: no excitation, 48~nrad or
+    96~nrad. The 4~bunches experiencing the same excitation amplitude
+    $n \, \Delta A$ ($n = 0, \ldots, 5$) are grouped by color. The
+    data are averaged over the 4~bunches, with the envelope
+    representing the standard deviation. The area with a blue
+    background highlights qualitatively the fast adjustment period of
+    the beam distribution transitioning into a new equilibrium state
+    (indicated by the gray background).}
+  \label{fig:10thexp}
 \end{figure*}

### 2018-03-05 resex.tex
Revised figure of beam profiles in 10th-turn experiments. Added general reference to Trad's thesis for BSRT. Updated to-do list.

@@
 current transformers (FBCTs). Transverse beam profiles and emittances
-were provided by the beam synchrotron radiation telescope (BSRT). All
-instruments were capable of delivering bunch-by-bunch data. The data
-analysis of the BSRT profiles is quite involved. In this paper, we
-focus on the direct comparison of the resulting profiles. A detailed
-description of the profile analysis can be found in
-Ref.~\cite{bsrtprofinj}, with individual experiments reported in
-Refs.~\cite{resexmd2016, resexmd2017}.
+were provided by the beam synchrotron radiation telescope
+(BSRT)~\cite{Trad:PhD:2015}. All instruments were capable of
+delivering bunch-by-bunch data. The data analysis of the BSRT profiles
+is quite involved. In this paper, we focus on the direct comparison of
+the resulting profiles. A detailed description of the profile analysis
+can be found in Ref.~\cite{bsrtprofinj}, with individual experiments
+reported in Refs.~\cite{resexmd2016, resexmd2017}.
 
@@
     background highlights qualitatively the fast adjustment period of
-    the beam distribution transitioning into a new equilibrium state
+    the beam distribution, transitioning into a new equilibrium state
     (indicated by the gray background).}
@@
 \begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch, no excitation
-		\includegraphics[width=1.0\linewidth]{profile_v_10thv_slot_50.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch, $10^{\mathrm{th}}$~turn V		\includegraphics[width=1.0\linewidth]{profile_v_10thv_slot_1300.png}
-	\end{minipage}
-	\caption{\label{fig:10thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2016 experiments.
-The profiles are taken at the end of the $10^{\mathrm{th}}$~turn pulsing in V.
-For both bunches the transverse damper is not active.
-The black line in the plot of the residual is the current profile minus the profile before applying the excitation and is a measure for the overall change of the distribution.
-The red line is the current profile minus its Gaussian fit, which gives and indication for the deviation of the distribution from a Gaussian distribution.
-For details on the analysis it is referred to \cite{bsrtprofinj}.
-The distribution of the reference bunch (left), here slot 50, stays unchanged, while the bunch experiencing the maximum excitation (right), here slot 1300, clearly shows a change to a non-Gaussian distribution.}
+  \begin{tabular}{cc}
+    Reference bunch, no excitation & Excited bunch, maximum amplitude \\
+    \includegraphics[width=0.48\textwidth]{profile_v_10thv_slot_50.png} &
+    \includegraphics[width=0.48\textwidth]{profile_v_10thv_slot_1300.png}\\
+    \end{tabular}
+    \caption{Vertical beam profiles measured with the Beam Synchrotron
+      Radiation Telescope (BSRT) during the 2016 experiments on
+      $10^{\mathrm{th}}$~turn pulsing in the vertical plane. The
+      transverse damping system was not active on these bunches. The
+      beam distributions at the end of the excitation period are shown
+      in black in the top plots, with a Gaussian fit in red. The
+      bottom plots show the residuals: final profile minus initial
+      profile (black); final profile minus its Gaussian fit
+      (red). Residuals are expressed as a fraction of the peak profile
+      amplitude. The black lines in the plots of the residuals are a
+      measure of the overall change of the distribution shape. The red
+      lines indicate how different the final distributions are from a
+      Gaussian shape. Details of the analysis are given in
+      Ref.~\cite{bsrtprofinj}. The distribution of the reference bunch
+      (left) is almost unchanged, whereas the bunch experiencing the
+      maximum excitation (right) shows a clear shift of particles from
+      the axis towards approximately $\pm 2\sigma$.}
+    \label{fig:10thexpprof}
 \end{figure*}
@@
 performance of the Beam Synchrotron Radiation Telescope
-(BSRT)~\cite{bsrtprofinj}. As an example, Fig.~\ref{fig:10thexpprof}
-shows the vertical profile of a reference bunch and of a bunch
-experiencing the maximum excitation
+(BSRT)~\cite{Trad:PhD:2015, bsrtprofinj}. As an example,
+Fig.~\ref{fig:10thexpprof} shows the vertical profile of a reference
+bunch and of a bunch experiencing the maximum excitation
 $A_{\mathrm{max}} = 5 \, \Delta A = \q{48}{nrad}$ and later

### 2018-03-05 resex.tex
Revised results section on 10th-turn pulsing.

@@
 \begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_intensity.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2_rel.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit1.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2.png}
-	\end{minipage}	
-	\caption{\label{fig:10thsim} Relative intensity (top left) and relative vert.
-emittance (top right), and hor.
-(bottom left) and vert.
-(bottom right) emittance obtained from simulations (distribution tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$.
-The solid line indicates an excitation with only $10^{\mathrm{th}}$~turn pulsing in V and the dotted line with $10^{\mathrm{th}}$~turn pulsing in V plus a random dipole noise component in H+V of 6~nrad.}
+  \begin{tabular}{cc}
+    \includegraphics[width=0.49\textwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_intensity.png} &
+    \includegraphics[width=0.49\textwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2_rel.png} \\
+    \includegraphics[width=0.49\textwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit1.png} &
+    \includegraphics[width=0.49\textwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2.png} \\
+  \end{tabular}	
+  \caption{Simulations (with the distribution-tracking method) based
+    on the 2016 injection optics with standard lattice errors and
+    tunes $(Q_x, Q_y) = (64.28, 59.31)$: relative beam intensity (top
+    left); normalized horizontal (bottom left) and vertical (bottom
+    right) emittances; vertical emittance relative to its initial
+    value (top right). The solid lines indicate the effect of
+    $10^{\mathrm{th}}$~turn pulsing in the vertical plane. The dotted
+    lines include both the resonant excitation and a random dipole
+    noise component of 6~nrad (in both horizontal and vertical
+    planes).}
+  \label{fig:10thsim}
 \end{figure*}
@@
   \centering
-%		\includegraphics[width=0.8\linewidth]{2016_scale_amp_10v_ran_lbllong_sim.png}
-\includegraphics[width=1.0\linewidth]{2016_scale_amp_10v_ran_lblshort_sim_prstab.png}
-	\caption{\label{fig:10thexploss} Comparison of scaling of loss rates $R$ as defined in Eqn.~\ref{eqn:lossrate} with the excitation amplitude for $10^{\mathrm{th}}$~turn pulsing in~V.
-The experimental results for a maximum excitation of $A_{\mathrm{max}}=48$~nrad are shown in black and for $A_{\mathrm{max}}=96$~nrad in green.
-The excitation amplitude error is 50\% and the error on the loss rate is the pure statistical error on the mean.
-The simulation results (see Fig.~\ref{fig:10thsim}) for 2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad are shown in red together with their statistical error.
-The loss rates follow approximately a quadratic behavior illustrated by a second order polynomial fit to the data shown as solid (for bunches with transverse damper off) or dashed (for bunches with transverse damper on) line.}
+  \includegraphics[width=\linewidth]{2016_scale_amp_10v_ran_lblshort_sim_prstab.png}
+  \caption{Comparison between experiments and simulations of loss
+    rates vs.\ resonant excitation amplitude for
+    $10^{\mathrm{th}}$~turn pulsing in the vertical plane. The
+    relative average loss rate~$R$ is defined in
+    Eq.~\ref{eqn:lossrate}. The experimental results for a maximum
+    excitation amplitude of 48~nrad are shown in black and those for
+    96~nrad are plotted in green (see also Fig.~\ref{fig:10thexp},
+    left, for instance). The results with damper off are represented
+    by solid dots, those with damper on are shown with open
+    circles. The simulation results, including random dipole noise,
+    are shown in red (see also Fig.~\ref{fig:10thsim}). Statistical
+    and systematic uncertainties are discussed in the text. The curves
+    represent empirical quadratic fits to the data with damper off
+    (solid) and with damper on (dashed).}
+  \label{fig:10thexploss}
 \end{figure}

### 2018-03-05 resex.tex
Intermediate step in revision of 7th-turn results.

@@
 
-\subsection{$7^{\mathrm{th}}$ turn pulsing\label{sec:simex7}}
-The $7^{\mathrm{th}}$~turn pulsing pattern has been tested during both experiments in 2016 and 2017.
-In 2016 the resonant excitation was tested for the first time in the LHC and the experiments were therefore still in the exploratory state.
-Based on the experience gained in 2016, the experiments could then be repeated more systematically with more bunches allowing to also test in addition the dependence on the excitation plane.
-Also in case of the $7^{\mathrm{th}}$~turn pulsing, the experiment in 2016 as well as in 2017 can be divided in different parts: the first part without excitation, the second part with a resonant excitation with a maximum amplitude of $A_{\mathrm{max}}=6$~nrad, a third part in which the excitation amplitude is further increased to $A_{\mathrm{max}}=12$~nrad and in case of the 2016 experiments even further to $A_{\mathrm{max}}=24$~nrad.
-As the machine tune was changed in standard operation from (64.28,59.31) in 2016 to (62.27,60.295) in 2017 accompanied by a slight change in optics, considered irrelevant in the context of the presented measurements, a direct comparison of the two experiments and thus check of reproducibility is however not possible.
-
-The described change in tune entailed a change in the driving resonances, from previously the $7Q_x$ resonances in 2016 to mainly the $7Q_x$ and $7Q_y$ resonance in 2017.
-This can be seen from the FMA analysis for the 2016 optics (see Fig.~\ref{fig:patternfma} bottom left and bottom right), which shows a strong increase in diffusion for $7^{\mathrm{th}}$~turn pulsing in H around the $7Q_x$ resonances and only small changes for pulsing in V.
-In case of the 2017 optics and tune, the $7Q_x$ as well as the $7Q_y$ resonances cross the tune footprint (see Fig.~\ref{fig:7th2017fma} top left) and are also excited by the $7^{\mathrm{th}}$~turn pulsing visible as an increase of the diffusion in tune around the corresponding resonance lines for pulsing only in H or only in V and thus also for H+V (see Fig.~\ref{fig:7th2017fma}).
-An increase in the diffusion in tune however much weaker than for the $7Q_x$ and $7Q_y$ resonances is also observed around the $7Q_x + 7Q_y$ resonances for pulsing in V and H+V suggesting also an excitation of this $14^{\mathrm{th}}$ order resonance in addition.
-Why it is not visible for pulsing only in H is not understood.
-\begin{figure*}[h]
+\subsection{Pulsing every $7^{\mathrm{th}}$ turn}
+\label{sec:simex7}
+
+\begin{figure*}
 	\begin{minipage}[t]{0.49\linewidth}
@@
 
-\begin{table}[b]
+\begin{table}
 	\caption{\label{tab:7thexp}%
@@
 
-\begin{figure*}[h]
+\begin{figure*}
 	\begin{minipage}[t]{0.49\linewidth}
@@
 \end{figure*}
+
+\begin{figure*}
+		2016 experiment, $7^\mathrm{th}$ H	\\	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
+	\end{minipage}	
+		2017 experiment, $7^\mathrm{th}$ H+V\\
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
+	\end{minipage}	
+	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude.
+For all bunches the damper was not active.}
+\end{figure*}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{2016+2017_scale_amp_7h_lblshort.png}
+	\caption{\label{fig:7thexploss} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in H and data obtained during the 2016 experiment (black and blue) and 2017 experiment (red and yellow).
+The excitation amplitude error is 50\% in 2016 and was reduced to 15\% in 2017.
+The solid and dashed line indicate the second order polynomial fit to the loss rate.}
+\end{figure}
+
+The $7^{\mathrm{th}}$~turn pulsing pattern was tested during both the
+2016 and 2017 experiments. In 2016, the resonant excitation was
+employed for the first time in the LHC and the experiments were
+therefore still in an exploratory stage. Based on the experience
+gained in 2016, the experiments were then repeated more
+systematically, with more bunches and on both excitation planes.
+
+The experiments were divided into different periods: the first period
+without excitation, the second period with a resonant excitation with
+a maximum amplitude of
+$A_{\mathrm{max}} = 5 \, \Delta A = \q{6}{nrad}$, and a third period
+in which the excitation amplitude was further increased to
+$A_{\mathrm{max}} = \q{12}{nrad}$. In~2016, experiments included a
+fourth period at $A_{\mathrm{max}} = \q{24}{nrad}$.
+
+The machine tunes were changed in standard operations from (64.28,
+59.31) in~2016 to (62.27, 60.295) in~2017. This change was accompanied
+by a small change in optics, considered negligible for these
+measurements. A direct comparison of the two experiments is therefore
+not possible. However, these differences provide a rough estimate of
+how the effects of resonant excitations are affected by lattice
+changes.
+
+An insight on how the change in tune entails a change in driving
+resonances is provided by frequency-map analysis.
+Fig.~\ref{fig:patternfma} (bottom left and bottom right plots), which
+was described previously, shows a strong increase in diffusion for
+$7^{\mathrm{th}}$~turn pulsing in the horizontal plane around the
+$7Q_x$ resonances, and only small changes for pulsing in the vertical
+plane. In the case of the 2017 optics and tune, both the $7Q_x$ and
+the $7Q_y$ resonances cross the tune footprint
+(Fig.~\ref{fig:7th2017fma}, top left). Both of these resonances are
+excited by $7^{\mathrm{th}}$~turn pulsing, as one can see from the
+increase in tune jitter around the corresponding resonance lines for
+each excitation plane separately, and thus also for their combined
+action (Fig.~\ref{fig:7th2017fma}, bottom left, bottom right, and top
+right plots). Although much weaker, an increase in tune diffusion is
+also observed around the $7Q_x + 7Q_y$ line when pulsing in the
+vertical plane or in both planes (diagonal yellow-green line in
+Fig.~\ref{fig:7th2017fma}, top right and bottom right plots). This
+suggests an excitation of this $14^{\mathrm{th}}$-order resonance. The
+reason why this phenomenon is much less pronounced for horizontal
+excitations is not understood.
 
@@
 \end{description}
-Similar changes of the emittance as predicted in simulations were also experimentally observed as depicted in Fig.~\ref{fig:7thexp}, where emittance growth was observed only in the horizontal plane for pulsing in H in 2016, and only in the vertical plane for pulsing in V and H+V in 2017 with no emittance growth at all for pulsing in H.
-Just as for the $10^{\mathrm{th}}$~turn pulsing, the emittance growth can be divided also for the $7^{\mathrm{th}}$~turn pulsing in a phase of fast adjustment of the beam distribution and the reach of a new equilibrium state indicated in blue (fast adjustment) and black (equilibrium state) in Fig.~\ref{fig:7thexp}.
-Also in this case, the distribution changes are directly visible on the beam profiles taken with the BSRT (see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexpprof}).
-\begin{figure*}[h]
-		2016 experiment, $7^\mathrm{th}$ H	\\	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
-	\end{minipage}	
-		2017 experiment, $7^\mathrm{th}$ H+V\\
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
-	\end{minipage}	
-	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude.
-For all bunches the damper was not active.}
-\end{figure*}
-
-The losses measured during the experiment however differ from the simulations in terms of for which case the largest rate is observed as well as quantitatively for the specific cases.
-For example smaller losses are expected for the 2016 experiments compared to the 2017 experiments from simulations, while experimentally higher loss rates were measured in 2016 (see Fig.~\ref{fig:7thexploss}).
-For the 2017 experiment, the highest loss rates are expected for pulsing in H+V, than in H and then with a considerably reduction in V.
-During the experiment comparable loss rates are however measured for all three planes (see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexploss2017}).
-Last but not least the loss rates as defined in Eqn.~\ref{eqn:lossrate} for the 2016 (red and yellow) and 2017 (black and gree) experiments are compared in Fig.~\ref{fig:7thexploss}.
-In both cases and as also observed for the $10^{\mathrm{th}}$~turn pulsing, the loss rate follows a quadratic scaling with the excitation amplitude.
-Loss rates were in general higher during the 2016 experiments.
-The loss rates from simulations are not shown in Fig.~\ref{fig:7thexploss} as losses were only observed for much higher excitation amplitudes than tried in the experiments.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1.0\linewidth]{2016+2017_scale_amp_7h_lblshort.png}
-	\caption{\label{fig:7thexploss} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in H and data obtained during the 2016 experiment (black and blue) and 2017 experiment (red and yellow).
-The excitation amplitude error is 50\% in 2016 and was reduced to 15\% in 2017.
-The solid and dashed line indicate the second order polynomial fit to the loss rate.}
-\end{figure}
-
-\clearpage
+Similar changes of the emittance as predicted in simulations were also
+experimentally observed as depicted in Fig.~\ref{fig:7thexp}, where
+emittance growth was observed only in the horizontal plane for pulsing
+in H in 2016, and only in the vertical plane for pulsing in V and H+V
+in 2017 with no emittance growth at all for pulsing in H. Just as for
+the $10^{\mathrm{th}}$~turn pulsing, the emittance growth can be
+divided also for the $7^{\mathrm{th}}$~turn pulsing in a phase of fast
+adjustment of the beam distribution and the reach of a new equilibrium
+state indicated in blue (fast adjustment) and black (equilibrium
+state) in Fig.~\ref{fig:7thexp}. Also in this case, the distribution
+changes are directly visible on the beam profiles taken with the BSRT
+(see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexpprof}).
+
+The losses measured during the experiment however differ from the
+simulations in terms of for which case the largest rate is observed as
+well as quantitatively for the specific cases. For example smaller
+losses are expected for the 2016 experiments compared to the 2017
+experiments from simulations, while experimentally higher loss rates
+were measured in 2016 (see Fig.~\ref{fig:7thexploss}). For the 2017
+experiment, the highest loss rates are expected for pulsing in H+V,
+than in H and then with a considerably reduction in V. During the
+experiment comparable loss rates are however measured for all three
+planes (see Appendix~\ref{app:sec:7},
+Fig.~\ref{fig:7thexploss2017}). Last but not least the loss rates as
+defined in Eqn.~\ref{eqn:lossrate} for the 2016 (red and yellow) and
+2017 (black and gree) experiments are compared in
+Fig.~\ref{fig:7thexploss}. In both cases and as also observed for the
+$10^{\mathrm{th}}$~turn pulsing, the loss rate follows a quadratic
+scaling with the excitation amplitude. Loss rates were in general
+higher during the 2016 experiments. The loss rates from simulations
+are not shown in Fig.~\ref{fig:7thexploss} as losses were only
+observed for much higher excitation amplitudes than tried in the
+experiments.
 

### 2018-03-13 resex.tex
Revised results on 7th-turn pulsing. Intermediate step in revision of figures for this section. Defined "nth-turn pulsing" commands. Moved figures from appendix to text.

@@
 \newcommand{\q}[2]{\ensuremath{#1\ \mathrm{#2}}} % quantity with units
+
+\newcommand{\seventhtp}{7th-turn pulsing}
+\newcommand{\eighthtp}{8th-turn pulsing}
+\newcommand{\tenthtp}{10th-turn pulsing}
 
@@
 
-\subsection{Pulsing every $7^{\mathrm{th}}$ turn}
+\subsection{Pulsing every 7th turn}
 \label{sec:simex7}
@@
 \begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		no excitation
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate_7th.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$7^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$7^{\mathrm{th}}$ H
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skh_dp0_ord7.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$7^{\mathrm{th}}$ V
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut7skv_dp0_ord7.png}
-	\end{minipage}	
-	\caption{\label{fig:7th2017fma} FMA analysis in frequency space without excitation (top left) and for $7^{\mathrm{th}}$~turn pulsing in H+V (top right), only in H (bottom left) and only in V (bottom right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
-The excitation is 96~nrad in both planes.
-The $7Q_x$ and $7Q_y$ resonance are both excited and in addition to the 14th order $7Q_x+7Q_y$.}
-\end{figure*}
-
-\begin{table}
-	\caption{\label{tab:7thexp}%
-		Overview of the effect of $7^{\mathrm{th}}$~turn pulsing during the two resonant excitation experiments in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}.
-The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V for horizontal and vertical at the same time.
-	}
-	\begin{ruledtabular}
-		\begin{tabular}{lcc}
-			Parameter & Experiment 2017 & Experiment 2016  \\
-			\colrule
-			optics & 2017 injection optics & 2016 injection optics \\
-			tune $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\\hline
-			excitation plane  & H,V and H+V&  H\\
-			losses & similar loss rates for  & large losses for \\
-			& pulsing in H,V &  pulsing in H \\
-			&  and H+V & \\
-			hor. emit. growth & no & yes, very small \\
-			vert. emit. growth & yes, for pulsing in & no\\
-			&  V and H+V & \\
+  \begin{tabular}{cc}
+    no excitation & 7th, H+V \\
+    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate_7th.png} &
+    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png} \\
+    7th, H & 7th, V \\
+    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
+    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
   \end{tabular}
-	\end{ruledtabular}
-\end{table}
+  \caption{FMA in betatron tune space based on the 2017 injection
+    optics with no machine errors and tunes (62.27, 60.295): without
+    excitation (top left) and for \seventhtp\ in H+V (top right), H
+    only (bottom left) and V only (bottom right). The excitation
+    amplitude is 96~nrad. The $7 Q_x$, $7 Q_y$, and 14th-order
+    $7 Q_x + 7 Q_y$ resonances are excited.}
+  \label{fig:7th2017fma} 
+\end{figure*}
 
 \begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png}
-	\end{minipage}
-	\caption{\label{fig:7thsim} Relative intensity (top left), relative vertical emittance (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$ and 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
-The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $7^{\mathrm{th}}$~turn pulsing with two different excitation amplitudes plus a random dipole noise component in H+V of 6~nrad.}
+  \begin{tabular}{cc}
+    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png} &
+    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png} \\
+    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png} &
+    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png} \\
+  \end{tabular}
+  \caption{Calculated bunch intensities and emittances from
+    distribution tracking based on the 2016 injection optics with
+    standard errors and $(Q_x, Q_y) = (64.28, 59.31)$ and 2017
+    injection optics with standard errors and
+    $(Q_x, Q_y) = (62.27, 60.295)$: relative intensity (top left);
+    horizontal emittance (bottom left); vertical emittance (bottom
+    right); the relative vertical emittance is also shown (top
+    right). The solid black line includes only a random dipole noise
+    component in H+V of 6~nrad. The dotted and dashed lines correspond
+    to \seventhtp\ with two different excitation amplitudes (24~nrad
+    and 48~nrad), plus a random dipole noise component in H+V of
+    6~nrad.}
+  \label{fig:7thsim}
 \end{figure*}
@@
 \begin{figure*}
-		2016 experiment, $7^\mathrm{th}$ H	\\	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
-	\end{minipage}	
-		2017 experiment, $7^\mathrm{th}$ H+V\\
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
-	\end{minipage}	
-	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude.
+  \begin{tabular}{ccc}
+    \multicolumn{3}{c}{2016 experiment, 7th, H}	\\
+    \includegraphics[width=0.32\textwidth]{2016_bunch_intensity_h7th_no_damper_avg.png} &
+    \includegraphics[width=0.32\textwidth]{2016_emith_avg_rel_h7th_no_damper.png} &
+    \includegraphics[width=0.32\textwidth]{2016_emitv_avg_rel_h7th_no_damper.png} \\
+    \multicolumn{3}{c}{2017 experiment, 7th, H+V} \\
+    \includegraphics[width=0.32\textwidth]{2017_bunch_intensity_hv7th_no_damper_avg.png} &
+    \includegraphics[width=0.32\textwidth]{2017_emith_avg_rel_hv7th_no_damper.png} &
+    \includegraphics[width=0.32\textwidth]{2017_emitv_avg_rel_hv7th_no_damper.png} \\
+  \end{tabular}
+  \caption{Measured losses and emittances during the 2016 and 2017
+    experiments: relative losses (left), relative horizontal
+    emittances (middle), and relative vertical emittances
+    (right). Measurements are averaged over the bunches experiencing
+    the same excitation amplitude. The transverse damping system was
+    not active in this set of measurements.}
+  \label{fig:7thexp}
+\end{figure*}
+
+% \begin{figure*}
+% 		2016 experiment, $7^\mathrm{th}$ H	\\	\begin{minipage}[t]{0.32\linewidth}
+% 		\centering
+% 		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
+% 	\end{minipage}	
+% 	\begin{minipage}[t]{0.32\linewidth}
+% 		\centering
+% 		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
+% 	\end{minipage}	
+% 	\begin{minipage}[t]{0.32\linewidth}
+% 		\centering
+% 		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
+% 	\end{minipage}	
+% 		2017 experiment, $7^\mathrm{th}$ H+V\\
+% 	\begin{minipage}[t]{0.32\linewidth}
+% 		\centering
+% 		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
+% 	\end{minipage}	
+% 	\begin{minipage}[t]{0.32\linewidth}
+% 		\centering
+% 		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
+% 	\end{minipage}	
+% 	\begin{minipage}[t]{0.32\linewidth}
+% 		\centering
+% 		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
+% 	\end{minipage}	
+% 	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude.
 For all bunches the damper was not active.}
-\end{figure*}
+% \end{figure*}
 
@@
 The solid and dashed line indicate the second order polynomial fit to the loss rate.}
+\end{figure}
+
+\begin{figure}
+	\begin{minipage}[t]{1.0\linewidth}
+	\centering
+	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
+	\end{minipage}
+	\begin{minipage}[t]{1.0\linewidth}
+		\centering
+		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
+	\end{minipage}
+	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by horizontal error bars.
+The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
+      \end{figure}
+
+\begin{figure}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch,\\ no excitation
+		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
+		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
+	\end{minipage}
+	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V.
+For both bunches the transverse damper is not active.
+The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
 \end{figure}
@@
 The experiments were divided into different periods: the first period
-without excitation, the second period with a resonant excitation with
+without excitation; the second period with a resonant excitation with
 a maximum amplitude of
-$A_{\mathrm{max}} = 5 \, \Delta A = \q{6}{nrad}$, and a third period
+$A_{\mathrm{max}} = 5 \, \Delta A = \q{6}{nrad}$; and a third period
 in which the excitation amplitude was further increased to
@@
 measurements. A direct comparison of the two experiments is therefore
-not possible. However, these differences provide a rough estimate of
-how the effects of resonant excitations are affected by lattice
-changes.
+not possible. However, these differences provide an estimate of how
+the effects of resonant excitations are affected by lattice changes.
 
@@
 resonances is provided by frequency-map analysis.
-Fig.~\ref{fig:patternfma} (bottom left and bottom right plots), which
-was described previously, shows a strong increase in diffusion for
-$7^{\mathrm{th}}$~turn pulsing in the horizontal plane around the
+Figure~\ref{fig:patternfma} (bottom left and bottom right plots),
+which was described previously, shows a strong increase in diffusion
+for $7^{\mathrm{th}}$~turn pulsing in the horizontal plane around the
 $7Q_x$ resonances, and only small changes for pulsing in the vertical
@@
 Fig.~\ref{fig:7th2017fma}, top right and bottom right plots). This
-suggests an excitation of this $14^{\mathrm{th}}$-order resonance. The
+suggests an excitation of the $14^{\mathrm{th}}$-order resonance. The
 reason why this phenomenon is much less pronounced for horizontal
@@
 
-The simulation results from the distribution tracking shown in Fig.~\ref{fig:7thsim} in general predict in terms of emittance for the two experiments:
+The simulated losses and emittances for both 2016 and 2017 conditions
+are shown in Fig.~\ref{fig:7thsim}. The experimental results are
+summarized in Figures~\ref{fig:7thexp}, \ref{fig:7thexploss},
+\ref{fig:7thexploss2017}, and~\ref{fig:7thexpprof}.
+
+Calculated losses for~2016 and~2017 are similar at 24~nrad and
+increase more rapidly with amplitude for 2017 conditions
+(Fig.~\ref{fig:7thsim}, top left). Horizontal pulsing is a few times
+stronger than vertical pulsing, and the combined H+V action appears
+slightly more effective than the sum of H and V separately.
+
+Experimental losses are plotted as a function of time in
+Fig.~\ref{fig:7thexp} and as a function of excitation amplitude in
+Figs.~\ref{fig:7thexploss}
+and~\ref{fig:7thexploss2017}. Figure~\ref{fig:7thexploss} shows that
+measured losses were actually larger in~2016 than they were
+in~2017. They were also a few times larger than those predicted by
+simulations. Another discrepancy was observed in the excitation plane:
+horizontal, vertical, and combined pulsing had similar effects
+(Fig.~\ref{fig:7thexploss2017}). The same considerations on systematic
+effects discussed in Section~\ref{sec:simex10} apply in this case.
+
+This pulsing pattern generated stronger losses than expected. In this
+case, we conclude that predictions are difficult, due to their
+sensitivity to lattice configuration, noise sources, and beam
+distributions.
+
+The simulated emittances from distribution tracking are shown in
+Fig.~\ref{fig:7thsim} (bottom left and right). The main calculated
+results are the following, for the two experimental conditions:
+%
 \begin{description}
-	\item[2016 experiment] Excitation amplitude dependent emittance growth in the horizontal but not in the vertical plane.
-	\item[2017 experiment] Excitation amplitude dependent emittance growth in the vertical plane but not in the horizontal plane for an excitation in V and H+V.
-For an excitation only in H no emittance growth is observed.
+\item[2016 conditions] In this case, excitations were only in the
+  horizontal plane. There is a fast increase of the horizontal
+  emittance, dependent on the excitation amplitude, after which a
+  constant emittance growth is observed. Vertical emittance growth is
+  almost negligible.
+\item[2017 conditions] Horizontal emittances are mostly affected by
+  the amplitude of H+V excitations and, to a much smaller extent, by
+  H~excitations; no effect is observed from V excitations. Vertical
+  emittances are mostly affected by the amplitude of H+V and V
+  excitations; the effect of H excitations is much smaller, if any,
+  with no clear dependence on amplitude.
 \end{description}
-Similar changes of the emittance as predicted in simulations were also
-experimentally observed as depicted in Fig.~\ref{fig:7thexp}, where
-emittance growth was observed only in the horizontal plane for pulsing
-in H in 2016, and only in the vertical plane for pulsing in V and H+V
-in 2017 with no emittance growth at all for pulsing in H. Just as for
-the $10^{\mathrm{th}}$~turn pulsing, the emittance growth can be
-divided also for the $7^{\mathrm{th}}$~turn pulsing in a phase of fast
-adjustment of the beam distribution and the reach of a new equilibrium
-state indicated in blue (fast adjustment) and black (equilibrium
-state) in Fig.~\ref{fig:7thexp}. Also in this case, the distribution
-changes are directly visible on the beam profiles taken with the BSRT
-(see Appendix~\ref{app:sec:7}, Fig.~\ref{fig:7thexpprof}).
-
-The losses measured during the experiment however differ from the
-simulations in terms of for which case the largest rate is observed as
-well as quantitatively for the specific cases. For example smaller
-losses are expected for the 2016 experiments compared to the 2017
-experiments from simulations, while experimentally higher loss rates
-were measured in 2016 (see Fig.~\ref{fig:7thexploss}). For the 2017
-experiment, the highest loss rates are expected for pulsing in H+V,
-than in H and then with a considerably reduction in V. During the
-experiment comparable loss rates are however measured for all three
-planes (see Appendix~\ref{app:sec:7},
-Fig.~\ref{fig:7thexploss2017}). Last but not least the loss rates as
-defined in Eqn.~\ref{eqn:lossrate} for the 2016 (red and yellow) and
-2017 (black and gree) experiments are compared in
-Fig.~\ref{fig:7thexploss}. In both cases and as also observed for the
-$10^{\mathrm{th}}$~turn pulsing, the loss rate follows a quadratic
-scaling with the excitation amplitude. Loss rates were in general
-higher during the 2016 experiments. The loss rates from simulations
-are not shown in Fig.~\ref{fig:7thexploss} as losses were only
-observed for much higher excitation amplitudes than tried in the
-experiments.
+
+A set of measured emittances is plotted in Fig.~\ref{fig:7thexp}
+(middle and right plots). In~2016, with only horizontal excitations,
+amplitude-dependent horizontal emittance growth was observed, with no
+effect on the vertical emittance. In~2017, V and H+V excitations
+generated vertical emittance growth, while horizontal excitations had
+no effect. These observations are consistent with the predicted
+behavior.
+
+During these experiments, emittances changed in two phases, as they
+did during the $10^{\mathrm{th}}$~turn pulsing studies
+(Section~\ref{sec:simex10}) and in simulations: a fast adjustment
+phase (shaded in blue in Fig.~\ref{fig:7thexp}) followed by a new
+equilibrium (in gray). Changes in the transverse beam distribution
+could be observed directly through changes in the synchrotron-light
+(BSRT) profiles (Fig.~\ref{fig:7thexpprof}).
 
@@
 
-\clearpage
-
 \subsection{random excitation\label{sec:simexran}}
@@
 \end{figure}
-
-\clearpage
 
@@
 Further studies are therefore needed in order to obtain a deeper understanding of the interaction of the transverse damper with a resonant excitation.
-\clearpage
 
@@
 
-\newpage
-
 \appendix
@@
 
-\newpage
-
-\section{Additional plots for $7^{\mathrm{th}}$~turn pulsing during 2017 experiments}
-\label{app:sec:7}
-\begin{figure}[h]
-	\begin{minipage}[t]{1.0\linewidth}
-	\centering
-	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
-	\end{minipage}
-	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
+% \section{Additional plots for $7^{\mathrm{th}}$~turn pulsing during 2017 experiments}
+% \label{app:sec:7}
+% \begin{figure}[h]
+% 	\begin{minipage}[t]{1.0\linewidth}
+% 	\centering
+% 	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
+% 	\end{minipage}
+% 	\begin{minipage}[t]{1.0\linewidth}
+% 		\centering
+% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
+% 	\end{minipage}
+% 	\begin{minipage}[t]{1.0\linewidth}
+% 		\centering
+% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
+% 	\end{minipage}
+% 	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
 The excitation amplitude error is 15\% indicated by horizontal error bars.
 The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
-\end{figure}
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
-		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
-	\end{minipage}
-	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+% \end{figure}
+% \begin{figure}[h]
+% 	\begin{minipage}[t]{0.49\linewidth}
+% 		\centering
+% 		reference bunch,\\ no excitation
+% 		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
+% 	\end{minipage}
+% 	\begin{minipage}[t]{0.49\linewidth}
+% 		\centering
+% 		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
+% 		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
+% 	\end{minipage}
+% 	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
 The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V.
@@
 The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
-\end{figure}
-
-\newpage
+% \end{figure}
+
+% \newpage
 
@@
 
-\clearpage
-
 \bibliography{resex-bibdesk}

### 2018-03-14 resex.tex
Revised results of 7th-turn pulsing, including figures and captions. Cropped 3 plots. Started using abbreviated commands for n-th turn pulsing.

@@
 \item \textbf{resonant excitation:} The electron beam is switched on
-  only every $k^{\mathrm{th}}$ turn. The excitation can be represented
+  only every $k$-th turn. The excitation can be represented
   by the following expression:
@@
   \end{equation}
-  with $f_n = n \cdot f_{\rm rev} / k$ and
-  where $\Sh_{kT}$ is the Dirac comb. In general,
-  $k^{\mathrm{th}}$~turn pulsing drives $k^{\mathrm{th}}$-order
+  with $f_n = n \cdot f_{\rm rev} / k$ and where $\Sh_{kT}$ is the
+  Dirac comb. In general, $k$-th~turn pulsing drives $k$-th order
   resonances~\cite{md_sim_hel_res_ex_fitterer}. This type of pulsing
@@
       chromaticity $(Q'_x, Q'_y)$ & \multicolumn{2}{c}{(+15, +15)}\\ \hline
-      pulsing patterns  & $7^{\mathrm{th}}$~turn H
-                                  &$7^{\mathrm{th}}$~turn H, V, H+V \\
+      pulsing patterns  & 7th~turn H
+                                  &7th~turn H, V, H+V \\
                 & $10^{\mathrm{th}}$~turn V & $8^{\mathrm{th}}$~turn
@@
 the two pulsing patterns with the largest calculated effects on the
-beam ($7^{\mathrm{th}}$- and $10^{\mathrm{th}}$-turn pulsing), one
-pattern with no effect ($8^{\mathrm{th}}$-turn pulsing), and the
-random excitation. In order to quantify the reproducibility of the
-results under different machine configurations, one pulsing pattern
-($7^{\mathrm{th}}$-turn pulsing) was tried first in~2016 and then
-repeated in~2017.
+beam (7th- and $10^{\mathrm{th}}$-turn pulsing), one pattern with no
+effect ($8^{\mathrm{th}}$-turn pulsing), and the random excitation. In
+order to quantify the reproducibility of the results under different
+machine configurations, one pulsing pattern (\seventhtp) was tried
+first in~2016 and then repeated in~2017.
 
@@
 tested experimentally, namely pulsing every $10^{\mathrm{th}}$~turn
-(Section~\ref{sec:simex10}), $7^{\mathrm{th}}$~turn
+(Section~\ref{sec:simex10}), 7th~turn
 (Section~\ref{sec:simex7}), $8^{\mathrm{th}}$~turn
@@
     \\
-    $7^{\mathrm{th}}$, H & $7^{\mathrm{th}}$, V \\
+    7th, H & 7th, V \\
     \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
@@
     pulsing in both horizontal and vertical planes (top left);
-    $7^{\mathrm{th}}$~turn pulsing in the horizontal (bottom left) and
+    \seventhtp\ in the horizontal (bottom left) and
     vertical (bottom right) planes. The excitation amplitude is
@@
     strong excitation of any resonance in case of vertical
-    $7^{\mathrm{th}}$~turn pulsing and the strong excitation in case
+    \seventhtp\ and the strong excitation in case
     of horizontal pulsing confirms the excitation of the $7 Q_x$
@@
     \includegraphics[width=0.49\textwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
-    $7^{\mathrm{th}}$, H+V & $10^{\mathrm{th}}$, H+V \\
+    7th, H+V & $10^{\mathrm{th}}$, H+V \\
     \includegraphics[width=0.49\textwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
@@
     lattice errors: no excitation (top left), random excitation (top
-    right), $7^{\mathrm{th}}$~turn pulsing (bottom left), and
-    $10^{\mathrm{th}}$~turn pulsing (bottom right). The excitations
+    right), \seventhtp\ (bottom left), and
+    \tenthtp\ (bottom right). The excitations
     are applied in both planes with an amplitude of 96~nrad. For each
@@
 and emittances on the pattern can be seen.  The largest losses are
-predicted for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and
+predicted for $3^{\mathrm{rd}}$, 7th and
 $10^{\mathrm{th}}$~turn pulsing, and for a uniform random
@@
 transverse aperture. Significant emittance growth is visible only for
-$7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and for the
+7th- and \tenthtp\ and for the
 random excitation. Compared to any resonant excitation, the random
@@
 
-The effects of $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing
-are also observed in the absence of machine lattice errors. In this
-case, the only sources of nonlinearity are sextupoles and
-octupoles~\cite{md_sim_hel_res_ex_fitterer}, suggesting that these
-nonlinearities are responsible for the pronounced beam
-sensitivity. The driven resonances are revealed by the frequency-map
-analysis, shown in Fig.~\ref{fig:patternfma}. The $7 Q_x$ resonance is
-excited by the $7^{\mathrm{th}}$~turn pulsing, and the $10 Q_x$ and
+The effects of 7th- and \tenthtp\ are also observed in the absence of
+machine lattice errors. In this case, the only sources of nonlinearity
+are sextupoles and octupoles~\cite{md_sim_hel_res_ex_fitterer},
+suggesting that these nonlinearities are responsible for the
+pronounced beam sensitivity. The driven resonances are revealed by the
+frequency-map analysis, shown in Fig.~\ref{fig:patternfma}. The
+$7 Q_x$ resonance is excited by the \seventhtp, and the $10 Q_x$ and
 $10 Q_y$ resonances by the $10^{\mathrm{th}}$~turn pulsing. As
@@
 
-In Fig.~\ref{fig:patternsim}, one can see that, in the cases of
-$7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing, the emittances
-start from an increased initial value and then stay almost constant
-for the duration of the simulation. This behavior is typical of the
-resonant excitation and it is not an artifact of the simulation. It
-was verified that it is due to the adjustment of the input beam
-distribution to a new equilibrium during the first $10^4$~turns. In
-all long-term simulations ($10^6$~turns, corresponding to 89~s)
-presented in this paper, the beam distribution is saved every
-$10^4$~turns. Therefore, this initial fast adjustment manifests itself
-in an increased initial emittance value.
+In Fig.~\ref{fig:patternsim}, one can see that, in the cases of 7th-
+and \tenthtp, the emittances start from an increased initial value and
+then stay almost constant for the duration of the simulation. This
+behavior is typical of the resonant excitation and it is not an
+artifact of the simulation. It was verified that it is due to the
+adjustment of the input beam distribution to a new equilibrium during
+the first $10^4$~turns. In all long-term simulations ($10^6$~turns,
+corresponding to 89~s) presented in this paper, the beam distribution
+is saved every $10^4$~turns. Therefore, this initial fast adjustment
+manifests itself in an increased initial emittance value.
 
@@
 at $10^4$~turns, and how they differ from each other and from a
-Gaussian distribution. For instance, for $7^{\mathrm{th}}$ turn
-pulsing (bottom left plot in Fig.~\ref{fig:patternhist}), simulations
-indicate a clear shift in beam population from the core ($< 1\sigma$)
-towards the regions at $\pm 1.5\sigma$ and towards the tails at
-$\pm 4\sigma$.
+Gaussian distribution. For instance, for \seventhtp\ (bottom left plot
+in Fig.~\ref{fig:patternhist}), simulations indicate a clear shift in
+beam population from the core ($< 1\sigma$) towards the regions at
+$\pm 1.5\sigma$ and towards the tails at $\pm 4\sigma$.
 
@@
       Radiation Telescope (BSRT) during the 2016 experiments on
-      $10^{\mathrm{th}}$~turn pulsing in the vertical plane. The
+      \tenthtp\ in the vertical plane. The
       transverse damping system was not active on these bunches. The
@@
     value (top right). The solid lines indicate the effect of
-    $10^{\mathrm{th}}$~turn pulsing in the vertical plane. The dotted
+    \tenthtp\ in the vertical plane. The dotted
     lines include both the resonant excitation and a random dipole
@@
     rates vs.\ resonant excitation amplitude for
-    $10^{\mathrm{th}}$~turn pulsing in the vertical plane. The
+    \tenthtp\ in the vertical plane. The
     relative average loss rate~$R$ is defined in
@@
 
-The $10^{\mathrm{th}}$~turn pulsing pattern was tested in 2016 with an
+The \tenthtp\ pattern was tested in 2016 with an
 excitation in the vertical plane only. As this was the first time a
@@
 experiment are shown in Fig.~\ref{fig:10thsim}. They are the results
-of distribution tracking for $10^{\mathrm{th}}$~turn pulsing without
+of distribution tracking for \tenthtp\ without
 additional noise (solid lines) and with an additional noise component
@@
 
-% \begin{figure*}
-% 		2016 experiment, $7^\mathrm{th}$ H	\\	\begin{minipage}[t]{0.32\linewidth}
-% 		\centering
-% 		\includegraphics[height=0.75\linewidth]{2016_bunch_intensity_h7th_no_damper_avg.png}
-% 	\end{minipage}	
-% 	\begin{minipage}[t]{0.32\linewidth}
-% 		\centering
-% 		\includegraphics[height=0.75\linewidth]{2016_emith_avg_rel_h7th_no_damper.png}
-% 	\end{minipage}	
-% 	\begin{minipage}[t]{0.32\linewidth}
-% 		\centering
-% 		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_h7th_no_damper.png}
-% 	\end{minipage}	
-% 		2017 experiment, $7^\mathrm{th}$ H+V\\
-% 	\begin{minipage}[t]{0.32\linewidth}
-% 		\centering
-% 		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hv7th_no_damper_avg.png}
-% 	\end{minipage}	
-% 	\begin{minipage}[t]{0.32\linewidth}
-% 		\centering
-% 		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hv7th_no_damper.png}
-% 	\end{minipage}	
-% 	\begin{minipage}[t]{0.32\linewidth}
-% 		\centering
-% 		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hv7th_no_damper.png}
-% 	\end{minipage}	
-% 	\caption{\label{fig:7thexp} Relative horizontal (top) and vertical (bottom) emittance obtained during the 2016 experiments (left) and the 2017 experiment (right) averaged over the bunches experiencing the same excitation amplitude.
-For all bunches the damper was not active.}
-% \end{figure*}
-
 \begin{figure}
   \centering
-	\includegraphics[width=1.0\linewidth]{2016+2017_scale_amp_7h_lblshort.png}
-	\caption{\label{fig:7thexploss} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing in H and data obtained during the 2016 experiment (black and blue) and 2017 experiment (red and yellow).
-The excitation amplitude error is 50\% in 2016 and was reduced to 15\% in 2017.
-The solid and dashed line indicate the second order polynomial fit to the loss rate.}
+  \includegraphics[width=\columnwidth]{2016+2017_scale_amp_7h_lblshort.png}
+  \caption{Measured loss rates vs.\ excitation amplitude for
+    \seventhtp\ in the horizontal plane during the 2016 experiment
+    (red and yellow) and the 2017 experiment (black and green). The
+    uncertainties on the excitation amplitude are dominated by the
+    calibration of the transverse feedback and damping system. The
+    uncertainties on the loss rates are statistical. An estimate of
+    the systematic uncertainties (due to changes in beam distribution,
+    for instance) is given by the difference between the two data sets
+    within each experiment. The lines indicate empirical second-order
+    polynomial fits, with damper off (solid) and with damper on
+    (dashed).}
+  \label{fig:7thexploss} 
 \end{figure}
 
-\begin{figure}
-	\begin{minipage}[t]{1.0\linewidth}
-	\centering
-	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
-	\end{minipage}
-	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by horizontal error bars.
-The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
-      \end{figure}
-
-\begin{figure}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
-		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
-	\end{minipage}
-	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V.
-For both bunches the transverse damper is not active.
-The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
-\end{figure}
-
-The $7^{\mathrm{th}}$~turn pulsing pattern was tested during both the
+\begin{figure*}
+  \begin{tabular}{ccc}
+    \includegraphics[width=0.32\textwidth]{2017_scale_amp_7h_lblshort.png} &
+    \includegraphics[width=0.32\textwidth]{2017_scale_amp_7v_lblshort.png} &
+    \includegraphics[width=0.32\textwidth]{2017_scale_amp_7hv_lblshort.png} \\
+  \end{tabular}
+  \caption{Comparison of the measured loss rates as a function of
+    excitation amplitude during the 2017 experiment for \seventhtp\ in
+    H only (left), V only (center), and in H+V (right). The three
+    excitations generate similar loss rates.}
+  \label{fig:7thexploss2017}
+\end{figure*}
+
+\begin{figure*}
+  \begin{tabular}{cc}
+    Reference bunch, no excitation & Excited bunch, 7th, H+V \\
+    \includegraphics[width=0.48\textwidth]{profile_v_7thhv_slot_2862.png} &
+    \includegraphics[width=0.48\textwidth]{profile_v_7thhv_slot_2696.png} \\
+  \end{tabular}
+  \caption{Vertical beam profiles measured with the Beam Synchrotron
+    Radiation Telescope (BSRT) during the 2017 experiments. The
+    profiles are taken at the end of \seventhtp\ in H+V. For these
+    bunches, the transverse damper was not active. The data is
+    presented in the same way as in Fig.~\ref{fig:10thexpprof}. The
+    profile changes in the bunch affected by the maximum excitation
+    (right) are larger than those displayed by the reference bunch
+    (left).}
+  \label{fig:7thexpprof}
+\end{figure*}
+
+The \seventhtp\ pattern was tested during both the
 2016 and 2017 experiments. In 2016, the resonant excitation was
@@
 which was described previously, shows a strong increase in diffusion
-for $7^{\mathrm{th}}$~turn pulsing in the horizontal plane around the
-$7Q_x$ resonances, and only small changes for pulsing in the vertical
-plane. In the case of the 2017 optics and tune, both the $7Q_x$ and
-the $7Q_y$ resonances cross the tune footprint
-(Fig.~\ref{fig:7th2017fma}, top left). Both of these resonances are
-excited by $7^{\mathrm{th}}$~turn pulsing, as one can see from the
+for \seventhtp\ in the horizontal plane around the $7Q_x$ resonances,
+and only small changes for pulsing in the vertical plane. In the case
+of the 2017 optics and tune, both the $7Q_x$ and the $7Q_y$ resonances
+cross the tune footprint (Fig.~\ref{fig:7th2017fma}, top left). Both
+of these resonances are excited by \seventhtp, as one can see from the
 increase in tune jitter around the corresponding resonance lines for
@@
 During these experiments, emittances changed in two phases, as they
-did during the $10^{\mathrm{th}}$~turn pulsing studies
+did during the \tenthtp\ studies
 (Section~\ref{sec:simex10}) and in simulations: a fast adjustment
@@
 \subsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
-Pulsing patterns which have little effect on the beam core are particularly interesting for the hollow electron lens operation if they in addition have a large effect on the halo.
-This is in general feasible considering the highly nonlinear field generated at the location of the halo particles by the HEL compared to no field or only a very small field at the location of the beam core.
-With this motivation in mind the $8^{\mathrm{th}}$~turn pulsing pattern was studied, for which a small effect in comparison to the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$ turn pulsing is expected (see Fig.~\ref{fig:8thsim}).
+
+Pulsing patterns which have little effect on the beam core are
+particularly interesting for the hollow electron lens operation if
+they in addition have a large effect on the halo. This is in general
+feasible considering the highly nonlinear field generated at the
+location of the halo particles by the HEL compared to no field or only
+a very small field at the location of the beam core. With this
+motivation in mind the \eighthtp\ pattern was studied, for which a
+small effect in comparison to the 7th- and \tenthtp\ is expected (see
+Fig.~\ref{fig:8thsim}).
+
 \begin{figure*}[h]
@@
 	\caption{\label{fig:8thsim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (distribution tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
-The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for $8^{\mathrm{th}}$~turn pulsing plus a random dipole noise component in H+V of 6~nrad.}
+The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for \eighthtp\ plus a random dipole noise component in H+V of 6~nrad.}
 \end{figure*}
@@
 	\end{minipage}	
-	\caption{\label{fig:8th2017fmaamp} FMA analysis in amplitude space without excitation (left) and for $8^{\mathrm{th}}$~turn pulsing (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
+	\caption{\label{fig:8th2017fmaamp} FMA analysis in amplitude space without excitation (left) and for \eighthtp\ (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
 The excitation is 96~nrad in both planes.
@@
 	\includegraphics[width=1.0\linewidth]{2017_scale_amp_8hv_lblshort.png}
-	\caption{\label{fig:8thexploss} Comparison of scaling of loss rates with excitation amplitude for $8^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
+	\caption{\label{fig:8thexploss} Comparison of scaling of loss rates with excitation amplitude for \eighthtp\ in H+V measured during the 2017 experiment.
 The excitation amplitude error is 15\% indicated by the horizontal error bars, the error on the loss rate is the error on the mean obtained through the average over the 6~bunches with the same excitation amplitude.}
@@
 On the other hand this pattern is also the most dangerous one for the proton beam core.
-The aim of this experiment was to test the effect on the core in comparison to the resonant excitation represented in the 2017 experiment by primarily the $7^{\mathrm{th}}$~turn pulsing as for $8^{\mathrm{th}}$~turn pulsing basically no effect was observed.
+The aim of this experiment was to test the effect on the core in comparison to the resonant excitation represented in the 2017 experiment by primarily the \seventhtp\ as for \eighthtp\ basically no effect was observed.
 The random excitation is in general so powerful as it excites all frequencies of the beam equally.
@@
 
-Qualitatively, this behavior could be also confirmed experimentally during the 2017 experiments.
-Here excitation amplitude dependent emittance growth was only recorded in the plane of excitation with all excitation amplitude dependent emittance growth rates in a comparable range.
-As example the relative horizontal and vertical emittance for pulsing in V and in H+V are depicted in Fig.~\ref{fig:ranexp}.
-Note that as expected from simulations and despite the large excitation amplitude dependent emittance growth, the growth rates stays constant.
-There is explicitly no fast adjustment phase succeeded by a new equilibrium state as observed for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing.
+Qualitatively, this behavior could be also confirmed experimentally
+during the 2017 experiments. Here excitation amplitude dependent
+emittance growth was only recorded in the plane of excitation with all
+excitation amplitude dependent emittance growth rates in a comparable
+range. As example the relative horizontal and vertical emittance for
+pulsing in V and in H+V are depicted in Fig.~\ref{fig:ranexp}. Note
+that as expected from simulations and despite the large excitation
+amplitude dependent emittance growth, the growth rates stays
+constant. There is explicitly no fast adjustment phase succeeded by a
+new equilibrium state as observed for the 7th- and \tenthtp.
+
 \begin{figure*}[h]
@@
 As an example Fig.~\ref{fig:ranexpprof} shows the profile for a random excitation only in V in comparison to a reference bunch.
-The distribution also stays Gaussian in case of a random excitation in contrast to the $10^{\mathrm{th}}$~turn pulsing in which case it clearly assumes a non-Gaussian shape (see Fig.~\ref{fig:10thexpprof}).
+The distribution also stays Gaussian in case of a random excitation in contrast to the \tenthtp\ in which case it clearly assumes a non-Gaussian shape (see Fig.~\ref{fig:10thexpprof}).
 
@@
 The loss rate is also considerably reduced for the bunches with transverse damper active independent of the plane the excitation is applied in (see Fig.~\ref{fig:ranexplossplane}).
-In comparison to the $7^{\mathrm{th}}$~turn pulsing, the loss rate for a random excitation is smaller for a maximum excitation amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}).
-This indicates an increase of the diffusion mostly in the beam tails for $7^{\mathrm{th}}$~turn pulsing, while for a random excitation mostly the core is affected.
+In comparison to the \seventhtp, the loss rate for a random excitation is smaller for a maximum excitation amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}).
+This indicates an increase of the diffusion mostly in the beam tails for \seventhtp, while for a random excitation mostly the core is affected.
 \begin{figure}[h]
@@
 	\includegraphics[width=1.0\linewidth]{2017_scale_amp_7th_ranhv_lblshort.png}
-	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and $7^{\mathrm{th}}$~turn pulsing in H+V measured during the 2017 experiment.
+	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and \seventhtp\ in H+V measured during the 2017 experiment.
 The excitation amplitude error is 15\% indicated by the horizontal error bars.}
@@
 		\centering
-		$7^{\mathrm{th}}$~turn V, 2017 exp.\\
+		7th~turn V, 2017 exp.\\
 		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
@@
 		\centering
-		$7^{\mathrm{th}}$~turn V, 2017 exp.\\
+		7th~turn V, 2017 exp.\\
 		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
@@
 	\end{minipage}	
-	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for a random excitation only in V (left) and for $7^{\mathrm{th}}$ (center) and $10^{\mathrm{th}}$~turn pulsing only in V (right).}
-\end{figure*}
-For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally, the transverse damper does in general not change any of the observables studied during the experiments, explicitly losses, emittance and beam distribution suggesting that indeed it is not damping any resonant excitation.
-On the other hand, the transverse damper considerably decreases any changes of the above parameters in case of a random excitation.
-As an example, the vertical emittance growth for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing in V is compared with a random excitation in V in Fig.~\ref{fig:damp}.
-The same observation is also made in case of losses as shown in Fig.~\ref{fig:ranexploss} comparing the $7^{\mathrm{th}}$~turn pulsing in H+V with a random excitation in H+V.
-For the $7^{\mathrm{th}}$~turn pulsing the two curves for the transverse damper active and not active almost coincide while for the random excitation a significant difference is visible.
+	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for a random excitation only in V (left) and for 7th- (center) and \tenthtp\ only in V (right).}
+\end{figure*}
+For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally,
+the transverse damper does in general not change any of the
+observables studied during the experiments, explicitly losses,
+emittance and beam distribution suggesting that indeed it is not
+damping any resonant excitation. On the other hand, the transverse
+damper considerably decreases any changes of the above parameters in
+case of a random excitation. As an example, the vertical emittance
+growth for the 7th- and \tenthtp\ in V is compared with a random excitation in V in Fig.~\ref{fig:damp}.
+The same observation is also made in case of losses as shown in Fig.~\ref{fig:ranexploss} comparing the \seventhtp\ in H+V with a random excitation in H+V.
+For the \seventhtp\ the two curves for the transverse damper active and not active almost coincide while for the random excitation a significant difference is visible.
 Until present, it is unknown why the transverse damper (ADT) appears to be incapable of damping any resonant excitation as it fulfills all requirements to detect as well as damp the caused oscillation:
@@
 As no hollow electron lens is currently installed in the LHC, the in reality non-linear kick was emulated to first order by a dipole kick with the corresponding excitation pattern.
-The experimental results in general suggest that for the kick amplitudes as expected from the HEL, considerable effects are to be expected for a random excitation and certain resonant excitations, explicitly $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing during the conducted experiments.
+The experimental results in general suggest that for the kick amplitudes as expected from the HEL, considerable effects are to be expected for a random excitation and certain resonant excitations, explicitly 7th- and \tenthtp\ during the conducted experiments.
 In addition, also one resonant excitation pattern, the $8^{\mathrm{th}}$~turn pulsing, was tested featuring a small effect on the beam core.
@@
 			 & & H+V & 1.2 \\\hline
-			$7^{\mathrm{th}}$~turn & 2016 & H & 2.4 \\
+			7th~turn & 2016 & H & 2.4 \\
 			 & 2017 & H & 2.4 \\
@@
 \end{equation}
-the tolerances for all experimentally tried pulsing pattern except for $8^{\mathrm{th}}$~turn pulsing in V lie below this value.
+the tolerances for all experimentally tried pulsing pattern except for \eighthtp\ in V lie below this value.
 As a next step, the experiment should be conducted at collision energy where the HEL is most needed.
@@
 We wish to thank Sergei Nagaitsev for the very useful discussion and profound question to find out why the transverse damper is not damping any resonant excitation.
-We would like to acknowledge Dmitry Shatilov for his help with the Lifetrac code, and St\'{e}phane Fartoukh, Riccardo De Maria and Rogelio Tomás for their help with generating the appropiate optics model.
+We would like to acknowledge Dmitry Shatilov for his help with the Lifetrac code, and St\'{e}phane Fartoukh, Riccardo De Maria and Rogelio Tom\'as for their help with generating the appropiate optics model.
 We are grateful also for the support from the beam instrumentation team, in particular Enrico Bravin and Georges Trad, with the BSRT and are thankful for the collaboration with St\'{e}phania Papadopoulou and Fanouria Antonio for the analysis of the BSRT profiles.
@@
 	\end{minipage}
-	\caption{\label{app:fig:fma:10} FMA analysis for $10^{\mathrm{th}}$~turn pulsing based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
+	\caption{\label{app:fig:fma:10} FMA analysis for \tenthtp\ based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
 The excitation is 120~nrad in the corresponding plane.
@@
 
-% \section{Additional plots for $7^{\mathrm{th}}$~turn pulsing during 2017 experiments}
-% \label{app:sec:7}
-% \begin{figure}[h]
-% 	\begin{minipage}[t]{1.0\linewidth}
-% 	\centering
-% 	\includegraphics[width=0.9\linewidth]{2017_scale_amp_7h_lblshort.png}
-% 	\end{minipage}
-% 	\begin{minipage}[t]{1.0\linewidth}
-% 		\centering
-% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7v_lblshort.png}
-% 	\end{minipage}
-% 	\begin{minipage}[t]{1.0\linewidth}
-% 		\centering
-% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_7hv_lblshort.png}
-% 	\end{minipage}
-% 	\caption{\label{fig:7thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $7^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by horizontal error bars.
-The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
-% \end{figure}
-% \begin{figure}[h]
-% 	\begin{minipage}[t]{0.49\linewidth}
-% 		\centering
-% 		reference bunch,\\ no excitation
-% 		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2862.png}
-% 	\end{minipage}
-% 	\begin{minipage}[t]{0.49\linewidth}
-% 		\centering
-% 		excited bunch,\\ $7^{\mathrm{th}}$~turn H+V	
-% 		\includegraphics[width=1.0\linewidth]{profile_v_7thhv_slot_2696.png}
-% 	\end{minipage}
-% 	\caption{\label{fig:7thexpprof} Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the $7^{\mathrm{th}}$~turn pulsing in H+V.
-For both bunches the transverse damper is not active.
-The distribution changes in the reference bunch (left), here slot 2862, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
-% \end{figure}
-
-% \newpage
-
-\section{Additional plots for $8^{\mathrm{th}}$~turn pulsing during 2017 experiments}
+\section{Additional plots for \eighthtp\ during 2017 experiments}
 \label{app:sec:8}
@@
 	\end{minipage}
-	\caption{\label{fig:8thexploss2017} Comparison of scaling of loss rates with excitation amplitude for $8^{\mathrm{th}}$~turn pulsing only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
+	\caption{\label{fig:8thexploss2017} Comparison of scaling of loss rates with excitation amplitude for \eighthtp\ only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
 The excitation amplitude error is 15\% indicated by horizontal error bars.
@@
 	\caption{\label{fig:8thexpprof} Horizontal beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the $8^{\mathrm{th}}$~turn pulsing in H.
+The profiles are taken at the end of the \eighthtp\ in H.
 For both bunches the transverse damper is not active.

### 2018-03-14 resex.tex
Intermediate step in revision of 8th-turn results. Moved figures.

@@
 \newcommand{\tenthtp}{10th-turn pulsing}
+
+\newcommand{\fma}{frequency-map analysis}
 
@@
 
-\subsection{$8^{\mathrm{th}}$ turn pulsing\label{sec:simex8}}
-
-Pulsing patterns which have little effect on the beam core are
-particularly interesting for the hollow electron lens operation if
-they in addition have a large effect on the halo. This is in general
-feasible considering the highly nonlinear field generated at the
-location of the halo particles by the HEL compared to no field or only
-a very small field at the location of the beam core. With this
-motivation in mind the \eighthtp\ pattern was studied, for which a
-small effect in comparison to the 7th- and \tenthtp\ is expected (see
-Fig.~\ref{fig:8thsim}).
-
-\begin{figure*}[h]
+\subsection{Pulsing every 8th turn}
+\label{sec:simex8}
+
+\begin{figure*}
 	\begin{minipage}[t]{0.49\linewidth}
@@
       \end{figure*}
-\begin{figure*}[t]
+
+\begin{figure*}
 	\begin{minipage}[t]{0.49\linewidth}
@@
 \end{figure*}
-In simulations no visible effect is indeed observed for excitation amplitudes smaller than 96~nrad.
-In the horizontal and vertical plane a small emittance growth due to the excitation is visible, but without any clear dependence on the amplitude or plane.
-The FMA analysis in amplitude space reveals also the driven resonances (Fig.~\ref{fig:8th2017fmaamp}), which are the $16Q_y$ and \mbox{$8Q_x-4Q_y$} resonances and several even higher orders.
-These are in general very high order resonances whose effect is thus expected to be small.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1.0\linewidth]{2017_scale_amp_8hv_lblshort.png}
-	\caption{\label{fig:8thexploss} Comparison of scaling of loss rates with excitation amplitude for \eighthtp\ in H+V measured during the 2017 experiment.
+
+% \begin{figure}
+% 	\centering
+% 	\includegraphics[width=1.0\linewidth]{2017_scale_amp_8hv_lblshort.png}
+% 	\caption{\label{fig:8thexploss} Comparison of scaling of loss rates with excitation amplitude for \eighthtp\ in H+V measured during the 2017 experiment.
 The excitation amplitude error is 15\% indicated by the horizontal error bars, the error on the loss rate is the error on the mean obtained through the average over the 6~bunches with the same excitation amplitude.}
+% \end{figure}
+
+\begin{figure*}
+  \begin{tabular}{ccc}
+  \includegraphics[width=0.32\textwidth]{2017_scale_amp_8h_lblshort.png} &
+  \includegraphics[width=0.32\textwidth]{2017_scale_amp_8v_lblshort.png} &
+  \includegraphics[width=0.32\textwidth]{2017_scale_amp_8hv_lblshort.png} \\
+  \end{tabular}
+  \caption{Comparison of scaling of loss rates with excitation
+    amplitude for \eighthtp\ only in H (top), only in V (center) and
+    in H+V (bottom) during the 2017 experiment. The excitation
+    amplitude error is 15\% indicated by horizontal error bars. The
+    vertical errorbars represent the pure statistical error based on
+    the averaging over several bunches with the same excitation
+    amplitude.}
+  \label{fig:8thexploss2017}
+\end{figure*}
+
+% \begin{figure}
+% 	\begin{minipage}[t]{1.0\linewidth}
+% 		\centering
+% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8h_lblshort.png}
+% 	\end{minipage}
+% 	\begin{minipage}[t]{1.0\linewidth}
+% 		\centering
+% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8v_lblshort.png}
+% 	\end{minipage}
+% 	\begin{minipage}[t]{1.0\linewidth}
+% 		\centering
+% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8hv_lblshort.png}
+% 	\end{minipage}
+% 	\caption{\label{fig:8thexploss2017} Comparison of scaling of loss rates with excitation amplitude for \eighthtp\ only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by horizontal error bars.
+The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
+% \end{figure}
+
+\begin{figure}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch,\\ no excitation
+		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_804.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ $8^{\mathrm{th}}$~turn H
+		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_632.png}
+	\end{minipage}
+	\caption{\label{fig:8thexpprof} Horizontal beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+The profiles are taken at the end of the \eighthtp\ in H.
+For both bunches the transverse damper is not active.
+The distribution changes in the reference bunch (left), here slot 804, are much less than for the bunch experiencing the maximum excitation (right), here slot 632.}
 \end{figure}
-The small effect on the beam expected from simulations could then be also confirmed experimentally during the 2017 experiment.
-For this purpose the excitation amplitude was increased to the maximum reachable amplitude of 96~nrad.
-As depicted in Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the excitation amplitude for an excitation in H+V.
-For an excitation only in V or only in H, no statistically relevant increase was observed (see Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexploss2017}).
-In the horizontal plane, a change in beam distribution was measured with the BSRT profiles (Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexpprof}) for an excitation only in~H.
-The distribution increased around $2.0~\sigma$ simultaneously with a depletion of the beam core.
-For an excitation in H+V one would expect to see a similar distribution change.
-However, this change is not visible as the emittance of the references bunches was much smaller, around $1.8~\mu$m normalized emittance, compared to around $2.6~\mu$m for the excited bunches \cite{resexmd2017}.
-This difference in emittance then resulted in a stronger emittance growth due to intra-beam scattering for the reference bunches in comparison to the excited bunches making the additional emittance growth and distribution changes due to the excitation undetectable.
-In the vertical plane, the beam distribution and emittance were not affected by the resonant excitation.
-
-\subsection{random excitation\label{sec:simexran}}
+
+Excitation patterns that have little effect on the beam core and a
+large effect on the halo are good candidates for extending the range
+of hollow electron lens operation. In principle, this is possible
+because of the highly nonlinear fields generated by the HEL at the
+transverse locations of the beam halo and the small fields at the beam
+core. For this reason, the \eighthtp\ pattern was chosen for the
+experimental studies, as simulations indicated much smaller core
+effects compared to 7th- and \tenthtp, as discussed in
+Section~\ref{sec:pattern}.
+
+Simulations of intensities, emittances, and bunch lengths for
+different amplitudes of the 8th-turn excitation are shown in
+Fig.~\ref{fig:8thsim}. No significant effects are predicted for
+excitation amplitudes up to 192~nrad. Horizontal and vertical
+emittances show small changes in growth rate, but without a clear
+dependence on the amplitude or plane of the excitation.
+
+The \fma\ in amplitude space is shown in
+Fig.~\ref{fig:8th2017fmaamp}. It reveals that the driven resonances
+are of high order, mainly $16 Q_y$ and \mbox{$8 Q_x - 4 Q_y$}, and
+their effect is therefore expected to be small.
+
+The small effect on the beam expected from simulations could then be
+also confirmed experimentally during the 2017 experiment. For this
+purpose the excitation amplitude was increased to the maximum
+reachable amplitude of 96~nrad. As depicted in
+Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the
+excitation amplitude for an excitation in H+V. For an excitation only
+in V or only in H, no statistically relevant increase was observed
+(see Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexploss2017}). In the
+horizontal plane, a change in beam distribution was measured with the
+BSRT profiles (Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexpprof})
+for an excitation only in~H. The distribution increased around
+$2.0~\sigma$ simultaneously with a depletion of the beam core. For an
+excitation in H+V one would expect to see a similar distribution
+change. However, this change is not visible as the emittance of the
+references bunches was much smaller, around $1.8~\mu$m normalized
+emittance, compared to around $2.6~\mu$m for the excited bunches
+\cite{resexmd2017}. This difference in emittance then resulted in a
+stronger emittance growth due to intra-beam scattering for the
+reference bunches in comparison to the excited bunches making the
+additional emittance growth and distribution changes due to the
+excitation undetectable. In the vertical plane, the beam distribution
+and emittance were not affected by the resonant excitation.
+
+% The small effect on the beam expected from simulations could then be
+% also confirmed experimentally during the 2017 experiment. For this
+% purpose the excitation amplitude was increased to the maximum
+% reachable amplitude of 96~nrad. As depicted in
+% Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the
+% excitation amplitude for an excitation in H+V. For an excitation only
+% in V or only in H, no statistically relevant increase was observed
+% (see Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexploss2017}). In the
+% horizontal plane, a change in beam distribution was measured with the
+% BSRT profiles (Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexpprof})
+% for an excitation only in~H. The distribution increased around
+% $2.0~\sigma$ simultaneously with a depletion of the beam core. For an
+% excitation in H+V one would expect to see a similar distribution
+% change. However, this change is not visible as the emittance of the
+% references bunches was much smaller, around $1.8~\mu$m normalized
+% emittance, compared to around $2.6~\mu$m for the excited bunches
+% \cite{resexmd2017}. This difference in emittance then resulted in a
+% stronger emittance growth due to intra-beam scattering for the
+% reference bunches in comparison to the excited bunches making the
+% additional emittance growth and distribution changes due to the
+% excitation undetectable. In the vertical plane, the beam distribution
+% and emittance were not affected by the resonant excitation.
+
+\subsection{Random excitation}
+\label{sec:simexran}
+
+\begin{figure*}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		no excitation
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		uniform random H+V
+		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerruranadthv_1nrad_dp0_amp.png}
+	\end{minipage}	
+	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space without excitation (left) and for a uniform random excitation with 1~nrad amplitude (right) in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).}
+\end{figure*}
+
+\begin{figure*}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_intensity.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_sigm.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_emit1.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_emit2.png}
+	\end{minipage}
+	\caption{\label{fig:ransim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+The solid line indicates the case with no excitation, and the dotted and dashed line the results for a uniform random excitation of 12~nrad and 24~nrad respectively.}
+      \end{figure*}
+
+\begin{figure*}
+	random excitation V\\
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_vran_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_vran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
+	\end{minipage}	
+		random excitation H+V\\
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hvran_no_damper_avg.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hvran_no_damper.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hvran_no_damper.png}
+	\end{minipage}	
+	\caption{\label{fig:ranexp} Relative intensity (left) and relative horizontal (center) and vertical (right) emittance for a random excitation only in V (top) and in H+V (bottom).
+Emittance growth is only caused in the plane of excitation.}
+\end{figure*}
+
+\begin{figure*}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		reference bunch,\\ no excitation
+		\includegraphics[width=1.0\linewidth]{profile_v_ranv_slot_1698.png}
+	\end{minipage}
+	\begin{minipage}[t]{0.49\linewidth}
+		\centering
+		excited bunch,\\ random excitation V
+		\includegraphics[width=1.0\linewidth]{profile_v_ranv_slot_1532.png}
+	\end{minipage}
+	\caption{\label{fig:ranexpprof}Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
+The profiles are taken at the end of the random excitation in V.
+For both bunches the transverse damper is not active.
+The distribution changes in the reference bunch (left), here slot 1698, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
+      \end{figure*}
+
+\begin{figure*}
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranh_lblshort.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranv_lblshort.png}
+	\end{minipage}	
+	\begin{minipage}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranhv_lblshort.png}
+	\end{minipage}	
+	\caption{\label{fig:ranexplossplane} Comparison of scaling of loss rates with excitation amplitude for a random excitation only in H (top), only in V (center) and in H+V (bottom) measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars.
+The loss rate is considerably reduced for the bunches with transverse damper active and is approximately additive with the excitation plane, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only in V.}
+      \end{figure*}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{2017_scale_amp_7th_ranhv_lblshort.png}
+	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and \seventhtp\ in H+V measured during the 2017 experiment.
+The excitation amplitude error is 15\% indicated by the horizontal error bars.}
+\end{figure}
+
 The random excitation presents in general the pulsing pattern leading to the fastest halo depletion rates achievable with an HEL.
@@
 The random excitation is in general so powerful as it excites all frequencies of the beam equally.
-\begin{figure*}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		no excitation
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		uniform random H+V
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerruranadthv_1nrad_dp0_amp.png}
-	\end{minipage}	
-	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space without excitation (left) and for a uniform random excitation with 1~nrad amplitude (right) in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).}
-\end{figure*}
-This can be seen clearly on the example of the FMA analysis in amplitude space depicted in Fig.~\ref{fig:ran2017fmaamp}.
-In the case of a random excitation, the diffusion in tune is increased in comparison to the $k^{\mathrm{th}}$~turn pulsing rather equally for all particle amplitudes.
-\begin{figure*}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_intensity.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_sigm.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_emit1.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_emit2.png}
-	\end{minipage}
-	\caption{\label{fig:ransim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
-The solid line indicates the case with no excitation, and the dotted and dashed line the results for a uniform random excitation of 12~nrad and 24~nrad respectively.}
-\end{figure*}
+
+This can be seen clearly on the example of the FMA analysis in
+amplitude space depicted in Fig.~\ref{fig:ran2017fmaamp}. In the case
+of a random excitation, the diffusion in tune is increased in
+comparison to the $k^{\mathrm{th}}$~turn pulsing rather equally for
+all particle amplitudes.
+
 The distribution tracking shown in Fig.~\ref{fig:ransim} also reveals a quite symmetric evolution of the beam parameters for a random excitation:
@@
         \end{itemize}
-\begin{figure*}[b]
-	random excitation V\\
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_vran_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_vran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
-	\end{minipage}	
-		random excitation H+V\\
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hvran_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hvran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hvran_no_damper.png}
-	\end{minipage}	
-	\caption{\label{fig:ranexp} Relative intensity (left) and relative horizontal (center) and vertical (right) emittance for a random excitation only in V (top) and in H+V (bottom).
-Emittance growth is only caused in the plane of excitation.}
-\end{figure*}
 
@@
 
-\begin{figure*}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_v_ranv_slot_1698.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ random excitation V
-		\includegraphics[width=1.0\linewidth]{profile_v_ranv_slot_1532.png}
-	\end{minipage}
-	\caption{\label{fig:ranexpprof}Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the random excitation in V.
-For both bunches the transverse damper is not active.
-The distribution changes in the reference bunch (left), here slot 1698, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
-\end{figure*}
 The changes in emittance are also directly visible in the beam profiles measured with the BSRT.
@@
 
-\begin{figure*}[h]
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranh_lblshort.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranv_lblshort.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranhv_lblshort.png}
-	\end{minipage}	
-	\caption{\label{fig:ranexplossplane} Comparison of scaling of loss rates with excitation amplitude for a random excitation only in H (top), only in V (center) and in H+V (bottom) measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.
-The loss rate is considerably reduced for the bunches with transverse damper active and is approximately additive with the excitation plane, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only in V.}
-\end{figure*}
-The experimental results for a maximum excitation amplitude of $A_{\mathrm{max}}=1$~nrad and $A_{\mathrm{max}}=6$~nrad also confirm the in general small losses predicted in simulations.
-The losses then increase more dramatically around the time when the maximum excitation amplitude is increased to $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexpprof}).
-This is most likely the point when, due to the large emittance growth, the physical aperture starts cutting into the more central region of the distribution.
-Furthermore, the dependence of the loss rate on the excitation plane is additive, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only V.
-The loss rate is also considerably reduced for the bunches with transverse damper active independent of the plane the excitation is applied in (see Fig.~\ref{fig:ranexplossplane}).
-In comparison to the \seventhtp, the loss rate for a random excitation is smaller for a maximum excitation amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for $A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}).
-This indicates an increase of the diffusion mostly in the beam tails for \seventhtp, while for a random excitation mostly the core is affected.
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=1.0\linewidth]{2017_scale_amp_7th_ranhv_lblshort.png}
-	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and \seventhtp\ in H+V measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.}
-\end{figure}
-
-\subsection{Effect of the transverse damper\label{sec:damp}}
-\begin{figure*}[b]
+The experimental results for a maximum excitation amplitude of
+$A_{\mathrm{max}}=1$~nrad and $A_{\mathrm{max}}=6$~nrad also confirm
+the in general small losses predicted in simulations. The losses then
+increase more dramatically around the time when the maximum excitation
+amplitude is increased to $A_{\mathrm{max}}=12$~nrad (see
+Fig.~\ref{fig:ranexpprof}). This is most likely the point when, due to
+the large emittance growth, the physical aperture starts cutting into
+the more central region of the distribution. Furthermore, the
+dependence of the loss rate on the excitation plane is additive,
+meaning that double the loss rate is observed for an excitation in H+V
+compared to only in H or only V. The loss rate is also considerably
+reduced for the bunches with transverse damper active independent of
+the plane the excitation is applied in (see
+Fig.~\ref{fig:ranexplossplane}). In comparison to the \seventhtp, the
+loss rate for a random excitation is smaller for a maximum excitation
+amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for
+$A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}). This
+indicates an increase of the diffusion mostly in the beam tails for
+\seventhtp, while for a random excitation mostly the core is affected.
+
+\subsection{Effect of the transverse damper}
+\label{sec:damp}
+
+\begin{figure*}
 	\centering
@@
       \end{figure*}
+
 For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally,
@@
 
-\section{Summary\label{sec:sum}}
+\section{Summary}
+\label{sec:sum}
+
 This paper intends to give an overview of the effects of the hollow electron lens (HEL) on the proton beam core and the experiments conducted in this context at the LHC in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}.
@@
 
-\section{Additional plots for \eighthtp\ during 2017 experiments}
-\label{app:sec:8}
-\begin{figure}[h]
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8h_lblshort.png}
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8v_lblshort.png}
-	\end{minipage}
-	\begin{minipage}[t]{1.0\linewidth}
-		\centering
-		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8hv_lblshort.png}
-	\end{minipage}
-	\caption{\label{fig:8thexploss2017} Comparison of scaling of loss rates with excitation amplitude for \eighthtp\ only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by horizontal error bars.
-The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
-\end{figure}
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_804.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ $8^{\mathrm{th}}$~turn H
-		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_632.png}
-	\end{minipage}
-	\caption{\label{fig:8thexpprof} Horizontal beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the \eighthtp\ in H.
-For both bunches the transverse damper is not active.
-The distribution changes in the reference bunch (left), here slot 804, are much less than for the bunch experiencing the maximum excitation (right), here slot 632.}
-\end{figure}
-
 \bibliography{resex-bibdesk}

### 2018-03-14 resex.tex
Revised text of 8th-turn results.

@@
 
-The small effect on the beam expected from simulations could then be
-also confirmed experimentally during the 2017 experiment. For this
-purpose the excitation amplitude was increased to the maximum
-reachable amplitude of 96~nrad. As depicted in
-Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the
-excitation amplitude for an excitation in H+V. For an excitation only
-in V or only in H, no statistically relevant increase was observed
-(see Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexploss2017}). In the
-horizontal plane, a change in beam distribution was measured with the
-BSRT profiles (Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexpprof})
-for an excitation only in~H. The distribution increased around
-$2.0~\sigma$ simultaneously with a depletion of the beam core. For an
-excitation in H+V one would expect to see a similar distribution
-change. However, this change is not visible as the emittance of the
-references bunches was much smaller, around $1.8~\mu$m normalized
-emittance, compared to around $2.6~\mu$m for the excited bunches
-\cite{resexmd2017}. This difference in emittance then resulted in a
-stronger emittance growth due to intra-beam scattering for the
-reference bunches in comparison to the excited bunches making the
-additional emittance growth and distribution changes due to the
-excitation undetectable. In the vertical plane, the beam distribution
-and emittance were not affected by the resonant excitation.
-
-% The small effect on the beam expected from simulations could then be
-% also confirmed experimentally during the 2017 experiment. For this
-% purpose the excitation amplitude was increased to the maximum
-% reachable amplitude of 96~nrad. As depicted in
-% Fig.~\ref{fig:8thexploss}, the loss rate increased slightly with the
-% excitation amplitude for an excitation in H+V. For an excitation only
-% in V or only in H, no statistically relevant increase was observed
-% (see Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexploss2017}). In the
-% horizontal plane, a change in beam distribution was measured with the
-% BSRT profiles (Appendix~\ref{app:sec:8}, Fig.~\ref{fig:8thexpprof})
-% for an excitation only in~H. The distribution increased around
-% $2.0~\sigma$ simultaneously with a depletion of the beam core. For an
-% excitation in H+V one would expect to see a similar distribution
-% change. However, this change is not visible as the emittance of the
-% references bunches was much smaller, around $1.8~\mu$m normalized
-% emittance, compared to around $2.6~\mu$m for the excited bunches
-% \cite{resexmd2017}. This difference in emittance then resulted in a
-% stronger emittance growth due to intra-beam scattering for the
-% reference bunches in comparison to the excited bunches making the
-% additional emittance growth and distribution changes due to the
-% excitation undetectable. In the vertical plane, the beam distribution
-% and emittance were not affected by the resonant excitation.
+The effects on the beam were tested during the 2017 experiment. The
+excitation amplitude was increased to a maximum of 96~nrad. The loss
+rates as a function of excitation amplitude are presented in
+Fig.~\ref{fig:8thexploss2017}. For H and V excitations, no significant
+increase of the loss rate was observed. A weak increase with amplitude
+was seen in the case of H+V excitation. Observed and predicted loss
+rates are of the same magnitude (about 0.2\%/h).
+
+Changes in beam distributions were also detected
+(Fig.~\ref{fig:8thexpprof}). The horizontal distribution responded to
+the H excitation with a slight depletion of the core and a
+corresponding increase around $2\sigma$. For an H+V excitation, one
+would expect a similar distribution change. However, this change could
+not be detected directly, because, in this case, the emittance of the
+control bunches was much smaller (\q{1.8}{\mu m}) than the emittance
+of the excited bunches (\q{2.6}{\mu m}), and intrabeam scattering
+masked the effects of the excitation~\cite{resexmd2017}. In the
+vertical plane, beam distributions and emittances were not affected
+by the resonant excitation.
 

### 2018-03-14 resex.tex
Revised results on 8th-turn pulsing. Defined a set of figure widths.

@@
 
+% abbreviations
 \newcommand{\seventhtp}{7th-turn pulsing}
@@
 \newcommand{\fma}{frequency-map analysis}
+
+% widths for figures
+\newlength{\halfwidth}
+\setlength{\halfwidth}{0.48\textwidth}
+\newlength{\thirdwidth}
+\setlength{\thirdwidth}{0.33\textwidth}
+\newlength{\bsrtwidth}
+\setlength{\bsrtwidth}{0.4\textwidth}
+\newlength{\fmawidth}
+\setlength{\fmawidth}{0.4\textwidth}
 
@@
 \begin{tabular}{cc}
-\includegraphics[width=0.49\textwidth]{2016injerra2b2u_pattern_3_5um_intensity.png}
+\includegraphics[width=\halfwidth]{2016injerra2b2u_pattern_3_5um_intensity.png}
   &
-\includegraphics[width=0.49\textwidth]{2016injerra2b2u_pattern_3_5um_sigm.png}
+\includegraphics[width=\halfwidth]{2016injerra2b2u_pattern_3_5um_sigm.png}
   \\
-\includegraphics[width=0.49\textwidth]{2016injerra2b2u_pattern_3_5um_emit1.png} &
-\includegraphics[width=0.49\textwidth]{2016injerra2b2u_pattern_3_5um_emit2.png}
+\includegraphics[width=\halfwidth]{2016injerra2b2u_pattern_3_5um_emit1.png} &
+\includegraphics[width=\halfwidth]{2016injerra2b2u_pattern_3_5um_emit2.png}
   \\
@@
     no excitation & $10^{\mathrm{th}}$, H+V \\
-    \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png} &
-    \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
+    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png} &
+    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
     \\
     7th, H & 7th, V \\
-    \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
-    \includegraphics[width=0.49\textwidth]{2016injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
+    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
+    \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
   \end{tabular}
@@
     no excitation & random, H+V \\
-    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_3_5um_hist_y.png} &
-    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
+    \includegraphics[width=\halfwidth]{2016injerra2b2u_3_5um_hist_y.png} &
+    \includegraphics[width=\halfwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
     7th, H+V & $10^{\mathrm{th}}$, H+V \\
-    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
+    \includegraphics[width=\halfwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
                   &
-    \includegraphics[width=0.49\textwidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png} \\
+    \includegraphics[width=\halfwidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png} \\
   \end{tabular}
@@
   \begin{tabular}{ccc}
-    \includegraphics[width=0.32\textwidth]{2016_bunch_intensity_v10th_no_damper_avg.png}&
-    \includegraphics[width=0.32\textwidth]{2016_emith_avg_rel_v10th_no_damper.png}&
-    \includegraphics[width=0.32\textwidth]{2016_emitv_avg_rel_v10th_no_damper.png}\\
+    \includegraphics[width=\thirdwidth]{2016_bunch_intensity_v10th_no_damper_avg.png}&
+    \includegraphics[width=\thirdwidth]{2016_emith_avg_rel_v10th_no_damper.png}&
+    \includegraphics[width=\thirdwidth]{2016_emitv_avg_rel_v10th_no_damper.png}\\
   \end{tabular}
@@
     Reference bunch, no excitation & Excited bunch, maximum amplitude \\
-    \includegraphics[width=0.48\textwidth]{profile_v_10thv_slot_50.png} &
-    \includegraphics[width=0.48\textwidth]{profile_v_10thv_slot_1300.png}\\
+    \includegraphics[width=\bsrtwidth]{profile_v_10thv_slot_50.png} &
+    \includegraphics[width=\bsrtwidth]{profile_v_10thv_slot_1300.png}\\
     \end{tabular}
@@
   \begin{tabular}{cc}
-    \includegraphics[width=0.49\textwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_intensity.png} &
-    \includegraphics[width=0.49\textwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2_rel.png} \\
-    \includegraphics[width=0.49\textwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit1.png} &
-    \includegraphics[width=0.49\textwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2.png} \\
+    \includegraphics[width=\halfwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_intensity.png} &
+    \includegraphics[width=\halfwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2_rel.png} \\
+    \includegraphics[width=\halfwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit1.png} &
+    \includegraphics[width=\halfwidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2.png} \\
   \end{tabular}	
@@
   \centering
-  \includegraphics[width=\linewidth]{2016_scale_amp_10v_ran_lblshort_sim_prstab.png}
+  \includegraphics[width=\columnwidth]{2016_scale_amp_10v_ran_lblshort_sim_prstab.png}
   \caption{Comparison between experiments and simulations of loss
@@
     no excitation & 7th, H+V \\
-    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate_7th.png} &
-    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png} \\
+    \includegraphics[width=\fmawidth]{2017injnocolc15o+19_6noerru_dp0_ord14_annotate_7th.png} &
+    \includegraphics[width=\fmawidth]{2017injnocolc15o+19_6noerrut7skhv_dp0_ord7.png} \\
     7th, H & 7th, V \\
-    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
-    \includegraphics[width=0.49\textwidth]{2017injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
+    \includegraphics[width=\fmawidth]{2017injnocolc15o+19_6noerrut7skh_dp0_ord7.png} &
+    \includegraphics[width=\fmawidth]{2017injnocolc15o+19_6noerrut7skv_dp0_ord7.png} \\
   \end{tabular}
@@
   \begin{tabular}{cc}
-    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png} &
-    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png} \\
-    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png} &
-    \includegraphics[width=0.49\textwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png} \\
+    \includegraphics[width=\halfwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_intensity.png} &
+    \includegraphics[width=\halfwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_rel_emit2.png} \\
+    \includegraphics[width=\halfwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit1.png} &
+    \includegraphics[width=\halfwidth]{2016+2017injerra2b2uran1_2e-3_7th_3_5um_emit2.png} \\
   \end{tabular}
@@
     distribution tracking based on the 2016 injection optics with
-    standard errors and $(Q_x, Q_y) = (64.28, 59.31)$ and 2017
-    injection optics with standard errors and
+    standard lattice errors and $(Q_x, Q_y) = (64.28, 59.31)$ and 2017
+    injection optics with standard lattice errors and
     $(Q_x, Q_y) = (62.27, 60.295)$: relative intensity (top left);
@@
     \multicolumn{3}{c}{2016 experiment, 7th, H}	\\
-    \includegraphics[width=0.32\textwidth]{2016_bunch_intensity_h7th_no_damper_avg.png} &
-    \includegraphics[width=0.32\textwidth]{2016_emith_avg_rel_h7th_no_damper.png} &
-    \includegraphics[width=0.32\textwidth]{2016_emitv_avg_rel_h7th_no_damper.png} \\
+    \includegraphics[width=\thirdwidth]{2016_bunch_intensity_h7th_no_damper_avg.png} &
+    \includegraphics[width=\thirdwidth]{2016_emith_avg_rel_h7th_no_damper.png} &
+    \includegraphics[width=\thirdwidth]{2016_emitv_avg_rel_h7th_no_damper.png} \\
     \multicolumn{3}{c}{2017 experiment, 7th, H+V} \\
-    \includegraphics[width=0.32\textwidth]{2017_bunch_intensity_hv7th_no_damper_avg.png} &
-    \includegraphics[width=0.32\textwidth]{2017_emith_avg_rel_hv7th_no_damper.png} &
-    \includegraphics[width=0.32\textwidth]{2017_emitv_avg_rel_hv7th_no_damper.png} \\
+    \includegraphics[width=\thirdwidth]{2017_bunch_intensity_hv7th_no_damper_avg.png} &
+    \includegraphics[width=\thirdwidth]{2017_emith_avg_rel_hv7th_no_damper.png} &
+    \includegraphics[width=\thirdwidth]{2017_emitv_avg_rel_hv7th_no_damper.png} \\
   \end{tabular}
@@
   \begin{tabular}{ccc}
-    \includegraphics[width=0.32\textwidth]{2017_scale_amp_7h_lblshort.png} &
-    \includegraphics[width=0.32\textwidth]{2017_scale_amp_7v_lblshort.png} &
-    \includegraphics[width=0.32\textwidth]{2017_scale_amp_7hv_lblshort.png} \\
+    \includegraphics[width=\thirdwidth]{2017_scale_amp_7h_lblshort.png} &
+    \includegraphics[width=\thirdwidth]{2017_scale_amp_7v_lblshort.png} &
+    \includegraphics[width=\thirdwidth]{2017_scale_amp_7hv_lblshort.png} \\
   \end{tabular}
@@
     Reference bunch, no excitation & Excited bunch, 7th, H+V \\
-    \includegraphics[width=0.48\textwidth]{profile_v_7thhv_slot_2862.png} &
-    \includegraphics[width=0.48\textwidth]{profile_v_7thhv_slot_2696.png} \\
+    \includegraphics[width=\bsrtwidth]{profile_v_7thhv_slot_2862.png} &
+    \includegraphics[width=\bsrtwidth]{profile_v_7thhv_slot_2696.png} \\
   \end{tabular}
@@
     profiles are taken at the end of \seventhtp\ in H+V. For these
-    bunches, the transverse damper was not active. The data is
+    bunches, the transverse damper was not active. The data are
     presented in the same way as in Fig.~\ref{fig:10thexpprof}. The
@@
 \begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_intensity.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_sigm.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit1.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit2.png}
-	\end{minipage}	
-	\caption{\label{fig:8thsim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulations (distribution tracking) based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
-The solid line indicates an excitation with only a random dipole noise component in H+V of 6~nrad, and the dotted and dashed line the results for \eighthtp\ plus a random dipole noise component in H+V of 6~nrad.}
-      \end{figure*}
-
-\begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		no excitation
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$8^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerrut8skhv_dp0_amp_annotate.png}
-	\end{minipage}	
-	\caption{\label{fig:8th2017fmaamp} FMA analysis in amplitude space without excitation (left) and for \eighthtp\ (right) based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).
-The excitation is 96~nrad in both planes.
-The $16Q_y$ and \mbox{$8Q_x-4Q_y$} are both excited.}
-\end{figure*}
-
-% \begin{figure}
-% 	\centering
-% 	\includegraphics[width=1.0\linewidth]{2017_scale_amp_8hv_lblshort.png}
-% 	\caption{\label{fig:8thexploss} Comparison of scaling of loss rates with excitation amplitude for \eighthtp\ in H+V measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars, the error on the loss rate is the error on the mean obtained through the average over the 6~bunches with the same excitation amplitude.}
-% \end{figure}
+  \begin{tabular}{cc}
+  \includegraphics[width=\halfwidth]{2017injerra2b2uran1_2e-3_8th_3_5um_intensity.png} &
+  \includegraphics[width=\halfwidth]{2017injerra2b2uran1_2e-3_8th_3_5um_sigm.png} \\
+  \includegraphics[width=\halfwidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit1.png} &
+  \includegraphics[width=\halfwidth]{2017injerra2b2uran1_2e-3_8th_3_5um_emit2.png} \\
+  \end{tabular}
+  \caption{Calculated relative intensities (top left), bunch lengths
+    (top right), horizontal emittances (bottom left), and vertical
+    emittances (bottom right) from distribution-tracking simulations
+    based on the 2017 injection optics with standard lattice errors
+    and tunes $(Q_x, Q_y) = (62.27, 60.295)$. The solid black line is
+    the reference case, including only a random dipole noise component
+    in H+V of 6~nrad. The dotted and dashed lines are the results for
+    \eighthtp, including the same random dipole noise component.}
+  \label{fig:8thsim}
+\end{figure*}
+
+\begin{figure*}
+  \begin{tabular}{cc}
+    No excitation & 8th, H+V \\
+    \includegraphics[width=\fmawidth]{2017injnocolc15o+19_6noerru_dp0_amp.png} &
+    \includegraphics[width=\fmawidth]{2017injnocolc15o+19_6noerrut8skhv_dp0_amp_annotate.png}
+  \end{tabular}
+  \caption{FMA in transverse amplitude space without excitation (left)
+    and for \eighthtp\ (right) based on the 2017 injection optics with
+    no lattice errors and tunes (62.27, 60.295). The excitation
+    amplitude is 96~nrad in both planes. The $16 Q_y$ and
+    \mbox{$8 Q_x - 4 Q_y$} resonances are excited.}
+  \label{fig:8th2017fmaamp}
+\end{figure*}
 
@@
   \begin{tabular}{ccc}
-  \includegraphics[width=0.32\textwidth]{2017_scale_amp_8h_lblshort.png} &
-  \includegraphics[width=0.32\textwidth]{2017_scale_amp_8v_lblshort.png} &
-  \includegraphics[width=0.32\textwidth]{2017_scale_amp_8hv_lblshort.png} \\
-  \end{tabular}
-  \caption{Comparison of scaling of loss rates with excitation
-    amplitude for \eighthtp\ only in H (top), only in V (center) and
-    in H+V (bottom) during the 2017 experiment. The excitation
-    amplitude error is 15\% indicated by horizontal error bars. The
-    vertical errorbars represent the pure statistical error based on
-    the averaging over several bunches with the same excitation
-    amplitude.}
+  \includegraphics[width=\thirdwidth]{2017_scale_amp_8h_lblshort.png} &
+  \includegraphics[width=\thirdwidth]{2017_scale_amp_8v_lblshort.png} &
+  \includegraphics[width=\thirdwidth]{2017_scale_amp_8hv_lblshort.png} \\
+  \end{tabular}
+  \caption{Measured loss rates vs.\ excitation amplitude for
+    \eighthtp\ in H only (left), in V only (center) and in H+V (right)
+    during the 2017 experiment. Uncertainties are statistical. The
+    differences between the three data sets within each excitation
+    mode (black, red, and yellow curves) provide an estimate of the
+    systematic errors.}
   \label{fig:8thexploss2017}
@@
 
-% \begin{figure}
-% 	\begin{minipage}[t]{1.0\linewidth}
-% 		\centering
-% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8h_lblshort.png}
-% 	\end{minipage}
-% 	\begin{minipage}[t]{1.0\linewidth}
-% 		\centering
-% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8v_lblshort.png}
-% 	\end{minipage}
-% 	\begin{minipage}[t]{1.0\linewidth}
-% 		\centering
-% 		\includegraphics[width=0.9\linewidth]{2017_scale_amp_8hv_lblshort.png}
-% 	\end{minipage}
-% 	\caption{\label{fig:8thexploss2017} Comparison of scaling of loss rates with excitation amplitude for \eighthtp\ only in H (top), only in V (center) and in H+V (bottom) during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by horizontal error bars.
-The vertical errorbars represent the pure statistical error based on the averaging over several bunches with the same excitation amplitude.}
-% \end{figure}
-
-\begin{figure}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_804.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ $8^{\mathrm{th}}$~turn H
-		\includegraphics[width=1.0\linewidth]{profile_h_8thh_slot_632.png}
-	\end{minipage}
-	\caption{\label{fig:8thexpprof} Horizontal beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the \eighthtp\ in H.
-For both bunches the transverse damper is not active.
-The distribution changes in the reference bunch (left), here slot 804, are much less than for the bunch experiencing the maximum excitation (right), here slot 632.}
-\end{figure}
+\begin{figure*}
+  \begin{tabular}{cc}
+    Reference bunch, no excitation & Excited bunch, 8th, H \\
+    \includegraphics[width=\bsrtwidth]{profile_h_8thh_slot_804.png} &
+    \includegraphics[width=\bsrtwidth]{profile_h_8thh_slot_632.png} \\
+   \end{tabular}
+   \caption{Horizontal beam profiles measured with the Beam
+     Synchrotron Radiation Telescope (BSRT) during the 2017
+     experiments. The profiles are taken at the end of the \eighthtp\
+     in H. For these bunches, the transverse damper was not
+     active. The data are presented in the same way as in
+     Fig.~\ref{fig:10thexpprof}. The profile changes in the bunch
+     affected by the maximum excitation (right) are larger than those
+     displayed by the control bunch (left).}
+   \label{fig:8thexpprof}
+\end{figure*}
 
@@
 	\end{minipage}
-	\caption{\label{fig:ransim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2017 injection optics with standard errors and $(Q_x,Q_y)=(62.27,60.295)$.
+	\caption{\label{fig:ransim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2017 injection optics with standard lattice errors and $(Q_x,Q_y)=(62.27,60.295)$.
 The solid line indicates the case with no excitation, and the dotted and dashed line the results for a uniform random excitation of 12~nrad and 24~nrad respectively.}

### 2018-03-15 resex.tex
Intermediate step in revision of results on random pulsing.

@@
 % abbreviations
+\newcommand{\kthtp}{$k$th-turn pulsing}
 \newcommand{\seventhtp}{7th-turn pulsing}
@@
 
+\newcommand{\Fma}{Frequency-map analysis}
 \newcommand{\fma}{frequency-map analysis}
@@
   \begin{equation}
+    \label{eq:random}
     U_{\mathrm{e-gun}} = (1-a) \cdot U_{\mathrm{max}} +
@@
 
-The random excitation presents in general the pulsing pattern leading to the fastest halo depletion rates achievable with an HEL.
-On the other hand this pattern is also the most dangerous one for the proton beam core.
-The aim of this experiment was to test the effect on the core in comparison to the resonant excitation represented in the 2017 experiment by primarily the \seventhtp\ as for \eighthtp\ basically no effect was observed.
-The random excitation is in general so powerful as it excites all frequencies of the beam equally.
-
-This can be seen clearly on the example of the FMA analysis in
-amplitude space depicted in Fig.~\ref{fig:ran2017fmaamp}. In the case
-of a random excitation, the diffusion in tune is increased in
-comparison to the $k^{\mathrm{th}}$~turn pulsing rather equally for
-all particle amplitudes.
-
-The distribution tracking shown in Fig.~\ref{fig:ransim} also reveals a quite symmetric evolution of the beam parameters for a random excitation:
+The random excitation is a powerful and flexible halo-cleaning
+pattern, but it is also potentially harmful for the proton beam
+core. A random pattern can be employed at full strength or added as a
+modulation to a continuous mode of operation (depending on the
+parameter $a$ in Eq.~\ref{eq:random}). During the 2017 experiment, we
+tested the effects of the random excitation at full strength ($a = 1$)
+and compared it with the resonant excitations.
+
+The random excitation excites practically all resonant frequencies in
+the beam.  This can be clearly seen in the FMA example in amplitude
+space illustrated in Fig.~\ref{fig:ran2017fmaamp}. Tune diffusion is
+increased rather uniformly at all particle amplitudes, and not on
+specific resonant lines, as it happens for \kthtp.
+
+The calculated results of distribution tracking, with excitation
+amplitudes up to 24~nrad, are shown in
+Fig.~\ref{fig:ransim}. They show the following features:
+%
 \begin{itemize}
-	\item Emittance growth is only caused by an excitation in the same plane.
-	\item The emittance growth is quantitatively comparable for an excitation in only the corresponding plane.
-Only a small increase or decrease is observed if the excitation is applied in both planes.
-The difference could be due to coupling.
-	\item The random excitation leads to an increase of the emittance growth rate. Compared to the $k^{\mathrm{th}}$~turn pulsing there is explicitly no initial adjustment phase of the distribution.
-	\item Negligible losses are observed independent of the plane of excitation.
+\item negligible losses are observed, independent of the plane of
+  excitation;
+\item there is no significant change in bunch length;
+\item emittance growth is only generated by excitations in the same
+  plane (or in both planes);
+\item horizontal and vertical emittance growth rates are large and
+  comparable in magnitude;
+\item when the excitation is applied in both planes, a small increase
+  or decrease in the emittance growth rate is observed with respect to
+  the cases of separate excitations, an effect likely due to lattice
+  coupling;
+\item emittance growth takes place at a constant rate, without the
+  initial adjustment phase characteristic of \kthtp.
 \end{itemize}
@@
 
-The changes in emittance are also directly visible in the beam profiles measured with the BSRT.
-As an example Fig.~\ref{fig:ranexpprof} shows the profile for a random excitation only in V in comparison to a reference bunch.
-The distribution also stays Gaussian in case of a random excitation in contrast to the \tenthtp\ in which case it clearly assumes a non-Gaussian shape (see Fig.~\ref{fig:10thexpprof}).
+The changes in emittance are also directly visible in the beam
+profiles measured with the BSRT. As an example
+Fig.~\ref{fig:ranexpprof} shows the profile for a random excitation
+only in V in comparison to a reference bunch. The distribution also
+stays Gaussian in case of a random excitation in contrast to the
+\tenthtp\ in which case it clearly assumes a non-Gaussian shape (see
+Fig.~\ref{fig:10thexpprof}).
 

### 2018-03-15 resex.tex
Another intermediate step in revision of random results. Emittance growth rates to be verified.

@@
 
-Qualitatively, this behavior could be also confirmed experimentally
-during the 2017 experiments. Here excitation amplitude dependent
-emittance growth was only recorded in the plane of excitation with all
-excitation amplitude dependent emittance growth rates in a comparable
-range. As example the relative horizontal and vertical emittance for
-pulsing in V and in H+V are depicted in Fig.~\ref{fig:ranexp}. Note
-that as expected from simulations and despite the large excitation
-amplitude dependent emittance growth, the growth rates stays
-constant. There is explicitly no fast adjustment phase succeeded by a
-new equilibrium state as observed for the 7th- and \tenthtp.
+The effects of random excitation were measured during the 2017
+experiments. Increasing emittance growth with excitation amplitude was
+recorded only in the plane of excitation. Horizontal and vertical
+emittance growth rates for a given amplitude were comparable. As an
+example, the relative horizontal and vertical emittances during V and
+H+V pulsing are illustrated in Fig.~\ref{fig:ranexp}. In addition, the
+excitations translated into constant growth rates. In particular,
+there was no adjustment phase followed by an equilibrium state, as
+observed during 7th- and \tenthtp. The qualitative behavior of
+emittances predicted in simulations was therefore confirmed. However,
+the calculations could only roughly reproduce the magnitude of the
+measured effects and were in general lower. For instance, at 6~nrad,
+the calculated emittance growth was 1\%/min, whereas in experiments we
+measured 5\%/min.
 

### 2018-03-16 resex.tex
Revised text of random excitation results. Added note in README to verify magnitude of loss rates.

@@
 
-The changes in emittance are also directly visible in the beam
-profiles measured with the BSRT. As an example
-Fig.~\ref{fig:ranexpprof} shows the profile for a random excitation
-only in V in comparison to a reference bunch. The distribution also
-stays Gaussian in case of a random excitation in contrast to the
-\tenthtp\ in which case it clearly assumes a non-Gaussian shape (see
-Fig.~\ref{fig:10thexpprof}).
-
-The experimental results for a maximum excitation amplitude of
-$A_{\mathrm{max}}=1$~nrad and $A_{\mathrm{max}}=6$~nrad also confirm
-the in general small losses predicted in simulations. The losses then
-increase more dramatically around the time when the maximum excitation
-amplitude is increased to $A_{\mathrm{max}}=12$~nrad (see
-Fig.~\ref{fig:ranexpprof}). This is most likely the point when, due to
-the large emittance growth, the physical aperture starts cutting into
-the more central region of the distribution. Furthermore, the
-dependence of the loss rate on the excitation plane is additive,
-meaning that double the loss rate is observed for an excitation in H+V
-compared to only in H or only V. The loss rate is also considerably
-reduced for the bunches with transverse damper active independent of
-the plane the excitation is applied in (see
-Fig.~\ref{fig:ranexplossplane}). In comparison to the \seventhtp, the
-loss rate for a random excitation is smaller for a maximum excitation
-amplitude of $A_{\mathrm{max}}=6$~nrad, but larger for
-$A_{\mathrm{max}}=12$~nrad (see Fig.~\ref{fig:ranexploss}). This
-indicates an increase of the diffusion mostly in the beam tails for
-\seventhtp, while for a random excitation mostly the core is affected.
+The changes in beam distributions were also directly detected in the
+BSRT profiles. As an example, Fig.~\ref{fig:ranexpprof} shows the
+profiles induced by a vertical random excitation, compared to those of
+a reference bunch. In this case, the distribution widens, but it
+retains a Gaussian form, in contrast to the case of \tenthtp, when it
+assumed a non-Gaussian shape (see Fig.~\ref{fig:10thexpprof}, for
+instance).
+
+The measured loss rates are presented in
+Fig.~\ref{fig:ranexplossplane} as a function of random excitation
+amplitude in H, V, and H+V planes.  Losses increased quadratically
+with excitation amplitude, contributing approximately 3\%/h at
+12~nrad. In simulations, the qualitative behavior was the same;
+however, the predicted magnitude was much lower (0.1\%/h at
+12~nrad). Moreover, the measured effect of the combined H+V excitation
+was approximately the sum of H and V separately.
+
+The systematic effects on losses during random excitations were
+smaller than during resonant pulsing, as shown by the agreement
+between the 3 data sets (black, red, yellow) within each of the
+excitation modes in Fig.~\ref{fig:ranexplossplane}. This fact may be
+an indication of the random excitation affecting the whole beam,
+whereas, in the cases under study, resonant pulsing patterns excited
+specific amplitude regions and were therefore more susceptible to
+their population and to the order in which experiments were
+performed. Another indication comes from the direct comparison of loss
+rates vs.\ excitation amplitude for random and \seventhtp\
+(Fig.~\ref{fig:ranexploss}). The first time it is applied (red
+points), the 7th-turn excitation grows quadratically with amplitude,
+whereas the second time it shows signs of saturation (yellow). In any
+case, although the dynamics of their action may be different, these
+two pulsing patterns were the most powerful, generating losses at the
+level of several percent per hour at an amplitude of 12~nrad.
+
+Another feature of random excitation was that the loss rates were
+reduced by about a factor~2 when the transverse damping system was
+active (dashed vs.\ solid lines in
+Fig.~\ref{fig:ranexplossplane}). This was true for all three modes of
+excitation (H, V, and H+V). On the other hand, for the resonant
+pulsing modes, the transverse damper seemed to have a negligible
+effect. This observation is discussed in more detail in
+Section~\ref{sec:damp}.
 

### 2018-03-16 resex.tex
Revised results on random excitation, including figures.

@@
 \begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		no excitation
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerru_dp0_amp.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		uniform random H+V
-		\includegraphics[width=1.0\linewidth]{2017injnocolc15o+19_6noerruranadthv_1nrad_dp0_amp.png}
-	\end{minipage}	
-	\caption{\label{fig:ran2017fmaamp} FMA analysis in amplitude space without excitation (left) and for a uniform random excitation with 1~nrad amplitude (right) in H+V based on the 2017 injection optics with no machine errors and a tune of (62.27,60.295).}
-\end{figure*}
-
-\begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_intensity.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_sigm.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_emit1.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017injerra2b2u_ranadt_3_5um_emit2.png}
-	\end{minipage}
-	\caption{\label{fig:ransim} Relative intensity (top left), bunch length (top right), and horizontal (bottom left) and vertical (bottom right) emittance obtained from distribution tracking based on the 2017 injection optics with standard lattice errors and $(Q_x,Q_y)=(62.27,60.295)$.
-The solid line indicates the case with no excitation, and the dotted and dashed line the results for a uniform random excitation of 12~nrad and 24~nrad respectively.}
-      \end{figure*}
-
-\begin{figure*}
-	random excitation V\\
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_vran_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_vran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
-	\end{minipage}	
-		random excitation H+V\\
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_bunch_intensity_hvran_no_damper_avg.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emith_avg_rel_hvran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_hvran_no_damper.png}
-	\end{minipage}	
-	\caption{\label{fig:ranexp} Relative intensity (left) and relative horizontal (center) and vertical (right) emittance for a random excitation only in V (top) and in H+V (bottom).
-Emittance growth is only caused in the plane of excitation.}
-\end{figure*}
-
-\begin{figure*}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		reference bunch,\\ no excitation
-		\includegraphics[width=1.0\linewidth]{profile_v_ranv_slot_1698.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		excited bunch,\\ random excitation V
-		\includegraphics[width=1.0\linewidth]{profile_v_ranv_slot_1532.png}
-	\end{minipage}
-	\caption{\label{fig:ranexpprof}Vertical beam profiles measured with the Beam Synchrotron Radiation Telescope (BSRT) during the 2017 experiments.
-The profiles are taken at the end of the random excitation in V.
-For both bunches the transverse damper is not active.
-The distribution changes in the reference bunch (left), here slot 1698, are much less than for the bunch experiencing the maximum excitation (right), here slot 2696.}
-      \end{figure*}
-
-\begin{figure*}
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranh_lblshort.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranv_lblshort.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		\includegraphics[width=1.0\linewidth]{2017_scale_amp_ranhv_lblshort.png}
-	\end{minipage}	
-	\caption{\label{fig:ranexplossplane} Comparison of scaling of loss rates with excitation amplitude for a random excitation only in H (top), only in V (center) and in H+V (bottom) measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.
-The loss rate is considerably reduced for the bunches with transverse damper active and is approximately additive with the excitation plane, meaning that double the loss rate is observed for an excitation in H+V compared to only in H or only in V.}
+  \begin{tabular}{cc}
+    No excitation & Random excitation, H+V \\
+    \includegraphics[width=\fmawidth]{2017injnocolc15o+19_6noerru_dp0_amp.png} &
+    \includegraphics[width=\fmawidth]{2017injnocolc15o+19_6noerruranadthv_1nrad_dp0_amp.png} \\
+  \end{tabular}
+  \caption{FMA in transverse amplitude space without excitation (left)
+    and with a random 1-nrad H+V excitation (right), based on the 2017
+    injection optics with no lattice errors and tunes (62.27,
+    60.295).}
+  \label{fig:ran2017fmaamp}
+\end{figure*}
+
+\begin{figure*}
+  \begin{tabular}{cc}
+    \includegraphics[width=\halfwidth]{2017injerra2b2u_ranadt_3_5um_intensity.png} &
+    \includegraphics[width=\halfwidth]{2017injerra2b2u_ranadt_3_5um_sigm.png}\\
+    \includegraphics[width=\halfwidth]{2017injerra2b2u_ranadt_3_5um_emit1.png} &
+    \includegraphics[width=\halfwidth]{2017injerra2b2u_ranadt_3_5um_emit2.png}\\
+  \end{tabular}
+  \caption{Simulations (distribution tracking) based on the 2017
+    injection optics with standard lattice errors and
+    $(Q_x, Q_y) = (62.27, 60.295)$: relative bunch intensity (top
+    left), bunch length (top right), horizontal emittance (bottom
+    left), and vertical emittance (bottom right). The solid black line
+    indicates the reference case with no excitation. The dashed and
+    dotted lines are the results of random excitations (H, V, or H+V)
+    with amplitudes 12~nrad and 24~nrad.}
+  \label{fig:ransim}
+\end{figure*}
+
+\begin{figure*}
+  \begin{tabular}{ccc}
+    \multicolumn{3}{c}{Random excitation, V} \\
+    \includegraphics[width=\thirdwidth]{2017_bunch_intensity_vran_no_damper_avg.png} &
+    \includegraphics[width=\thirdwidth]{2017_emith_avg_rel_vran_no_damper.png} &
+    \includegraphics[width=\thirdwidth]{2017_emitv_avg_rel_vran_no_damper.png} \\
+    \multicolumn{3}{c}{Random excitation, H+V} \\
+    \includegraphics[width=\thirdwidth]{2017_bunch_intensity_hvran_no_damper_avg.png} &
+    \includegraphics[width=\thirdwidth]{2017_emith_avg_rel_hvran_no_damper.png} &
+    \includegraphics[width=\thirdwidth]{2017_emitv_avg_rel_hvran_no_damper.png}\\
+  \end{tabular}
+  \caption{Measured effects of the random excitation (in V, top row;
+    and H+V, bottom row) during the 2017 experiment: relative
+    intensity losses (left), relative horizontal emittance (center),
+    and relative vertical emittance (right).}
+  \label{fig:ranexp}
+\end{figure*}
+
+\begin{figure*}
+  \begin{tabular}{cc}
+    Reference bunch, no excitation & Excited bunch, random excitation, V \\
+    \includegraphics[width=\bsrtwidth]{profile_v_ranv_slot_1698.png} &
+    \includegraphics[width=\bsrtwidth]{profile_v_ranv_slot_1532.png} \\
+  \end{tabular}
+  \caption{Vertical beam profiles measured with the Beam Synchrotron
+    Radiation Telescope (BSRT) during the 2017 experiments. The
+    profiles are taken at the end of the random excitation in V. For
+    these bunches, the transverse damper was not active. The data are
+    presented in the same way as in Fig.~\ref{fig:10thexpprof}. The
+    distribution changes in the bunch experiencing the maximum
+    excitation (right) were larger than those in the reference bunch
+    (left). In both cases, distributions retained a Gaussian shape.}
+  \label{fig:ranexpprof}
+\end{figure*}
+
+\begin{figure*}
+  \begin{tabular}{ccc}
+  \includegraphics[width=\thirdwidth]{2017_scale_amp_ranh_lblshort.png} &
+  \includegraphics[width=\thirdwidth]{2017_scale_amp_ranv_lblshort.png} &
+  \includegraphics[width=\thirdwidth]{2017_scale_amp_ranhv_lblshort.png} \\
+  \end{tabular}
+  \caption{Measured loss rates as a function of amplitude for random
+    excitations in H only (left), V only (center), and H+V
+    (right). For each of the three excitation modes, three consecutive
+    data sets were taken (black, red, and yellow), with increasing
+    maximum amplitude. Data were taken simultaneously with no
+    transverse damper on some bunches (filled circles and solid lines)
+    and with the damper active on other bunches (empty circles and
+    dashed lines). The lines represent empirical second-order
+    polynomial fits.}
+  \label{fig:ranexplossplane}
 \end{figure*}
@@
   \centering
-	\includegraphics[width=1.0\linewidth]{2017_scale_amp_7th_ranhv_lblshort.png}
-	\caption{\label{fig:ranexploss}  Comparison of scaling of loss rates with excitation amplitude for a random excitation and \seventhtp\ in H+V measured during the 2017 experiment.
-The excitation amplitude error is 15\% indicated by the horizontal error bars.}
+  \includegraphics[width=\columnwidth]{2017_scale_amp_7th_ranhv_lblshort.png}
+  \caption{Comparison of measured loss rates vs.\ excitation amplitude
+    between random and \seventhtp\ in H+V, obtained during the 2017
+    experiment.}
+  \label{fig:ranexploss}
 \end{figure}

### 2018-03-19 resex.tex
Revised text of transverse damper observations.

@@
 
-For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally,
-the transverse damper does in general not change any of the
-observables studied during the experiments, explicitly losses,
-emittance and beam distribution suggesting that indeed it is not
-damping any resonant excitation. On the other hand, the transverse
-damper considerably decreases any changes of the above parameters in
-case of a random excitation. As an example, the vertical emittance
-growth for the 7th- and \tenthtp\ in V is compared with a random excitation in V in Fig.~\ref{fig:damp}.
-The same observation is also made in case of losses as shown in Fig.~\ref{fig:ranexploss} comparing the \seventhtp\ in H+V with a random excitation in H+V.
-For the \seventhtp\ the two curves for the transverse damper active and not active almost coincide while for the random excitation a significant difference is visible.
-Until present, it is unknown why the transverse damper (ADT) appears to be incapable of damping any resonant excitation as it fulfills all requirements to detect as well as damp the caused oscillation:
+During \kthtp\ experiments, the action of the transverse feedback and
+damping system (ADT) did not significantly change any of the
+observables, namely losses, emittances, and beam distributions. On the
+other hand, the ADT considerably reduced any changes of the above
+parameters in case of random excitations.
+
+An example is shown in Fig.~\ref{fig:ranexploss}, where losses from
+\seventhtp\ in H+V are compared with those measured during random
+excitations in H+V. For \seventhtp, the results with and without
+damper almost coincide, whereas a significant reduction was observed
+for random excitations. As another example, vertical emittances for
+7th- and \tenthtp\ in V are compared to those obtained during vertical
+random excitations in Fig.~\ref{fig:damp}.
+
+The reason why the transverse damper appeared to be capable of damping
+random but not resonant excitations is unknown. In principle, it
+fulfills the requirements needed to detect and damp the resulting
+oscillations:
+%
 \begin{itemize}
-	\item The closed orbit distortion caused by the resonant excitation during the experiments is large enough to be detectable with the ADT pickups.
-	\item The ADT compares the position of each bunch with its position in the previous turn in order to detect an orbit distortion to be damped.
-Therefore an oscillation due to any $k^{\mathrm{th}}$~turn pulsing with $k>1$ would be noticed and damped.
-	\item The ADT is capable of damping the oscillation of each bunch individually.
-The ADT should therefore be capable of applying the correct corrective kick for each group of bunches seeing the same excitation amplitude.
+\item the closed-orbit distortion caused by the resonant excitations
+  was large enough to be detectable by the ADT pickups;
+\item in order to detect the orbit distortion, the system compares the
+  position of each bunch with its position in the previous turn;
+  therefore, an oscillation due to \kthtp, with $k > 1$, should be
+  detected and damped;
+\item the transverse damper acts on each bunch individually, so it
+  should be capable of applying the appropriate corrective kick to
+  each group of bunches subject to the same excitation amplitude.
 \end{itemize}
-Further studies are therefore needed in order to obtain a deeper understanding of the interaction of the transverse damper with a resonant excitation.
+
+We report these observations in an attempt to advance the
+understanding of the interaction between resonant excitations and
+damping systems. Further studies are obviously needed.
 

### 2018-03-19 resex.tex
Revised section on transverse damper (including figures).

@@
 \begin{figure*}
-	\centering
-	damper not active\\
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-	random excitation V, 2017\\
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_no_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		7th~turn V, 2017 exp.\\
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		$10^{\mathrm{th}}$~turn V, 2016 exp.\\
-		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_no_damper_no_text.png}
-	\end{minipage}	
-	damper active\\
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		random excitation V, 2017 exp.\\
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_vran_with_damper.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		7th~turn V, 2017 exp.\\
-		\includegraphics[height=0.75\linewidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.32\linewidth}
-		\centering
-		$10^{\mathrm{th}}$~turn V, 2016 exp.\\
-		\includegraphics[height=0.75\linewidth]{2016_emitv_avg_rel_v10th_with_damper_no_text.png}
-	\end{minipage}	
-	\caption{\label{fig:damp} Relative intensity for bunches with transverse damper not active (top) and transverse damper active (bottom) for a random excitation only in V (left) and for 7th- (center) and \tenthtp\ only in V (right).}
+  \begin{tabular}{ccc}
+    \hline
+    \multicolumn{3}{c}{Transverse damper off} \\ \hline
+    Random excitation, V (2017) & \seventhtp, V (2017) & \tenthtp, V
+                                                         (2016) \\
+    \includegraphics[width=\thirdwidth]{2017_emitv_avg_rel_vran_no_damper.png}&
+    \includegraphics[width=\thirdwidth]{2017_emitv_avg_rel_v7th_no_damper_no_text.png} &
+    \includegraphics[width=\thirdwidth]{2016_emitv_avg_rel_v10th_no_damper_no_text.png}\\
+    \hline
+    \multicolumn{3}{c}{Transverse damper active} \\ \hline
+    Random excitation, V (2017) & \seventhtp, V (2017) & \tenthtp, V
+                                                         (2016) \\
+    \includegraphics[width=\thirdwidth]{2017_emitv_avg_rel_vran_with_damper.png} &
+    \includegraphics[width=\thirdwidth]{2017_emitv_avg_rel_v7th_with_damper_no_text.png} &
+    \includegraphics[width=\thirdwidth]{2016_emitv_avg_rel_v10th_with_damper_no_text.png} \\
+    \end{tabular}
+    \caption{Comparison of measured vertical emittances for bunches
+      with transverse damper off (top row) and with transverse damper
+      on (bottom row), during random V excitations (left), \seventhtp\
+      in V (center) and \tenthtp\ in V (right). (Because of a change
+      in experimental setup, for the bunches represented by the yellow
+      line in the bottom right plot, the excitation was zero instead
+      of the maximum value.)}
+    \label{fig:damp}
 \end{figure*}

### 2018-03-19 resex.tex
Revised discussion and conclusions

@@
 
-\section{Summary}
+\section{Discussion and conclusions}
 \label{sec:sum}
 
-This paper intends to give an overview of the effects of the hollow electron lens (HEL) on the proton beam core and the experiments conducted in this context at the LHC in 2016~\cite{resexmd2016} and 2017~\cite{resexmd2017}.
-For a perfect HEL without any imperfections no effect on the core is to be expected as the field at the center vanishes by construction.
-Only in the case of imperfections in the guidance and profile of the electron beam, a residual kick on the proton beam core can occur.
-In this case and assuming HL-LHC parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}), the contribution from the electron lens bends with approximately $0.5$~nrad to first order is negligible with respect to the kick expected from the central region of the HEL of approximatley $20$~nrad to first order caused by profile imperfections in the electron beam.
-In DC operation these kicks, even if nonlinear in reality, are negligible compared to the nonlinearities present in the LHC and HL-LHC and are thus of no concern.
-The picture changes if the HEL is operated in pulsed mode, in which case noise is introduced on the halo particles (intended) as well as on the beam core (unintended) in the presence of imperfections of the HEL entailing much more stringent tolerances.
-
-To study the effect on the beam core in case of the much more dangerous pulsed operation, beam experiments were conducted in 2016 and 2017 at injection.
-As no hollow electron lens is currently installed in the LHC, the in reality non-linear kick was emulated to first order by a dipole kick with the corresponding excitation pattern.
-The experimental results in general suggest that for the kick amplitudes as expected from the HEL, considerable effects are to be expected for a random excitation and certain resonant excitations, explicitly 7th- and \tenthtp\ during the conducted experiments.
-In addition, also one resonant excitation pattern, the $8^{\mathrm{th}}$~turn pulsing, was tested featuring a small effect on the beam core.
-Assuming that this pattern has a large effect on the beam halo, it would be an optimal candidate for a feasible HEL excitation pattern.
-
-The resonant excitation in general features a rich nonlinear dynamics leading to a diversity in beam distribution and diffusion changes as can be seen for example from the very diverse effects of the different $k^{\mathrm{th}}$~turn pulsing patterns observed in simulations and experiments.
-The origin of this diversity is due to the fact that only certain frequencies and thus resonances are excited.
-Typical for the resonant excitation is also the fast adjustment of the beam distribution to a new equilibrium state which could be for the first time measured in the experiments and predicted in the simulations presented in this paper.
-The random excitation on the other hand excites all beam frequencies equally, leading to the highest increase in halo removal rates for the HEL \cite{hel_halo_hllhc_fitterer} but also the strongest effects on the core, mainly in terms of strong emittance growth as predicted in simulations, experiments and also a multitude of earlier publications on the subject of random noise.
-From the experiments the tolerances summarized in Table~\ref{tab:tolexp} could be obtained, where for each excitation pattern and plane the smallest kick amplitude is listed for which an increase in emittance or losses was observed.
-\begin{table}
-	\caption{\label{tab:tolexp}%
-		Minimum kick amplitudes $\Delta z'$ for which either an increase in emittance or losses was observed during the LHC experiments.
-The tolerances are given for the case of the transverse damper not active.
-The same tolerances are obtained for the case with the damper active in case of all resonant excitation.
-In case of a random excitation the tolerance can be increased by roughly a factor~2.
-	}
-	\begin{ruledtabular}
-		\begin{tabular}{lccc}
-		pulsing pattern & exp. year & plane & $\Delta z' \ [\mathrm{nrad}]$  \\
-			\colrule
-			random & 2017 & H or V & 2.4 \\
-			 & & H+V & 1.2 \\\hline
-			7th~turn & 2016 & H & 2.4 \\
-			 & 2017 & H & 2.4 \\
-			 & 2017 & V & 1.2 \\
-			 & 2017 & H+V & 1.2 \\\hline
-			$8^{\mathrm{th}}$~turn & 2017 & H & 14.4 \\
-			& 2017 & V & $>96$ \\
-			& 2017 & H+V & 12 \\\hline
-			$10^{\mathrm{th}}$~turn & 2016 & V & 9.6 \\
-		\end{tabular}
-	\end{ruledtabular}
-\end{table}
-Tolerances in case of a random excitation are extremely stringent.
-For a resonant excitation the effect depends strongly on the order $k$ of the $k^{\mathrm{th}}$~turn pulsing.
-For a future HEL, the pattern could therefore be tuned to the machine and beam configuration in order to yield a strong effect on the halo particles and small effect on the beam core.
-Compared to the estimated kick amplitude of the HEL on the beam core of
-\begin{equation}
-\Delta x', \Delta y' = 20 \ \mathrm{nrad}
-\end{equation}
-the tolerances for all experimentally tried pulsing pattern except for \eighthtp\ in V lie below this value.
-As a next step, the experiment should be conducted at collision energy where the HEL is most needed.
-At collision energy the inherent noise level of the LHC is smaller which could lead to more relaxed tolerances than the ones obtained at injection energy.
+We described the results of numerical simulations and experimental
+studies conducted in~2016 and in~2017 on the effects of random and
+resonant excitations on proton losses, emittances, and beam
+distributions in the Large Hadron Collider.
+
+These studies were motivated in part by the need to assess the effects
+of a hollow electron lens (HEL), which is being proposed for active
+halo control in HL-LHC. Hollow electron beam collimation was
+demonstrated in the Tevatron in continuous
+mode~\cite{hel_tevatron_stancari}, i.e.\ with the same electron beam
+current acting on a given bunch every turn. In this case, the unwanted
+residual kicks, which arise from profile asymmetries, injection and
+extraction bends, solenoid field misalignments, etc. have a negligible
+effect on the circulating beam. An electron lens that is pulsed
+resonantly can have much stronger effects, and this fact was used in
+the Tevatron for abort-gap
+cleaning~\cite{hel_tevatron_abortgap_zhang}. The resonant-pulsing
+capability (having different currents every turn for a given set of
+bunches) will be incorporated in the HL-LHC hollow electron lens to
+enhance its range of achievable removal rates. However, due to time
+constraints, only a very limited set of experimental tests with
+resonant kicks could be done in the Tevatron. More studies are
+currently being planned using the electron lenses in the Relativistic
+Heavy Ion Collider (RHIC) at Brookhaven National
+Laboratory~\cite{Fischer:PRL:2015}.
+
+The effects of resonant excitations depend on several factors,
+including the machine lattice and the distribution of the beam in
+phase space and in betatron tunes. For this reason, we decided to
+study resonant transverse kicks on proton beams directly in the
+LHC. We relied on the flexibility of the LHC transverse feedback and
+damping system (ADT) to emulate the nonlinear residual kicks of a
+hollow electron lens with transverse dipole kicks, using various
+excitation patterns. We focused on the effects on the beam core,
+namely losses, emittances and beam distributions, as these determine
+the magnitude of the residual fields that can be tolerated. Obviously,
+the choice of pulsing patterns to be used in operations depends also
+on their effects on the beam halo. These effects are outside the scope
+of the present work (and they are challenging to measure directly).
+
+The dipolar component of the residual kick from the electron lens is
+approximately 0.5~nrad from the injection/extraction bends and about
+20~nrad from profile imperfections in the main overlap region. These
+estimates are based on current HL-LHC parameters and on electron-beam
+measurements on a test stand. These values can be reduced by improving
+cathode emission, solenoid-field uniformity, and transport of the
+intense magnetized electron beams.
+
+The most effective pulsing patterns were identified by numerical
+tracking simulations and by frequency-map analysis. These calculations
+provided insights on the rich nonlinear dynamics of resonant
+excitations. Kicking the beam every 7th turn and every 10th turn were
+chosen as examples of strong excitations, whereas \eighthtp\ had very
+weak consequences.
+
+These patterns were tested in experiments and compared with random
+excitations. Bunch filling schemes were devised so that reference
+bunches and different excitation amplitudes could be measured at the
+same time, including the presence or absence of transverse
+feedback. The influence of collective effects was minimized by
+reducing the bunch charge. Because these were the first LHC
+experiments on resonant excitations, and to reduce turn-around time,
+all studies were conducted at injection energy.
+
+The experiments confirmed the relative strength of the pulsing
+modes. For instance, random and \seventhtp\ generated losses at the
+level of a few percent per hour at 12~nrad, whereas the same level of
+losses was obtained at 96~nrad for \tenthtp; even at 96~nrad, losses
+from \eighthtp\ did not exceed 0.6\%/h. A similar hierarchy was
+observed with respect to emittance growth. In many cases, simulations
+underestimated losses and emittance growth. One important unknown
+factor was the magnitude of machine noise and its sources, which had
+to be extrapolated from collision to injection energy. The comparison
+between \seventhtp\ in 2016 and 2017 emphasized the sensitivity of
+horizontal and vertical excitations to working point and tune
+footprint.
+
+A clear distinction between resonant and random excitations was their
+effect on beam distributions. Random noise caused smooth emittance
+growth and widening Gaussian distributions. Resonant excitations
+generated a fast adjustment of the beam distribution to a new,
+non-Gaussian form, followed by a phase of steady evolution. This
+behavior was predicted in simulations and was clearly observed in
+measured synchrotron-radiation profiles. Another difference between
+the two types of excitation was the repeatability of the strength of
+the random excitation vs.\ the dependence of \kthtp\ on the order in
+which the excitations were implemented --- later applications being
+weaker. This confirmed the hypothesis, emphasized by frequency-map
+analysis, that random kicks excite most of the beam, whereas resonant
+pulsing drives specific subsets of particles in tune or amplitude
+space. Finally, the transverse damping system strongly mitigated
+losses and emittance growth generated by random excitations, but had
+negligible effect on resonant pulsing patterns.
+
+For active halo control in HL-LHC, resonant pulsing of the electron
+lenses, or the addition of a random noise component to continuous
+operation, may significantly extend the achievable range of halo
+removal rates. However, further work is needed to identify, both in
+simulations and experimentally, candidate excitation patterns that are
+effective for halo removal and that preserve the beam core (such as
+\eighthtp\ in these studies). Because of the complex dynamics,
+tolerances on residual kicks vary widely, from a few nanoradians for
+the most powerful (such as random at full strength or \seventhtp) to
+about a hundred nanoradians for the more benign (such as
+\eighthtp). The magnitude of the acceptable kicks depends on the
+excitation pattern, on machine lattice, and on the type of application
+(fast scraping before collisions vs.\ long-term tail suppression
+during the course of a physics fill, for instance). Of course,
+sensitivity to the experimental conditions may make resonant patterns
+less attractive than a small random noise component added to a
+constant kick.
+
+These studies are the first systematic investigation of the effects of
+resonant excitations on the proton beam in the LHC.  For practical
+reasons, the studies were done at injection energy. Because of the
+different lattices, working point, noise sources, etc. results at
+collision energy may be different and should be investigated.
 
 \begin{acknowledgments}
-We wish to acknowledge Giulia Papotti with her help with the preparations for the 2016 experiments and her general support from the operations side, Roderik Bruce and Gianluca Valentino for there very helpful suggestions in preparing the experiment at the LHC and we would like to thank all participants of the experiment for helping acquire the data presented in this paper.
-We wish to thank Sergei Nagaitsev for the very useful discussion and profound question to find out why the transverse damper is not damping any resonant excitation.
-We would like to acknowledge Dmitry Shatilov for his help with the Lifetrac code, and St\'{e}phane Fartoukh, Riccardo De Maria and Rogelio Tom\'as for their help with generating the appropiate optics model.
-We are grateful also for the support from the beam instrumentation team, in particular Enrico Bravin and Georges Trad, with the BSRT and are thankful for the collaboration with St\'{e}phania Papadopoulou and Fanouria Antonio for the analysis of the BSRT profiles.
+  We wish to acknowledge Giulia Papotti with her help with the
+  preparations for the 2016 experiments and her general support from
+  the operations side, Roderik Bruce and Gianluca Valentino for there
+  very helpful suggestions in preparing the experiment at the LHC and
+  we would like to thank all participants of the experiment for
+  helping acquire the data presented in this paper. We wish to thank
+  Sergei Nagaitsev for the very useful discussion and profound
+  question to find out why the transverse damper is not damping any
+  resonant excitation. We would like to acknowledge Dmitry Shatilov
+  for his help with the Lifetrac code, and St\'{e}phane Fartoukh,
+  Riccardo De Maria and Rogelio Tom\'as for their help with generating
+  the appropiate optics model. We are grateful also for the support
+  from the beam instrumentation team, in particular Enrico Bravin and
+  Georges Trad, with the BSRT and are thankful for the collaboration
+  with St\'{e}phania Papadopoulou and Fanouria Antonio for the
+  analysis of the BSRT profiles.
 \end{acknowledgments}

### 2018-03-20 resex.tex
Revised acknowledgements. Updated corresponding author. Defined command for computer code.

@@
 \newcommand{\q}[2]{\ensuremath{#1\ \mathrm{#2}}} % quantity with units
+\newcommand{\code}[1]{\textsc{#1}} % computer code
 
@@
 
-\thanks{Fermilab is operated by Fermi Research Alliance, LLC under
-	Contract No.~DE-AC02-07CH11359 with the United States Department of
-	Energy. This work was partially supported by the US DOE LHC
-	Accelerator Research Program (LARP) and by the European FP7 HiLumi
-	LHC Design Study, Grant Agreement 284404.}
+\thanks{This manuscript has been authored by Fermi Research Alliance,
+  LLC under Contract No.~DE-AC02-07CH11359 with the U.S.\ Department
+  of Energy (DOE), Office of Science, Office of High Energy
+  Physics. This work was partially supported by the U.S.\ DOE LHC
+  Accelerator Research Program (LARP), by the European FP7 HiLumi LHC
+  Design Study, Grant Agreement 284404, and by the High Luminosity LHC
+  (HL-LHC) Project.}
 
 \author{Miriam Fitterer}
- \email{mfittere@fnal.gov}
 \author{Giulio Stancari}
+\email[Corresponding author: ]{stancari@fnal.gov}
 \author{Alexander Valishev}
@@
                               different implementation of the
-                              $a_1, b_1$ coefficients in Lifetrac and
-                              MAD-X. $b_2$ errors are adjusted to
+                              $a_1, b_1$ coefficients in \code{Lifetrac} and
+                              \code{MAD-X}. $b_2$ errors are adjusted to
                               yield an average
@@
 %
-For both simulation types, the tracking code Lifetrac~\cite{lifetrac}
-was used. The simulation parameters are summarized in
-Table~\ref{tab:sim_param}. Further details are given in
-Refs.~\cite{md_sim_hel_res_ex_fitterer, resexmd2017}.
+For both simulation types, the tracking code
+\code{Lifetrac}~\cite{lifetrac} was used. The simulation parameters
+are summarized in Table~\ref{tab:sim_param}. Further details are given
+in Refs.~\cite{md_sim_hel_res_ex_fitterer, resexmd2017}.
 
@@
 \begin{acknowledgments}
-  We wish to acknowledge Giulia Papotti with her help with the
-  preparations for the 2016 experiments and her general support from
-  the operations side, Roderik Bruce and Gianluca Valentino for there
-  very helpful suggestions in preparing the experiment at the LHC and
-  we would like to thank all participants of the experiment for
-  helping acquire the data presented in this paper. We wish to thank
-  Sergei Nagaitsev for the very useful discussion and profound
-  question to find out why the transverse damper is not damping any
-  resonant excitation. We would like to acknowledge Dmitry Shatilov
-  for his help with the Lifetrac code, and St\'{e}phane Fartoukh,
-  Riccardo De Maria and Rogelio Tom\'as for their help with generating
-  the appropiate optics model. We are grateful also for the support
-  from the beam instrumentation team, in particular Enrico Bravin and
-  Georges Trad, with the BSRT and are thankful for the collaboration
-  with St\'{e}phania Papadopoulou and Fanouria Antonio for the
-  analysis of the BSRT profiles.
+  There are several people we would like to acknowledge for their
+  support, insights, participation in studies, and assistance in data
+  collection. In particular, Giulia Pa\-potti offered essential
+  support for machine operations and helped prepare the 2016
+  experiments. Roderik Bruce and Gianluca Va\-len\-ti\-no made
+  important suggestions on the design of the experiments. Dmitry
+  Sha\-ti\-lov shared his deep understanding of numerical tracking
+  simulations with the \code{Lifetrac} code. St\'{e}\-phane
+  Far\-toukh, Ric\-car\-do De Maria and Ro\-ge\-lio Tom\'{a}s offered
+  invaluable assistance in generating the appropiate accelerator
+  lattice models. We wish to thank Sergei Na\-gai\-tsev for the
+  insightful discussions on the interaction between the transverse
+  damping system and resonant excitations. We are grateful to
+  En\-ri\-co Bra\-vin, Georges Trad, and the whole beam
+  instrumentation team for their crucial support on
+  synchrotron-radiation diagnostics. We thank Ste\-fania
+  Pa\-pa\-do\-pou\-lou and Fa\-nou\-ria An\-to\-niou for the fruitful
+  collaboration on the analysis of the synchrotron-radiation profiles.
+
+  This work was made possible by the High Luminosity LHC (HL-LHC)
+  Project, the HiLumi LHC Design Study, and by the United States
+  Department of Energy LHC Accelerator Research Program (LARP).
 \end{acknowledgments}

### 2018-03-20 resex.tex
Removed appendix on 10th-turn FMA. Minor changes to figures.

@@
 \documentclass[%
+%preprint,
 reprint,
@@
 %frontmatterverbose, 
-%preprint,
 %showpacs,preprintnumbers,
@@
 %prstper,
-%floatfix,
+floatfix,
+%endfloats,
 longbibliography
@@
 \newlength{\halfwidth}
-\setlength{\halfwidth}{0.48\textwidth}
+\setlength{\halfwidth}{0.47\textwidth}
 \newlength{\thirdwidth}
@@
 \newlength{\bsrtwidth}
-\setlength{\bsrtwidth}{0.4\textwidth}
+\setlength{\bsrtwidth}{0.38\textwidth}
 \newlength{\fmawidth}
-\setlength{\fmawidth}{0.4\textwidth}
+\setlength{\fmawidth}{0.38\textwidth}
+\newlength{\smallfmawidth}
+\setlength{\smallfmawidth}{0.24\textwidth}
 
@@
 \begin{figure*}
-  \begin{minipage}[c]{\columnwidth}
-    \includegraphics[width=\columnwidth]{passive_halo_control.png}
-    \includegraphics[width=\columnwidth]{active_halo_control.png}
+  \hspace*{\fill}
+  \begin{minipage}[c]{0.45\textwidth}
+    \includegraphics[width=\linewidth]{passive_halo_control.png}
+    \includegraphics[width=\linewidth]{active_halo_control.png}
+  \end{minipage} \hfill
+  \begin{minipage}[c]{0.4\textwidth}
+    \includegraphics[width=\linewidth]{diffusion_illustration}
   \end{minipage}
-  \hfill
-  \begin{minipage}[c]{\columnwidth}
-  \includegraphics[width=\columnwidth]{diffusion_illustration}
-  \end{minipage}
+  \hspace*{\fill}
   \caption{Left: Sketch of passive halo control with a conventional
@@
 \begin{figure*}
-  \includegraphics[width=0.75\textwidth]{hel_layout_epbeam}
+  \includegraphics[width=0.8\textwidth]{hel_layout_epbeam}
   \caption{Layout of the hollow electron lens for HL-LHC. (Courtesy of
@@
 \begin{figure*}
-  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} \hfill
-  \includegraphics[height=0.9\columnwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}
+  \begin{tabular}{cc}
+    \includegraphics[width=0.48\textwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_2D} &
+    \includegraphics[width=0.4\textwidth]{CHG1b_170523_8p75A_2-4-2kG_500V_75mA_hires_j-vs-r_binned_nogrid_nolabel}\\
+  \end{tabular}
   \caption{Example of current-density distribution measurements for
@@
 \begin{figure*}
-  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} \hfill
-  \includegraphics[width=\columnwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}
+  \begin{tabular}{cc}
+    \includegraphics[width=0.4\textwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Emap} &
+    \includegraphics[width=0.4\textwidth]{CHG1b_170512_8p75A_2-4-2p7kG_500V_76mA_hires_Eslice}\\
+  \end{tabular}
   \caption{Calculated relative electric field for the hollow electron
@@
     \multicolumn{1}{l}{\emph{2016 experiment}} \\
-    \includegraphics[width=0.95\textwidth]{bunchfilling_2016.png} \\
+    \includegraphics[width=0.9\textwidth]{bunchfilling_2016.png} \\
     \\
     \multicolumn{1}{l}{\emph{2017 experiment}} \\
-    \includegraphics[width=0.95\textwidth]{bunchfilling_2017.png}
+    \includegraphics[width=0.9\textwidth]{bunchfilling_2017.png}
   \end{tabular}
@@
 \begin{figure*}
-  \includegraphics[width=\textwidth]{bunchfilling_measured.png}	
+  \includegraphics[width=0.9\textwidth]{bunchfilling_measured.png}	
   \caption{Example of bunch-by-bunch beam centroid motion for Beam~1
@@
     pulsing in both horizontal and vertical planes (top left);
-    \seventhtp\ in the horizontal (bottom left) and
-    vertical (bottom right) planes. The excitation amplitude is
-    120~nrad in the corresponding plane. The colors (blue to red)
-    represent the tune jitter of tracked particles starting at each
-    given location in tune space~\cite{fmalaskar}. The absence of a
-    strong excitation of any resonance in case of vertical
-    \seventhtp\ and the strong excitation in case
-    of horizontal pulsing confirms the excitation of the $7 Q_x$
-    resonance. For $10^{\mathrm{th}}$~turn pulsing, there is in
-    contrast no significant difference between H, V, or H+V (see
-    Appendix~\ref{app:sec:10}, Fig.~\ref{app:fig:fma:10}).}
+    \seventhtp\ in the horizontal (bottom left) and vertical (bottom
+    right) planes. The excitation amplitude is 120~nrad in the
+    corresponding plane. The colors (blue to red) represent the tune
+    jitter of tracked particles starting at each given location in
+    tune space~\cite{fmalaskar}. The absence of a strong excitation of
+    any resonance in case of vertical \seventhtp\ and the strong
+    excitation in case of horizontal pulsing confirms the excitation
+    of the $7 Q_x$ resonance. For $10^{\mathrm{th}}$~turn pulsing,
+    there is in contrast no significant difference between H, V, or
+    H+V (Fig.~\ref{fig:fma:10}).}
   \label{fig:patternfma}
@@
 
-\begin{figure*}
+\begin{figure}
   \begin{tabular}{cc}
-    no excitation & random, H+V \\
-    \includegraphics[width=\halfwidth]{2016injerra2b2u_3_5um_hist_y.png} &
-    \includegraphics[width=\halfwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
-    7th, H+V & $10^{\mathrm{th}}$, H+V \\
-    \includegraphics[width=\halfwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
+    No excitation & 10th, H+V \\
+    \includegraphics[width=\smallfmawidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png} &
+    \includegraphics[width=\smallfmawidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png} \\
+    10th, H & 10th, V \\
+    \includegraphics[width=\smallfmawidth]{2016injnocolc15o+19_6noerrut10skh_dp0_ord10.png} &
+    \includegraphics[width=\smallfmawidth]{2016injnocolc15o+19_6noerrut10skv_dp0_ord10.png} \\
+  \end{tabular}
+  \caption{FMA for \tenthtp\ based on the 2016 injection
+    optics with no lattice errors and tunes (64.28, 59.31). The
+    excitation is 120~nrad in the corresponding plane. There is no
+    significant difference between pulsing in H only, V only, or in H+V.}
+  \label{fig:fma:10}
+\end{figure}
+
+\begin{figure*}
+  \begin{tabular}{cc}
+    No excitation & Random, H+V \\
+    \includegraphics[width=\bsrtwidth]{2016injerra2b2u_3_5um_hist_y.png} &
+    \includegraphics[width=\bsrtwidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png} \\
+    7th, H+V & 10th, H+V \\
+    \includegraphics[width=\bsrtwidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
                   &
-    \includegraphics[width=\halfwidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png} \\
+    \includegraphics[width=\bsrtwidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png} \\
   \end{tabular}
@@
 
-\appendix
-\section{Additional plots for 10th turn pulsing during the 2016 experiments}
-\label{app:sec:10}
-\begin{figure}[h]
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		no excitation
-		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png}
-	\end{minipage}
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$10^{\mathrm{th}}$ H+V
-		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skhv_dp0_ord10.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$10^{\mathrm{th}}$ H
-		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skh_dp0_ord10.png}
-	\end{minipage}	
-	\begin{minipage}[t]{0.49\linewidth}
-		\centering
-		$10^{\mathrm{th}}$ V
-		\includegraphics[width=1.0\linewidth]{2016injnocolc15o+19_6noerrut10skv_dp0_ord10.png}
-	\end{minipage}
-	\caption{\label{app:fig:fma:10} FMA analysis for \tenthtp\ based on the 2016 injection optics with no machine errors and a tune of (64.28,59.31).
-The excitation is 120~nrad in the corresponding plane.
-There is no significant difference in the FMA analysis between pulsing only in H, only in V or in H+V.}
-\end{figure}
-
+% bibliography
 \bibliography{resex-bibdesk}

### 2018-03-21 resex.tex
Intermediate step in revision of bibliography. Minor changes in front matter.
Another step in revision of bibliography. Re-introduced references on noise.

@@
-\documentclass[%
-%preprint,
-reprint,
-%superscriptaddress,
-%groupedaddress,
-%unsortedaddress,
-%runinaddress,
-%frontmatterverbose, 
-%showpacs,preprintnumbers,
-%nofootinbib,
-%nobibnotes,
-%bibnotes,
- amsmath,amssymb,
- aps,
-%pra,
-%prb,
-%rmp,
-prstab,
-%prstper,
-floatfix,
-%endfloats,
-longbibliography
+\documentclass[
+prstab,reprint
+%,draft
+%,preprint,tightenlines
+,linenumbers
+,longbibliography
+,preprintnumbers
+,showpacs
+,showkeys
+,byrevtex
+,amsfonts,amssymb,amsmath
+,floatfix
+%,endfloats
 ]{revtex4-1}
 
-\usepackage{graphicx}% Include figure files
-\usepackage{dcolumn}% Align table columns on decimal point
-\usepackage{bm}% bold math
-\usepackage{hyperref}% add hypertext capabilities
-%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
-%\linenumbers\relax % Commence numbering lines
-\usepackage{color}
-%\usepackage{siunitx}
-\usepackage{url}
-
-%\usepackage[showframe,%Uncomment any one of the following lines to test 
-%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
-%%text={7in,10in},centering,
-%%margin=1.5in,
-%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
-%%height=10in,a5paper,hmargin={3cm,0.8in},
-%]{geometry}
-
-\DeclareFontFamily{U}{wncy}{}
-\DeclareFontShape{U}{wncy}{m}{n}{<->wncyr10}{}
-\DeclareSymbolFont{mcy}{U}{wncy}{m}{n}
-\DeclareMathSymbol{\Sh}{\mathord}{mcy}{"58}
+\usepackage{bm}
+\usepackage{dcolumn}
+\usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 \newcommand{\fma}{frequency-map analysis}
+
+% symbol for Dirac comb
+\DeclareFontFamily{U}{wncy}{}
+\DeclareFontShape{U}{wncy}{m}{n}{<->wncyr10}{}
+\DeclareSymbolFont{mcy}{U}{wncy}{m}{n}
+\DeclareMathSymbol{\Sh}{\mathord}{mcy}{"58}
 
@@
 \author{Giulio Stancari}
-\email[Corresponding author: ]{stancari@fnal.gov}
+\email[Corresponding author. E-mail: ]{stancari@fnal.gov}
 \author{Alexander Valishev}
@@
 pattern, but it is also potentially harmful for the proton beam
-core. A random pattern can be employed at full strength or added as a
+core. The effects of random noise in various machines have been
+studied extensively in the past (see, for instance,
+Refs.~\cite{noise_lebedev_ssc, noise_alexahin_lhc, noise_2007_ohmi,
+  noise_2014_ohmi}).
+
+A random pattern can be employed at full strength or added as a
 modulation to a continuous mode of operation (depending on the

### 2018-03-21 resex.tex
One more step in revision of bibliography. Replaced Chamonix reference (not accessible) on 2016 LHC performance with Wenninger at IPAC17. Added Warp reference. Renamed BibTeX file.

@@
     \begin{tabular}{lccccc}
-      Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{chamonix2017param}	& LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{fcc_param_2017} \\
+      Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{Wenninger:IPAC:2017}	& LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{fcc_param_2017} \\
       \colrule
@@
     calculations are based on measurements at the Fermilab
-    electron-lens test stand combined with Warp calculations of the
-    electric potentials and fields in a cylindrical beam pipe.}
+    electron-lens test stand combined with \code{Warp} calculations of
+    the electric potentials and fields in a cylindrical beam pipe.}
   \label{core:fig:1}
@@
 resulting electric field is calculated, with the vacuum pipe as
-boundary, using the solvers of the Warp particle-in-cell
-code~\cite{warp}. (The contribution of the magnetic field generated by
-the electron current was neglected in this study.) The field,
-integrated over the trajectory of the protons, is then translated into
-a symplectic kick map.
+boundary, using the solvers of the \code{Warp} particle-in-cell
+code~\cite{warp, Friedman:IEEE-plasma:2014}. (The contribution of the
+magnetic field generated by the electron current was neglected in this
+study.) The field, integrated over the trajectory of the protons, is
+then translated into a symplectic kick map.
 
@@
 $b = \q{30}{mm}$ is used. The potential and fields are then calculated
-with Warp~\cite{warp}. The resulting relative electric field strengths
-are shown in Fig.~\ref{core:fig:1}. The electric field is obviously
-proportional to the charge density of electrons. Its relative strength
-is rather insensitive to the hollow beam radii, as long as these
-remain small compared to the beam pipe radius. Further details on the
-measurements and Warp simulations can be found in
+with \code{Warp}~\cite{warp, Friedman:IEEE-plasma:2014}. The resulting
+relative electric field strengths are shown in
+Fig.~\ref{core:fig:1}. The electric field is obviously proportional to
+the charge density of electrons. Its relative strength is rather
+insensitive to the hollow beam radii, as long as these remain small
+compared to the beam pipe radius. Further details on the measurements
+and \code{Warp} simulations can be found in
 Ref.~\cite{hel_res_field_stancari_2017}.
@@
 Refs.~\cite{noise_lebedev_ssc, noise_alexahin_lhc, noise_2007_ohmi,
-  noise_2014_ohmi}).
+  noise_2014_ohmi, md400_noise_injection}).
 
@@
 % bibliography
-\bibliography{resex-bibdesk}
+\bibliography{resex}
 

### 2018-03-22 resex.tex
First complete revision of bibliography

@@
 ,showkeys
-,byrevtex
 ,amsfonts,amssymb,amsmath
 ,floatfix
-%,endfloats
 ]{revtex4-1}
@@
     calculations are based on measurements at the Fermilab
-    electron-lens test stand combined with \code{Warp} calculations of
+    electron-lens test stand combined with \code{warp} calculations of
     the electric potentials and fields in a cylindrical beam pipe.}
@@
 resulting electric field is calculated, with the vacuum pipe as
-boundary, using the solvers of the \code{Warp} particle-in-cell
+boundary, using the solvers of the \code{warp} particle-in-cell
 code~\cite{warp, Friedman:IEEE-plasma:2014}. (The contribution of the
@@
 $b = \q{30}{mm}$ is used. The potential and fields are then calculated
-with \code{Warp}~\cite{warp, Friedman:IEEE-plasma:2014}. The resulting
+with \code{warp}~\cite{warp, Friedman:IEEE-plasma:2014}. The resulting
 relative electric field strengths are shown in
@@
 compared to the beam pipe radius. Further details on the measurements
-and \code{Warp} simulations can be found in
+and \code{warp} simulations can be found in
 Ref.~\cite{hel_res_field_stancari_2017}.
@@
                               different implementation of the
-                              $a_1, b_1$ coefficients in \code{Lifetrac} and
-                              \code{MAD-X}. $b_2$ errors are adjusted to
+                              $a_1, b_1$ coefficients in \code{lifetrac} and
+                              \code{mad-x}. $b_2$ errors are adjusted to
                               yield an average
@@
 For both simulation types, the tracking code
-\code{Lifetrac}~\cite{lifetrac} was used. The simulation parameters
+\code{lifetrac}~\cite{lifetrac} was used. The simulation parameters
 are summarized in Table~\ref{tab:sim_param}. Further details are given
@@
 Refs.~\cite{noise_lebedev_ssc, noise_alexahin_lhc, noise_2007_ohmi,
-  noise_2014_ohmi, md400_noise_injection}).
+  noise_2014_ohmi, md_noise_bbLHC}).
 
@@
   Sha\-ti\-lov shared his deep understanding of numerical tracking
-  simulations with the \code{Lifetrac} code. St\'{e}\-phane
+  simulations with the \code{lifetrac} code. St\'{e}\-phane
   Far\-toukh, Ric\-car\-do De Maria and Ro\-ge\-lio Tom\'{a}s offered

### 2018-03-22 resex.tex
Minor changes in layout and fonts. Fixed the rest of the ordinal numbers.

@@
 \documentclass[
-prstab,reprint
+prstab
+,reprint
 %,draft
@@
 \usepackage{graphicx}
-\usepackage{hyperref}
+\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue,
+            filecolor=blue, urlcolor=blue]{hyperref}
+
+\usepackage{mathptmx}
+\linespread{1.1}            
+\usepackage[T1]{fontenc}
 
@@
       chromaticity $(Q'_x, Q'_y)$ & \multicolumn{2}{c}{(+15, +15)}\\ \hline
-      pulsing patterns  & 7th~turn H
-                                  &7th~turn H, V, H+V \\
-                & $10^{\mathrm{th}}$~turn V & $8^{\mathrm{th}}$~turn
-                                              H, V, H+V \\
+      pulsing patterns  & 7th~turn H & 7th~turn H, V, H+V \\
+                & 10th~turn V & 8th~turn H, V, H+V \\
                 & &  random  H, V, H+V\\
@@
     detected by the real-time transverse activity monitor during the
-    2017 experiment, when the beam was excited every
-    $8^\mathrm{th}$~turn.}
+    2017 experiment, when the beam was excited every 8th~turn.}
   \label{fig:fill_meas} 
@@
 the two pulsing patterns with the largest calculated effects on the
-beam (7th- and $10^{\mathrm{th}}$-turn pulsing), one pattern with no
-effect ($8^{\mathrm{th}}$-turn pulsing), and the random excitation. In
-order to quantify the reproducibility of the results under different
-machine configurations, one pulsing pattern (\seventhtp) was tried
-first in~2016 and then repeated in~2017.
+beam (7th- and \tenthtp), one pattern with no effect (\eighthtp), and
+the random excitation. In order to quantify the reproducibility of the
+results under different machine configurations, one pulsing pattern
+(\seventhtp) was tried first in~2016 and then repeated in~2017.
 
@@
 during the 2017 experiments, when the beam was excited every
-$8^{\mathrm{th}}$~turn. The measured activity shows the typical
-amplitude of bunch centroid excursions and reflects the expected
-excitation pattern.
+8th~turn. The measured activity shows the typical amplitude of bunch
+centroid excursions and reflects the expected excitation pattern.
 
@@
 results for each of the specific excitation patterns that could be
-tested experimentally, namely pulsing every $10^{\mathrm{th}}$~turn
-(Section~\ref{sec:simex10}), 7th~turn
-(Section~\ref{sec:simex7}), $8^{\mathrm{th}}$~turn
-(Section~\ref{sec:simex8}), and the random excitation
+tested experimentally, namely pulsing every 10th~turn
+(Section~\ref{sec:simex10}), 7th~turn (Section~\ref{sec:simex7}),
+8th~turn (Section~\ref{sec:simex8}), and the random excitation
 (Section~\ref{sec:simexran}).  In Section~\ref{sec:damp}, we describe
@@
   \begin{tabular}{cc}
-    no excitation & $10^{\mathrm{th}}$, H+V \\
+    No excitation & 10th, H+V \\
     \includegraphics[width=\fmawidth]{2016injnocolc15o+19_6noerru_dp0_ord10_annotate.png} &
@@
     2016 injection optics with no machine errors and tunes (64.28,
-    59.31): without excitation (top left); $10^{\mathrm{th}}$~turn
-    pulsing in both horizontal and vertical planes (top left);
-    \seventhtp\ in the horizontal (bottom left) and vertical (bottom
-    right) planes. The excitation amplitude is 120~nrad in the
-    corresponding plane. The colors (blue to red) represent the tune
-    jitter of tracked particles starting at each given location in
-    tune space~\cite{fmalaskar}. The absence of a strong excitation of
-    any resonance in case of vertical \seventhtp\ and the strong
+    59.31): without excitation (top left); \tenthtp\ in both
+    horizontal and vertical planes (top left); \seventhtp\ in the
+    horizontal (bottom left) and vertical (bottom right) planes. The
+    excitation amplitude is 120~nrad in the corresponding plane. The
+    colors (blue to red) represent the tune jitter of tracked
+    particles starting at each given location in tune
+    space~\cite{fmalaskar}. The absence of a strong excitation of any
+    resonance in case of vertical \seventhtp\ and the strong
     excitation in case of horizontal pulsing confirms the excitation
-    of the $7 Q_x$ resonance. For $10^{\mathrm{th}}$~turn pulsing,
-    there is in contrast no significant difference between H, V, or
-    H+V (Fig.~\ref{fig:fma:10}).}
+    of the $7 Q_x$ resonance. For \tenthtp, there is in contrast no
+    significant difference between H, V, or H+V
+    (Fig.~\ref{fig:fma:10}).}
   \label{fig:patternfma}
@@
 and emittances on the pattern can be seen.  The largest losses are
-predicted for $3^{\mathrm{rd}}$, 7th and
-$10^{\mathrm{th}}$~turn pulsing, and for a uniform random
+predicted for 3rd-, 7th- and \tenthtp, and for a uniform random
 excitation. Because the bunch length decreases with the number of lost
@@
 transverse aperture. Significant emittance growth is visible only for
-7th- and \tenthtp\ and for the
-random excitation. Compared to any resonant excitation, the random
-excitation shows by far the strongest effect.
+7th- and \tenthtp\ and for the random excitation. Compared to any
+resonant excitation, the random excitation shows by far the strongest
+effect.
 
@@
 $7 Q_x$ resonance is excited by the \seventhtp, and the $10 Q_x$ and
-$10 Q_y$ resonances by the $10^{\mathrm{th}}$~turn pulsing. As
-octupoles can only drive even resonances, the sources of the $7 Q_x$
-resonances are the sextupoles, while the octupoles generate the tune
-footprint. The other pulsing patterns do not exhibit any increase in
-losses or emittance growth without magnetic errors. Their effect can
-be thus attributed to an interaction of the excitation with the
-lattice errors, implying also a sensitivity to the chosen random seed
-in simulations and to the specific lattice in experiments.
+$10 Q_y$ resonances by the \tenthtp. As octupoles can only drive even
+resonances, the sources of the $7 Q_x$ resonances are the sextupoles,
+while the octupoles generate the tune footprint. The other pulsing
+patterns do not exhibit any increase in losses or emittance growth
+without magnetic errors. Their effect can be thus attributed to an
+interaction of the excitation with the lattice errors, implying also a
+sensitivity to the chosen random seed in simulations and to the
+specific lattice in experiments.
 
@@
 
-\subsection{Pulsing every $10^{\mathrm{th}}$ turn}
+\subsection{Pulsing every 10th turn}
 \label{sec:simex10}
@@
   \end{tabular}
-  \caption{Summary of the 2016 experiments on $10^{\mathrm{th}}$~turn
-    pulsing in the vertical plane: losses (left), horizontal
-    emittances (middle) and vertical emittances (right), relative to
-    their initial values. The transverse damping system was not active
-    in this case. The 3 excitation periods are labeled in black
-    according to the value of the maximum excitation amplitude
+  \caption{Summary of the 2016 experiments on \tenthtp\ in the
+    vertical plane: losses (left), horizontal emittances (middle) and
+    vertical emittances (right), relative to their initial values. The
+    transverse damping system was not active in this case. The 3
+    excitation periods are labeled in black according to the value of
+    the maximum excitation amplitude
     $A_{\mathrm{max}} = 5 \, \Delta A$: no excitation, 48~nrad or
@@
 Fig.~\ref{fig:7th2017fma}, top right and bottom right plots). This
-suggests an excitation of the $14^{\mathrm{th}}$-order resonance. The
-reason why this phenomenon is much less pronounced for horizontal
-excitations is not understood.
+suggests an excitation of the 14th-order resonance. The reason why
+this phenomenon is much less pronounced for horizontal excitations is
+not understood.
 

### 2018-03-23 resex.tex
Minor changes to bibliography.
Revisions in title and abstract

@@
 \usepackage{graphicx}
-\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue,
-            filecolor=blue, urlcolor=blue]{hyperref}
+\usepackage[dvipsnames]{xcolor}
+\usepackage[colorlinks=true, allcolors=NavyBlue]{hyperref}
 
@@
 
-\title{Experimental and numerical studies of the effects of resonant
-  and random excitations of the proton beam in the LHC, with
-  applications to the design of a pulsed hollow electron lens for
-  active halo control}
+\title{Effects of resonant and random excitations on the proton beam
+  in the Large Hadron Collider, with applications to the design of a
+  pulsed hollow electron lens for active halo control}
 
@@
 \begin{abstract}
-  In the High Luminosity Large Hadron Collider (HL-LHC), a
-  considerable amount of energy will be stored in the beam tails, due
-  to the high beam intensity and to an overpopulation of the tails
-  compared to a Gaussian distribution. To control and clean the beam
-  halo, the installation of two hollow electron lenses (HEL), one per
-  beam, is being considered. In standard electron-lens operation, a
-  proton bunch sees the same electron current at every
-  revolution. Pulsed operation (i.e., different current for different
-  turns) is also considered, because it can widen the range of
-  achievable halo removal rates. For an axially symmetric electron
-  lens, only the halo particles are excited, leaving the core
-  unperturbed. If a residual field is present at the location of the
-  beam core, these particles are exposed to transverse kicks and to
-  noise. In this paper, we present the results of numerical
-  simulations and of the LHC experiments conducted in~2016 and~2017 to
-  study the effects of random noise and of resonant excitations on the
-  beam core. The excitation patterns were generated by the transverse
-  damper system, which is a very flexible source of transverse dipole
-  kicks. The main observables were losses, emittances, and transverse
+  We present the results of numerical simulations and experimental
+  studies about the effects of random and resonant excitations on
+  proton losses, emittances, and beam distributions in the Large
+  Hadron Collider (LHC). In addition to shedding light on complex
+  nonlinear effects, these studies are applied to the design of hollow
+  electron lenses (HEL) for active beam halo control. In the
+  High-Luminosity Large Hadron Collider (HL-LHC), a considerable
+  amount of energy will be stored in the beam tails. To control and
+  clean the beam halo, the installation of two hollow electron lenses,
+  one per beam, is being considered. In standard electron-lens
+  operation, a proton bunch sees the same electron current at every
+  revolution. Pulsed electron beam operation (i.e., different currents
+  for different turns) is also considered, because it can widen the
+  range of achievable halo removal rates. For an axially symmetric
+  electron beam, only protons in the halo are excited. If a residual
+  field is present at the location of the beam core, these particles
+  are exposed to time-dependent transverse kicks and to noise. We
+  discuss the numerical simulations and the experiments conducted
+  in~2016 and~2017 at injection energy in the LHC. The excitation
+  patterns were generated by the transverse feedback and damping
+  system, which acted as a very flexible source of dipole kicks. The
+  main observables were losses, emittances, and transverse
   distributions as a function of excitation patterns and
-  strengths. The resonant excitations showed rich nonlinear dynamics
-  and nontrivial changes of the beam distribution, which, to our
+  strengths. The resonant excitations induced rich dynamical effects
+  and nontrivial changes of the beam distributions, which, to our
   knowledge, have not previously been observed and studied in this
-  detail. We also give estimates, based upon existing measurements of
-  prototype electron guns, of the residual fields that may be expected
-  in the HL-LHC hollow electron lens.
+  detail. We conclude with a discussion of the tolerable and
+  achievable residual fields and proposals for further studies.
 \end{abstract}

### 2018-03-25 resex.tex
Changes to title, abstract, and intro.

@@
 \title{Effects of resonant and random excitations on the proton beam
-  in the Large Hadron Collider, with applications to the design of a
-  pulsed hollow electron lens for active halo control}
+  in the Large Hadron Collider, with applications to the design of
+  pulsed hollow electron lenses for active halo control}
 
@@
   We present the results of numerical simulations and experimental
-  studies about the effects of random and resonant excitations on
+  studies about the effects of resonant and random excitations on
   proton losses, emittances, and beam distributions in the Large
@@
   patterns were generated by the transverse feedback and damping
-  system, which acted as a very flexible source of dipole kicks. The
-  main observables were losses, emittances, and transverse
-  distributions as a function of excitation patterns and
-  strengths. The resonant excitations induced rich dynamical effects
-  and nontrivial changes of the beam distributions, which, to our
-  knowledge, have not previously been observed and studied in this
-  detail. We conclude with a discussion of the tolerable and
-  achievable residual fields and proposals for further studies.
+  system, which acted as a flexible source of dipole kicks. Proton
+  beam losses, emittances, and transverse distributions were recorded
+  as a function of excitation patterns and strengths. The resonant
+  excitations induced rich dynamical effects and nontrivial changes of
+  the beam distributions, which, to our knowledge, have not previously
+  been observed and studied in this detail. We conclude with a
+  discussion of the tolerable and achievable residual fields and
+  proposals for further studies.
 \end{abstract}
@@
 
+In circular accelerators and storage rings, beam quality can be
+affected by the interplay of external excitations with machine
+lattice. This work, through calculations and experiments, focuses on
+how a certain class of resonant excitations influences beam dynamics
+and to which extent these excitations can cause beam losses, emittance
+growth, or changes in the particle beam distributions. Besides their
+general relevance to the topic of complex nonlinear dynamics, these
+studies were motivated by the need to assess the effects of a pulsed
+hollow electron lens for active beam halo control.
+
 Considering past, current and future high energy colliders, each new
 machine has represented a considerable leap in stored beam energy
-(Table~\ref{tab:stored_energy}).
-
-Furthermore, recent measurements at the LHC show that the tails of the
-transverse beam distribution are overpopulated compared to a Gaussian
-distribution. This results in a considerable amount of energy being
-stored in the beam tails. In particular, in the case of the LHC, about
-5\% of the beam population is stored in the tails (i.e., above
-3.5$\sigma$, where $\sigma$ is the Gaussian rms core beam size),
-compared to 0.22\% in an ideal Gaussian distribution, leading to 19~MJ
-of stored energy for nominal LHC parameters and 34~MJ in the case of
+(Table~\ref{tab:stored_energy}). Furthermore, recent measurements at
+the LHC show that the tails of the transverse beam distribution are
+overpopulated compared to a Gaussian distribution. This results in a
+considerable amount of energy being stored in the beam tails. In
+particular, in the case of the LHC, about 5\% of the beam population
+is stored in the tails (i.e., above 3.5$\sigma$, where $\sigma$ is the
+Gaussian rms core beam size), compared to 0.22\% in an ideal Gaussian
+distribution, leading to 19~MJ of stored energy for nominal LHC
+parameters and 34~MJ in the case of
 HL-LHC~\cite{helreview_valentino}. This leads to the conclusion that a
@@
 simulations and measurements are given in Section~\ref{sec:simex},
-with discussion and summary in Section~\ref{sec:sum}. Auxiliary data
-and plots are collected in the Appendices.
+with discussion and summary in Section~\ref{sec:sum}.
 

### 2018-03-25 resex.tex
Minor changes in preparation for submission: PhySH keywords instead of PACS, kth command, preprint number, spelled out rms, no horizontal lines in damper on/off figure, comma after i.e.

@@
 ,preprintnumbers
-,showpacs
+%,showpacs
 ,showkeys
@@
 
+\usepackage[american]{babel}
 \usepackage{bm}
@@
 % abbreviations
+\newcommand{\kth}{$k$th}
 \newcommand{\kthtp}{$k$th-turn pulsing}
@@
 
-% PACS 2008:
-% 29.20.db Storage rings and colliders
-
-\pacs{29.20.D-}% PACS, the Physics and Astrtheonomy
-                             % Classification Scheme.
-%\keywords{Suggested keywords}%Use showkeys class option if keyword
-                              %display desired
+%\pacs{}
+
+\keywords{High-energy accelerators and colliders; Nonlinear beam
+  dynamics; Beam resonances; Beam control; Collimation}
+
+\preprint{FERMILAB-PUB-18-084-AD-APC}
+
 \maketitle
@@
 is stored in the tails (i.e., above 3.5$\sigma$, where $\sigma$ is the
-Gaussian rms core beam size), compared to 0.22\% in an ideal Gaussian
-distribution, leading to 19~MJ of stored energy for nominal LHC
-parameters and 34~MJ in the case of
+Gaussian root-mean-square beam size), compared to 0.22\% in an ideal
+Gaussian distribution, leading to 19~MJ of stored energy for nominal
+LHC parameters and 34~MJ in the case of
 HL-LHC~\cite{helreview_valentino}. This leads to the conclusion that a
@@
 \item \textbf{resonant excitation:} The electron beam is switched on
-  only every $k$-th turn. The excitation can be represented
+  only every \kth\ turn. The excitation can be represented
   by the following expression:
@@
   with $f_n = n \cdot f_{\rm rev} / k$ and where $\Sh_{kT}$ is the
-  Dirac comb. In general, $k$-th~turn pulsing drives $k$-th order
+  Dirac comb. In general, \kthtp\ drives \kth-order
   resonances~\cite{md_sim_hel_res_ex_fitterer}. This type of pulsing
@@
   \begin{tabular}{ccc}
-    \hline
-    \multicolumn{3}{c}{Transverse damper off} \\ \hline
+    \multicolumn{3}{c}{\emph{Transverse damper off}} \\
     Random excitation, V (2017) & \seventhtp, V (2017) & \tenthtp, V
@@
     \includegraphics[width=\thirdwidth]{2016_emitv_avg_rel_v10th_no_damper_no_text.png}\\
-    \hline
-    \multicolumn{3}{c}{Transverse damper active} \\ \hline
+    \multicolumn{3}{c}{\emph{Transverse damper active}} \\
     Random excitation, V (2017) & \seventhtp, V (2017) & \tenthtp, V
@@
 demonstrated in the Tevatron in continuous
-mode~\cite{hel_tevatron_stancari}, i.e.\ with the same electron beam
+mode~\cite{hel_tevatron_stancari}, i.e., with the same electron beam
 current acting on a given bunch every turn. In this case, the unwanted

### 2018-03-25 resex.tex
Added aps class option, removed babel, replaced rms with standard deviation in introduction.

@@
-\documentclass[
-prstab
+\documentclass[aps
+,prstab
 ,reprint
@@
 
-\usepackage[american]{babel}
 \usepackage{bm}
@@
 is stored in the tails (i.e., above 3.5$\sigma$, where $\sigma$ is the
-Gaussian root-mean-square beam size), compared to 0.22\% in an ideal
-Gaussian distribution, leading to 19~MJ of stored energy for nominal
-LHC parameters and 34~MJ in the case of
+standard deviation of the Gaussian beam core), compared to 0.22\% in
+an ideal Gaussian distribution, leading to 19~MJ of stored energy for
+nominal LHC parameters and 34~MJ in the case of
 HL-LHC~\cite{helreview_valentino}. This leads to the conclusion that a

### 2018-04-19 resex.tex
Implemented Miriam's comments on circulated draft

@@
       Beam energy [TeV] & 0.98 & 6.5 & 7.0 & 7.0 & 50.0\\
-      Number of bunches & 36 & 2220 & 2808 & 2748 & ? \\
+      Number of bunches & 36 & 2220 & 2808 & 2748 & $\sim$10500 \\
       Number of particles per bunch & $2.90\times 10^{11}$ & $1.15\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
@@
     59.31): without excitation (top left); \tenthtp\ in both
-    horizontal and vertical planes (top left); \seventhtp\ in the
+    horizontal and vertical planes (top right); \seventhtp\ in the
     horizontal (bottom left) and vertical (bottom right) planes. The
@@
 without magnetic errors. Their effect can be thus attributed to an
-interaction of the excitation with the lattice errors, implying also a
+interaction of the excitation with the machine errors, implying also a
 sensitivity to the chosen random seed in simulations and to the
-specific lattice in experiments.
+specific error distribution in experiments.
 
@@
 Although on a different time scale (minutes instead of seconds), a
-similar re-adjustment of the beam distribution in response to an
-external excitation was directly observed experimentally for the first
-time, to our knowledge, during the 2016 and 2017 LHC experiments
-presented in this paper, as discussed below.
+similar re-adjustment of the beam distribution was directly observed
+for the first time, to our knowledge, during the 2016 and 2017 LHC
+experiments presented in this paper, as discussed below.
 
@@
 amplitude of the excitation. As discussed below, an increase of the
-emittance-growth rates with the applied excitation amplitude is what
-is observed experimentally.
+emittance-growth rates with the applied excitation amplitude was also
+observed experimentally.
 
@@
 of measurements. Overall, we conclude that the effects of this
-resonant excitation could be measured in the LHC within a factor~2 and
-could be predicted with similar accuracy.
+resonant excitation on loss rates could be measured in the LHC within
+a factor~2 and could be predicted with similar accuracy.
 
@@
 measurements. A direct comparison of the two experiments is therefore
-not possible. However, these differences provide an estimate of how
-the effects of resonant excitations are affected by lattice changes.
+not possible. However, these differences show how the effects of
+resonant excitations are affected by small changes in fractional tune.
 
@@
 was seen in the case of H+V excitation. Observed and predicted loss
-rates are of the same magnitude (about 0.2\%/h).
+rates were roughly of the same magnitude (about 0.1--0.4\%/h).
 
@@
 control bunches was much smaller (\q{1.8}{\mu m}) than the emittance
-of the excited bunches (\q{2.6}{\mu m}), and intrabeam scattering
+of the excited bunches (\q{2.6}{\mu m}), and intra-beam scattering
 masked the effects of the excitation~\cite{resexmd2017}. In the
@@
 observed during 7th- and \tenthtp. The qualitative behavior of
-emittances predicted in simulations was therefore confirmed. However,
-the calculations could only roughly reproduce the magnitude of the
-measured effects and were in general lower. For instance, at 6~nrad,
-the calculated emittance growth was 1\%/min, whereas in experiments we
-measured 5\%/min.
+emittances predicted in simulations was therefore confirmed.
+Calculations could also reproduce the magnitude of the measured
+effects. For instance, at 6~nrad, the calculated emittance growth
+rates were 3.6\%/min in the horizontal plane and 2.7\%/min vertically,
+whereas in experiments we measured 5.0\%/min.
 
@@
 The effects of resonant excitations depend on several factors,
-including the machine lattice and the distribution of the beam in
-phase space and in betatron tunes. For this reason, we decided to
-study resonant transverse kicks on proton beams directly in the
-LHC. We relied on the flexibility of the LHC transverse feedback and
-damping system (ADT) to emulate the nonlinear residual kicks of a
-hollow electron lens with transverse dipole kicks, using various
-excitation patterns. We focused on the effects on the beam core,
-namely losses, emittances and beam distributions, as these determine
-the magnitude of the residual fields that can be tolerated. Obviously,
-the choice of pulsing patterns to be used in operations depends also
-on their effects on the beam halo. These effects are outside the scope
-of the present work (and they are challenging to measure directly).
+including machine optics and the distribution of the beam in phase
+space and in betatron tunes. For this reason, we decided to study
+resonant transverse kicks on proton beams directly in the LHC. We
+relied on the flexibility of the LHC transverse feedback and damping
+system (ADT) to emulate the nonlinear residual kicks of a hollow
+electron lens with transverse dipole kicks, using various excitation
+patterns. We focused on the effects on the beam core, namely losses,
+emittances and beam distributions, as these determine the magnitude of
+the residual fields that can be tolerated. Obviously, the choice of
+pulsing patterns to be used in operations depends also on their
+effects on the beam halo. These effects are outside the scope of the
+present work (and they are challenging to measure directly).
 
@@
 
-The most effective pulsing patterns were identified by numerical
+The effects of the pulsing patterns were evaluated by numerical
 tracking simulations and by frequency-map analysis. These calculations
@@
 from \eighthtp\ did not exceed 0.6\%/h. A similar hierarchy was
-observed with respect to emittance growth. In many cases, simulations
-underestimated losses and emittance growth. One important unknown
-factor was the magnitude of machine noise and its sources, which had
-to be extrapolated from collision to injection energy. The comparison
-between \seventhtp\ in 2016 and 2017 emphasized the sensitivity of
-horizontal and vertical excitations to working point and tune
-footprint.
+observed with respect to emittance growth; in this case, \tenthtp\ was
+stronger than \seventhtp. In many cases, simulations underestimated
+losses and emittance growth. One important unknown factor was the
+magnitude of machine noise and its sources, which had to be
+extrapolated from collision to injection energy. Other factors were
+collective effects such as intra-beam scattering and electron cloud,
+which were minimized in experiments but could not be entirely
+avoided. The comparison between \seventhtp\ in 2016 and 2017
+emphasized the sensitivity of horizontal and vertical excitations to
+working point and tune footprint.
 

### 2018-04-20 exclude-unused-figs.sh
Removed unused figures. Removed line numbers.

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+for image_file in $(ls *.png)
+do
+    occ=`grep $image_file tmp-file-list.txt -c`
+#    echo $image_file $occ
+    if [ "$occ" -eq "0" ]
+    then
+	echo "File $image_file is NOT in use."
+	cp $image_file tmp/
+	git rm $image_file
+    fi
+done

### 2018-04-20 resex_with_bbl_for_arXiv.tex
Version for arXiv: combined LaTeX source with bbl file; included PDF output.

--- 2018-04-19 resex.tex
@@
 %,preprint,tightenlines
-,linenumbers
+%,linenumbers
 ,longbibliography
@@
 % bibliography
-\bibliography{resex}
+%\bibliography{resex}
+
+%merlin.mbs apsrev4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
+%Control: key (0)
+%Control: author (0) dotless jnrlst
+%Control: editor formatted (1) identically to author
+%Control: production of article title (0) allowed
+%Control: page (1) range
+%Control: year (0) verbatim
+%Control: production of eprint (0) enabled
+\begin{thebibliography}{44}%
+\makeatletter
+\providecommand \@ifxundefined [1]{%
+ \@ifx{#1\undefined}
+}%
+\providecommand \@ifnum [1]{%
+ \ifnum #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \@ifx [1]{%
+ \ifx #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \natexlab [1]{#1}%
+\providecommand \enquote  [1]{``#1''}%
+\providecommand \bibnamefont  [1]{#1}%
+\providecommand \bibfnamefont [1]{#1}%
+\providecommand \citenamefont [1]{#1}%
+\providecommand \href@noop [0]{\@secondoftwo}%
+\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
+\providecommand \@href[1]{\@@startlink{#1}\@@href}%
+\providecommand \@@href[1]{\endgroup#1\@@endlink}%
+\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
+  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
+\providecommand \@@startlink[1]{}%
+\providecommand \@@endlink[0]{}%
+\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
+\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
+\providecommand \urlprefix  [0]{URL }%
+\providecommand \Eprint [0]{\href }%
+\providecommand \doibase [0]{http://dx.doi.org/}%
+\providecommand \selectlanguage [0]{\@gobble}%
+\providecommand \bibinfo  [0]{\@secondoftwo}%
+\providecommand \bibfield  [0]{\@secondoftwo}%
+\providecommand \translation [1]{[#1]}%
+\providecommand \BibitemOpen [0]{}%
+\providecommand \bibitemStop [0]{}%
+\providecommand \bibitemNoStop [0]{.\EOS\space}%
+\providecommand \EOS [0]{\spacefactor3000\relax}%
+\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
+\let\auto@bib@innerbib\@empty
+%</preamble>
+\bibitem [{\citenamefont {Holmes}\ \emph {et~al.}(2011)\citenamefont {Holmes},
+  \citenamefont {Moore},\ and\ \citenamefont {Shiltsev}}]{tevatron}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Holmes}}, \bibinfo {author} {\bibfnamefont {R.~S.}\ \bibnamefont {Moore}}, \
+  and\ \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Shiltsev}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Overview of the {T}evatron
+  collider complex: Goals, operations and performance},}\ }\href {\doibase
+  10.1088/1748-0221/6/08/T08001} {\bibfield  {journal} {\bibinfo  {journal}
+  {JINST}\ }\textbf {\bibinfo {volume} {6}},\ \bibinfo {pages} {T08001}
+  (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Wenninger}(2017)}]{Wenninger:IPAC:2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Wenninger}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Approaching
+  nominal performance at {LHC}},}\ }in\  \cite{ipac17},\ pp.\ \bibinfo {pages}
+  {13--18}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Br\"{u}ning}\ \emph {et~al.}(2004)\citenamefont
+  {Br\"{u}ning}, \citenamefont {Collier}, \citenamefont {Lebrun}, \citenamefont
+  {Myers}, \citenamefont {Ostojic}, \citenamefont {Poole},\ and\ \citenamefont
+  {Proudlock}}]{lhc_design}%
+  \BibitemOpen
+  \bibinfo {editor} {\bibfnamefont {O.~S.}\ \bibnamefont {Br\"{u}ning}},
+  \bibinfo {editor} {\bibfnamefont {P.}~\bibnamefont {Collier}}, \bibinfo
+  {editor} {\bibfnamefont {P.}~\bibnamefont {Lebrun}}, \bibinfo {editor}
+  {\bibfnamefont {S.}~\bibnamefont {Myers}}, \bibinfo {editor} {\bibfnamefont
+  {R.}~\bibnamefont {Ostojic}}, \bibinfo {editor} {\bibfnamefont
+  {J.}~\bibnamefont {Poole}}, \ and\ \bibinfo {editor} {\bibfnamefont
+  {P.}~\bibnamefont {Proudlock}},\ eds.,\ \href {\doibase
+  10.5170/CERN-2004-003-V-1} {\emph {\bibinfo {title} {{LHC} Design Report}}},\
+  \bibinfo {series} {Yellow Reports: Monographs}, Vol.~\bibinfo {volume} {1}\
+  (\bibinfo  {publisher} {CERN},\ \bibinfo {address} {Geneva, Switzerland},\
+  \bibinfo {year} {2004})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Apollinari}\ \emph {et~al.}(2017)\citenamefont
+  {Apollinari}, \citenamefont {B{\'{e}}jar~Alonso}, \citenamefont
+  {Br{\"{u}}ning}, \citenamefont {Fessia}, \citenamefont {Lamont},
+  \citenamefont {Rossi},\ and\ \citenamefont {Tavian}}]{hlcdr}%
+  \BibitemOpen
+  \bibinfo {editor} {\bibfnamefont {G.}~\bibnamefont {Apollinari}}, \bibinfo
+  {editor} {\bibfnamefont {I.}~\bibnamefont {B{\'{e}}jar~Alonso}}, \bibinfo
+  {editor} {\bibfnamefont {O.}~\bibnamefont {Br{\"{u}}ning}}, \bibinfo {editor}
+  {\bibfnamefont {P.}~\bibnamefont {Fessia}}, \bibinfo {editor} {\bibfnamefont
+  {M.}~\bibnamefont {Lamont}}, \bibinfo {editor} {\bibfnamefont
+  {L.}~\bibnamefont {Rossi}}, \ and\ \bibinfo {editor} {\bibfnamefont
+  {L.}~\bibnamefont {Tavian}},\ eds.,\ \href {\doibase 10.23731/CYRM-2017-004}
+  {\emph {\bibinfo {title} {High-Luminosity {L}arge {H}adron {C}ollider
+  ({HL-LHC}): Technical Design Report V.~0.1}}},\ \bibinfo {series} {Yellow
+  Reports: Monographs}\ No.\ \bibinfo {number} {CERN-2017-007-M}\ (\bibinfo
+  {publisher} {CERN},\ \bibinfo {address} {Geneva, Switzerland},\ \bibinfo
+  {year} {2017})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Benedikt}\ and\ \citenamefont
+  {Zimmermann}(2015)}]{fcc_param_2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Benedikt}}\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
+  {Zimmermann}},\ }\href {https://cds.cern.ch/record/2110736} {\emph {\bibinfo
+  {title} {Future Circular Colliders}}},\ \bibinfo {type} {Tech. Rep.}\
+  \bibinfo {number} {CERN-ACC-2015-0164}\ (\bibinfo  {institution} {CERN},\
+  \bibinfo {address} {Geneva, Switzerland},\ \bibinfo {year}
+  {2015})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Valentino}(2016)}]{helreview_valentino}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Valentino}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {What did we
+  learn about halo population during {MD}s and regular operation?}}\ }in\
+  \cite{helreview},\ \bibinfo {note} {\url{https://indico.cern.ch/event/567839}
+  [Online; accessed 21-Mar-2018]}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Br\"uning}(2016)}]{helreview}%
+  \BibitemOpen
+  \bibinfo {editor} {\bibfnamefont {O.~S.}\ \bibnamefont {Br\"uning}},\ ed.,\
+  \href {https://indico.cern.ch/event/567839} {\emph {\bibinfo {title} {Review
+  of the Needs for a Hollow E-lens for the {HL-LHC}}}}\ (\bibinfo
+  {organization} {CERN},\ \bibinfo {address} {Geneva, Switzerland},\ \bibinfo
+  {year} {2016})\ \bibinfo {note} {\url{https://indico.cern.ch/event/567839}
+  [Online; accessed 21-Mar-2018]}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Br{\"{u}}ning}\ \emph {et~al.}(2011)\citenamefont
+  {Br{\"{u}}ning}, \citenamefont {Dominguez}, \citenamefont {Myers},
+  \citenamefont {Rossi}, \citenamefont {Todesco},\ and\ \citenamefont
+  {Zimmermann}}]{helhcparam2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.}~\bibnamefont
+  {Br{\"{u}}ning}}, \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont
+  {Dominguez}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Myers}},
+  \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Rossi}}, \bibinfo
+  {author} {\bibfnamefont {E.}~\bibnamefont {Todesco}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {F.}~\bibnamefont {Zimmermann}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {{HE-LHC} beam parameters, optics and
+  beam-dynamics issues},}\ }in\ \href {\doibase 10.5170/CERN-2011-003.7} {\emph
+  {\bibinfo {booktitle} {{EuCARD-AccNet-EuroLumi} Workshop: The High-Energy
+  {L}arge {H}adron {C}ollider ({HE-LHC}10)}}},\ \bibinfo {series and number}
+  {\bibinfo {series} {Yellow Reports: Conference Proceedings}\ No.\ \bibinfo
+  {number} {CERN-2011-003}},\ \bibinfo {editor} {edited by\ \bibinfo {editor}
+  {\bibfnamefont {E.}~\bibnamefont {Todesco}}\ and\ \bibinfo {editor}
+  {\bibfnamefont {F.}~\bibnamefont {Zimmermann}}}\ (\bibinfo  {publisher}
+  {CERN},\ \bibinfo {address} {Geneva, Switzerland},\ \bibinfo {year} {2011})\
+  pp.\ \bibinfo {pages} {7--12},\ \Eprint {http://arxiv.org/abs/1108.1617}
+  {arXiv:1108.1617 [physics.acc-ph]} \BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Molson}\ \emph {et~al.}(2017)\citenamefont {Molson},
+  \citenamefont {Bruce}, \citenamefont {Faus-Golfe}, \citenamefont {Fiascaris},
+  \citenamefont {Krainer},\ and\ \citenamefont {Redaelli}}]{fcc_coll_ipac2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Molson}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Bruce}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Faus-Golfe}}, \bibinfo
+  {author} {\bibfnamefont {M.}~\bibnamefont {Fiascaris}}, \bibinfo {author}
+  {\bibfnamefont {A.~M.}\ \bibnamefont {Krainer}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Redaelli}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Status of the {FCC-hh} collimation system},}\ }in\
+  \cite{ipac17},\ pp.\ \bibinfo {pages} {64--67}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bruce}(2016)}]{helreview_bruce}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Bruce}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Alternative
+  methods for halo depletion (damper, tune modulation, and wire), long-range
+  beam-beam and comparison of their performance / reliability to that of a
+  hollow electron lens},}\ }in\  \cite{helreview},\ \bibinfo {note}
+  {\url{https://indico.cern.ch/event/567839} [Online; accessed
+  21-Mar-2018]}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Stancari}\ \emph {et~al.}(2011)\citenamefont
+  {Stancari}, \citenamefont {Valishev}, \citenamefont {Annala}, \citenamefont
+  {Kuznetsov}, \citenamefont {Shiltsev}, \citenamefont {Still},\ and\
+  \citenamefont {Vorobiev}}]{hel_tevatron_stancari}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Stancari}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Valishev}},
+  \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Annala}}, \bibinfo
+  {author} {\bibfnamefont {G.}~\bibnamefont {Kuznetsov}}, \bibinfo {author}
+  {\bibfnamefont {V.}~\bibnamefont {Shiltsev}}, \bibinfo {author}
+  {\bibfnamefont {D.~A.}\ \bibnamefont {Still}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {L.~G.}\ \bibnamefont {Vorobiev}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Collimation with hollow electron beams},}\
+  }\href {\doibase 10.1103/PhysRevLett.107.084802} {\bibfield  {journal}
+  {\bibinfo  {journal} {Phys.\ Rev.\ Lett.}\ }\textbf {\bibinfo {volume}
+  {107}},\ \bibinfo {pages} {084802} (\bibinfo {year} {2011})}\BibitemShut
+  {NoStop}%
+\bibitem [{hlv()}]{hlv13}%
+  \BibitemOpen
+  \href
+  {http://abpdata.web.cern.ch/abpdata/lhc_optics_web/www/hllhc13/round15/index.html}
+  {\enquote {\bibinfo {title} {{LHC} optics web: {HLLHCV1.3} --- collision
+  round 15~cm optics},}\ }\bibinfo {howpublished}
+  {\url{http://abpdata.web.cern.ch/abpdata/lhc_optics_web/www/hllhc13/round15/index.html}
+  [Online; accessed 22-Mar-2018]}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Stancari}\ \emph {et~al.}(2015)\citenamefont
+  {Stancari}, \citenamefont {Previtali}, \citenamefont {Valishev},
+  \citenamefont {Bruce}, \citenamefont {Redaelli}, \citenamefont {Rossi},\ and\
+  \citenamefont {Salvachua~Ferrando}}]{hel_cdr}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Stancari}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Previtali}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Valishev}}, \bibinfo
+  {author} {\bibfnamefont {R.}~\bibnamefont {Bruce}}, \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Redaelli}}, \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Rossi}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {B.}~\bibnamefont {Salvachua~Ferrando}},\ }\href
+  {https://arxiv.org/abs/1405.2033} {\emph {\bibinfo {title} {Conceptual Design
+  of Hollow Electron Lenses for Beam Halo Control in the {L}arge {H}adron
+  {C}ollider}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo {number}
+  {FERMILAB-TM-2572-APC, CERN-ACC-2014-0248}\ (\bibinfo  {institution}
+  {Fermilab and CERN},\ \bibinfo {year} {2015})\ \Eprint
+  {http://arxiv.org/abs/1405.2033} {arXiv:1405.2033 [physics.acc-ph]}
+  \BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Shiltsev}\ \emph {et~al.}(2008)\citenamefont
+  {Shiltsev}, \citenamefont {Bishofberger}, \citenamefont {Kamerdzhiev},
+  \citenamefont {Kozub}, \citenamefont {Kufer}, \citenamefont {Kuznetsov},
+  \citenamefont {Martinez}, \citenamefont {Olson}, \citenamefont {Pfeffer},
+  \citenamefont {Saewert}, \citenamefont {Scarpine}, \citenamefont {Seryi},
+  \citenamefont {Solyak}, \citenamefont {Sytnik}, \citenamefont {Tiunov},
+  \citenamefont {Tkachenko}, \citenamefont {Wildman}, \citenamefont {Wolff},\
+  and\ \citenamefont {Zhang}}]{Shiltsev:PRSTAB:2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
+  {Shiltsev}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
+  {Bishofberger}}, \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont
+  {Kamerdzhiev}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Kozub}},
+  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Kufer}}, \bibinfo
+  {author} {\bibfnamefont {G.}~\bibnamefont {Kuznetsov}}, \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Martinez}}, \bibinfo {author}
+  {\bibfnamefont {M.}~\bibnamefont {Olson}}, \bibinfo {author} {\bibfnamefont
+  {H.}~\bibnamefont {Pfeffer}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Saewert}}, \bibinfo {author} {\bibfnamefont
+  {V.}~\bibnamefont {Scarpine}}, \bibinfo {author} {\bibfnamefont
+  {A.}~\bibnamefont {Seryi}}, \bibinfo {author} {\bibfnamefont
+  {N.}~\bibnamefont {Solyak}}, \bibinfo {author} {\bibfnamefont
+  {V.}~\bibnamefont {Sytnik}}, \bibinfo {author} {\bibfnamefont
+  {M.}~\bibnamefont {Tiunov}}, \bibinfo {author} {\bibfnamefont
+  {L.}~\bibnamefont {Tkachenko}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Wildman}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Wolff}}, \ and\ \bibinfo {author} {\bibfnamefont {X.-L.}\
+  \bibnamefont {Zhang}},\ }\bibfield  {title} {\enquote {\bibinfo {title}
+  {{T}evatron electron lenses: Design and operation},}\ }\href {\doibase
+  10.1103/PhysRevSTAB.11.103501} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys.\ Rev.\ ST Accel.\ Beams}\ }\textbf {\bibinfo {volume} {11}},\ \bibinfo
+  {pages} {103501} (\bibinfo {year} {2008})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Fischer}\ \emph {et~al.}(2015)\citenamefont
+  {Fischer}, \citenamefont {Gu}, \citenamefont {Altinbas}, \citenamefont
+  {Costanzo}, \citenamefont {Hock}, \citenamefont {Liu}, \citenamefont {Luo},
+  \citenamefont {Marusic}, \citenamefont {Michnoff}, \citenamefont {Miller},
+  \citenamefont {Pikin}, \citenamefont {Schoefer}, \citenamefont {Thieberger},\
+  and\ \citenamefont {White}}]{Fischer:PRL:2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
+  {Fischer}}, \bibinfo {author} {\bibfnamefont {X.}~\bibnamefont {Gu}},
+  \bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont {Altinbas}}, \bibinfo
+  {author} {\bibfnamefont {M.}~\bibnamefont {Costanzo}}, \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Hock}}, \bibinfo {author} {\bibfnamefont
+  {C.}~\bibnamefont {Liu}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Luo}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Marusic}},
+  \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Michnoff}}, \bibinfo
+  {author} {\bibfnamefont {T.~A.}\ \bibnamefont {Miller}}, \bibinfo {author}
+  {\bibfnamefont {A.~I.}\ \bibnamefont {Pikin}}, \bibinfo {author}
+  {\bibfnamefont {V.}~\bibnamefont {Schoefer}}, \bibinfo {author}
+  {\bibfnamefont {P.}~\bibnamefont {Thieberger}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {S.~M.}\ \bibnamefont {White}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Operational head-on beam-beam compensation with electron
+  lenses in the {R}elativistic {H}eavy {I}on {C}ollider},}\ }\href {\doibase
+  10.1103/PhysRevLett.115.264801} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys.\ Rev.\ Lett.}\ }\textbf {\bibinfo {volume} {115}},\ \bibinfo {pages}
+  {264801} (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Shiltsev}(2016)}]{Shiltsev:elens-book:2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Vladimir~D.}\
+  \bibnamefont {Shiltsev}},\ }\href {\doibase 10.1007/978-1-4939-3317-4} {\emph
+  {\bibinfo {title} {Electron Lenses for Super-Colliders}}}\ (\bibinfo
+  {publisher} {Springer},\ \bibinfo {year} {2016})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Fitterer}\ \emph
+  {et~al.}(2016{\natexlab{a}})\citenamefont {Fitterer}, \citenamefont
+  {Stancari}, \citenamefont {Valishev},\ and\ \citenamefont
+  {Redaelli}}]{hel_halo_hllhc_fitterer}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Fitterer}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Stancari}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Valishev}}, \ and\
+  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Redaelli}},\ }\href
+  {\doibase 10.2172/1408322} {\emph {\bibinfo {title} {Simulation Study of
+  Hollow Electron Beam Collimation in {HL-LHC}}}},\ \bibinfo {type} {Tech.
+  Rep.}\ \bibinfo {number} {FERMILAB-TM-2636-AD}\ (\bibinfo  {institution}
+  {Fermilab},\ \bibinfo {year} {2016})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Fitterer}\ \emph {et~al.}(2017)\citenamefont
+  {Fitterer}, \citenamefont {Stancari}, \citenamefont {Valishev}, \citenamefont
+  {De~Maria}, \citenamefont {Redaelli}, \citenamefont {Sjobak},\ and\
+  \citenamefont {Wagner}}]{hl_halo_ipac2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Fitterer}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Stancari}},
+  \bibinfo {author} {\bibfnamefont {a.}~\bibnamefont {Valishev}}, \bibinfo
+  {author} {\bibfnamefont {R.}~\bibnamefont {De~Maria}}, \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Redaelli}}, \bibinfo {author}
+  {\bibfnamefont {K.}~\bibnamefont {Sjobak}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.~F.}\ \bibnamefont {Wagner}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Implementation of hollow electron lenses in
+  \code{sixtrack} and first simulation results for the {HL-LHC}},}\ }in\
+  \cite{ipac17},\ pp.\ \bibinfo {pages} {3795--3798}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Fitterer}\ \emph
+  {et~al.}(2016{\natexlab{b}})\citenamefont {Fitterer}, \citenamefont
+  {Stancari},\ and\ \citenamefont {Valishev}}]{md_sim_hel_res_ex_fitterer}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Fitterer}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Stancari}},
+  \ and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Valishev}},\
+  }\href {http://inspirehep.net/record/1496416} {\emph {\bibinfo {title}
+  {Effect of Pulsed Hollow Electron-Lens Operation on the Proton Beam Core in
+  {LHC}}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo {number}
+  {FERMILAB-TM-2635-AD}\ (\bibinfo  {institution} {Fermilab},\ \bibinfo {year}
+  {2016})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zhang}\ \emph {et~al.}(2008)\citenamefont {Zhang},
+  \citenamefont {Bishofberger}, \citenamefont {Kamerdzhiev}, \citenamefont
+  {Lebedev}, \citenamefont {Shiltsev}, \citenamefont {Thurman-Keup},\ and\
+  \citenamefont {Tollestrup}}]{hel_tevatron_abortgap_zhang}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {X.-L.}\ \bibnamefont
+  {Zhang}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Bishofberger}},
+  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Kamerdzhiev}}, \bibinfo
+  {author} {\bibfnamefont {V.}~\bibnamefont {Lebedev}}, \bibinfo {author}
+  {\bibfnamefont {V.}~\bibnamefont {Shiltsev}}, \bibinfo {author}
+  {\bibfnamefont {R.}~\bibnamefont {Thurman-Keup}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Tollestrup}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Generation and diagnostics of uncaptured beam in
+  the {F}ermilab {T}evatron and its control by electron lenses},}\ }\href
+  {\doibase 10.1103/PhysRevSTAB.11.051002} {\bibfield  {journal} {\bibinfo
+  {journal} {Phys.\ Rev.\ ST Accel.\ Beams}\ }\textbf {\bibinfo {volume}
+  {11}},\ \bibinfo {pages} {051002} (\bibinfo {year} {2008})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Stancari}(2014)}]{hel_bends_stancari}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Stancari}},\ }\href {https://arxiv.org/abs/1403.6370} {\emph {\bibinfo
+  {title} {Calculation of the Transverse Kicks Generated by the Bends of a
+  Hollow Electron Lens}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo {number}
+  {FERMILAB-FN-0972-APC}\ (\bibinfo  {institution} {Fermilab},\ \bibinfo {year}
+  {2014})\ \Eprint {http://arxiv.org/abs/1403.6370} {arXiv:1403.6370
+  [physics.acc-ph]} \BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Morozov}\ \emph {et~al.}(2012)\citenamefont
+  {Morozov}, \citenamefont {Stancari}, \citenamefont {Valishev},\ and\
+  \citenamefont {Shatilov}}]{hel_model_polynomial_morozov}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {I.~A.}\ \bibnamefont
+  {Morozov}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Stancari}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Valishev}}, \ and\
+  \bibinfo {author} {\bibfnamefont {D.~N.}\ \bibnamefont {Shatilov}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Simulation of hollow
+  electron beam collimation in the {Fermilab} {Tevatron} collider},}\ }in\
+  \href {http://accelconf.web.cern.ch/AccelConf/IPAC2012/papers/moeppb008.pdf}
+  {\emph {\bibinfo {booktitle} {Proceedings of the 3rd International Particle
+  Accelerator Conference ({IPAC12})}}},\ \bibinfo {series and number} {\bibinfo
+  {number} {FERMILAB-CONF-12-126-APC}}\ (\bibinfo  {publisher} {IPAC'12 /
+  IEEE},\ \bibinfo {year} {2012})\ pp.\ \bibinfo {pages} {94--96},\ \bibinfo
+  {note} {\url{http://accelconf.web.cern.ch/AccelConf/IPAC2012} [Online;
+  accessed 21-Mar-2018]}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Stancari}(2018)}]{hel_res_field_stancari_2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Stancari}},\ }\href
+  {https://indico.cern.ch/event/698949/contributions/2868279/attachments/1590019/2515871/Stancari_LHC_e-lens_meeting_2018-01-26_ebeam_sim_K09.pdf}
+  {\enquote {\bibinfo {title} {Electron-beam simulations: Electron-gun emission
+  and residual fields from measured profiles},}\ }\bibinfo {howpublished}
+  {\url{https://indico.cern.ch/event/698949} [Online; accessed 22-Mar-2018]}
+  (\bibinfo {year} {2018}),\ \bibinfo {note} {{CERN} meeting on hollow electron
+  lens beam results}\BibitemShut {NoStop}%
+\bibitem [{war()}]{warp}%
+  \BibitemOpen
+  \href@noop {} {\emph {\bibinfo {title} {The \code{warp} particle-in-cell
+  code}}},\ \bibinfo {organization} {Lawrence Berkeley National Laboratory},\
+  \bibinfo {address} {\url{http://warp.lbl.gov} [Online; accessed
+  21-Mar-2018]}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Friedman}\ \emph {et~al.}(2014)\citenamefont
+  {Friedman}, \citenamefont {Cohen}, \citenamefont {Grote}, \citenamefont
+  {Lund}, \citenamefont {Sharp}, \citenamefont {Vay}, \citenamefont {Haber},\
+  and\ \citenamefont {Kishek}}]{Friedman:IEEE-plasma:2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Friedman}}, \bibinfo {author} {\bibfnamefont {R.~H.}\ \bibnamefont {Cohen}},
+  \bibinfo {author} {\bibfnamefont {D.~P.}\ \bibnamefont {Grote}}, \bibinfo
+  {author} {\bibfnamefont {S.~M.}\ \bibnamefont {Lund}}, \bibinfo {author}
+  {\bibfnamefont {W.~M.}\ \bibnamefont {Sharp}}, \bibinfo {author}
+  {\bibfnamefont {J.~L.}\ \bibnamefont {Vay}}, \bibinfo {author} {\bibfnamefont
+  {I.}~\bibnamefont {Haber}}, \ and\ \bibinfo {author} {\bibfnamefont {R.~A.}\
+  \bibnamefont {Kishek}},\ }\bibfield  {title} {\enquote {\bibinfo {title}
+  {Computational methods in the {W}arp code framework for kinetic simulations
+  of particle beams and plasmas},}\ }\href {\doibase 10.1109/TPS.2014.2308546}
+  {\bibfield  {journal} {\bibinfo  {journal} {IEEE Trans.\ Plasma Sci.}\
+  }\textbf {\bibinfo {volume} {42}},\ \bibinfo {pages} {1321--1334} (\bibinfo
+  {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{hel()}]{hel_test_stand_fnal}%
+  \BibitemOpen
+  \href {https://cdcvs.fnal.gov/redmine/projects/elens/wiki/Test_Stand}
+  {\enquote {\bibinfo {title} {The {F}ermilab electron-lens test stand},}\
+  }\bibinfo {howpublished}
+  {\url{https://cdcvs.fnal.gov/redmine/projects/elens/wiki/Test_Stand} [Online;
+  accessed 22-Mar-2018]}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Jo}\ \emph {et~al.}(2018)\citenamefont {Jo},
+  \citenamefont {Kim}, \citenamefont {Stancari}, \citenamefont {Chung},\ and\
+  \citenamefont {Lee}}]{Jo:PoP:2018}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.~H.}\ \bibnamefont
+  {Jo}}, \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont {Kim}},
+  \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Stancari}}, \bibinfo
+  {author} {\bibfnamefont {M.}~\bibnamefont {Chung}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {H.~J.}\ \bibnamefont {Lee}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Control of the diocotron instability of a hollow electron
+  beam with periodic dipole magnets},}\ }\href {\doibase 10.1063/1.5018425}
+  {\bibfield  {journal} {\bibinfo  {journal} {Physics of Plasmas}\ }\textbf
+  {\bibinfo {volume} {25}},\ \bibinfo {pages} {011607} (\bibinfo {year}
+  {2018})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Davidson}(2001)}]{Davidson:nonneutral-plasmas:2001}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Ronald~C.}\
+  \bibnamefont {Davidson}},\ }\href@noop {} {\emph {\bibinfo {title} {Physics
+  of Nonneutral Plasmas}}}\ (\bibinfo  {publisher} {World Scientific},\
+  \bibinfo {year} {2001})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Fitterer}\ \emph
+  {et~al.}(2016{\natexlab{c}})\citenamefont {Fitterer}, \citenamefont
+  {Stancari}, \citenamefont {Valishev}, \citenamefont {Bruce}, \citenamefont
+  {Papadopoulou}, \citenamefont {Papotti}, \citenamefont {Pellegrini},
+  \citenamefont {Redaelli}, \citenamefont {Trad}, \citenamefont {Valuch},
+  \citenamefont {Valentino}, \citenamefont {Wagner},\ and\ \citenamefont
+  {Xu}}]{resexmd2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Fitterer}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Stancari}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Valishev}}, \bibinfo
+  {author} {\bibfnamefont {R.}~\bibnamefont {Bruce}}, \bibinfo {author}
+  {\bibfnamefont {P.~S.}\ \bibnamefont {Papadopoulou}}, \bibinfo {author}
+  {\bibfnamefont {G.}~\bibnamefont {Papotti}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Pellegrini}}, \bibinfo {author} {\bibfnamefont
+  {S.}~\bibnamefont {Redaelli}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Trad}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
+  {Valuch}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Valentino}},
+  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Wagner}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {C.}~\bibnamefont {Xu}},\ }\href
+  {https://cds.cern.ch/record/2264616} {\emph {\bibinfo {title} {Effect of a
+  Resonant Excitation on the Evolution of the Beam Emittance and Halo
+  Population}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo {number}
+  {CERN-ACC-NOTE-2017-0037}\ (\bibinfo  {institution} {CERN},\ \bibinfo {year}
+  {2016})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Fitterer}\ \emph {et~al.}(2018)\citenamefont
+  {Fitterer}, \citenamefont {Stancari}, \citenamefont {Valishev}, \citenamefont
+  {Bruce}, \citenamefont {Redaelli}, \citenamefont {Salvachua~Ferrando},
+  \citenamefont {Valuch}, \citenamefont {Valentino}, \citenamefont {Wagner},\
+  and\ \citenamefont {Xu}}]{resexmd2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Fitterer}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Stancari}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Valishev}}, \bibinfo
+  {author} {\bibfnamefont {R.}~\bibnamefont {Bruce}}, \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Redaelli}}, \bibinfo {author}
+  {\bibfnamefont {B.~M.}\ \bibnamefont {Salvachua~Ferrando}}, \bibinfo {author}
+  {\bibfnamefont {D.}~\bibnamefont {Valuch}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Valentino}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Wagner}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {C.}~\bibnamefont {Xu}},\ }\href {https://cds.cern.ch/record/2305121} {\emph
+  {\bibinfo {title} {Effect of a Resonant Excitation on the Evolution of the
+  Beam Emittance and Halo Population}}},\ \bibinfo {type} {Tech. Rep.}\
+  \bibinfo {number} {CERN-ACC-NOTE-2018-0007}\ (\bibinfo  {institution}
+  {CERN},\ \bibinfo {year} {2018})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Trad}(2015)}]{Trad:PhD:2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Trad}},\ }\emph {\bibinfo {title} {Development and Optimisation of the {SPS}
+  and {LHC} Beam Diagnostics Based on Synchrotron Radiation Monitors}},\ \href
+  {http://cds.cern.ch/record/2266055} {Ph.D. thesis},\ \bibinfo  {school}
+  {Universit\'e de Grenoble, France} (\bibinfo {year} {2015})\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Fitterer}\ and\ \citenamefont
+  {Stancari}(2017)}]{bsrtprofinj}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Fitterer}}\ and\ \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Stancari}},\ }\href@noop {} {\emph {\bibinfo {title} {Analysis of {BSRT}
+  Profiles in the {LHC} at Injection}}},\ \bibinfo {type} {Tech. Rep.}\
+  \bibinfo {number} {FERMILAB-TM-2657-AD-APC}\ (\bibinfo  {institution}
+  {Fermilab},\ \bibinfo {year} {2017})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Linnecar}\ \emph {et~al.}(2008)\citenamefont
+  {Linnecar}, \citenamefont {Angoletta}, \citenamefont {Arnaudon},
+  \citenamefont {Baudrenghien}, \citenamefont {Bohl}, \citenamefont {Brunner},
+  \citenamefont {Butterworth}, \citenamefont {Ciapala}, \citenamefont
+  {Dubouchet}, \citenamefont {Ferreira~Bento}, \citenamefont {Glenat},
+  \citenamefont {Hagmann}, \citenamefont {H{\"{o}}fle}, \citenamefont {Julie},
+  \citenamefont {Killing}, \citenamefont {Kotzian}, \citenamefont {Landre},
+  \citenamefont {Louwerse}, \citenamefont {Maesen}, \citenamefont
+  {Martinez~Yanez}, \citenamefont {Molendijk}, \citenamefont {Montesinos},
+  \citenamefont {Nicou}, \citenamefont {Noirjean}, \citenamefont {Papotti},
+  \citenamefont {Pashnin}, \citenamefont {Pechaud}, \citenamefont {Pradier},
+  \citenamefont {Rossi}, \citenamefont {Sanchez~Quesada}, \citenamefont
+  {Schokker}, \citenamefont {Shaposhnikova}, \citenamefont {Sorokoletev},
+  \citenamefont {Stellfeld}, \citenamefont {T{\"{u}}ckmantel}, \citenamefont
+  {Valuch}, \citenamefont {Wehrle},\ and\ \citenamefont
+  {Weierud}}]{adt_sum_2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Linnecar}}, \bibinfo {author} {\bibfnamefont {M.~E.}\ \bibnamefont
+  {Angoletta}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Arnaudon}},
+  \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Baudrenghien}}, \bibinfo
+  {author} {\bibfnamefont {T.}~\bibnamefont {Bohl}}, \bibinfo {author}
+  {\bibfnamefont {O.}~\bibnamefont {Brunner}}, \bibinfo {author} {\bibfnamefont
+  {A.}~\bibnamefont {Butterworth}}, \bibinfo {author} {\bibfnamefont
+  {E.}~\bibnamefont {Ciapala}}, \bibinfo {author} {\bibfnamefont
+  {F.}~\bibnamefont {Dubouchet}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Ferreira~Bento}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Glenat}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Hagmann}}, \bibinfo {author} {\bibfnamefont
+  {W.}~\bibnamefont {H{\"{o}}fle}}, \bibinfo {author} {\bibfnamefont
+  {C.}~\bibnamefont {Julie}}, \bibinfo {author} {\bibfnamefont
+  {F.}~\bibnamefont {Killing}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Kotzian}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Landre}}, \bibinfo {author} {\bibfnamefont
+  {R.}~\bibnamefont {Louwerse}}, \bibinfo {author} {\bibfnamefont
+  {P.}~\bibnamefont {Maesen}}, \bibinfo {author} {\bibfnamefont
+  {P.}~\bibnamefont {Martinez~Yanez}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Molendijk}}, \bibinfo {author} {\bibfnamefont
+  {E.}~\bibnamefont {Montesinos}}, \bibinfo {author} {\bibfnamefont
+  {C.}~\bibnamefont {Nicou}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Noirjean}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Papotti}}, \bibinfo {author} {\bibfnamefont
+  {A.}~\bibnamefont {Pashnin}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Pechaud}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Pradier}}, \bibinfo {author} {\bibfnamefont
+  {V.}~\bibnamefont {Rossi}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Sanchez~Quesada}}, \bibinfo {author} {\bibfnamefont
+  {M.}~\bibnamefont {Schokker}}, \bibinfo {author} {\bibfnamefont
+  {E.}~\bibnamefont {Shaposhnikova}}, \bibinfo {author} {\bibfnamefont
+  {R.}~\bibnamefont {Sorokoletev}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Stellfeld}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {T{\"{u}}ckmantel}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Valuch}}, \bibinfo {author} {\bibfnamefont
+  {U.}~\bibnamefont {Wehrle}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {F.}~\bibnamefont {Weierud}},\ }\href {http://cds.cern.ch/record/1176380}
+  {\emph {\bibinfo {title} {Hardware and Initial Beam Commissioning of the
+  {LHC} {RF} Systems}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo {number}
+  {LHC-PROJECT-Report-1172}\ (\bibinfo  {institution} {CERN},\ \bibinfo {year}
+  {2008})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {H{\"{o}}fle}\ \emph {et~al.}(2011)\citenamefont
+  {H{\"{o}}fle}, \citenamefont {Kotzian}, \citenamefont {Schokker},\ and\
+  \citenamefont {Valuch}}]{adt_sum_2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
+  {H{\"{o}}fle}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Kotzian}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Schokker}}, \
+  and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Valuch}},\ }\href
+  {http://cds.cern.ch/record/1378463} {\emph {\bibinfo {title} {{LHC} Damper
+  Beam Commissioning in 2010}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo
+  {number} {CERN-ATS-2011-033}\ (\bibinfo  {institution} {CERN},\ \bibinfo
+  {year} {2011})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Buffat}\ \emph
+  {et~al.}(2017{\natexlab{a}})\citenamefont {Buffat}, \citenamefont
+  {Barranco~Garc{\'{\i}}a}, \citenamefont {Pieloni}, \citenamefont {Tambasco},\
+  and\ \citenamefont {Valuch}}]{adt_noise_emit_2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
+  {Buffat}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Barranco~Garc{\'{\i}}a}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Pieloni}}, \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Tambasco}}, \
+  and\ \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Valuch}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Observations of emittance
+  growth in the presence of external noise in the {LHC}},}\ }in\
+  \cite{ipac17},\ pp.\ \bibinfo {pages} {2117--2120}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Laskar}(2003)}]{fmalaskar}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Laskar}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Frequency map
+  analysis and particle accelerators},}\ }in\ \href
+  {http://accelconf.web.cern.ch/AccelConf/p03/PAPERS/WOAB001.PDF} {\emph
+  {\bibinfo {booktitle} {Proceedings of the 2003 Particle Accelerator
+  Conference ({PAC03})}}},\ \bibinfo {editor} {edited by\ \bibinfo {editor}
+  {\bibfnamefont {J.}~\bibnamefont {Chew}}, \bibinfo {editor} {\bibfnamefont
+  {P.}~\bibnamefont {Lucas}}, \ and\ \bibinfo {editor} {\bibfnamefont
+  {S.}~\bibnamefont {Webber}}}\ (\bibinfo  {publisher} {IEEE},\ \bibinfo {year}
+  {2003})\ p.\ \bibinfo {pages} {378}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Shatilov}\ \emph {et~al.}(2005)\citenamefont
+  {Shatilov}, \citenamefont {Alexahin}, \citenamefont {Lebedev},\ and\
+  \citenamefont {Valishev}}]{lifetrac}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.~N.}\ \bibnamefont
+  {Shatilov}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Alexahin}},
+  \bibinfo {author} {\bibfnamefont {V.}~\bibnamefont {Lebedev}}, \ and\
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Valishev}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {\code{lifetrac} code for the weak-strong
+  simulation of the beam-beam effects in {Tevatron}},}\ }in\ \href
+  {http://accelconf.web.cern.ch/AccelConf/p05/PAPERS/TPAT084.PDF} {\emph
+  {\bibinfo {booktitle} {Proceedings of the 2005 Particle Accelerator
+  Conference ({PAC05})}}},\ \bibinfo {series and number} {\bibinfo {number}
+  {FERMILAB-CONF-05-246-AD}},\ \bibinfo {editor} {edited by\ \bibinfo {editor}
+  {\bibfnamefont {C.}~\bibnamefont {Horak}}}\ (\bibinfo  {publisher} {IEEE},\
+  \bibinfo {year} {2005})\ p.\ \bibinfo {pages} {4138}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Buffat}\ \emph
+  {et~al.}(2017{\natexlab{b}})\citenamefont {Buffat}, \citenamefont
+  {Biancacci}, \citenamefont {Furuseth}, \citenamefont {Jacquet}, \citenamefont
+  {Metral}, \citenamefont {Pellegrini}, \citenamefont {Pojer}, \citenamefont
+  {Trad}, \citenamefont {Valuch}, \citenamefont {Barranco~Garcia},
+  \citenamefont {Pieloni}, \citenamefont {Tambasco},\ and\ \citenamefont
+  {Li}}]{md1433_noise_top_energy}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
+  {Buffat}}, \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Biancacci}},
+  \bibinfo {author} {\bibfnamefont {S.~V.}\ \bibnamefont {Furuseth}}, \bibinfo
+  {author} {\bibfnamefont {D.}~\bibnamefont {Jacquet}}, \bibinfo {author}
+  {\bibfnamefont {E.}~\bibnamefont {Metral}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Pellegrini}}, \bibinfo {author} {\bibfnamefont
+  {M.}~\bibnamefont {Pojer}}, \bibinfo {author} {\bibfnamefont
+  {G.}~\bibnamefont {Trad}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
+  {Valuch}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Barranco~Garcia}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Pieloni}}, \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Tambasco}}, \
+  and\ \bibinfo {author} {\bibfnamefont {Q.}~\bibnamefont {Li}},\ }\href
+  {https://cds.cern.ch/record/2261037} {\emph {\bibinfo {title} {Probing the
+  Behaviour of High Brightness Bunches in Collision at 6.5~{TeV} and the
+  Interplay with an External Source of Noise ({MD} 1433)}}},\ \bibinfo {type}
+  {Tech. Rep.}\ \bibinfo {number} {CERN-ACC-NOTE-2017-0030}\ (\bibinfo
+  {institution} {CERN},\ \bibinfo {year} {2017})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Barranco~Garcia}\ \emph {et~al.}(2016)\citenamefont
+  {Barranco~Garcia}, \citenamefont {Buffat}, \citenamefont {Pieloni},
+  \citenamefont {Tambasco}, \citenamefont {Trad}, \citenamefont {Valuch},
+  \citenamefont {Betz}, \citenamefont {Wendt}, \citenamefont {Pojer},
+  \citenamefont {Solfaroli~Camillocci}, \citenamefont {Salvachua~Ferrando},
+  \citenamefont {Fuchsberger}, \citenamefont {Albert},\ and\ \citenamefont
+  {Qiang}}]{md_noise_bbLHC}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Barranco~Garcia}}, \bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
+  {Buffat}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Pieloni}},
+  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Tambasco}}, \bibinfo
+  {author} {\bibfnamefont {G.}~\bibnamefont {Trad}}, \bibinfo {author}
+  {\bibfnamefont {D.}~\bibnamefont {Valuch}}, \bibinfo {author} {\bibfnamefont
+  {M.}~\bibnamefont {Betz}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Wendt}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Pojer}},
+  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Solfaroli~Camillocci}},
+  \bibinfo {author} {\bibfnamefont {B.~M.}\ \bibnamefont {Salvachua~Ferrando}},
+  \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Fuchsberger}}, \bibinfo
+  {author} {\bibfnamefont {M.}~\bibnamefont {Albert}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Qiang}},\ }\href
+  {https://cds.cern.ch/record/2125228} {\emph {\bibinfo {title} {{MD} 400:
+  {LHC} Emittance Growth in Presence of an External Source of Noise during
+  Collision}}},\ \bibinfo {type} {Tech. Rep.}\ \bibinfo {number}
+  {CERN-ACC-NOTE-2016-0020}\ (\bibinfo  {institution} {CERN},\ \bibinfo {year}
+  {2016})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Lebedev}(1995)}]{noise_lebedev_ssc}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {V.~A.}\ \bibnamefont
+  {Lebedev}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Emittance
+  growth due to noise and its suppression with the feedback system in large
+  hadron colliders},}\ }\href {\doibase 10.1063/1.47298} {\bibfield  {journal}
+  {\bibinfo  {journal} {AIP Conf.\ Proc.}\ }\textbf {\bibinfo {volume} {326}},\
+  \bibinfo {pages} {396--423} (\bibinfo {year} {1995})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Alexahin}(1997)}]{noise_alexahin_lhc}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.~I.}\ \bibnamefont
+  {Alexahin}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {On the
+  emittance growth due to noise in hadron colliders and methods of its
+  suppression},}\ }\href {\doibase 10.1016/S0168-9002(96)01190-4} {\bibfield
+  {journal} {\bibinfo  {journal} {Nucl.\ Instrum.\ Methods Phys.\ Res.\ A}\
+  }\textbf {\bibinfo {volume} {391}},\ \bibinfo {pages} {73--76} (\bibinfo
+  {year} {1997})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Ohmi}\ \emph {et~al.}(2007)\citenamefont {Ohmi},
+  \citenamefont {Calaga}, \citenamefont {Tom{\'{a}}s}, \citenamefont
+  {H{\"{o}}fle},\ and\ \citenamefont {Zimmermann}}]{noise_2007_ohmi}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
+  {Ohmi}}, \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Calaga}},
+  \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont {Tom{\'{a}}s}}, \bibinfo
+  {author} {\bibfnamefont {W.}~\bibnamefont {H{\"{o}}fle}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {F.}~\bibnamefont {Zimmermann}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Beam-beam effects with an external noise in
+  {LHC}},}\ }in\ \href
+  {http://accelconf.web.cern.ch/AccelConf/p07/PAPERS/TUPAN048.PDF} {\emph
+  {\bibinfo {booktitle} {Proceedings of the 22nd {P}article {A}ccelerator
+  {C}onference ({PAC07})}}},\ \bibinfo {editor} {edited by\ \bibinfo {editor}
+  {\bibfnamefont {C.}~\bibnamefont {Petit-Jean-Genaz}}}\ (\bibinfo {year}
+  {2007})\ p.\ \bibinfo {pages} {1496}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Ohmi}(2014)}]{noise_2014_ohmi}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
+  {Ohmi}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Beam-beam effects
+  under the influence of external noise},}\ }in\ \href {\doibase
+  10.5170/CERN-2014-004.69} {\emph {\bibinfo {booktitle} {Proceedings of the
+  {ICFA} Mini-Workshop on Beam-Beam Effects in Hadron Colliders ({BB2013})}}},\
+  \bibinfo {series and number} {\bibinfo {series} {Yellow Reports: Conference
+  Proceedings}\ No.\ \bibinfo {number} {CERN-2014-004}},\ \bibinfo {editor}
+  {edited by\ \bibinfo {editor} {\bibfnamefont {W.}~\bibnamefont {Herr}}\ and\
+  \bibinfo {editor} {\bibfnamefont {G.}~\bibnamefont {Papotti}}}\ (\bibinfo
+  {organization} {CERN},\ \bibinfo {address} {Geneva, Switzerland},\ \bibinfo
+  {year} {2014})\ pp.\ \bibinfo {pages} {69--74},\ \Eprint
+  {http://arxiv.org/abs/1410.4092} {arXiv:1410.4092 [physics.acc-ph]}
+  \BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Arduini}\ \emph {et~al.}(2017)\citenamefont
+  {Arduini}, \citenamefont {Lindroos}, \citenamefont {Pranke}, \citenamefont
+  {Schaa},\ and\ \citenamefont {Seidel}}]{ipac17}%
+  \BibitemOpen
+  \bibinfo {editor} {\bibfnamefont {G.}~\bibnamefont {Arduini}}, \bibinfo
+  {editor} {\bibfnamefont {M.}~\bibnamefont {Lindroos}}, \bibinfo {editor}
+  {\bibfnamefont {J.}~\bibnamefont {Pranke}}, \bibinfo {editor} {\bibfnamefont
+  {V.~R.~W.}\ \bibnamefont {Schaa}}, \ and\ \bibinfo {editor} {\bibfnamefont
+  {M.}~\bibnamefont {Seidel}},\ eds.,\ \href
+  {http://accelconf.web.cern.ch/AccelConf/ipac2017} {\emph {\bibinfo {title}
+  {Proceedings of the 2017 International Particle Accelerator Conference
+  ({IPAC17})}}}\ (\bibinfo  {publisher} {JACoW},\ \bibinfo {address} {Geneva,
+  Switzerland},\ \bibinfo {year} {2017})\BibitemShut {NoStop}%
+\end{thebibliography}%
 

### 2018-06-21 response.tex
Added rough template of response to referees

\documentclass[aps
,prstab
%,reprint
%,draft
,preprint,tightenlines
%,linenumbers
%,longbibliography
%,preprintnumbers
%,showpacs
%,showkeys
,amsfonts,amssymb,amsmath
%,floatfix
]{revtex4-1}

\usepackage{bm}
\usepackage{dcolumn}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true, allcolors=NavyBlue]{hyperref}

\usepackage{mathptmx}
\linespread{1.1}            
\usepackage[T1]{fontenc}

\newcommand{\q}[2]{\ensuremath{#1\ \mathrm{#2}}} % quantity with units
\newcommand{\code}[1]{\textsc{#1}} % computer code
\newcommand{\comm}[1]{\emph{#1}} % editorial or referee comment
\newcommand{\note}[1]{\emph{\textcolor{red}{#1}}} % note, to-do

% abbreviations
\newcommand{\kth}{$k$th}
\newcommand{\kthtp}{$k$th-turn pulsing}
\newcommand{\seventhtp}{7th-turn pulsing}
\newcommand{\eighthtp}{8th-turn pulsing}
\newcommand{\tenthtp}{10th-turn pulsing}

\newcommand{\Fma}{Frequency-map analysis}
\newcommand{\fma}{frequency-map analysis}

\begin{document}

\title{Responses to the comments of editors and referees}

\date{\today}

\preprint{Phys. Rev. Accel. Beams, manuscript ZR10120 Fitterer}

\maketitle

Dear Editors,

thank you for the editorial comments and referee reports.

The paper was thoroughly revised, taking those suggestions into account.

Responses to all recommendations and criticisms are reported below.

\section*{Editorial comments}

\comm{
We also suggest that Reference [5] should be updated to a journal article:

[5] M. Benedikt, F. Zimmermann, ``Towards future circular colliders,''
Journal of the Korean Physical Society (2016)

and perhaps in addition to include the recent IPAC paper

[5b] M. Benedikt and F. Zimmermann,
``FCC: Colliders at the Energy Frontier,'' 
Proc. IPAC2018, Vancouver.
}

\note{(Giulio)}

\comm{In reviewing the figures of your paper, we note that the following
changes would be needed in order for your figures to conform to the
style of the Physical Review.  Please check all figures for the
following problems and make appropriate changes in the text of the
paper itself wherever needed for consistency.

Figure(s) [3,13,21,25,29]
              The abbreviation "a.u." is used to indicate atomic units only;
          arbitrary units are abbreviated "arb.  units"; "A.U." is
          for astronomical units; "amu" is for atomic mass units.
          Please amend your figures and manuscript text accordingly.

Figure(s) [4]
               The size and font of the letters and numbers appearing in
          figures should be uniform (or close to it).  Please make the
          letters and numbers more consistent.

Figure(s) [4,11]
              Please change E1, E2, etc., to power of 10 notation (i.e.,
          10 with appropriate superscript).  We do not print computer
          notation.

Figure(s) [9,10,16,23,26]
              Please modify quantities so that superscripts and subscripts are
          set as they would appear in text. Subscripts should be printed
          below the line. Also, we do not print computer notation. For
          example, set powers as superscript numbers; do not use ^2,
          **2, etc., for squares of quantities. Please adjust the text
          of the paper accordingly.

Figure(s) [9,10,16,23,26]
               The lettering in the axis labels and/or numbering size should
          be increased.

Figure(s) [13,21,25,29]
               Please rearrange power of 10 in axis label for clarity:  Either
          (i) place the power of 10 as a factor, without parentheses, in
          front of the axis label quantity, changing the sign of the power
          as needed; or (ii) incorporate the power of 10 in the topmost
          or rightmost number on the scale.  Please refer to the URL
          http://journals.aps.org/authors/axis-labels-and-scales-on-graphs-h18
          for a pictorial representation of the preferred forms for
          axis labels.

Figure(s) [23,26]
              Please insert post-decimal zeros (1.0, not 1.) or delete the
          decimal point.
        }

\note{(Miriam)}

TITLE:

This journal does print title footnotes.  Move informatio into your
acknowledgments section.

The following problems were noted in your manuscript.

* Please check the changes made to the title of your manuscript.
  We propose a modified and somewhat shorter title.

----------------------------------------------------------------------
Report of Referee A -- ZR10120/Fitterer
----------------------------------------------------------------------

REFEREE A - ZR10120

The paper describes quite extensive simulation and experimental work 
in estimating the effects of either resonant or random dipole 
excitations on the LHC proton beam at injection. Quantities studied 
include beam loss, transverse beam profiles, and bunch length. The 
studies are motivated by the planned use of a hollow electron lens in 
the HL-LHC in pulsed mode (either resonantly with the betatron tune, 
or randomly). The pulsed mode should deplete the beam halo but not 
affect the beam core.

The paper is well written, with an excellent introduction, high 
quality tables and figure (a few issues are mentioned below), and very 
few typos. It reports in substantial detail on work performed in 2016 
and 2017.

My main criticism is with the discussion and conclusion. In this 
section, the previous sections are summarized as one would expect, but 
the authors may be able to come to more pronounced conclusions. The 
following sentence is somewhat disappointing after such a large body 
of work: "However, further work is needed to identify, both in 
simulations and experimentally, candidate excitation patterns that are 
effective for halo removal and that preserve the beam core ...".

Would it be possible, for example, to summarize the beam losses of all 
cases studied experimentally and in simulation in a single table? What 
conclusions could be drawn from such a direct comparison for the cases 
studied? Are any of these excitation pattern viable for a hollow 
electron lens in the HL-LHC - which was the motivation for this work 
after all?

Typos and minor comments:

- Table I: "... for past, present and future colliders." Here the 
authors only show the colliders at their labs. The Tevatron is really 
only an example of 3 colliders with comparable stores energy, and the 
ISR (not mentioned at all) had up to several MJ of stored energy too.

- Table III: "$\mu perv$" is not an SI unit. Use 
"$\mu$\text{AV}^{-2/3}$".

- Table IV: MOF and MOD are not defined

- line 472: "can thus attributed" => "can thus be attributed"

- all figures: Since the paper reports on both measurements and 
simulations it would help the reader if the captions would start with 
either "Measurement ..." or "Simulation of ..." or "Calculation of 
..."

- Figure 8: Can "random" be a dashed line for better distinction? 
Also, the emittance growth for "random" should be calculable (e.g. 
Handbook 1999, Syphers, Sec. 4.5.6 "Emittance growth").

- Fig. 30, 31: it is difficult to distinguish the symbols for "damper 
on" and "damper off"

- Ref. [15] - A better or additional reference may be X. Gu, PRAB 20, 
023501, 2017.

- Ref. [41] is not used in the text - there may be more unused 
references

- there may be more references on emittance growth - although I could 
not find a specific one on resonant excitation, as the authors have 
pointed out

----------------------------------------------------------------------
Report of Referee B -- ZR10120/Fitterer
----------------------------------------------------------------------

REFEREE B - ZR10120

In this article various types of beam excitations are numerically and 
experimentally investigated, which act on proton losses, beam 
emittances, and beam distributions in the LHC. These studies are 
important for the design of hollow electron lenses (HEL) for active 
beam halo control of HL-LHC, where a considerable amount of energy 
will be stored in the beam tails.

Given the need of active halo control for HL-LHC and for future 
high-power accelerators such as HE-LHC and FCC-hh these investigations 
are very essential. The experimental work and simulation results 
presented in this article are excellent. In my view, the article 
contains too many technical details, especially too many excitation 
schemes that lead to confusing plots (see Figs. 17, 18, 22, 27, 28 and 
32), with many detailed results that may not be of great interest for 
the general audience. This detailed information will certainly be very 
useful for colleagues working on high-power hadron colliders and 
should be complied in a technical note or elsewhere. Also in some of 
the FMA diagrams (see Figs. 9, 10 and 16) the difference between the 
various excitation modes are barely visible.

Sometimes plot sizes are just too small and in other cases, the 
difference between various settings seem to be marginal and does not 
contain any further findings.

Since this work is very relevant for the design and upgrade of future 
high-power hadron accelerators and the performed work is excellent, I 
suggest reducing the length and content of the paper to a more compact 
and comprehensible level. In particular, the number of excitation 
schemes displayed in individual plots should be reduced to a level 
that the content of the various plots is clearly recognizable.

\end{document}

### 2018-09-15 response.tex
Formatted response letter and updated gitignore

@@
 ,amsfonts,amssymb,amsmath
-%,floatfix
+% ,floatfix
 ]{revtex4-1}
@@
 \usepackage[T1]{fontenc}
+
+\setlength{\parindent}{0ex}
+\setlength{\parskip}{0.5\baselineskip}
 
@@
 
-\title{Responses to the comments of editors and referees}
+Manuscript ZR10120/Fitterer, submitted to Phys. Rev. Accel. Beams  \hfill \today
 
-\date{\today}
-
-\preprint{Phys. Rev. Accel. Beams, manuscript ZR10120 Fitterer}
-
-\maketitle
+\section*{Responses to the comments of editors and referees}
 
@@
 
+Best regards,
+
+Giulio Stancari\\
+\emph{Corresponding author}\\
+\texttt{stancari@fnal.gov}
+
 \section*{Editorial comments}
 
-\comm{
-We also suggest that Reference [5] should be updated to a journal article:
+{\em We [\ldots] suggest that Reference [5] should be updated to a journal
+  article:
 
-[5] M. Benedikt, F. Zimmermann, ``Towards future circular colliders,''
-Journal of the Korean Physical Society (2016)
+  [5] M. Benedikt, F. Zimmermann, ``Towards future circular
+  colliders,'' Journal of the Korean Physical Society (2016)
 
@@
 
-[5b] M. Benedikt and F. Zimmermann,
-``FCC: Colliders at the Energy Frontier,'' 
-Proc. IPAC2018, Vancouver.
+  [5b] M. Benedikt and F. Zimmermann, ``FCC: Colliders at the Energy
+  Frontier,'' Proc. IPAC2018, Vancouver.
 }
 
-\note{(Giulio)}
+\section*{Problems with manuscript}
 
-\comm{In reviewing the figures of your paper, we note that the following
-changes would be needed in order for your figures to conform to the
-style of the Physical Review.  Please check all figures for the
-following problems and make appropriate changes in the text of the
-paper itself wherever needed for consistency.
+{\em In reviewing the figures of your paper, we note that the
+  following changes would be needed in order for your figures to
+  conform to the style of the Physical Review.  Please check all
+  figures for the following problems and make appropriate changes in
+  the text of the paper itself wherever needed for consistency.
 
-Figure(s) [3,13,21,25,29]
-              The abbreviation "a.u." is used to indicate atomic units only;
-          arbitrary units are abbreviated "arb.  units"; "A.U." is
-          for astronomical units; "amu" is for atomic mass units.
-          Please amend your figures and manuscript text accordingly.
+\begin{description}
 
-Figure(s) [4]
-               The size and font of the letters and numbers appearing in
-          figures should be uniform (or close to it).  Please make the
-          letters and numbers more consistent.
+\item[Figure(s) 3, 13, 21, 25, 29] The abbreviation ``a.u.'' is used
+  to indicate atomic units only; arbitrary units are abbreviated
+  ``arb.  units''; ``A.U.'' is for astronomical units; ``amu'' is for
+  atomic mass units.  Please amend your figures and manuscript text
+  accordingly.
 
-Figure(s) [4,11]
-              Please change E1, E2, etc., to power of 10 notation (i.e.,
-          10 with appropriate superscript).  We do not print computer
-          notation.
+\item[Figure(s) 4] The size and font of the letters and numbers
+  appearing in figures should be uniform (or close to it).  Please
+  make the letters and numbers more consistent.
 
-Figure(s) [9,10,16,23,26]
-              Please modify quantities so that superscripts and subscripts are
-          set as they would appear in text. Subscripts should be printed
-          below the line. Also, we do not print computer notation. For
-          example, set powers as superscript numbers; do not use ^2,
-          **2, etc., for squares of quantities. Please adjust the text
-          of the paper accordingly.
+\item[Figure(s) 4, 11] Please change E1, E2, etc., to power of 10
+  notation (i.e., 10 with appropriate superscript).  We do not print
+  computer notation.
 
-Figure(s) [9,10,16,23,26]
-               The lettering in the axis labels and/or numbering size should
-          be increased.
+\item[Figure(s) 9, 10, 16, 23, 26] Please modify quantities so that
+  superscripts and subscripts are set as they would appear in
+  text. Subscripts should be printed below the line. Also, we do not
+  print computer notation. For example, set powers as superscript
+  numbers; do not use \verb!^2!, \verb!**2!, etc., for squares of
+  quantities. Please adjust the text of the paper accordingly.
 
-Figure(s) [13,21,25,29]
-               Please rearrange power of 10 in axis label for clarity:  Either
-          (i) place the power of 10 as a factor, without parentheses, in
-          front of the axis label quantity, changing the sign of the power
-          as needed; or (ii) incorporate the power of 10 in the topmost
-          or rightmost number on the scale.  Please refer to the URL
-          http://journals.aps.org/authors/axis-labels-and-scales-on-graphs-h18
-          for a pictorial representation of the preferred forms for
-          axis labels.
+\item[Figure(s) 9, 10, 16, 23, 26] The lettering in the axis labels
+  and/or numbering size should be increased.
 
-Figure(s) [23,26]
-              Please insert post-decimal zeros (1.0, not 1.) or delete the
-          decimal point.
+\item[Figure(s) 13, 21, 25, 29] Please rearrange power of 10 in axis
+  label for clarity: Either (i)~place the power of 10 as a factor,
+  without parentheses, in front of the axis label quantity, changing
+  the sign of the power as needed; or (ii)~incorporate the power of 10
+  in the topmost or rightmost number on the scale.  Please refer to
+  the URL
+  \url{http://journals.aps.org/authors/axis-labels-and-scales-on-graphs-h18}
+  for a pictorial representation of the preferred forms for axis
+  labels.
+
+\item[Figure(s) 23, 26] Please insert post-decimal zeros (1.0, not 1.)
+  or delete the decimal point.
+
+\end{description}
 }
 
-\note{(Miriam)}
-
+{\em
 TITLE:
@@
 
-* Please check the changes made to the title of your manuscript.
+\begin{itemize}
+\item Please check the changes made to the title of your manuscript.
   We propose a modified and somewhat shorter title.
+\end{itemize}
+}
 
-----------------------------------------------------------------------
-Report of Referee A -- ZR10120/Fitterer
-----------------------------------------------------------------------
+\section*{Report of Referee A}
 
-REFEREE A - ZR10120
-
+{\em
 The paper describes quite extensive simulation and experimental work 
@@
 quality tables and figure (a few issues are mentioned below), and very 
-few typos. It reports in substantial detail on work performed in 2016 
-and 2017.
+few typos. It reports in substantial detail on work performed in~2016 
+and~2017.
 
@@
 following sentence is somewhat disappointing after such a large body 
-of work: "However, further work is needed to identify, both in 
+of work: ``However, further work is needed to identify, both in 
 simulations and experimentally, candidate excitation patterns that are 
-effective for halo removal and that preserve the beam core ...".
+effective for halo removal and that preserve the beam core ...''.
 
@@
 studied? Are any of these excitation pattern viable for a hollow
-electron lens in the HL-LHC - which was the motivation for this work 
+electron lens in the HL-LHC --- which was the motivation for this work
 after all?
@@
 
-- Table I: "... for past, present and future colliders." Here the 
+\begin{itemize}
+
+\item Table I: ``... for past, present and future colliders.'' Here the 
 authors only show the colliders at their labs. The Tevatron is really 
-only an example of 3 colliders with comparable stores energy, and the 
+only an example of 3~colliders with comparable stores energy, and the 
 ISR (not mentioned at all) had up to several MJ of stored energy too.
 
-- Table III: "$\mu perv$" is not an SI unit. Use 
-"$\mu$\text{AV}^{-2/3}$".
+\item Table III: ``\verb!$\mu perv$!'' is not an SI unit. Use 
+``\verb!$\mu$\text{AV}^{-2/3}$!''.
 
-- Table IV: MOF and MOD are not defined
+\item Table IV: MOF and MOD are not defined
 
-- line 472: "can thus attributed" => "can thus be attributed"
+\item line 472: ``can thus attributed'' => ``can thus be attributed''
 
-- all figures: Since the paper reports on both measurements and 
+\item all figures: Since the paper reports on both measurements and 
 simulations it would help the reader if the captions would start with 
-either "Measurement ..." or "Simulation of ..." or "Calculation of 
-..."
+either ``Measurement ...'' or ``Simulation of ...'' or ``Calculation of 
+...''
 
-- Figure 8: Can "random" be a dashed line for better distinction? 
-Also, the emittance growth for "random" should be calculable (e.g. 
-Handbook 1999, Syphers, Sec. 4.5.6 "Emittance growth").
+\item Figure~8: Can ``random'' be a dashed line for better distinction? 
+Also, the emittance growth for ``random'' should be calculable (e.g. 
+Handbook 1999, Syphers, Sec. 4.5.6 ``Emittance growth'').
 
-- Fig. 30, 31: it is difficult to distinguish the symbols for "damper 
-on" and "damper off"
+\item Fig.~30, 31: it is difficult to distinguish the symbols for ``damper 
+on'' and ``damper off''
 
-- Ref. [15] - A better or additional reference may be X. Gu, PRAB 20, 
+\item Ref.~[15] --- A better or additional reference may be X.~Gu, PRAB 20, 
 023501, 2017.
 
-- Ref. [41] is not used in the text - there may be more unused 
+\item Ref.~[41] is not used in the text --- there may be more unused 
 references
 
-- there may be more references on emittance growth - although I could 
+\item there may be more references on emittance growth --- although I could 
 not find a specific one on resonant excitation, as the authors have 
@@
 
-----------------------------------------------------------------------
-Report of Referee B -- ZR10120/Fitterer
-----------------------------------------------------------------------
+\end{itemize}
+}
 
-REFEREE B - ZR10120
+\section*{Report of Referee B}
 
-In this article various types of beam excitations are numerically and 
-experimentally investigated, which act on proton losses, beam 
+{\em
+
+  In this article various types of beam excitations are numerically
+  and experimentally investigated, which act on proton losses, beam
   emittances, and beam distributions in the LHC. These studies are
@@
   Given the need of active halo control for HL-LHC and for future
-high-power accelerators such as HE-LHC and FCC-hh these investigations 
-are very essential. The experimental work and simulation results 
-presented in this article are excellent. In my view, the article 
-contains too many technical details, especially too many excitation 
-schemes that lead to confusing plots (see Figs. 17, 18, 22, 27, 28 and 
-32), with many detailed results that may not be of great interest for 
-the general audience. This detailed information will certainly be very 
-useful for colleagues working on high-power hadron colliders and 
-should be complied in a technical note or elsewhere. Also in some of 
-the FMA diagrams (see Figs. 9, 10 and 16) the difference between the 
-various excitation modes are barely visible.
+  high-power accelerators such as HE-LHC and FCC-hh these
+  investigations are very essential. The experimental work and
+  simulation results presented in this article are excellent. In my
+  view, the article contains too many technical details, especially
+  too many excitation schemes that lead to confusing plots (see
+  Figs.~17, 18, 22, 27, 28 and~32), with many detailed results that
+  may not be of great interest for the general audience. This detailed
+  information will certainly be very useful for colleagues working on
+  high-power hadron colliders and should be complied in a technical
+  note or elsewhere. Also in some of the FMA diagrams (see Figs.~9, 10
+  and 16) the difference between the various excitation modes are
+  barely visible.
 
@@
 
-Since this work is very relevant for the design and upgrade of future 
-high-power hadron accelerators and the performed work is excellent, I 
-suggest reducing the length and content of the paper to a more compact 
-and comprehensible level. In particular, the number of excitation 
-schemes displayed in individual plots should be reduced to a level 
-that the content of the various plots is clearly recognizable.
+  Since this work is very relevant for the design and upgrade of
+  future high-power hadron accelerators and the performed work is
+  excellent, I suggest reducing the length and content of the paper to
+  a more compact and comprehensible level. In particular, the number
+  of excitation schemes displayed in individual plots should be
+  reduced to a level that the content of the various plots is clearly
+  recognizable.
+}
 

### 2018-09-15 resex.tex
Removed unused figures. Removed line numbers.
Updated FCC references and table of collider parameters

@@
     \begin{tabular}{lccccc}
-      Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{Wenninger:IPAC:2017}	& LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{fcc_param_2017} \\
+      Collider& Tevatron (protons) \cite{tevatron} & LHC 2016
+                                                     \cite{Wenninger:IPAC:2017}
+      & LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{Benedikt:JKPS:2016, Benedikt:IPAC:2018} \\
       \colrule
       Beam energy [TeV] & 0.98 & 6.5 & 7.0 & 7.0 & 50.0\\
-      Number of bunches & 36 & 2220 & 2808 & 2748 & $\sim$10500 \\
+      Number of bunches & 36 & 2220 & 2808 & 2748 & 10600 \\
       Number of particles per bunch & $2.90\times 10^{11}$ & $1.15\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
-      Stored beam energy [MJ] & 1.6 & 265.9 & 362.2 & 678.0 & 8400 \\
+      Stored beam energy [MJ] & 1.6 & 265.9 & 362.2 & 678.0 & 8480 \\
     \end{tabular}

### 2018-09-15 response.tex
Modified title and acknowledgments. Updated response letter.

@@
 \newcommand{\code}[1]{\textsc{#1}} % computer code
-\newcommand{\comm}[1]{\emph{#1}} % editorial or referee comment
 \newcommand{\note}[1]{\emph{\textcolor{red}{#1}}} % note, to-do
+
+% environment for editorial or referee comments
+\newenvironment{co}{\em}{}
+\newcommand{\bco}{\begin{co}}
+\newcommand{\eco}{\end{co}}
+
+% environment for responses
+\newenvironment{re}{\color{NavyBlue}}{}
+\newcommand{\bre}{\begin{re}}
+\newcommand{\ere}{\end{re}}
 
@@
 
-{\em We [\ldots] suggest that Reference [5] should be updated to a journal
-  article:
+\bco
+We [\ldots] suggest that Reference [5] should be updated to a
+journal article:
 
-  [5] M. Benedikt, F. Zimmermann, ``Towards future circular
-  colliders,'' Journal of the Korean Physical Society (2016)
+[5] M. Benedikt, F. Zimmermann, ``Towards future circular colliders,''
+Journal of the Korean Physical Society (2016)
 
@@
 Frontier,'' Proc. IPAC2018, Vancouver.
-}
+\eco
+
+\bre
+The references were updated.
+\ere
 
@@
 
-{\em In reviewing the figures of your paper, we note that the
+\bco In reviewing the figures of your paper, we note that the
   following changes would be needed in order for your figures to
@@
   figures for the following problems and make appropriate changes in
-  the text of the paper itself wherever needed for consistency.
+  the text of the paper itself wherever needed for consistency.\eco
 
@@
 
+\bco
 \item[Figure(s) 3, 13, 21, 25, 29] The abbreviation ``a.u.'' is used
@@
   accordingly.
+\eco
 
+\note{(Miriam.)}
+
+\bco
 \item[Figure(s) 4] The size and font of the letters and numbers
@@
   make the letters and numbers more consistent.
+\eco
 
+\bre Done.\ere
+
+\bco
 \item[Figure(s) 4, 11] Please change E1, E2, etc., to power of 10
@@
   computer notation.
+\eco
 
+\bre Done for Fig.~4. \ere \note{Miriam, could you please take care of Fig.~11?}
+
+\bco
 \item[Figure(s) 9, 10, 16, 23, 26] Please modify quantities so that
@@
   quantities. Please adjust the text of the paper accordingly.
+\eco
 
+\note{This comment is a bit extreme. Miriam, could you emphasize the
+  difference between text and subscripts?}
+
+\bco
 \item[Figure(s) 9, 10, 16, 23, 26] The lettering in the axis labels
   and/or numbering size should be increased.
+\eco
 
+\note{Miriam.}
+
+\bco
 \item[Figure(s) 13, 21, 25, 29] Please rearrange power of 10 in axis
@@
   labels.
+\eco
 
+\note{Miriam.}
+
+\bco
 \item[Figure(s) 23, 26] Please insert post-decimal zeros (1.0, not 1.)
   or delete the decimal point.
+\eco
+
+\note{Miriam.}
 
 \end{description}
-}
 
-{\em
+\bco
 TITLE:
@@
 acknowledgments section.
+\eco
 
+\bre Done.\ere
+
+\bco
 The following problems were noted in your manuscript.
@@
 \end{itemize}
-}
+\eco
+
+\bre We slightly shortened the title. However, the authors feel that
+``active halo control'' is an important concept that should be
+emphasized in the title. \ere
 
@@
 
-{\em
+\bco
 The paper describes quite extensive simulation and experimental work 
@@
 \end{itemize}
-}
+\eco
 
@@
 
-{\em
+\bco
 
@@
   recognizable.
-}
+\eco
 

### 2018-09-15 resex.tex
Modified title and acknowledgments. Updated response letter.

@@
 \title{Effects of resonant and random excitations on the proton beam
-  in the Large Hadron Collider, with applications to the design of
-  pulsed hollow electron lenses for active halo control}
-
-\thanks{This manuscript has been authored by Fermi Research Alliance,
-  LLC under Contract No.~DE-AC02-07CH11359 with the U.S.\ Department
-  of Energy (DOE), Office of Science, Office of High Energy
-  Physics. This work was partially supported by the U.S.\ DOE LHC
-  Accelerator Research Program (LARP), by the European FP7 HiLumi LHC
-  Design Study, Grant Agreement 284404, and by the High Luminosity LHC
-  (HL-LHC) Project.}
+  in the Large Hadron Collider with applications to active halo
+  control with pulsed hollow electron lenses}
 
@@
 
-  This work was made possible by the High Luminosity LHC (HL-LHC)
-  Project, the HiLumi LHC Design Study, and by the United States
-  Department of Energy LHC Accelerator Research Program (LARP).
+  This manuscript has been authored by Fermi Research Alliance,
+  LLC under Contract No.~DE-AC02-07CH11359 with the U.S.\ Department
+  of Energy (DOE), Office of Science, Office of High Energy
+  Physics. This work was partially supported by the U.S.\ DOE LHC
+  Accelerator Research Program (LARP), by the European FP7 HiLumi LHC
+  Design Study, Grant Agreement 284404, and by the High Luminosity LHC
+  (HL-LHC) Project.
 \end{acknowledgments}

### 2018-09-16 response.tex
Started responses to Referee A.

@@
 effective for halo removal and that preserve the beam core ...''. 
+\eco
 
+\bre We agree that the sentence may sound too weak. We wanted to
+stress that this work focuses on core effects, and further studies are
+needed on halo removal. The sentence was reworded to emphasize the
+conclusions of these studies.  \ere
+
+\bco
 Would it be possible, for example, to summarize the beam losses of all
@@
 after all?
+\eco
 
+\note{Giulio, Miriam}
+
+\bco
 Typos and minor comments:

### 2018-09-16 resex.tex
Started responses to Referee A.

@@
 operation, may significantly extend the achievable range of halo
-removal rates. However, further work is needed to identify, both in
-simulations and experimentally, candidate excitation patterns that are
-effective for halo removal and that preserve the beam core (such as
-\eighthtp\ in these studies). Because of the complex dynamics,
-tolerances on residual kicks vary widely, from a few nanoradians for
-the most powerful (such as random at full strength or \seventhtp) to
-about a hundred nanoradians for the more benign (such as
-\eighthtp). The magnitude of the acceptable kicks depends on the
-excitation pattern, on machine lattice, and on the type of application
-(fast scraping before collisions vs.\ long-term tail suppression
-during the course of a physics fill, for instance). Of course,
-sensitivity to the experimental conditions may make resonant patterns
-less attractive than a small random noise component added to a
-constant kick.
+removal rates.  In these studies, we identified candidate excitation
+patterns (such as \eighthtp) that preserve the beam core, and their
+effects were quantified. As mentioned above, complementary
+calculations and experiments are needed to evaluate their efficacy for
+halo removal. Because of the complex dynamics, tolerances on residual
+kicks vary widely, from a few nanoradians for the most powerful (such
+as random at full strength or \seventhtp) to about a hundred
+nanoradians for the more benign (such as \eighthtp). The magnitude of
+the acceptable kicks depends on the excitation pattern, on machine
+lattice, and on the type of application (fast scraping before
+collisions vs.\ long-term tail suppression during the course of a
+physics fill, for instance). Of course, sensitivity to the
+experimental conditions may make resonant patterns less attractive
+than a small random noise component added to a constant kick.
 

### 2018-09-16 resex.tex
Continued with corrections from Referee A: colliders, perveance, MOF/MOD, thus be.

@@
 \begin{table*}
-  \caption{ Stored beam energy for different past, present and future
-    colliders. Each new machine represents a leap in stored beam
+  \caption{Stored beam energy for a few examples of past, present and future
+    colliders. New machines represents a leap in stored beam
     energy.}
@@
       Peak yield $I_e$ at 10~keV & 5.0 & A\\
-      Gun perveance, $P$ & 5 & $\mu$perv\\
+      Gun perveance, $P$ & 5 & $\mathrm{\mu A / V^{3/2}}$\\
       Inner/outer cathode radii, $R_1/R_2$ & 6.75/12.7 & mm\\
@@
     excitation is abbreviated as H for horizontal, V for vertical and
-    H+V for horizontal and vertical at the same time.}
+    H+V for horizontal and vertical at the same time. MOF and MOD
+    refer to the focusing and defocusing Landau octupoles.}
   \label{tab:md_param}
@@
 patterns do not exhibit any increase in losses or emittance growth
-without magnetic errors. Their effect can be thus attributed to an
+without magnetic errors. Their effect can thus be attributed to an
 interaction of the excitation with the machine errors, implying also a
